### 2016-07-18 QPLIB-1.0.tex
hello world

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%  QPLIB-1.0.tex
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
\section{Introduction}\label{sec:intro}

Quadratic Programming (QP) problems, where both the objective function and the constraints contain (at most) expressions in the variables of degree $2$, contain a surprisingly diverse set of rather different problems. According to the fine details of the formulation, solving a QP may require employing either fundamentally combinatorial techniques, or ideas rooted in nonlinear optimization principles, or a mix of the two. In this sense, QP is likely one of the classes of problems where the collaboration between the communities interested in combinatorial and nonlinear optimization is more necessary, and potentially fruitful.

However, this diversity also implies that QP means very different things to different researchers. It is perhaps therefore not surprising that, unlike for ``simpler'' problems classes \cite{Koch2011}, so far there has never been a single library containing all different kinds of instances of QP. Several libraries devoted to special cases of QP are indeed available; however, each of them is either focussed on one application (a specific problem that can be modeled as QP), or on QPs with specific structural properties that make them suitable to be solved with some given class of algorithmic approaches. To the best of our knowledge,  collecting a set of instances of QP that is at the same time not overwhelmingly numerous and significant for the many different interested communities has not been attempted, yet. This work constitutes a first step in this direction.

In this paper we report the steps that have been done to collect a (hopefully) significant library of QP instances, filtering the large set of available (or specifically provided) instances in order to end up with a manageable set that still contains a meaningful sample of all possible QP types. A particularly thorny issue in this process is how to select instances that are ``interesting''. Our solution is to take this to mean ``challenging for a significant set of solution methods''. Our filtering process has then been in part based on the idea that if a significant fraction of the solvers that can solve a QP instance do so in a ``short'' time, then the instance is not challenging enough to be included in the library. This also takes into account the fact that if very few (maybe one) solver can solve it very efficiently by exploiting some specific structure, but most other approaches cannot, then the instance can still be deemed ``interesting''. Putting this approach in practice requires a nontrivial number of technical steps and decisions that are detailed in the paper. We hope that our work can provide useful guidelines for other interested researchers.

A consequence of our focus is that this paper is \emph{not} concerned about the different performance of the very diverse QP solvers; we will \emph{not} report any data comparing them. The only reason why solvers are used (and, therefore, described) in this context is to ensure that the instances of the library are nontrivial at least for a significant fraction of the current solution methods; providing guidance about which solver is most suited to some specific class of QPs is entirely outside the scope of our work.

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
\subsection{Motivation}\label{subsec:motiv}

\framebox{TASK 2 : write motivation}

Optimization problems with quadratic constraints and/or objective function (QP) have been the subject of a considerable amount of research in the last decade. At least some of the rationale for this interest is likely due to the fact that they are the ``least nonlinear nonlinear problems''. Hence, in particular for the convex case, tools and techniques that have been honed during decades of research for Linear Programming (LP), typically with integrality constraints (MILP), can often be extended to the quadratic case with at least less effort than what would be required for tackling general Non Linear Programming (NLP) problems, without or with integrality constraints (MINLP). This has indeed happened over and over again with different algorithmic techniques, such as Interior Point methods, active-set methods (of which the simplex method is a prototypical example), enumeration methods, cut-generation techniques, reformulation techniques, and many others ({\bf citations?}). Similarly, nonconvex continuous QP are perhaps the ``simplest'' class of problems that require features like spatial enumeration techniques to be solved. Hence, they are both the natural basis for the development of general techniques for nonconvex NLP, and a very specific class so that specialized approaches can be developed ({\bf citations, like copositive/standard QP?}).

On the other hand, (MI)QP is, in some sense, a considerably more expressive class than (MI)LP. Quadratic expressions are found, either naturally or after appropriate reformulations, in very many practical problems ({\bf citations?}). In general, any continuous function can be approximated with arbitrary accuracy (over a compact set) by a polynomial of arbitrary degree; in turn, every polynomial can be broken down to a system of quadratic expressions. Hence, (MI)QP is, in some sense, roughly as expressive as the whole of (MI)NLP. Of course this is, in principle, true for (MI)LP as well, but at the cost of much larger and much more complex formulations ({\bf citations?}). Hence, for many applications QP may represent the ``sweet spot'' between the effectiveness, but lower expressive power, of (MI)LP and the higher expressive power, but much higher computational cost, of (MI)NLP.

\ldots

The structure of this paper is the following. In \S \ref{sec:QPbasic} we review the basic notions about QP. In particular, \S \ref{subsec:notation} sets out the notation, \S \ref{sec:classification} proposes a---to the best of our knowledge, new---taxonomy of QP that helps us in discussing the (very) different classes of QPs, \S \ref{sec:algo} very briefly reviews the solution methods for QP upon which the solvers we have employed are based, and \S \ref{subsec:solver} describes the solvers. Then, \S \ref{sec:lib} describes the process used to obtain the library and its results; the software tools that we have used, and that are freely released together with the library, are discussed in \S \ref{subsec:tools}. Some conclusions are drawn in \S \ref{sec:conclusions}, after which in the Appendix a complete description of all the instances of the library is provided.

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%  End QPLIB-1.0.tex
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

### 2016-07-18 QPLIB-2.1.tex
hello world

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%  QPLIB-2.1.tex
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
\section{Quadratic Programming in a nutshell}\label{sec:QPbasic}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
\subsection{Notation}\label{subsec:notation}

In mathematical optimization, a Quadratic Program (QP) is an optimization problem in which either the objective function or some of the constraints are quadratic functions. More specifically, the problem has the form
%
\begin{align*}	
 \min \;\;
 & x^\top Q^0 x + L^0 x \\
 %
 & x^\top Q^i x + L^i x \leq C^i & i \in \Mcm \\ 
 %
 & l_j \leq x_j \leq u_j & j \in \Mcn  \\
 %
 & x_j \in \mathbb{Z} & j \in \Mcz
\end{align*}
%
where:
%
\begin{itemize}
 \item $\Mcn = \{ 1, \ldots, n \}$ is the set of (indices) of variables, and $\Mcm = \{ 1, \ldots, m \}$ is the set of (indices) of constraints;
 %
 \item $x = [x_j]_{j = 1}^n \in \RR^n$ is a finite vector of real variables;
 %
 \item $Q^i$ for $i \in \{ 0 \} \cup \Mcm$ are symmetric $n \times n$ real matrices: because one is always only interested in the value of quadratic functions of the type $x^\top Q^i x$, symmetry can be assumed without loss of generality by just replacing both $Q^i_{hk}$ and $Q^i_{kh}$ with their average $(Q^i_{hk} + Q^i_{kh}) / 2$;
 %
 \item $L^i$ and $C^i$ for $i \in \{ 0 \} \cup \Mcm$ are, respectively, real $n$-vectors and real constants;
 %
 \item $-\infty \leq l_j < u_j \leq \infty$ are the (extended) real upper and lower bounds on each variable $x_j$ for $j \in \Mcn$, assumed different because otherwise the variable is fixed;
 %
 \item $\Mcm = \Mcq \cup \Mcl$ where $Q^i = 0$ for all $i \in \Mcl$ (i.e., these are the linear constraints, as opposed to the truly quadratic ones);
 %
 \item the variables in $\Mcz \subseteq \Mcm$ are restricted to only attain integer values.
\end{itemize}
%
Due to the presence of integral requirements on the variables, this class of problems is often referred to as Mixed-Integer Quadratic Program (MIQP). It will be sometimes useful to refer to the (sub)set $\Mcb =  \{ \, i \in \Mcz : l_j = 0 , u_j = 1 \, \} \subseteq \Mcz$ of the binary variables, and to $\Mcr = \Mcn \setminus \Mcz$ as the set of continuous ones. Similarly, it will be sometimes useful to distinguish the (sub)set $\Mcx = \{ \, j : l_j > -\infty \lor u_j < \infty \, \}$ of the box-constrained variables from $\Mcu = \Mcn \setminus \Mcx$ of the unconstrained ones (in the sense that finite bounds are not explicitly provided in the data of the problem, although they may be implied by the other constraints).

The relative flexibility offered by quadratic functions, as opposed e.g.~to linear ones, allows several reformulation techniques to be applicable to this family of problems in order to emphasize different properties of the various components. Some of these reformulation techniques will be commented later on; here we remark that, for instance, integrality requirements, 
in particular in the form of binary variables could always be ``hidden'' 
by introducing (non convex) quadratic constraints utilizing the celebrated relationship
%in particular under the form of binary variables, could always be ``hidden'' under the presence of (nonconvex) quadratic constraints due to the celebrated relationship 
$x_j \in \{0, 1\} \iff x_j^2 = x_j$. Therefore, when discussing these problems an effort has to be made to distinguish between features that come from the original model, and those that can be introduced by reformulation techniques in order to extract (and algorithmically exploit) specific properties.

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%  End QPLIB-2.1.tex
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

### 2016-07-18 QPLIB-2.2.tex
hello world

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%  QPLIB-2.2.tex
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
\subsection{Classification}\label{sec:classification}

Despite the apparent simplicity of the definition given in \S \ref{subsec:notation}, Quadratic Programming instances can be of several rather different ``types'' in practice, depending on the fine details of the data. In particular, many algorithmic approaches can only be applied to QP when the data of the problem has specific properties. A taxonomy of QP instances should therefore strive to identify the set of properties that an instance should have in order to apply the most relevant computational methods. However, the sheer number of different existing approaches, and the fact that new ones are proposed, makes it hard to provide a taxonomy that is both simple and covers all possible special cases. This is why, in this paper, we propose an approach that aims at finding a good balance between simplicity and coverage of the main families of computational methods.

% -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -
\subsubsection{Classification}\label{ssec:taxonomy}

Our taxonomy is based on a three-fields code of the form ``\textit{FVC}'', where \textit{F} indicates the objective function, \textit{V} the variables, and \textit{C} the constraints of the problem. The fields can be given the following values:
%
\begin{itemize}
 \item objective function: (L)inear, (D)iagonal convex quadratic, (C)onvex quadra-tic, nonconvex (Q)uadratic;
 %
 \item variables: (C)ontinuous only, (B)inary only, (M)ixed binary and continuous, (I)nteger only, (G)eneral (all threee types);
 %
 \item constraints: (N)one, (B)ox, (L)inear, (C)onvex quadratic, nonconvex (Q)ua-dratic.
\end{itemize}
%
The wildcard ``*'' will be used to indicate any possible choice, and lists of the form ``\{X, Y, Z\}'' will indicate that the value of the given field can freely attain any of the specified values.

The ordering of the values in the previous lists is not irrelevant; in general, problems become ``harder'' when going from left to right. More specifically, for the \textit{F} and \textit{C} fields the order is that of \emph{strict} containment between problem classes: for instance, Linear objective functions are strictly contained in Diagonal convex quadratic ones (by just allowing the diagonal elements to be all-zero), which are strictly contained into general Convex quadratic ones (by allowing the off-diagonal elements to be all-zero), which in turn are strictly contained into general nonconvex Quadratic ones (by allowing any symmetric $Q^0$, hence possibly SDP ones as well). The only field for which the containment relationship is not a total order is \textit{V}, for which only the partial orderings
\[
 \mbox{C} \subset \mbox{M} \subset \mbox{G}
 \qquad,\qquad
 \mbox{B} \subset \mbox{M} \subset \mbox{G}
 \qquad,\qquad
 \mbox{B} \subset \mbox{I} \subset \mbox{G}
\]
hold. In the following discussion we will repeatedly exploit this by assuming that, unless otherwise mentioned, when a method can be applied to a given problem, it can be applied as well to all simpler problems where the value of each field is arbitrarily replaced with a value denoting a less general class.

% -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -
\subsubsection{Examples and reformulations}\label{ssec:reform}

We will now provide a first general discussion about the different problem classes that our proposed taxonomy defines. Some of them are actually ``too simple'' to make sense in our context. For instance, D*B problems have only diagonal quadratic (hence separable) objective function and bound constraints; as such, they read
\[
 \textstyle
 \min \; \Big\{ \,
 \sum_{j \in \Mcn} \big( \, Q^0_j x_j^2 + L^0_j x_j \, \big) \;:\;
 l_j \leq x_j \leq u_j \quad j \in \Mcn \;\;,\;\;
 x_j \in \mathbb{Z} \quad j \in \Mcz \, \Big\}
 \;\; .
\]
Hence, their solution only requires the independent minimization of a convex quadratic univariate function in each single variable $x_j$ over a box constraint and possibly integrality requirements: this can be attained trivially in $O(1)$ by closed-form formul{\ae}, projection and rounding arguments. \emph{A fortiori}, the even simpler cases L*B, D*N and L*N (the latter obviously unbounded unless $L^0 = 0$) will not be discussed here. Similarly, CCB (and, a fortiori, CCN) are immediately solved by linear algebra techniques, and therefore are of no interest in this context. On the other end of the spectrum, in general QP is a hard problem. Actually, LIQ---linear objective function and quadratic constraints in integer variables with no finite bounds, i.e. 
\[
 \textstyle
 \min \; \Big\{ \, L^0 x \;:\;
 x^\top Q^i x + L^i x \leq C^i \quad i \in \Mcm \;\;,\;\;
 x_j \in \mathbb{Z} \quad j \in \Mcn \, \Big\}
 \;\; ,
\]
is not only $\mathcal{NP}$-hard, but downright undecidable \cite{jeroslow}. Hence so are the ``harder'' \{C,Q\}IQ.

\smallskip
It is important to note that the relationships between the different classes can be somehow blurred because reformulation techniques may allow to move one instance from one class to the other. The example in the introduction, for instance, says that *M*---instances with only binary and continuous variables---can be recast as *CQ: nonconvex quadratic constraints can always take the place of binary variables. Actually, this is also true for *G* as long as $\Mcu = \emptyset$, as bounded general integer variables can be represented by binary ones.

Another relevant reformulation trick concerns the fact that, as soon as quadratic constraints are allowed, then a linear objective function can be assumed w.l.o.g.. Indeed, any Q** (C*C) problem can always be rewritten as
%
\begin{align*}	
 \min \;\;
 & x^0 \\
 %
 & x^\top Q^0 x + L^0 x \leq x^0 & \\ 
 %
 & x^\top Q^i x + L^i x \leq C^i & i \in \Mcm \\ 
 %
 & l_j \leq x_j \leq u_j & j \in \Mcn  \\
 %
 & x_j \in \mathbb{Z} & j \in \Mcz
\end{align*}
%
i.e., a L*Q (L*C). Hence, it is clear that quadratic constraints are, in a well-defined sense, the most general situation (cf.~also the result above about hardness of LIQ).

When a $Q^i$ is positive semidefinite (SDP), i.e., the corresponding constraint/objective function is convex, general quadratic constraints are in fact equivalent to diagonal ones. In fact, every SDP matrix can be factorized as $Q^i = L^i (L^i)^\top$, e.g.~by the (incomplete) Cholesky factorization, $f^i(x) = x^\top Q^i x = \sum_{j \in \Mcn} z^2_jz$ where $z = x^\top L^i$. Hence, one could think that D** problems need not be distinguished from C** ones; however, this is true only for ``complicated'' constraints, but not for ``simple'' ones, because the above reformulation technique introduces linear constraints. Indeed, while C*L (and, a fortiori, C*\{C,Q\}) can always be brought to D*L (D*\{C,Q\}), using the same technique C*B becomes D*L, which is significantly different from D*B. In practice, a diagonal convex objective function under linear constraints is found in many applications ({\bf citations?}), so that D*L still makes sense to distinguish the case where the objective function is ``naturally'' separable from that where separability is artificially introduced, although this is in theory always possible. Conversely, there is usually little gain in distinguishing between **C problems and ``**D'' ones where the constraints are restricted to be diagonal, as there are much fewer cases where this naturally occurs ({\bf to the best of my knowledge}). This is why we did not include a ``D'' option for the constraints.

% -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -
\subsubsection{QP classes}\label{ssec:classes}

The proposed taxonomy can then be used to describe the main classes of QP according to the type of algorithms that can be applied for their solution:
%
\begin{itemize}
 \item \emph{Linear Programs} LCL and \emph{Mixed-Integer Linear Programs} LGL have been subject of an enormous amount of research and have their well-established instance libraries \cite{Koch2011}, so they won't be explicitly addressed here.
 %
 \item \emph{Convex Continuous Quadratic Programs} CCC can be solved in polynomial time by Interior-Point techniques; the simpler CCL can also be solved by means of ``simplex-like'' techniques ({\bf citations?}). Actually, a slightly larger class of problems can be solved with Interior-Point methods: those that can be represented by Second-Order Cone Programs. When written as QPs the corresponding $Q^i$ may not be positive semidefinite, but still the problems can be efficiently solved. Of course, these problems can still require considerable time, like LCL, when the size of the instance grows. In this sense, like in the linear case, a significant divide is from solvers that need all the data of QP to work, and those that are ``matrix-free'', i.e., only require the solution of simple operations (typically, matrix-vector products) with the data of the problem to work ({\bf citations?}). While in our library we have never exploited such a characteristic, which is not suitable to the use of standard modeling tools, this may be relevant for the solution of very-large-scale CIC.
 %
 \item \emph{Nonconvex Continuous Quadratic Programs} QCQ are instead in general $\mathcal{NP}$-hard, even if the constraints are very specific (QCB) and only one eigenvalue of $Q^0$ is negative \cite{Hemmecke2010}. They therefore require enumerative techniques like the spatial Branch\&Bound ({\bf citations?}), to be solved to optimality. Of course, local approaches 
are available that are able to
% are available  that can be able to 
efficiently provide local optima (or at least saddle points) of the CQC, but providing global guarantees about the quality of the obtained solutions is challenging. In our library we have specifically focussed on  \emph{exact} solution of the instances.
  %
 \item \emph{Convex Integer Quadratic Programs} CGC are in general $\mathcal{NP}$-hard, and therefore require enumerative techniques to be solved. However, convexity of the objective function and constraints implies that efficient techniques (see CCC) can be used at least to solve the continuous relaxation. The general view is that CGC are not, all other things being equal, 
% not 
 substantially more difficult than LGL to solve, especially if the objective function and/or the constraints have specific properties (e.g., DGL, CGL). Often integer variables are in fact binary ones, so several CCC models are C\{B,M\}C ones. In practice binary variables are considered to lead to somewhat easier problems than general integer ones (cf.~the cited result about hardness of unbounded integer quadratic programs), and several algorithmic techniques have been specifically developed for this special case. However, the general approaches for CBC are basically the same as for CGC, so there is seldom the need to distinguish between the two classes as far as solvability is concerned, although matters can be different regarding actual solution cost. Programs with only binary variables CBC can be easier than mixed-binary or integer ones C\{M,I\}C because some techniques are specifically known for the binary-only case, cf.~the next point ({\bf citations?}). Absence of continuous variables, even in the presence of integer ones CIC, can also lead to specific techniques ({\bf citations??}).  
  %
  \item \emph{Nonconvex Binary Quadratic Programs} QB\{B, N, L\} obviously are $\mathcal{NP}$-hard. However, the special nature of binary variables combined with quadratic forms allows for quite specific techniques to be developed, among which is the reformulation of the problem as a LBL. Also, many well-known combinatorial problems can be naturally reformulated as problems of this class, and therefore a considerable number of results have been obtained by exploiting specific properties of the set of constraints ({\bf some citations? some specific problem to mention apart from Max-Cut?}).
  %
  \item \emph{Nononvex Integer Quadratic Programs} QGQ is the more general, and therefore is the most general, class. Due to the lack of convexity even when integrality requirements are removed, solution methods must typically combine several algorithmic ideas, such as enumeration (distinguishing the role of integral variables from that of continuous ones involved into nonconvex terms) and techniques (e.g., outer approximation, SDP relaxation, \ldots) that allow to efficiently compute bounds. As in the convex case, QBQ, QMQ, and QIQ can benefit from more specific properties of the variables ({\bf citations?}).
\end{itemize}
%
This description is purposely coarse; each of these classes can be subdivided into several other ones on the grounds of more detailed information about structures present in their constraints/objective function. These can have a significant algorithmic impact, and therefore can be of interest to researchers. Common structures are, e.g., network flow or knapsack-type linear constraints, semi-continuous variables, or the fact that a nonconvex quadratic objective function/constraint can be reformulated as a second-order cone (hence, convex) one. It would be rather hard to collect a comprehensive list of all types of structures that may be of interest of some researcher, since these are as varied as the different possible approaches for specialized sub-classes of QP. For this reason we do not attempt such a more refined classification, and limit ourselves to the coarser one described in this paragraph.

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%  End QPLIB-2.2.tex
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

### 2016-07-18 QPLIB-2.3.tex
hello world

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%  QPLIB-2.3.tex
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{Solution Methods}\label{sec:algo}

In this section we provide a quick overview of existing solution methods for QP, restricting ourselves to these implemented by the set of solvers considered in this paper (Table \ref{t:solvers}). For each approach we briefly describe the main algorithmic ideas and point out the formulation they address according to the classification set out in Sect.~\ref{sec:classification}. We remark that many solvers implement more than one algorithm, among which the user can choose at runtime. Moreover, algorithms are typically implemented by different solvers in different ways, so that the same conceptual algorithm can sometimes yield wildly different results or performance measures on the same instances.

% antigone/glomiqo: sBB
% baron: sBB
% couenne: sBB
% knitro: IPM with CG, SLQP, convexBB + outerapproxBB
% lindo/lindoglobal: sBB
% scip: sBB
% msnlp/oqnlp: multistart
% alphaecp: cutting plane for pseudoconvex MINLP
% bonmin: convexBB
% dicopt: alternating projection NLP/MILP
% sbb: convexBB
% conopt: GRG (active set -- like SQP)
% ipopt: IPM w/ line search filter
% lgo: Lipschitz BB, random search, multistart (default) [cassato?]
% minos: active set method
% snopt: active set SQP
% xpress: convexBB
% gurobi: convexBB
% cplex: sBB

The methods implemented by the solvers in Table~\ref{t:solvers} can, following \cite{neumaier}, be broadly organized in three categories: {\it incomplete}, {\it asymptotically complete}, and {\it complete}. Incomplete methods are only able to identify solutions, often locally optimal according to a suitable notion, and may even fail to find one even when one exists; in particular, they are typically not able to determine that an instance is empty. Asymptotically complete methods can find a globally optimal solution with probability one in infinite time, but again they cannot prove that a given instance is infeasible. Complete methods find an approximate globally optimal solution to within a prescribed optimality tolerance within finite time, or prove that none such exists (but see Sect.~\ref{s:complete} below); they are often referred to as \emph{exact} methods in the computational optimization community. According to \cite{neumaier}, {\it rigorous} methods also exist which find global within given tolerances even in the presence of rounding errors, except in ``near-degenerate cases''. Since it is debatable whether any of the solver we are using can be classified as rigorous, we have elected to limit ourselves to declaring solvers complete.

Incomplete methods are usually realized as local search algorithms, asymptotically complete methods are usually realized by meta-heuristic methods such as multi-start or simulated annealing, and complete methods for $\mathcal{NP}$-hard problems such as QP are typically realized as implicit exhaustive exploration algorithms. However, these three categories may exhibit some overlap. For example, any deterministic method for solving QCQ locally is incomplete in general, but becomes complete for CCC, since any local optimum of a convex QP is also global. Therefore, when we state that a given algorithm is incomplete or (asymptotically) complete we mean that it is so the largest problem class that the solver naturally targets, although it may be complete on specific sub-classes. For example, {\sc SnOpt} naturally targets continuous nonconvex NLPs and it is incomplete on this class, and therefore on QCQ, but becomes complete for CCC. In general, all complete methods for a problem $P$ must be complete for any problem $Q \subseteq P$, while a complete method for $P$ might be incomplete for $R \supset P$.

% -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -
\subsubsection{Incomplete methods}\label{s:incomplete}

Local search methods typically require as an input a solution $x'$ and attempt to improve it---either towards feasibility, or towards optimality, or both---using only information that is available in a neighborhood of $x'$. In general, local search methods are incomplete. As remarked above, however, local search methods deployed on a convex problem behave like complete methods, possibly via devices that that allow to find a feasible starting solution if there is one (the so-called ``phase 0'').

Most local search methods for *C* are iterative in nature, and need a feasible starting point $x^0$ as input. The ($k+1$)-st iterate is obtained as $x^{k+1} = x^k + \alpha_k d^k$, where $\alpha_k$ (a scalar) is the {\it step length}, and $d^k$ (a vector) is the {\it search direction}, e.g.~belonging to a tangent manifold and having a negative directional derivative at $x^k$, in order to improving optimality while reducing infeasibility. Local search methods for *I* are also iterative, but since defining a useful notion of descent direction is harder in presence of integer variables, $d^k$ and $\alpha_k$ are usually computed in different ways, depending on the application at hand.

The solvers in Table \ref{t:solvers} which implement incomplete methods are {\sc ConOpt}, {\sc IpOpt}, {\sc Minos}, {\sc SnOpt} and {\sc KNitro}. The former four solvers implement local search algorithms for continuous nonconvex NLPs (a problem class containing QCQ). Note that traditionally {\sc KNitro} used to be a local (nonconvex) NLP solver. All of these solvers implement essentially two types of local search methods for NLP:
%
\begin{itemize}
 \item active set methods ({\sc ConOpt}, {\sc Minos}, {\sc SnOpt});
 %
 \item interior point methods ({\sc IpOpt}, {\sc KNitro}).
\end{itemize}
%
Active set and interior point methods have been defined for *C* but also for general (continuous) NLPs. In fact, all of the solvers named above target this larger problem class.

\paragraph{Active set methods.}
%
At each iteration $k$ the algorithm forms the {\it active set} $\mathcal{A}^k$ containing the (indices of the) {\it active constraints}, i.e., all of the equality constraints as well as the inequality constraints that are satisfied at equality at the current iterate $x^k$. A subproblem, consisting of a minimization of a certain auxiliary objective function subject to all active constraints, is then solved to identify a good search direction $d^k$. An appropriate step length $\alpha_k$ is then found by checking for the inactive constraints (since these must be satisfied too). If at $x^{k+1}$ some of the previously inactive constraints have become active, or vice versa, $\mathcal{A}^k$ is updated accordingly. This general scheme works because the subproblem is defined in such a way as to be (a) easier to solve than the original problem, and (b) helping the sequence $x^k$ to converge to a Karush-Kuhn-Tucker (KKT) point.

\paragraph{Interior point methods.}
%
These approaches can be seen as recasting the original constrained problem QP as a parametrized family of unconstrained ones QP$_\mu$, where the constraints are moved in the objective function via a \emph{barrier term}---that goes to $+\infty$ as the boundary of the feasible region is approached---weighted with the \emph{barrier parameter} $\mu \in (0, \infty)$. In the convex case, the \emph{central path}---the continuous line formed from the optimal solutions of QP$_\mu$ for all varying $\mu$, typically unique e.g.~if the classical barrier based on the logarithmic function is employed---leads to a (central) optimal solution of QP when $\mu \to 0$. Starting from an appropriately constructed ``central'' point (close to the solution of QP$_\mu$ for ``large'' $\mu$), these algorithms strive to follow the central path by performing $O(1)$ Newton steps before (substantially) reducing $\mu$. The algorithms can be shown to converge in a small number of iterations, each of which can be costly due to the need of solving an appropriately modified version of the KKT conditions for QP (``slackened'' with $\mu$), a (possibly) large-scale linear system. Actually, nowadays the algorithm is most often implemented in the primal-dual version, where the nonlinear KKT system is iteratively solved with Newton-like iterations; this has the extra advantage of allowing to remove the need of a feasible (central) starting point.

% -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -
\subsubsection{Asymptotically complete methods}\label{sss:asymc}

Asymptotically complete methods do not usually require a starting point, and, if given sufficient time (infinite in the worst case) will identify a globally optimal solution with probability one. Most often, these methods are meta-heuristics, involving an element of random choice, which exploit a given (heuristic) local search procedure.

The solvers in Table \ref{t:solvers} which implement asymptotically complete methods are {\sc oqnlp}, {\sc msnlp} and certain sub-solvers of {\sc lgo}. Specifically, we consider the following methods:
%
\begin{itemize}
 \item global adaptive random search ({\sc lgo\_gars});
 %
 \item multi-start ({\sc lgo\_ms}, {\sc msnlp}, {\sc oqnlp}); specifically, the former two apply to QCQ whereas the latter to QGQ.
\end{itemize}

\paragraph{Global Adaptive Random Search.}
%
This is a modification of an algorithm called {\it pure random search}, which consists in sampling a random point $x'$ from a given compact set known to contain a global optimum, and then sampling a new candidate solution $y$ in a neighborhood of $x'$, setting $x' \leftarrow y$ if $y$ improves $x'$, and repeating as long as a termination condition is not satisfied. The adaptivity stems from changing the distribution for sampling $y$ at run-time, depending on the quality of the solutions identified by the method. Since this method only depends on sampling and function evaluation, it is usually fast. In the {\sc lgo\_gars} solver, it provides a useful starting point for a subsequent local search procedure. Asymptotic global convergence is attained by restarting the random search from different initial points $x'$.

\paragraph{Multi-start.}
%
Multi-start methods define a loop around a given local search procedure so that it starts from many different starting points, perform local search, and record the best optimum found so far as they explore the search space randomly. For example, any of the methods described in Sect.~\ref{s:incomplete} can be embedded in a multi-start framework as follows:
%
\begin{enumerate}
 \item initialize a ``best solution so far'' $x^\ast$ 
 %
 \item sample a starting point $x'$ uniformly at random from a given compact set known to contain a global optimum; \label{step:ms2}
 %
 \item run a local search method from $x'$ to yield an improved (feasible) point $x$
 %
 \item if $x$ improves on $x^\ast$ with respect to the objective function value, replace $x^\ast$ with $x$
 %
 \item repeat from Step \ref{step:ms2} until a given termination condition is satisfied. \label{step:ms5}
\end{enumerate}
%
The method is asymptotically complete if the termination condition in Step \ref{step:ms2} is a certificate of global optimality for $x^\ast$, which is usually hard to obtain. However, typically some bound on the total CPU time, or number of function evaluations, or any other criteria that makes sense for the application at hand, is needed, which renders the method incomplete.

In general, the applicability of meta-heuristics to a given problem depends 
on whether the local search they utilize
%on what whether the local search they rest on 
addresses that problem or not. 
{\bf Antonio: the concept of ``addresses that problem'' is not very clear to me, this sentence may benefit from rephrasing (or the paragraph from deleting if it's not deemed to be utterly necessary).} Depending on the local search employed, Multi-start methods can address MINLP of the most general class.

% -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -
\subsubsection{Complete methods}\label{s:complete}

Complete methods are often referred to as \emph{exact} in a large part of the mathematical optimization community. This nomenclature has to be used with care, as it implicitly makes assumptions on the underlying computational model that may not be acceptable in all cases. To see that, consider that, as already mentioned, QPs (more precisely, LIQ) are generally undecidable \cite{jeroslow}; and yet, there exists a general decision method for deciding feasibility of systems of polynomial equations and inequalities \cite{tarski-reals}, including the solution of LCQ with zero objective function. This apparent contradiction is due to the fact that the two statements refer to different computational models: the former is based on the Turing Machine (TM), whereas the latter is based on the Real RAM (RRAM) machine \cite{blum}. Due to the potentially infinite nature of exact real arithmetic computations, exact computations on the RRAM necessarily end up being approximate on the TM. Analogously, a complete method may reasonably be called ``exact'' on a RRAM; however, the computers we use in practice are more akin to TMs than RRAMs, and therefore 
calling \emph{exact} a solver that employs
%dubbing \emph{exact} a solver that employ 
floating point computations is, technically speaking, stretching the meaning of the word. However, because the term is well understood in the computational optimization community, in the following we shall loosen the distinction between complete and exact methods, with either properties intended to mean ``complete'' in the sense of \cite{neumaier}.

\paragraph{Branch-and-Bound.}
%
Nearly all of the complete solvers in Table \ref{t:solvers} that address $\mathcal{NP}$-hard problems (i.e.~those in QGQ$\smallsetminus$CCC) are based on Branch-and-Bound (BB). This is an implicit but exhaustive search process based on exploring a \emph{branching tree} of the problem, where each node in the tree represents a subset of the feasible region. Guaranteed lower and upper bounds to the objective function value relative to nodes are computed in various ways. Nodes are discarded when: (a) they can be shown to be empty; (b) their bound in the optimization direction is worse than an opposite global bound; (c) a global optimum limited to the node can be found (this happens when the two bounds are closer than a given $\varepsilon$ tolerance); (d) they are selected for branching, which means expanding the tree constructing at least two new nodes, children of the current one. Branching takes place by identifying one or more branching directions, which are usually a coordinate axes, and one or more branching point per direction, in various common sense fashions. The algorithm is driven by a queue of active nodes, usually endowed with a priority to select the most promising node from which to continue exploration of the tree (such as ``most promising bound''); the BB algorithm terminates when the queue is empty.

Typically, bounds in the optimization direction are computed by means of convex relaxations \cite{mccormick,couenne}, which replace nonconvex terms $t(x)$ with linearization variables $\hat{t}$, and then replace the corresponding defining constraints $\hat{t} = t(x)$ by means of lower and upper (respectively, convex and concave) bounding functions $\hat{t} \ge \underline{t}(x)$ and $\hat{t} \le\bar{t}(x)$. This is actually where finite (and tight) bounds on the variables are crucial, which differentiate also in practice the bounded case from the unbounded one. Different strategies are used when the nonconvexities are only quadratic \cite{glomiqo,bliek}.

When the BB algorithm is allowed to select coordinate directions corresponding to continuous variables, it is called {\it spatial} BB (sBB). Branching on continuous (rather than integer or binary) variables becomes necessary in the presence of nonconvex nonlinearities, as it happens e.g.~in QCQ, since the quality of the bounds improves as the feasible set in the current node gets smaller.

BB algorithms are exponential time in the worst case, and their exponential behavior unfortunately often shows up in practice. They can also be used heuristically (forsaking their completeness guarantee) by either terminating them early, or by using non-guaranteed bounds.

The solvers in Table \ref{t:solvers} BB type methods are:
%
\begin{itemize}
 \item {\sc Antigone}, {\sc Baron}, {\sc Couenne}, {\sc Lindo}, {\sc LindoGlobal}, {\sc SCIP}, {\sc lgo\_bb}, which implement complete BB algorithms for QGQ;
 %
\item {\sc CPLEX}, which implements a complete BB algorithm for QGL;
%
 \item {\sc KNitro\_bb}, {\sc BonMin}, {\sc sBB}, {\sc XPress}, {\sc GuRoBi}, which implement complete BB algorithms for CGC. {\bf how about {\sc mosek}?}
\end{itemize}
%
We remark that the latter category can be used as incomplete solvers for QGQ. We also remark that CPLEX can currently only target problems with linear constraints when the objective function is nonconvex \cite{bliek}.

\paragraph{Cutting plane approaches}.
%
The two remaining solvers in Table \ref{t:solvers} are {\sc AlphaECP} \cite{westerlund} and {\sc DICOpt} \cite{quesada}, which are complete cutting plane methods based on different principles. Characteristic of both solvers is that they need to employ a complete method for solving MILP (LIL) sub-problems at each iteration; in turn, this is typically based on BB, which is therefore a crucial technique also for this class of approaches. Additionally, incomplete methods can be used to provide local solutions. Other solvers, like {\sc BonMin}, also offer this kind of approach among their algorithmic options.

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%  End QPLIB-2.3.tex
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

### 2016-07-18 QPLIB-2.4.tex
hello world

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%  QPLIB-2.4.tex
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
\subsection{Solvers}\label{subsec:solver}

We now provide a succinct list of the solvers we have tested, using the approaches described in \S \ref{sec:algo}. In Table \ref{t:solvers}, we mark with ``I'' a pair (solver, problem) if the solver accepts the problem in input but it is an incomplete solver for the problem, with ``A'' if it is asymptotically complete, with ``C'' if it is complete, and leave it blank if the solver won't accept the problem in input.

{\bf The table has to be checked, as I've extrapolated from the text but I'm not 100\% sure. Also, ``?''s have to be removed.}

\begin{table}
{
 \centering                
 \scriptsize                
 \setlength{\tabcolsep}{3pt}                
% \renewcommand \arraystretch{1}

\begin{tabular}{lccccccc}
\toprule  
                 & CGL & QGL & CGC & QGQ & CCC & QCQ \\
\hline
{\sc Antigone}    &  C  &  C  &  C  &  C  &  C  &  C  \\
{\sc Baron}       &  C  &  C  &  C  &  C  &  C  &  C  \\
{\sc Couenne}     &  C  &  C  &  C  &  C  &  C  &  C  \\
{\sc KNitro}      &  C  &  I  &  C  &  I  &  C  &  A  \\
{\sc lindo api}   &  C  &  C  &  C  &  C  &  C  &  C  \\
{\sc SCIP}        &  C  &  C  &  C  &  C  &  C  &  C  \\
{\sc oqnlp}       &  A  &  A  &  A  &  A  &  C  &  A  \\
{\sc AlphaECP}    &  C  &  I  &  C  &  I  &  C  &  I  \\
{\sc BonMin}      &  C  &  I  &  C  &  I  &  C  &  I  \\
{\sc DICOpt}      &  C  &  I  &  C  &  I  &  C  &  I  \\
{\sc sBB}         &  C  &  I  &  C  &  I  &  C  &  I  \\
{\sc ConOpt}      &     &     &     &     &  C  &  I  \\
{\sc IpOpt}       &     &     &     &     &  C  &  I  \\
{\sc lgo}         &     &     &     &     &  C? &  A  \\  %try it out
{\sc Minos}       &     &     &     &     &  C  &  I  \\
{\sc msnlp}       &     &     &     &     &  C  &  A  \\
{\sc SnOpt}       &     &     &     &     &  C  &  I  \\
{\sc XPress}      &  C  &     &  C  &     &  C  &  ?  \\ %try it out
{\sc Gurobi}      &  C  &     &  C  &     &  C  &  ?  \\ %try it out
{\sc CPLEX}       &  C  &  C  &  C  &     &  C  &  I? \\ %try it out
{\sc mosek}       &  C  &     &  C  &     &  C  &     \\
\hline
\end{tabular}                 
\caption{Families of QP problems that can be tackled each solver} \label{t:solvers}             
}

\end{table}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%  End QPLIB-2.4.tex
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

### 2016-07-18 QPLIB-3.tex
hello world

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%  QPLIB-3.tex
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
\section{Library Construction}\label{sec:lib}

In this Section we present all the steps performed in order to build the library. In particular we describe the first set of gathered instances (Section \ref{subsec:instColl}), we discuss the issues concerning the format of the instances (Section \ref{subsec:format}), we present the feature used to classified the instances (Section \ref{subsec:feature}), and finally we then describe the selection process that we have used to filter the instances in order to construct the final library (Section \ref{subsec:selection}).

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
\subsection{Instance Collection}\label{subsec:instColl}

In this section we describe the procedure we adopted to gather the instances. In January $2014$, we issued an online call for instances to the main international mailing lists of the mathematical optimization and numerical analysis community, in order to reach the largest possible set of interested researchers, both in academia and in industry. The call remained open for 10 months, during which we received a large number of contributions with different characteristics. The instances we received  are both artificial and coming from real-world applications.

In addition to spontaneous contribution we scanned the other known libraries of instances and we selected all the QP ones. In particular we cite here the libraries of ``generic'' QP instances from which we draw material:

\begin{itemize}
 \item {\tt POLIP} \url{http://polip.zib.de/pipformat.php}
 %
 \item {\tt MINLP} \url{http://www.gamsworld.org/minlp/minlplib.htm}
 %
 \item {\tt MacMINLP} \url{https://wiki.mcs.anl.gov/leyffer/index.php/MacMINLP}
 %
 \item {\tt Meszaros} \url{http://www.doc.ic.ac.uk/~im/00README.QP}
 %
 \item {\tt BARON Library} \url{http://www.minlp.com/nlp-and-minlp-test-problems}
 %
 \item {\tt GAMS Library} \url{http://www.gamsworld.org/performance/performlib.htm}
\end{itemize}
%
Other QP instances were found in libraries devoted to specific optimization problems that can be modeled as QP, such as Max-Cut, the Quadratic Assignment Problem, Portfolio Optimization problems and several others; for reason of space we refrain from listing all these sources.

At the end of this process we had gathered more than eight thousand instances. Three fourths of them contained discrete variables, while the remaining ones contained only continuous variables.

\begin{center}
\begin{table}[]
 \centering

 \setlength{\tabcolsep}{5pt}
%  \arraystretch{1}
\begin{tabular}{cccc}
Starting set& \multicolumn{ 2}{c}{ $\approx$ 8500 Instances }& \\
& \multicolumn{ 2}{c}{$\Downarrow$}& \\
& $\approx$ 6000 Discr. Inst.  & $\approx$ 2500 Cont. inst. & \\
First Filter  & $\Downarrow$  &  & \\
 & $\approx$ 3000 Discr. Inst.  & $\Downarrow$ & \\
Second Filter & $\Downarrow$  &  & \\
 & 600 Discr. Inst.  & 250  Cont. inst. & \\
\end{tabular}  
%\begin{center}\end{center} 
\caption{Instance filter steps} \label{tab:filters}
\end{table}
\end{center}

\paragraph{First Instances Filter.} As already mentioned an important characteristic of an instances is its computational difficulty, i.e., the CPU time needed by a complete solver (cf.~\S \ref{sec:algo}) to solve the instance to global optimality. Accordingly, for each of the gathered instance we run the solvers in {\tt GAMS} (see Table \ref{}) able to solve it to global optimality. The number of solvers depends on the category of the instances under consideration. Then in order to reduce the number of instances we performed a first filter based on a relative measure of computational difficulty, i.e., we discarded all ``easy'' instances which are solved by at least 30\% of the complete solvers within a time limit of 30 seconds. Thanks to this first filter we obtained what we call the {\it filtered test bed}.

%In Table \ref{tab:initialSet}, we report the number of instances in the filtered set subdivided according to the classification proposed in Section \ref{sec:classification}. We present in the table two different options for the field {\it Objective Functions}, i.e., Quadratic and Convex; three different options for the field {\it Variables}, i.e., Binary, Continuous and General and three different options for the field {\it Constraints}, i.e., Linear, Quadratic and Convex. The other categories of instances were not represented in the initial test bed. In particular the gathered a large number of Quadratic Binary Linear (QBL), Quadratic Continuous Quadratic (QCQ) and Quadratic General Quadratic (QGQ) instances, i.e., 1791 instances, 1957 instances and 2524 instances respectively. We gathered 901 Convex General Convex (CGC) instances. We gathered relatively few Quadratic Binary Quadratic (QBQ), Convex Continuous Convex (CCC) and Convex Mixed Convex (CMC) instances, i.e., 175 instances, 195 instances and 201 instances respectively. Finally we gathered only 17 Quadratic Mixed Linear (QML) instances.

\bigskip
The characteristics of the instances in the final library are presented in Table \ref{tab:FinalSet-D} for \emph{discrete} instances (*\{B,M,I,G\}*) and in Table \ref{tab:FinalSet-C} for \emph{continuous} ones (*C*).

\begin{table}
 \centering
 %\scriptsize
 \setlength{\tabcolsep}{18pt}
 \renewcommand \arraystretch{1.1}
\begin{tabular}{lllr}
\toprule
Obj. Fun. & Variables & Constraints & \#\\
\cmidrule(lr){1-4}
%
\multirow{7}*{Linear}
          & Binary  & Quadratic &  71 \\
          & \multirow{2}*{Mixed}
                    & Convex    &   5 \\
          &         & Quadratic & 277 \\
%          & \multirow{2}*{Integer}
%                    & Convex    &  15 \\
%          &         & Quadratic &   5 \\
          & \multirow{2}*{General}
                    & Convex    &  22 \\
          &         & Quadratic &   8 \\
%\hline
\\
\multirow{4}*{Convex}
          & Binary  & Linear    &  12 \\
          & \multirow{2}*{Mixed}
                    & Linear    &  37 \\
          &         & Quadratic &   3 \\
          & General & Linear    &   1 \\
%\hline
\\
\multirow{7}*{Quadratic}
          & \multirow{3}*{Binary}
                    & None      &  29 \\
          &         & Linear    & 180 \\
%          &         & Quadratic &   8 \\
          & \multirow{2}*{Mixed}
                    & Linear    &  12 \\
          &         & Quadratic &   9 \\
          & Integer & Linear    &   2 \\
          & \multirow{2}*{General}
                    & Linear    &   1 \\
          &         & Quadratic &   2 \\
\hline
Total     &         &           & 600 \\
%
\bottomrule
\end{tabular}  
\label{tab:FinalSet-D}
\caption{Classification of the final set of discrete instances} 
\end{table}

\begin{table}
 \centering
 %\scriptsize
 \setlength{\tabcolsep}{18pt}
 \renewcommand \arraystretch{1.1}
\begin{tabular}{llr}
\toprule
Obj. Fun. & Constraints & \#\\
\cmidrule(lr){1-3}
%
\multirow{2}*{Linear}    & Convex    &  14 \\
                         & Quadratic &  71 \\
\multirow{3}*{Convex}    & None      &   6 \\
                         & Box       &   6 \\
                         & Linear    &  50 \\
                         & Convex    &   3 \\
                         & Quadratic &   6 \\
\multirow{3}*{Quadratic} & Box       &   8 \\
                         & Linear    &  19 \\
                         & Convex    &  27 \\
                         & Quadratic &  40 \\
\hline
Total                    &           & 250 \\
%
\bottomrule
\end{tabular}  
\label{tab:FinalSet-C}
\caption{Classification of the final set of continuous instances} 
\end{table}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
\subsection{Instance Format}\label{subsec:format}
\framebox{TASK X : write }
In the call for instances no specific formats were imposed 
%we did not impose a specific format 
for the submissions. 
%In this phase we also eliminated multiples copies of the same instance.
To evaluate the instances we decided, for practical reasons, to use GAMS as common platform for all the experiments involving commercial solvers.
For this reason, we decided to translate all instances into the \texttt{.gms} gams format.
 In a preliminary phase, all the instances received were divided according to their format and subsequently translated.
In Sect.
%~\ref{subsec:CP_convex} 
the tools used 
to translate an instance from a given format to the \texttt{.gms} format
%to translate an instance from and to a given format to the \texttt{.gms} format 
are described more in detail.\\

As second format, we introduced 
a specific format \texttt{.qplib}. This new format is capable to describe all the instances of the library in a sparse form. 
In comparison to a more \emph{high level} format like \texttt{.gms}, the new format presents two advantages:  
it is easier to read by a self-made parser and it produces smaller files.
\textcolor{red}{ToDo: describe \texttt{.qplib} format. Maybe in an appendix?}

% is lighter in terms of sie
%  significantly 
%less memory 
%in comparison and

% for this reason we used the \texttt{.gms} format as one of the format of the library. %checked if one instance was present copies of the same instance, potentially written in different formats

\subsection{Instance Features}\label{subsec:feature}
\framebox{TASK X : write }

%Characteristics that should help to analyze the instances
%
%General characteristics:

%Each QP instance can be classified accordingly to several features.
For each instance of the starting set, the following features have been collected:
\begin{itemize}
\item Objective function characteristics:
\begin{itemize}
\item Type of objective function:  Linear, Convex or Quadratic.
\item Density of the objective function, i.e. the percentage of nonzero entries of $Q_0$. 
\item Percentage of negative eigenvalues of $Q_0$.  
\end{itemize}
\item Variables characteristics:
\begin{itemize}
\item Number of Continuous, Binary and Integer variables.
%\item number of binary variables
%\item number of integer variables
\end{itemize}
\item Constraints characteristics:
\begin{itemize}
\item Number of Linear, Convex and Quadratic constraints.
\item Density of the constraints, i.e. the percentage of nonzero entries of the coefficients of the Linear, Continuous and Quadratic constraints. 
\end{itemize}
\end{itemize}

The mentioned features 
\begin{itemize}
\item constraint-types: big-M constraints? box-constraints? combinatorial constraints? linear network?
\item nonconvex: whether they can be turned into cones through appropriate decomposition
\item quadratic: are discrete variables contained in the quadratic part or in the linear part only?
\item note: some remark like "max-cut", "quadratic knapsack", etc.
\end{itemize}

Solver-dependent characteristics:
\begin{itemize}
\item time (?) [how measure time? what architecture will be used for tests? how to deal with parallel machines/algorithms?]
\item number of nodes in case of branch-and-bound algorithms
\item memory (?)
\item matrix-free access (?)
\item lower bound at root node
\item time of finding first feasible solution
\end{itemize}

\begin{itemize}
\item The static analysis is not enough to identify the hardness of the instances.
\item \textcolor{red}{An empirical way for testing the hardness of one instance is the time needed to solve it}.
\item We decided to use a \textcolor{red}{broad set of solvers} to test the computational hardness of the instances.
\end{itemize}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
\subsection{Instance Selection}\label{subsec:selection}
\framebox{TASK X : write }

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%  End QPLIB-3.tex
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

### 2016-07-18 QPLIB.tex
hello world

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%  QPLIB.tex
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\newif\ifMPC
\global\MPCtrue

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\ifMPC
 \documentclass[smallextended]{svjour3}
 \smartqed
 \journalname{Mathematical Programming Computation} 
\else
 \documentclass[12pt,A4wide]{article}
\fi

\usepackage{url}
\usepackage{secdot}
\usepackage{color}
\usepackage{booktabs}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{multirow}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
% page size

\ifMPC
\else
 \setlength{\textwidth}{16 cm}
 \setlength{\topmargin}{0 cm}
 \setlength{\headsep}{0 cm}
 \setlength{\textheight}{24 cm}
 \setlength{\evensidemargin}{0cm}
 \setlength{\oddsidemargin}{0cm}
\fi

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\def\NN{\mathbb{N}}
\def\RR{\mathbb{R}}
\def\Mcq{\mathcal{Q}}
\def\Mcl{\mathcal{L}}
\def\Mcm{\mathcal{M}}
\def\Mcn{\mathcal{N}}
\def\Mcb{\mathcal{B}}
\def\Mcr{\mathcal{R}}
\def\Mcu{\mathcal{U}}
\def\Mcx{\mathcal{X}}
\def\Mcz{\mathcal{Z}}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
% Theorem-like environment definitions

\ifMPC
\else
\newtheorem{theorem}{Theorem}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{remark}{Remark}

\newenvironment{example}{\begin{EX}\rm}{\end{EX}}

\newenvironment{proof}[1][Proof]{\textbf{#1.} }{\ \rule{0.5em}{0.5em}}
%\renewenvironment{proof}{{\bfseries Proof}}{\qed}

\def\remark{\stepcounter{theorem}\par{\bf Remark}.
\ignorespaces}
\def\endremark{ \vbox{\hrule\hbox{%
    \vrule height1.3ex\hskip0.8ex\vrule}\hrule
   }\par}

\newenvironment{keywords}{\vspace{0.4in}\begin{quote}\small \em
{\bf Keywords\/}:}{\end{quote}}
\fi

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\begin{document}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\title{QPLIB: a Library of a Quadratic Programming Instances}

\ifMPC
\titlerunning{QPLIB}
\fi

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\ifMPC
\author{Fabio Furini \and
        Emiliano Traversi \and
        Alper Atamturk \and
        Pietro Belotti \and
        Pierre Bonami \and 
        Sourour Elloumi \and
        Antonio Frangioni \and 
        Ambros Gleixner \and
        Nick Gould \and 
        Leo Liberti \and
        Andrea Lodi \and
        Ruth Misener \and
        Hans Mittelmann \and
        Nick Sahinidis \and 
        Frederic Roupin \and 
        Stefan Vigerske \and 
        Angelika Wiegele
        }

\institute{Fabio Furini \at
           LAMSADE, Universit{\'e} Paris Dauphine, 75775 Paris, France,
           \email{fabio.furini@dauphine.fr}
%
\and       
           Emiliano Traversi \at
           LIPN, Universit{\'e} de Paris 13, 93430 Villetaneuse, France.
           \email{emiliano.traversi@lipn.fr}
%
\and       Alper Atamturk \at
%
\and       Pietro Belotti \at
%
\and       Pierre Bonami \at 
%
\and       Sourour Elloumi \at
%
\and       Antonio Frangioni \at
           Dipartimento di Informatica, Universit\`a di Pisa,
           Largo B. Pontecorvo 2, 56127 Pisa, Italy,
           \email{frangio@di.unipi.it}
%
\and       Ambros Gleixner \at
           Zuse Institute Berlin, Takustr. 7, 14195 Berlin, Germany. 
           \email{gleixner@zib.de}
%
\and       Nick Gould \at 
%
\and       Leo Liberti \at
           CNRS LIX, Ecole Polytechnique, 91128 Palaiseau, France.
           \email{liberti@lix.polytechnique.fr}
%
\and       Andrea Lodi \at
           Ecole Polytechnique de Montréal, Canada.               \email{andrea.lodi@polymtl.ca}
%
\and       Ruth Misener \at
%
\and       Hans Mittelmann \at
           School of Mathematical and Statistical Sciences, Arizona State University,      Tempe, AZ 85287-1804, U.S.A.
\email{mittelmann@asu.edu}
%
\and       Nick Sahinidis \at 
%
\and       Frederic Roupin \at 
%
\and       Stefan Vigerske \at 
           GAMS Software GmbH, P.O. Box 40 59,50216 Frechen, Germany
           \email{stefan@gams.com}
%
\and       Angelika Wiegele \at
           }

\else

\author{Fabio Furini\\
        LAMSADE, Universit{\'e} Paris Dauphine, 75775 Paris, France\\
        {\tt fabio.furini@dauphine.fr}
%
\and   
        Emiliano Traversi\\
        LIPN, Universit{\'e} de Paris 13, 93430 Villetaneuse, France\\
        {\tt emiliano.traversi@lipn.fr}
%
\and    Alper Atamturk\\
%
\and    Pietro Belotti\\
%
\and    Pierre Bonami\\
%
\and    Sourour Elloumi\\
%
\and    Antonio Frangioni\\
        Dipartimento di Informatica, Universit\`a di Pisa\\
        Largo B. Pontecorvo 2, 56127 Pisa, Italy\\
        {\tt frangio@di.unipi.it}
%
\and    Ambros Gleixner\\
%
\and    Nick Gould\\
%
\and    Leo Liberti\\
        CNRS LIX, Ecole Polytechnique, 91128 Palaiseau, France\\
        {\tt liberti@lix.polytechnique.fr}
%
\and    Andrea Lodi\\
%
\and    Ruth Misener\\
%
\and    Hans Mittelmann\\
%
\and    Nick Sahinidis\\
%
\and    Frederic Roupin\\
%
\and    Stefan Vigerske\\
%
\and    Angelika Wiegele
        }
\fi

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\ifMPC
 \date{Received: date / Accepted: date}
\else
 \date{}
\fi

\maketitle

\begin{abstract}
 This paper describes a new instance library for Quadratic Programming (QP). QP is a very ``varied'' class, comprising sub-classes of problems ranging from trivial to indecidable. Correspondingly, solution methods for QP are very diverse, ranging from entirely combinatorial ones to completely continuous ones, to many where both aspects are fundamental. Selecting a set of instances of QP that is at the same time not overwhelmingly numerous and significant for the many different interested communities is therefore challenging. In order to help our selection process we propose a simple taxonomy for QP instances. We then briefly survey the field of QP, giving an overview of theory, methods and solvers. Finally we describe how the library was put together, and how the final results look.
\end{abstract}

\ifMPC
 \keywords{Instance Library, Quadratic Programming}

 \subclass{90C06 \and 90C25}
\else
 \begin{keywords}
  Instance Library, Quadratic Programming
 \end{keywords}
\fi

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\input{QPLIB-1.0.tex}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\input{QPLIB-2.1.tex}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\input{QPLIB-2.2.tex}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\input{QPLIB-2.3.tex}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\input{QPLIB-2.4.tex}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\input{QPLIB-3.tex}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\input{QPLIB-4.tex}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\input{conclusion.tex}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
\section{Acknowledgements}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\bibliography{biblio}

\ifMPC
 \bibliographystyle{spmpsci}
\else
 \bibliographystyle{plain}
\fi

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
\appendix

\input{Appendix.tex}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\end{document}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%  End QPLIB.tex
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

### 2016-07-19 QPLIB.tex
Nikolaos Sahinidis' comments on the abstract

@@
 
-\title{QPLIB: a Library of a Quadratic Programming Instances}
+\title{QPLIB: A Library of Quadratic Programming Instances}
 
@@
  This paper describes a new instance library for Quadratic Programming (QP).
-QP is a very ``varied'' class, comprising sub-classes of problems ranging from trivial to indecidable.
-Correspondingly, solution methods for QP are very diverse, ranging from entirely combinatorial ones to completely continuous ones, to many where both aspects are fundamental.
-Selecting a set of instances of QP that is at the same time not overwhelmingly numerous and significant for the many different interested communities is therefore challenging.
-In order to help our selection process we propose a simple taxonomy for QP instances.
+QP is a very ``varied'' class, comprising sub-classes of problems ranging from trivial to undecidable.
+Solution methods for QP are very diverse, ranging from entirely combinatorial ones to completely continuous ones, to many where both aspects are fundamental.
+Selecting a set of instances of QP that is at the same time not overwhelmingly difficult and sufficiently challenging for the many different interested communities is therefore important.
+ %In order to help our selection process 
+ We propose a simple taxonomy for QP instances that leads to a systematic problem selection mechanism.
 We then briefly survey the field of QP, giving an overview of theory, methods and solvers.
-Finally we describe how the library was put together, and how the final results look.
+Finally we describe how the library was put together, and detail its final contents.
 \end{abstract}

### 2016-07-19 QPLIB-2.2.tex
comments and minor corrections from first proofreading

@@
 Our taxonomy is based on a three-fields code of the form ``\textit{FVC}'', where \textit{F} indicates the objective function, \textit{V} the variables, and \textit{C} the constraints of the problem.
-The fields can be given the following values:
+The fields can be given the following values:\todo{\small Ambros: maybe OVC instead of FVC?}
 %
@@
 \subsubsection{Examples and reformulations}\label{ssec:reform}
+\todo{\small Ambros: maybe switch this with the next section? the next one contains more familiar material to the reader and might be an easier first introduction to the notation.}
 
@@
 When a $Q^i$ is positive semidefinite (SDP), i.e., the corresponding constraint/objective function is convex, general quadratic constraints are in fact equivalent to diagonal ones.
-In fact, every SDP matrix can be factorized as $Q^i = L^i (L^i)^\top$, e.g.~by the (incomplete) Cholesky factorization, $f^i(x) = x^\top Q^i x = \sum_{j \in \Mcn} z^2_jz$ where $z = x^\top L^i$.
+In fact, every SDP matrix can be factorized as $Q^i = L^i (L^i)^\top$,\todo{\small Ambros: did you really mean the same $L^i$ as above?} e.g.~by the (incomplete) Cholesky factorization, $f^i(x) = x^\top Q^i x = \sum_{j \in \Mcn} z^2_jz$ where $z = x^\top L^i$.
 Hence, one could think that D** problems need not be distinguished from C** ones; however, this is true only for ``complicated'' constraints, but not for ``simple'' ones, because the above reformulation technique introduces linear constraints.
@@
   %
-  \item \emph{Nononvex Integer Quadratic Programs} QGQ is the more general, and therefore is the most general, class.
+  \item \emph{Nonconvex Integer Quadratic Programs} QGQ is the more general, and therefore is the most general, class.
 Due to the lack of convexity even when integrality requirements are removed, solution methods must typically combine several algorithmic ideas, such as enumeration (distinguishing the role of integral variables from that of continuous ones involved into nonconvex terms) and techniques (e.g., outer approximation, SDP relaxation, \ldots) that allow to efficiently compute bounds.

### 2016-07-19 QPLIB.tex
comments and minor corrections from first proofreading

@@
 \usepackage{multirow}
+\usepackage{todonotes}
 
@@
 QP is a very ``varied'' class, comprising sub-classes of problems ranging from trivial to undecidable.
-Solution methods for QP are very diverse, ranging from entirely combinatorial ones to completely continuous ones, to many where both aspects are fundamental.
+Solution methods for QP are very diverse, ranging from entirely combinatorial ones to completely continuous ones, including many for which both aspects are fundamental.
 Selecting a set of instances of QP that is at the same time not overwhelmingly difficult and sufficiently challenging for the many different interested communities is therefore important.

### 2016-07-21 QPLIB-3.tex
comments and minor corrections from first proofreading
nimg: added QPLIB format section, currently in appendix       Updated main document to reflect this, and added nimg coordinates       Alightly updated Section 3 to reflect QPLIB format

@@
 
-In this Section we present all the steps performed in order to build the library.
+In this section we present all the steps performed in order to build the library.
 In particular we describe the first set of gathered instances (Section \ref{subsec:instColl}), we discuss the issues concerning the format of the instances (Section \ref{subsec:format}), we present the feature used to classified the instances (Section \ref{subsec:feature}), and finally we then describe the selection process that we have used to filter the instances in order to construct the final library (Section \ref{subsec:selection}).
@@
 
-As second format, we introduced 
-a specific format \texttt{.qplib}. This new format is capable to describe all the instances of the library in a sparse form. 
-In comparison to a more \emph{high level} format like \texttt{.gms}, the new format presents two advantages:  
+In addition, we have introduced
+a specific format \texttt{.qplib}. This new format is capable of describing
+all the instances of the library in a sparse form.
+In comparison to a more \emph{high level} format like \texttt{.gms}, the new
+format presents two advantages:
 it is easier to read by a self-made parser and it produces smaller files.
-\textcolor{red}{ToDo: describe \texttt{.qplib} format. Maybe in an appendix?}
+See Appendix~\label{sec:format} for details.
+%\textcolor{red}{ToDo: describe \texttt{.qplib} format. Maybe in an appendix?}
 

### 2016-07-21 format.tex
nimg: added QPLIB format section, currently in appendix       Updated main document to reflect this, and added nimg coordinates       Alightly updated Section 3 to reflect QPLIB format

\section{The file format}\label{sec:format}

\newcommand{\vect}[1]{\left(\begin{array}{c}#1\end{array}\right)}
\newcommand{\mat}[2]{\left(\begin{array}{#1}#2\end{array}\right)}
\newcommand{\arr}[2]{\begin{array}{#1}#2\end{array}}
\newcommand{\mininx}[1]{ {\renewcommand{\arraystretch}{0.8}
                     \begin{array}[t]{c}
                     \mbox{minimize} \vspace*{-1mm} \\
                     \mbox{ $\scriptstyle x \in \smallRe^n #1 $ }
                     \end{array} \;} }
\newcommand{\s}[1]{^{\mbox{\protect\tiny #1}}}
\newcommand{\sfrac}[2]{{\scriptstyle \frac{#1}{#2}}}
\newcommand{\half}{\sfrac{1}{2}}
\newcommand{\smallRe}
     {\mbox{\raisebox{-0.8pt}{\footnotesize I\hskip -1.5pt R\hskip -0.5pt}}}
\newcommand{\tim}[1]{\;\; \mbox{#1} \;\;}
\newcommand{\st}{\mbox{subject to}\;\;}
\newcommand{\lbox}[1]{\parbox[t]{105mm}{#1}\vspace*{1mm}}

The QPLIB format is designed to cope with a variety of
linear and quadratic programs. We distinguish, in particular,
linear programs (LP)
\begin{align*}
 \mininx{} & l^T x + f \\
 \st & c\s{L} \leq A x \leq c\s{U} \tim{and} x\s{L} \leq x \leq x\s{U}, \\
\intertext{linear programs with quadratic constraints (LPQC)}
 \mininx{} & l^T x + f \\
 \st & c\s{L} \leq A x + \half  [x,Q\s{C},x] \leq c\s{U}
     \tim{and} x\s{L} \leq x \leq x\s{U}, \\
\intertext{bound-constrained quadratic programs (BQP)}
 \mininx{} & \half x^T Q x + l^T x + f \\
 \st & x\s{L} \leq x \leq x\s{U}, \\
\intertext{quadratic programs (QP)}
 \mininx{} & \half x^T Q x + l^T x + f \\
 \st & c\s{L} \leq A x \leq c\s{U} \tim{and} x\s{L} \leq x \leq x\s{U}, \\
\intertext{and quadratic program with quadratic constraints (QPQC)}
 \mininx{} & \half x^T Q x + l^T x + f \\
 \st & c\s{L} \leq A x + \half [x,Q\s{C},x] \leq c\s{U}
     \tim{and} x\s{L} \leq x \leq x\s{U},
\end{align*}
where $Q\s{C}$ is the diagonal matrix with blocks $Q\s{C}_k$, $k=1,\ldots,m$,
$Q$ and the $Q\s{C}_k$ are symmetric $n$ by $n$ Hessian matrices,
$A$ is a rectangular $m$ by $n$ constraint matrix,
$l$, $x\s{L}$, $x\s{U}$, $c\s{L}$, $c\s{U}$
are vectors of appropriate sizes and $f$ is a scalar; the notation
$[ x, Q\s{C},x]$ denotes the vector whose $k$-th component is
$x^T Q\s{C}_k x$.
In addition, all or a subset of the variables $x$ may be required to be
integer, while further restrictions (such as a variable being binary)
may be imposed via explicit constraints.

The QPLIB format is defined in Table~\ref{tab-qplib-format}.
The data is in free format (blanks separate values), but must occur in
the order given here. Any blank lines, or lines starting with any of the
characters !, \% or \# are ignored. Each term in the first column of
Table~\ref{tab-qplib-format} denotes a required value. Any strings beyond
those required on a given line will be regarded as comments and ignored.
Real values may either by in decimal or exponential formats; for the latter,
the exponent may be preceded by either the character {\tt D} or {\tt E},
e.g. {\tt 12.56D+2} or {\tt 12.56E+2}.

\vspace*{2mm}

%\vspace*{-10mm}
%{\tiny
%\LTcapwidth=16.17cm
\begin{longtable}{|llr|}
\caption{\label{tab-qplib-format}{The QPLIB file format: refer to the notes
after the table for more details.}}\\
\hline
data & description & note \\
\hline
\endfirsthead
\caption{The QPLIB file format (continued)}\\
\hline
data & description & note \\
\hline
\endhead
\hline
\endfoot
\hline
\endlastfoot
%\begin{tabular}{|llr|}
%\hline
name & problem name (character string) & \\
type & problem type (character string) & [1] \\
\hline
$n$  & number of variables (integer) & \\
$m$  & number of general linear constraints (integer) & [2] \\
\hline
$n\s{Q}$ & number of nonzeros  (integer) in lower triangle of $Q$  & [3] \\
$i$\; $j$\; $Q_{i,j}$ & \lbox{row and column indices (integers) and value (real)
for each nonzero entry of $Q$, if $n\s{Q} > 0$, one triple on each line} & \\
\hline
$l_d$ & default value (real) for entries in $l$ & \\
$n\s{l}$ & number of non-default entries (integer) in $l$ & \\
$i$\; $l_i$ & \lbox{index (integer) and value (real) for each non-default
term in $l$, if $n\s{l} > 0$, one pair per line} & \\
\hline
$f$ & value (real) of f & \\ \hline
$n\s{Q${}_c$}$ & number of nonzeros (integer) in lower triangle of
$Q\s{C}$  & [2,4] \\
$k$\; $i$\; $j$\; $Q\s{C}_{k,i,j}$
& \lbox{constraint, row and column indices (integers) and value (real) for
 each entry of $Q\s{C}$, if $n\s{Q${}_c$} > 0$, one quadruple on each line} & \\
\hline
$n\s{A}$ & number of nonzeros  (integer) in $A$                     & [2] \\
$i$\; $j$\; $A_{i,j}$ & \lbox{row and column indices (integers) and value (real)
for each nonzero
entry of $A$, if $n\s{A} > 0$, one triple on each line} & \\
\hline
$\infty$ & \lbox{value (real) for infinity for constraint or variable
bounds---any bound greater than or equal to this in, absolute value,
is infinite} & \\
\hline
$c\s{L}_d$ & default value (real) for entries in $c\s{L}$            & [2] \\
$n^{c\s{L}}$ & number of non-default entries (integer) in $c\s{L}$    & [2] \\
$i$\;\ $c\s{L}_i$ & \lbox{index (integer) and value (real) for each non-default
term
in $c\s{L}$, if $n^{c\s{L}} > 0$, one pair per line}        & [2] \\
\hline
$c\s{U}_d$ & default value (real) for entries in $c\s{U}$            & [2] \\
$n^{c\s{U}}$ & number of non-default entries (integer) in $c\s{U}$    & [2] \\
$i$\;\ $c\s{U}_i$ & \lbox{index (integer) and value (real) for each non-default
term
in $c\s{U}$, if $n^{c\s{U}} > 0$, one pair per line}        & [2] \\
\hline
$x\s{L}_d$ & default value (real) for entries in $x\s{L}$            &     \\
$n^{x\s{L}}$ & number of non-default entries (integer) in $x\s{L}$    &     \\
$i$\;\ $x\s{L}_i$ & \lbox{index (integer) and value (real) for each non-default
term in $x\s{L}$, if $n^{x\s{L}} > 0$, one pair per line} & \\
\hline
$x\s{U}_d$ & default value (real) for entries in $x\s{U}$            &     \\
$n^{x\s{U}}$ & number of non-default entries (integer) in $x\s{U}$    &     \\
$i$\;\ $x\s{U}_i$ & \lbox{index (integer) and value (real) for each non-default
term in $x\s{U}$, if $n^{x\s{U}} > 0$, one pair per line} & \\
\hline
$v_d$ & \lbox{default variable type  (integer, 0 for continuous variables,
 1 for integer variables)} & [5] \\
$n^v$ & number of non-default variables (integer)  & [5] \\
$i$\; $v_i$ & \lbox{index  and type (integers) for each non-default
variable type, if  $n^v > 0$, one pair per line} & [5] \\
\hline
$x\s{0}_d$ & \lbox{default value (real) for the components of the starting
point  $x^0$ for the  variables $x$} & \\
$n^{x^0}$ & number of non-default starting entries (integer) in $x$ & \\
$i$\; $x_i^0$ & \lbox{index (integer) and value (real) for each non-default
starting value in $x^0$, if $n^{x^0} > 0$, one pair per line} &  \\
\hline
$y\s{0}_d$ & \lbox{default value (real) for the components of the starting
point $y^0$ for the Lagrange multipliers $y$ for the general
constraints} &  [2] \\
$n^{y^0}$ & number of non-default starting entries  (integer) in $y$ &  [2] \\
$i$\; $y_i^0$ & \lbox{index (integer) and value (real) for each non-default
starting value in $y^0$, if $n^{y^0} > 0$, one pair per line} &  [2] \\
\hline
$z\s{0}_d$ & \lbox{default value (real) for the components of the starting
point $z^0$ for the dual variables $z$ for the simple bound constraints} & \\
$n^{z^0}$ & number of non-default starting entries (integer) in $z$ & \\
$i$\; $z_i^0$ & \lbox{index (integer) and value (real) for each
non-default starting value in $z^0$, if $n^{z^0} > 0$, one pair per line} &  \\
\hline
$n^x_d$ & \lbox{number of non-default names (integer) of variables---default
for variable $i$ is the character string representing the numerical
value $i$} &  \\
$i$\; $x_i$ & \lbox{index (integer) and name (character string)
for each non-default variable name, if $n^x_d > 0$,  one pair per line} & \\
\hline
$n^c_d$ & \lbox{number of non-default names (integer) of general
constraints---default for constraint $i$ is the character string representing
the numerical value $i$} &  \\
$i$\; $c_i$ & \lbox{index  (integer) and name (character string) for each
non-default constraint name, if $n^c_d > 0$,  one pair per line} & \\
\hline
\end{longtable}
%}

%\end{tabular}

\renewcommand{\descriptionlabel}[1]{\hspace{\labelsep}\texttt{#1}}

\begin{description}
\item [{[1]}]
The problem type is represented by a character string as one of the following:
\begin{description}
\item For continuous problems (i.e., all variables are continuous):
\begin{description}[leftmargin=!,labelwidth=\widthof{\ttfamily SMILQPC}]
\item [LP]      a linear program,
\item [LPQC]    a linear program with quadratic constraints,
\item [BQP]     a bound-constrained quadratic program,
\item [QP]      a quadratic program, and
\item [QPQC]    a quadratic program with quadratic constraints.
\end{description}
\item For integer problems (i.e., all variables are integer valued):
\begin{description}[leftmargin=!,labelwidth=\widthof{\ttfamily SMILQPC}]
\item [ILP]     an integer linear program,
\item [ILPQC]   an integer linear program with quadratic constraints,
\item [IBQP]    an integer bound-constrained quadratic program,
\item [IQP]     an integer quadratic program, and
\item [IQPQC]   an integer quadratic program with quadratic constraints.
\end{description}
\item For mixed-integer problems (i.e., there is a mix of continuous and
integer variables):
\begin{description}[leftmargin=!,labelwidth=\widthof{\ttfamily SMILQPC}]
\item [MILP]    a mixed-integer linear program,
\item [MILPQC]  a mixed-integer linear program with quadratic constraints,
\item [MIBQP]   a mixed-integer bound-constrained quadratic program,
\item [MIQP]    a mixed-integer quadratic program, and
\item [MIQPQC]  a mixed-integer quadratic program with quadratic constraints.
\end{description}
\end{description}

\item [{[2]}]
for bound-constrained QPs, these sections are omitted.
\item [{[3]}]
for linear program with quadratic constraints, this section is omitted.
\item [{[4]}]
for problems without quadratic constraints, this section is omitted.
\item [{[5]}]
for purely-continuous or purely-integer problems, this section is omitted.
\end{description}

As a simple example, consider the mixed-integer QP
\[\arr{l}{\mininx{} x_1^2 + x_2^2 + x_3^2 - x_1 x_2 - x_2 x_3
  - 0.2 x_1  -0.4 x_2 - 0.2 x_3 \\
\st 1 \leq x_1 + x_2, \; 1 \leq x_1 + x_3, \; 0 \leq x_1 \leq 1,
\; 0 \leq x_2 \leq 2, \; \mbox{and binary} \; x_3.}
\]
This may then be represent in QPLIB format as follows:

{\small
\begin{verbatim}
! ---------------
! example problem
! ---------------
MIPBAND   # problem name
MIQP      # problem is a mixed-integer quadratic program
3         # variables
2         # general linear constraints
5         # nonzeros in lower triangle of Q
1 1 2.0   5 lines of row & column index & value of nonzeros in lower triangle Q
2 1 -1.0  |
2 2 2.0   |
3 2 -1.0  |
3 3 2.0   |
-0.2      default value for entries in l
1         # non default entries in l
2 -0.4    1 lines of index & value of non-default values in l
0.0       value of f
4         # nonzeros in A
1 1 1.0   4 lines of row & column index & value of nonzeros in A
1 2 1.0   |
2 1 1.0   |
2 3 1.0   |
1.0E+20   infinity
1.0       default value for entries in c^L
0         # non default entries in c^L
1.0E+20   default value for entries in c^U
0         # non default entries in c^U
0.0       default value for entries in x^L
0         # non default entries in x^L
1.0       default value for entries in x^U
1         # non default entries in x^U
2 2.0     1 line of non-default indices and values in x^U
0         default variable type is continuous
1         # non default variable types
3 1       variable 3 is integer
1.0       default value for initial values for x
0         # non default entries in x
0.0       default value for initial values for y
0         # non default entries in y
0.0       default value for initial values for z
0         # non default entries in z
0         # non default names for variables
0         # non default names for constraints
\end{verbatim}
}

### 2016-07-22 format.tex
add my comments to the format

@@
 \end{align*}
+\todo[inline]{\small Stefan: Are LPQC and QPQC frequently used abbreviations?
+If a linear program has quadratic constraints, how can it then still be a linear program?
+I am more familiar with the abbreviations QCP instead of LPQC and QCQP instead of QPQC.
+I further think that one could skip QCP and consider them as part of QCQP.}
 where $Q\s{C}$ is the diagonal matrix with blocks $Q\s{C}_k$, $k=1,\ldots,m$,
@@
 
+\todo[inline]{\small Stefan: I would suggest to make it simpler, even though that
+would reduce flexibility. Always fix the default value for coefficients to 0
+(otherwise, how you parse this? First initialize a n x n matrix with the non-zero
+default and then overwrite with the given entries, which may then be zero?)
+and have no default bounds for variables. If you then also adapt the suggestion
+to use a constraint sense ($\leq$,$\geq$,$=$) and a right-hand-side instead of
+a left- and right-hand-side, then there is also no need to specify a value for infinity.
+This would leave out ranged constraints - but are they really used? One could
+do it like CPLEX and have an extra constraint sense for ranges and an extra
+section to specify the range.}
+
 %\end{tabular}
@@
 \end{description}
+\todo[inline]{\small Stefan: I agree that leaving out information on
+numerical properties of the matrices and vectors (convexity, etc.) is good here.
+Such properties may also depend on tolerances, etc.
+As above, I would suggest to eliminate LPQC and rename QPQC into QCP.
+Having a string for ``unspecified'' may be nice, too, or one just uses MIQCQP in this case.}
 \end{description}

### 2016-07-27 QPLIB-2.3.tex
typos corrected and Mosek added

@@
 In general, local search methods are incomplete.
-As remarked above, however, local search methods deployed on a convex problem behave like complete methods, possibly via devices that that allow to find a feasible starting solution if there is one (the so-called ``phase 0'').
+As remarked above, however, local search methods deployed on a convex problem behave like complete methods, possibly via devices that that allow one to find a feasible starting solution if there is one (the so-called ``phase 0'').
 
@@
 %
- \item {\sc KNitro\_bb}, {\sc BonMin}, {\sc sBB}, {\sc XPress}, {\sc GuRoBi}, which implement complete BB algorithms for CGC. {\bf how about {\sc mosek}?}
+ \item {\sc Knitro\_bb}, {\sc BonMin}, {\sc sBB}, {\sc Xpress}, {\sc Gurobi}, and {\sc Mosek} which implement complete BB algorithms for CGC.}
 \end{itemize}

### 2016-07-27 QPLIB-3.tex
Gams instead of GAMS

@@
 \paragraph{First Instances Filter.} As already mentioned an important characteristic of an instances is its computational difficulty, i.e., the CPU time needed by a complete solver (cf.~\S \ref{sec:algo}) to solve the instance to global optimality.
-Accordingly, for each of the gathered instance we run the solvers in {\tt GAMS} (see Table \ref{}) able to solve it to global optimality.
+Accordingly, for each of the gathered instance we run the solvers in {\tt Gams} (see Table \ref{}) able to solve it to global optimality.
 The number of solvers depends on the category of the instances under consideration.
@@
 %In this phase we also eliminated multiples copies of the same instance.
-To evaluate the instances we decided, for practical reasons, to use GAMS as common platform for all the experiments involving commercial solvers.
-For this reason, we decided to translate all instances into the \texttt{.gms} gams format.
+To evaluate the instances we decided, for practical reasons, to use Gams as common platform for all the experiments involving commercial solvers.
+For this reason, we decided to translate all instances into the \texttt{.gms} Gams format.
  In a preliminary phase, all the instances received were divided according to their format and subsequently translated.

### 2016-07-28 QPLIB.tex
nimg: added QPLIB format section, currently in appendix       Updated main document to reflect this, and added nimg coordinates       Alightly updated Section 3 to reflect QPLIB format
Pietro's affiliation
included Nick's comments

@@
 \usepackage{todonotes}
+
+\usepackage{longtable}
+\usepackage{enumitem}
+\usepackage{calc}
 
@@
 \and       Pietro Belotti \at
+           Xpress-Optimizer team, FICO, Birmingham UK
+           \email{pietrobelotti@fico.com}
 %
@@
 \and       Nick Gould \at
+           STFC-Rutherford Appleton Laboratory, Chilton, Oxfordshire OX11 0QX,
+           England.
+           \email{nick.gould@stfc.ac.uk}
 %
@@
  This paper describes a new instance library for Quadratic Programming (QP).
-QP is a very ``varied'' class, comprising sub-classes of problems ranging from trivial to undecidable.
-Solution methods for QP are very diverse, ranging from entirely combinatorial ones to completely continuous ones, including many for which both aspects are fundamental.
-Selecting a set of instances of QP that is at the same time not overwhelmingly difficult and sufficiently challenging for the many different interested communities is therefore important.
- %In order to help our selection process 
- We propose a simple taxonomy for QP instances that leads to a systematic problem selection mechanism.
-We then briefly survey the field of QP, giving an overview of theory, methods and solvers.
-Finally we describe how the library was put together, and detail its final contents.
+ QP is a very ``varied'' class, comprising sub-classes of problems ranging from
+ trivial to undecidable. Solution methods for QP are very diverse, ranging from
+ entirely combinatorial ones to completely continuous ones, including many for
+ which both aspects are fundamental. Selecting a set of instances of QP that is
+ at the same time not overwhelmingly numerous and sufficiently challenging for the
+ many different interested communities is therefore important. We propose a simple
+ taxonomy for QP instances that leads to a systematic problem selection mechanism.
+ We then briefly survey the field of QP, giving an overview of theory, methods and
+ solvers. Finally we describe how the library was put together, and detail its final
+ contents.
 \end{abstract}
@@
 
+%  QPLIB format. Move to main text?
+
+\input{format.tex}
+
 %- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

### 2016-07-28 QPLIB-1.0.tex
implemented Nick's changes, a few additions (QP = MI-QCQP ...)

@@
 
-Quadratic Programming (QP) problems, where both the objective function and the constraints contain (at most) expressions in the variables of degree $2$, contain a surprisingly diverse set of rather different problems.
+Quadratic Programming (QP) problems, where both the objective function and the constraints contain (at most) expressions in the variables of degree $2$, include a surprisingly diverse set of rather different instances.
 According to the fine details of the formulation, solving a QP may require employing either fundamentally combinatorial techniques, or ideas rooted in nonlinear optimization principles, or a mix of the two.
@@
 However, this diversity also implies that QP means very different things to different researchers.
+This is illustrated by the simple fact that the class of problems that we simply refer to here as ``QP'' is called in different ways, among which Mixed-Integer Quadratically Constrained Quadratic Problems (MI-QCQP).
 It is perhaps therefore not surprising that, unlike for ``simpler'' problems classes \cite{Koch2011}, so far there has never been a single library containing all different kinds of instances of QP.
 Several libraries devoted to special cases of QP are indeed available; however, each of them is either focussed on one application (a specific problem that can be modeled as QP), or on QPs with specific structural properties that make them suitable to be solved with some given class of algorithmic approaches.
-To the best of our knowledge,  collecting a set of instances of QP that is at the same time not overwhelmingly numerous and significant for the many different interested communities has not been attempted, yet.
+To the best of our knowledge, collecting a set of QP instances that is at the same time not overwhelmingly numerous and significant for the many different interested communities has not been attempted, yet.
 This work constitutes a first step in this direction.
@@
 A particularly thorny issue in this process is how to select instances that are ``interesting''.
-Our solution is to take this to mean ``challenging for a significant set of solution methods''.
+Our choica has been to take this to mean ``challenging for a significant set of solution methods''.
 Our filtering process has then been in part based on the idea that if a significant fraction of the solvers that can solve a QP instance do so in a ``short'' time, then the instance is not challenging enough to be included in the library.
-This also takes into account the fact that if very few (maybe one) solver can solve it very efficiently by exploiting some specific structure, but most other approaches cannot, then the instance can still be deemed ``interesting''.
+Yet, if very few (maybe one) solvers can solve it very efficiently by exploiting some specific structure, but most other approaches cannot, then the instance can still be deemed ``interesting''.
 Putting this approach in practice requires a nontrivial number of technical steps and decisions that are detailed in the paper.
@@
 
-Optimization problems with quadratic constraints and/or objective function (QP) have been the subject of a considerable amount of research in the last decade.
-At least some of the rationale for this interest is likely due to the fact that they are the ``least nonlinear nonlinear problems''.
+Optimization problems with quadratic constraints and/or objective function (QP) have been the subject of a considerable amount of research over the last 60 years.
+At least some of the rationale for this interest is likely due to the fact that they are the ``least-nonlinear nonlinear problems''.
 Hence, in particular for the convex case, tools and techniques that have been honed during decades of research for Linear Programming (LP), typically with integrality constraints (MILP), can often be extended to the quadratic case with at least less effort than what would be required for tackling general Non Linear Programming (NLP) problems, without or with integrality constraints (MINLP).

### 2016-07-28 QPLIB-2.1.tex
comments and minor corrections from first proofreading
allowed to fix variables (Nick)

@@
 In mathematical optimization, a Quadratic Program (QP) is an optimization problem in which either the objective function or some of the constraints are quadratic functions.
-More specifically, the problem has the form
+More specifically, the problem has the form\todo{\small Ambros: I would use $b^i$ instead of $L^i$ and $c^i$ instead of $C^i$ (might avoid $L$ vs.
+$l$ confusion); plus capitals are often reserved for matrices)}
 %
@@
  %
- \item $-\infty \leq l_j < u_j \leq \infty$ are the (extended) real upper and lower bounds on each variable $x_j$ for $j \in \Mcn$, assumed different because otherwise the variable is fixed;
+ \item $-\infty \leq l_j \leq u_j \leq \infty$ are the (extended) real upper and lower bounds on each variable $x_j$ for $j \in \Mcn$;
  %
@@
 The relative flexibility offered by quadratic functions, as opposed e.g.~to linear ones, allows several reformulation techniques to be applicable to this family of problems in order to emphasize different properties of the various components.
-Some of these reformulation techniques will be commented later on; here we remark that, for instance, integrality requirements,
-in particular in the form of binary variables could always be ``hidden'' 
-by introducing (non convex) quadratic constraints utilizing the celebrated relationship
-%in particular under the form of binary variables, could always be ``hidden'' under the presence of (nonconvex) quadratic constraints due to the celebrated relationship 
-$x_j \in \{0, 1\} \iff x_j^2 = x_j$.
+Some of these reformulation techniques will be commented later on; here we remark that, for instance, integrality requirements, in particular in the form of binary variables could always be ``hidden'' by introducing (non convex) quadratic constraints utilizing the celebrated relationship $x_j \in \{0, 1\} \iff x_j^2 = x_j$.
 Therefore, when discussing these problems an effort has to be made to distinguish between features that come from the original model, and those that can be introduced by reformulation techniques in order to extract (and algorithmically exploit) specific properties.

### 2016-07-28 QPLIB-2.2.tex
typo corrected
included Nick's comments

@@
  %
- \item variables: (C)ontinuous only, (B)inary only, (M)ixed binary and continuous, (I)nteger only, (G)eneral (all threee types);
+ \item variables: (C)ontinuous only, (B)inary only, (M)ixed binary and continuous, (I)nteger only, (G)eneral (all three types);
  %
- \item constraints: (N)one, (B)ox, (L)inear, (C)onvex quadratic, nonconvex (Q)ua-dratic.
+ \item constraints: (N)one, (B)ox, (L)inear, (D)iagonal convex quadratic, (C)onvex quadratic, nonconvex (Q)ua-dratic.
 \end{itemize}
@@
 hold.
-In the following discussion we will repeatedly exploit this by assuming that, unless otherwise mentioned, when a method can be applied to a given problem, it can be applied as well to all simpler problems where the value of each field is arbitrarily replaced with a value denoting a less general class.
+In the following discussion we will repeatedly exploit this by assuming that, unless otherwise mentioned, when a method can be applied to a given problem, it can be applied as well to all simpler problems where the value of each field is arbitrarily replaced with a value denoting a less-general class.
 
@@
 \]
-Hence, their solution only requires the independent minimization of a convex quadratic univariate function in each single variable $x_j$ over a box constraint and possibly integrality requirements: this can be attained trivially in $O(1)$ by closed-form formul{\ae}, projection and rounding arguments.
+Hence, their solution only requires the independent minimization of a convex quadratic univariate function in each single variable $x_j$ over a box constraint and possibly integrality requirements, which can be attained trivially in $O(1)$ (per variable) by closed-form formul{\ae}, projection and rounding arguments.
 \emph{A fortiori}, the even simpler cases L*B, D*N and L*N (the latter obviously unbounded unless $L^0 = 0$) will not be discussed here.
-Similarly, CCB (and, a fortiori, CCN) are immediately solved by linear algebra techniques, and therefore are of no interest in this context.
+Similarly, CCN are immediately solved by linear algebra techniques, and therefore are of no interest in this context.
 On the other end of the spectrum, in general QP is a hard problem.
@@
 It is important to note that the relationships between the different classes can be somehow blurred because reformulation techniques may allow to move one instance from one class to the other.
-The example in the introduction, for instance, says that *M*---instances with only binary and continuous variables---can be recast as *CQ: nonconvex quadratic constraints can always take the place of binary variables.
+The already recalled fact that $x^2 = x \iff x \in \{0, 1\}$, for instance, says that *M*---instances with only binary and continuous variables---can be recast as *CQ: nonconvex quadratic constraints can always take the place of binary variables.
 Actually, this is also true for *G* as long as $\Mcu = \emptyset$, as bounded general integer variables can be represented by binary ones.
@@
 In practice, a diagonal convex objective function under linear constraints is found in many applications ({\bf citations?}), so that D*L still makes sense to distinguish the case where the objective function is ``naturally'' separable from that where separability is artificially introduced, although this is in theory always possible.
-Conversely, there is usually little gain in distinguishing between **C problems and ``**D'' ones where the constraints are restricted to be diagonal, as there are much fewer cases where this naturally occurs ({\bf to the best of my knowledge}).
-This is why we did not include a ``D'' option for the constraints.
 
@@
  %
- \item \emph{Convex Continuous Quadratic Programs} CCC can be solved in polynomial time by Interior-Point techniques; the simpler CCL can also be solved by means of ``simplex-like'' techniques ({\bf citations?}).
+ \item \emph{Convex Continuous Quadratic Programs} CCC can be solved in polynomial time by Interior-Point techniques; the simpler CCL can also be solved by means of ``simplex-like'' techniques, usually referred to as active-set methods ({\bf citations?}).
 Actually, a slightly larger class of problems can be solved with Interior-Point methods: those that can be represented by Second-Order Cone Programs.
@@
 They therefore require enumerative techniques like the spatial Branch\&Bound ({\bf citations?}), to be solved to optimality.
-Of course, local approaches
-are available that are able to
-% are available  that can be able to 
-efficiently provide local optima (or at least saddle points) of the CQC, but providing global guarantees about the quality of the obtained solutions is challenging.
+Of course, local approaches are available that are able to efficiently provide saddle points (hopefully, local optima) of the CQC, but providing global guarantees about the quality of the obtained solutions is challenging.
 In our library we have specifically focussed on  \emph{exact} solution of the instances.
@@
 However, convexity of the objective function and constraints implies that efficient techniques (see CCC) can be used at least to solve the continuous relaxation.
-The general view is that CGC are not, all other things being equal,
-% not 
- substantially more difficult than LGL to solve, especially if the objective function and/or the constraints have specific properties (e.g., DGL, CGL).
+The general view is that CGC are not, all other things being equal,  substantially more difficult than LGL to solve, especially if the objective function and/or the constraints have specific properties (e.g., DGL, CGL).
 Often integer variables are in fact binary ones, so several CCC models are C\{B,M\}C ones.
@@
   %
-  \item \emph{Nonconvex Integer Quadratic Programs} QGQ is the more general, and therefore is the most general, class.
+  \item \emph{Nonconvex Integer Quadratic Programs} QGQ is the most general, and therefore is the most difficul, class.
 Due to the lack of convexity even when integrality requirements are removed, solution methods must typically combine several algorithmic ideas, such as enumeration (distinguishing the role of integral variables from that of continuous ones involved into nonconvex terms) and techniques (e.g., outer approximation, SDP relaxation, \ldots) that allow to efficiently compute bounds.
@@
 %
-This description is purposely coarse; each of these classes can be subdivided into several other ones on the grounds of more detailed information about structures present in their constraints/objective function.
+This description is purposely coarse; each of these classes can be subdivided into several others on the grounds of more detailed information about structures present in their constraints/objective function.
 These can have a significant algorithmic impact, and therefore can be of interest to researchers.
 Common structures are, e.g., network flow or knapsack-type linear constraints, semi-continuous variables, or the fact that a nonconvex quadratic objective function/constraint can be reformulated as a second-order cone (hence, convex) one.
-It would be rather hard to collect a comprehensive list of all types of structures that may be of interest of some researcher, since these are as varied as the different possible approaches for specialized sub-classes of QP.
+It would be rather hard to collect a comprehensive list of all types of structures that may be of interest to any individual researcher, since these are as varied as the different possible approaches for specialized sub-classes of QP.
 For this reason we do not attempt such a more refined classification, and limit ourselves to the coarser one described in this paragraph.

### 2016-07-28 QPLIB-2.3.tex
included Nick's comments

@@
 In general, local search methods are incomplete.
-As remarked above, however, local search methods deployed on a convex problem behave like complete methods, possibly via devices that that allow one to find a feasible starting solution if there is one (the so-called ``phase 0'').
+As remarked above, however, local search methods deployed on a convex problem behave like complete methods, possibly via devices that that allow one to find a feasible starting solution if there is one (like what was used to be called ``phase -1'' in the simplex method).
 
@@
 The ($k+1$)-st iterate is obtained as $x^{k+1} = x^k + \alpha_k d^k$, where $\alpha_k$ (a scalar) is the {\it step length}, and $d^k$ (a vector) is the {\it search direction}, e.g.~belonging to a tangent manifold and having a negative directional derivative at $x^k$, in order to improving optimality while reducing infeasibility.
+Alternatively, $x^{k+1} = x^k + d^k(\alpha_k)$ where $d^k(\cdot)$ is an arc satisfying $d^k(0) = 0$.
 Local search methods for *I* are also iterative, but since defining a useful notion of descent direction is harder in presence of integer variables, $d^k$ and $\alpha_k$ are usually computed in different ways, depending on the application at hand.
@@
 At each iteration $k$ the algorithm forms the {\it active set} $\mathcal{A}^k$ containing the (indices of the) {\it active constraints}, i.e., all of the equality constraints as well as the inequality constraints that are satisfied at equality at the current iterate $x^k$.
-A subproblem, consisting of a minimization of a certain auxiliary objective function subject to all active constraints, is then solved to identify a good search direction $d^k$.
+A subproblem, consisting of a minimization of a certain auxiliary objective function and including only active constraints, is then solved to identify a good search direction $d^k$.
 An appropriate step length $\alpha_k$ is then found by checking for the inactive constraints (since these must be satisfied too).
 If at $x^{k+1}$ some of the previously inactive constraints have become active, or vice versa, $\mathcal{A}^k$ is updated accordingly.
-This general scheme works because the subproblem is defined in such a way as to be (a) easier to solve than the original problem, and (b) helping the sequence $x^k$ to converge to a Karush-Kuhn-Tucker (KKT) point.
+This general scheme works because the step is chosen so that the objective decreases, and the active set cannot repeat.
 

### 2016-07-28 QPLIB-2.1.tex
changed notation as Ambros asked

@@
 In mathematical optimization, a Quadratic Program (QP) is an optimization problem in which either the objective function or some of the constraints are quadratic functions.
-More specifically, the problem has the form\todo{\small Ambros: I would use $b^i$ instead of $L^i$ and $c^i$ instead of $C^i$ (might avoid $L$ vs.
-$l$ confusion); plus capitals are often reserved for matrices)}
+More specifically, the problem has the form
 %
@@
  \min \;\;
- & x^\top Q^0 x + L^0 x \\
+ & x^\top Q^0 x + b^0 x \\
  %
- & x^\top Q^i x + L^i x \leq C^i & i \in \Mcm \\ 
+ & x^\top Q^i x + b^i x \leq c^i & i \in \Mcm \\ 
  %
@@
  %
- \item $L^i$ and $C^i$ for $i \in \{ 0 \} \cup \Mcm$ are, respectively, real $n$-vectors and real constants;
+ \item $b^i$ and $c^i$ for $i \in \{ 0 \} \cup \Mcm$ are, respectively, real $n$-vectors and real constants;
  %

### 2016-07-28 QPLIB-2.2.tex
changed notation as Ambros asked

@@
 
-Our taxonomy is based on a three-fields code of the form ``\textit{FVC}'', where \textit{F} indicates the objective function, \textit{V} the variables, and \textit{C} the constraints of the problem.
-The fields can be given the following values:\todo{\small Ambros: maybe OVC instead of FVC?}
+Our taxonomy is based on a three-fields code of the form ``\textit{OVC}'', where \textit{O} indicates the objective function, \textit{V} the variables, and \textit{C} the constraints of the problem.
+The fields can be given the following values:
 %
@@
  \min \; \Big\{ \,
- \sum_{j \in \Mcn} \big( \, Q^0_j x_j^2 + L^0_j x_j \, \big) \;:\;
+ \sum_{j \in \Mcn} \big( \, Q^0_j x_j^2 + b^0_j x_j \, \big) \;:\;
  l_j \leq x_j \leq u_j \quad j \in \Mcn \;\;,\;\;
@@
 Hence, their solution only requires the independent minimization of a convex quadratic univariate function in each single variable $x_j$ over a box constraint and possibly integrality requirements, which can be attained trivially in $O(1)$ (per variable) by closed-form formul{\ae}, projection and rounding arguments.
-\emph{A fortiori}, the even simpler cases L*B, D*N and L*N (the latter obviously unbounded unless $L^0 = 0$) will not be discussed here.
+\emph{A fortiori}, the even simpler cases L*B, D*N and L*N (the latter obviously unbounded unless $b^0 = 0$) will not be discussed here.
 Similarly, CCN are immediately solved by linear algebra techniques, and therefore are of no interest in this context.
@@
  \textstyle
- \min \; \Big\{ \, L^0 x \;:\;
- x^\top Q^i x + L^i x \leq C^i \quad i \in \Mcm \;\;,\;\;
+ \min \; \Big\{ \, b^0 x \;:\;
+ x^\top Q^i x + b^i x \leq c^i \quad i \in \Mcm \;\;,\;\;
  x_j \in \mathbb{Z} \quad j \in \Mcn \, \Big\}
@@
  %
- & x^\top Q^0 x + L^0 x \leq x^0 & \\ 
+ & x^\top Q^0 x + b^0 x \leq x^0 & \\ 
  %
- & x^\top Q^i x + L^i x \leq C^i & i \in \Mcm \\ 
+ & x^\top Q^i x + b^i x \leq c^i & i \in \Mcm \\ 
  %
@@
 When a $Q^i$ is positive semidefinite (SDP), i.e., the corresponding constraint/objective function is convex, general quadratic constraints are in fact equivalent to diagonal ones.
-In fact, every SDP matrix can be factorized as $Q^i = L^i (L^i)^\top$,\todo{\small Ambros: did you really mean the same $L^i$ as above?} e.g.~by the (incomplete) Cholesky factorization, $f^i(x) = x^\top Q^i x = \sum_{j \in \Mcn} z^2_jz$ where $z = x^\top L^i$.
+In fact, every SDP matrix can be factorized as $Q^i = L^i (L^i)^\top$, e.g.~by the (incomplete) Cholesky factorization, $f^i(x) = x^\top Q^i x = \sum_{j \in \Mcn} z^2_jz$ where $z = x^\top L^i$.
 Hence, one could think that D** problems need not be distinguished from C** ones; however, this is true only for ``complicated'' constraints, but not for ``simple'' ones, because the above reformulation technique introduces linear constraints.

### 2016-07-28 QPLIB-2.3.tex
changed notation as Ambros asked

@@
 Due to the potentially infinite nature of exact real arithmetic computations, exact computations on the RRAM necessarily end up being approximate on the TM.
-Analogously, a complete method may reasonably be called ``exact'' on a RRAM; however, the computers we use in practice are more akin to TMs than RRAMs, and therefore
-calling \emph{exact} a solver that employs
-%dubbing \emph{exact} a solver that employ 
-floating point computations is, technically speaking, stretching the meaning of the word.
+Analogously, a complete method may reasonably be called ``exact'' on a RRAM; however, the computers we use in practice are more akin to TMs than RRAMs, and therefore calling \emph{exact} a solver that employs floating point computations is, technically speaking, stretching the meaning of the word.
 However, because the term is well understood in the computational optimization community, in the following we shall loosen the distinction between complete and exact methods, with either properties intended to mean ``complete'' in the sense of \cite{neumaier}.

### 2016-07-29 QPLIB.tex
affiliation updated (please check)

@@
         Emiliano Traversi \and
-        Alper Atamturk \and
         Pietro Belotti \and
-        Pierre Bonami \and
-        Sourour Elloumi \and
         Antonio Frangioni \and
@@
         Nick Sahinidis \and
-        Frederic Roupin \and
         Stefan Vigerske \and
@@
 \institute{Fabio Furini \at
-           LAMSADE, Universit{\'e} Paris Dauphine, 75775 Paris, France,
+           LAMSADE, Universit{\'e} Paris Dauphine, 75775 Paris, France.
            \email{fabio.furini@dauphine.fr}
@@
 %
-\and       Alper Atamturk \at
-%
 \and       Pietro Belotti \at
@@
            \email{pietrobelotti@fico.com}
-%
-\and       Pierre Bonami \at
-%
-\and       Sourour Elloumi \at
 %
@@
 \and       Ruth Misener \at
+           Department of Computing,Imperial College London, UK. \email{r.misener@imperial.ac.uk}
 %
@@
 \and       Nick Sahinidis \at
-%
-\and       Frederic Roupin \at
+Chemical Engineering, Carnegie Mellon University, U.S.A. \email{sahinidis@cmu.edu}
 %
 \and       Stefan Vigerske \at
-           GAMS Software GmbH, P.O. Box 40 59,50216 Frechen, Germany
+           GAMS Software GmbH, P.O. Box 40 59,50216 Frechen, Germany.
            \email{stefan@gams.com}
@@
 \and       Angelika Wiegele \at
+Institut für Mathematik, Alpen-Adria-Universität Klagenfurt, Austria.  \email{angelika.wiegele@aau.at}
            }
@@
 \author{Fabio Furini\\
-        LAMSADE, Universit{\'e} Paris Dauphine, 75775 Paris, France\\
-        {\tt fabio.furini@dauphine.fr}
 %
@@
         Emiliano Traversi\\
-        LIPN, Universit{\'e} de Paris 13, 93430 Villetaneuse, France\\
-        {\tt emiliano.traversi@lipn.fr}
-%
-\and    Alper Atamturk\\
 %
@@
 %
-\and    Pierre Bonami\\
-%
-\and    Sourour Elloumi\\
-%
 \and    Antonio Frangioni\\
-        Dipartimento di Informatica, Universit\`a di Pisa\\
-        Largo B. Pontecorvo 2, 56127 Pisa, Italy\\
-        {\tt frangio@di.unipi.it}
 %
@@
 \and    Leo Liberti\\
-        CNRS LIX, Ecole Polytechnique, 91128 Palaiseau, France\\
-        {\tt liberti@lix.polytechnique.fr}
 %
@@
 \and    Nick Sahinidis\\
-%
-\and    Frederic Roupin\\
 %

### 2016-07-29 QPLIB-2.2.tex
changed SDP to PSD

@@
 The ordering of the values in the previous lists is not irrelevant; in general, problems become ``harder'' when going from left to right.
-More specifically, for the \textit{F} and \textit{C} fields the order is that of \emph{strict} containment between problem classes: for instance, Linear objective functions are strictly contained in Diagonal convex quadratic ones (by just allowing the diagonal elements to be all-zero), which are strictly contained into general Convex quadratic ones (by allowing the off-diagonal elements to be all-zero), which in turn are strictly contained into general nonconvex Quadratic ones (by allowing any symmetric $Q^0$, hence possibly SDP ones as well).
+More specifically, for the \textit{F} and \textit{C} fields the order is that of \emph{strict} containment between problem classes: for instance, Linear objective functions are strictly contained in Diagonal convex quadratic ones (by just allowing the diagonal elements to be all-zero), which are strictly contained into general Convex quadratic ones (by allowing the off-diagonal elements to be all-zero), which in turn are strictly contained into general nonconvex Quadratic ones (by allowing any symmetric $Q^0$, hence possibly positive semidefinite ones as well).
 The only field for which the containment relationship is not a total order is \textit{V}, for which only the partial orderings
@@
 
-When a $Q^i$ is positive semidefinite (SDP), i.e., the corresponding constraint/objective function is convex, general quadratic constraints are in fact equivalent to diagonal ones.
-In fact, every SDP matrix can be factorized as $Q^i = L^i (L^i)^\top$, e.g.~by the (incomplete) Cholesky factorization, $f^i(x) = x^\top Q^i x = \sum_{j \in \Mcn} z^2_jz$ where $z = x^\top L^i$.
+When a $Q^i$ is positive semidefinite (PSD), i.e., the corresponding constraint/objective function is convex, general quadratic constraints are in fact equivalent to diagonal ones.
+In fact, every PSD matrix can be factorized as $Q^i = L^i (L^i)^\top$, e.g.~by the (incomplete) Cholesky factorization, $f^i(x) = x^\top Q^i x = \sum_{j \in \Mcn} z^2_jz$ where $z = x^\top L^i$.
 Hence, one could think that D** problems need not be distinguished from C** ones; however, this is true only for ``complicated'' constraints, but not for ``simple'' ones, because the above reformulation technique introduces linear constraints.

### 2016-07-29 QPLIB-2.1.tex
minor edit

@@
 
-In mathematical optimization, a Quadratic Program (QP) is an optimization problem in which either the objective function or some of the constraints are quadratic functions.
+In mathematical optimization, a Quadratic Program (QP) is an optimization problem in which either the objective function, or some of the constraints, or both, are quadratic functions.
 More specifically, the problem has the form

### 2016-07-29 QPLIB-1.0.tex
added Nick's citations

@@
 
-Quadratic Programming (QP) problems, where both the objective function and the constraints contain (at most) expressions in the variables of degree $2$, include a surprisingly diverse set of rather different instances.
+Quadratic Programming (QP) problems, where either the objective function \cite{wiki:qp}, or the constraints \cite{wiki:qcqp}, or both contain (at most) expressions in the variables of degree $2$, include a surprisingly diverse set of rather different instances.
+This is not surprising, given the vast scope of practical applications of these problems, and of solution methods designed to solve them \cite{GoulToin00a}.
 According to the fine details of the formulation, solving a QP may require employing either fundamentally combinatorial techniques, or ideas rooted in nonlinear optimization principles, or a mix of the two.

### 2016-07-29 QPLIB-2.3.tex
added projected gradient on Nick's request

@@
 Note that traditionally {\sc KNitro} used to be a local (nonconvex) NLP solver.
-All of these solvers implement essentially two types of local search methods for NLP:
+All of these solvers implement essentially three types of local search methods for NLP:
 %
 \begin{itemize}
- \item active set methods ({\sc ConOpt}, {\sc Minos}, {\sc SnOpt});
+ \item active set methods ({\sc ConOpt}, {\sc Minos}, {\sc SnOpt}) \cite{Dost09};
  %
- \item interior point methods ({\sc IpOpt}, {\sc KNitro}).
+ \item interior point methods ({\sc IpOpt}, {\sc KNitro}) \cite{Wright97};
+ %
+ \item projected gradient methods \cite{CalaMore87:mp, ChenGui13:coap}.
 \end{itemize}
@@
 Actually, nowadays the algorithm is most often implemented in the primal-dual version, where the nonlinear KKT system is iteratively solved with Newton-like iterations; this has the extra advantage of allowing to remove the need of a feasible (central) starting point.
+
+\paragraph{Projected gradient methods.}
+%
+These approaches are similar to active-set methods in that at each iteration they consider the gradient projected onto the set of active constraints, in order to (try to) guarantee feasibility of the next iterate.
+They differ from active-set methods in that the set of active constraints can change dramatically on each iteration, and they have stronger convergence properties, more akin to those of interior-point methods.
 

### 2016-07-29 QPLIB-2.2.tex
added Nick's citations

@@
  %
- \item \emph{Convex Continuous Quadratic Programs} CCC can be solved in polynomial time by Interior-Point techniques; the simpler CCL can also be solved by means of ``simplex-like'' techniques, usually referred to as active-set methods ({\bf citations?}).
+ \item \emph{Convex Continuous Quadratic Programs} CCC can be solved in polynomial time by Interior-Point techniques \cite{Wright97}; the simpler CCL can also be solved by means of ``simplex-like'' techniques, usually referred to as active-set methods \cite{Dost09}.
 Actually, a slightly larger class of problems can be solved with Interior-Point methods: those that can be represented by Second-Order Cone Programs.

### 2016-07-29 QPLIB-1.0.tex
added copositive/standard quadratic citations

@@
 Similarly, nonconvex continuous QP are perhaps the ``simplest'' class of problems that require features like spatial enumeration techniques to be solved.
-Hence, they are both the natural basis for the development of general techniques for nonconvex NLP, and a very specific class so that specialized approaches can be developed ({\bf citations, like copositive/standard QP?}).
+Hence, they are both the natural basis for the development of general techniques for nonconvex NLP, and a very specific class so that specialized approaches can be developed \cite{Dur2010, Burer2012}.
 

### 2016-08-04 QPLIB-1.0.tex
added a few citations, corrected a few typoes

@@
 A particularly thorny issue in this process is how to select instances that are ``interesting''.
-Our choica has been to take this to mean ``challenging for a significant set of solution methods''.
+Our choice has been to take this to mean ``challenging for a significant set of solution methods''.
 Our filtering process has then been in part based on the idea that if a significant fraction of the solvers that can solve a QP instance do so in a ``short'' time, then the instance is not challenging enough to be included in the library.
@@
 Similarly, nonconvex continuous QP are perhaps the ``simplest'' class of problems that require features like spatial enumeration techniques to be solved.
-Hence, they are both the natural basis for the development of general techniques for nonconvex NLP, and a very specific class so that specialized approaches can be developed \cite{Dur2010, Burer2012}.
+Hence, they are both the natural basis for the development of general techniques for nonconvex NLP, and a very specific class so that specialized approaches can be developed \cite{Dur2010,Burer2012}.
 
 On the other hand, (MI)QP is, in some sense, a considerably more expressive class than (MI)LP.
-Quadratic expressions are found, either naturally or after appropriate reformulations, in very many practical problems ({\bf citations?}).
+Quadratic expressions are found, either naturally or after appropriate reformulations, in very many practical problems (\cite{Kochenberger2014}, {\bf citations?}).
 In general, any continuous function can be approximated with arbitrary accuracy (over a compact set) by a polynomial of arbitrary degree; in turn, every polynomial can be broken down to a system of quadratic expressions.
@@
 The structure of this paper is the following.
-In \S \ref{sec:QPbasic} we review the basic notions about QP.
-In particular, \S \ref{subsec:notation} sets out the notation, \S \ref{sec:classification} proposes a---to the best of our knowledge, new---taxonomy of QP that helps us in discussing the (very) different classes of QPs, \S \ref{sec:algo} very briefly reviews the solution methods for QP upon which the solvers we have employed are based, and \S \ref{subsec:solver} describes the solvers.
-Then, \S \ref{sec:lib} describes the process used to obtain the library and its results; the software tools that we have used, and that are freely released together with the library, are discussed in \S \ref{subsec:tools}.
-Some conclusions are drawn in \S \ref{sec:conclusions}, after which in the Appendix a complete description of all the instances of the library is provided.
+In \S~\ref{sec:QPbasic} we review the basic notions about QP.
+In particular, \S~\ref{subsec:notation} sets out the notation, \S~\ref{sec:classification} proposes a---to the best of our knowledge, new---taxonomy of QP that helps us in discussing the (very) different classes of QPs, \S~\ref{sec:algo} very briefly reviews the solution methods for QP upon which the solvers we have employed are based, and \S~\ref{subsec:solver} describes the solvers.
+Then, \S~\ref{sec:lib} describes the process used to obtain the library and its results; the software tools that we have used, and that are freely released together with the library, are discussed in \S~\ref{subsec:tools}.
+Some conclusions are drawn in \S~\ref{sec:conclusions}, after which in the Appendix a complete description of all the instances of the library is provided.
 

### 2016-08-04 QPLIB-2.2.tex
added a few citations, corrected a few typoes

@@
 
-Despite the apparent simplicity of the definition given in \S \ref{subsec:notation}, Quadratic Programming instances can be of several rather different ``types'' in practice, depending on the fine details of the data.
+Despite the apparent simplicity of the definition given in \S~\ref{subsec:notation}, Quadratic Programming instances can be of several rather different ``types'' in practice, depending on the fine details of the data.
 In particular, many algorithmic approaches can only be applied to QP when the data of the problem has specific properties.
@@
 The ordering of the values in the previous lists is not irrelevant; in general, problems become ``harder'' when going from left to right.
-More specifically, for the \textit{F} and \textit{C} fields the order is that of \emph{strict} containment between problem classes: for instance, Linear objective functions are strictly contained in Diagonal convex quadratic ones (by just allowing the diagonal elements to be all-zero), which are strictly contained into general Convex quadratic ones (by allowing the off-diagonal elements to be all-zero), which in turn are strictly contained into general nonconvex Quadratic ones (by allowing any symmetric $Q^0$, hence possibly positive semidefinite ones as well).
+More specifically, for the \textit{O} and \textit{C} fields the order is that of \emph{strict} containment between problem classes: for instance, Linear objective functions are strictly contained in Diagonal convex quadratic ones (by just allowing the diagonal elements to be all-zero), which are strictly contained into general Convex quadratic ones (by allowing the off-diagonal elements to be all-zero), which in turn are strictly contained into general nonconvex Quadratic ones (by allowing any symmetric $Q^0$, hence possibly positive semidefinite ones as well).
 The only field for which the containment relationship is not a total order is \textit{V}, for which only the partial orderings
@@
 However, the special nature of binary variables combined with quadratic forms allows for quite specific techniques to be developed, among which is the reformulation of the problem as a LBL.
-Also, many well-known combinatorial problems can be naturally reformulated as problems of this class, and therefore a considerable number of results have been obtained by exploiting specific properties of the set of constraints ({\bf some citations?
+Also, many well-known combinatorial problems can be naturally reformulated as problems of this class, and therefore a considerable number of results have been obtained by exploiting specific properties of the set of constraints ( \cite{Rendl2008}, {\bf some citations?
 some specific problem to mention apart from Max-Cut?}).
@@
 Due to the lack of convexity even when integrality requirements are removed, solution methods must typically combine several algorithmic ideas, such as enumeration (distinguishing the role of integral variables from that of continuous ones involved into nonconvex terms) and techniques (e.g., outer approximation, SDP relaxation, \ldots) that allow to efficiently compute bounds.
-As in the convex case, QBQ, QMQ, and QIQ can benefit from more specific properties of the variables ({\bf citations?}).
+As in the convex case, QBQ, QMQ, and QIQ can benefit from more specific properties of the variables (\cite{Buchheim2013,Dong2016}, {\bf citations?}).
 \end{itemize}

### 2016-08-04 QPLIB-2.3.tex
added a few citations, corrected a few typoes

@@
  %
- \item projected gradient methods \cite{CalaMore87:mp, ChenGui13:coap}.
+ \item projected gradient methods \cite{CalaMore87:mp,ChenGui13:coap}.
 \end{itemize}
@@
 %
- \item {\sc Knitro\_bb}, {\sc BonMin}, {\sc sBB}, {\sc Xpress}, {\sc Gurobi}, and {\sc Mosek} which implement complete BB algorithms for CGC.}
+ \item {\sc Knitro\_bb}, {\sc BonMin}, {\sc sBB}, {\sc Xpress}, {\sc Gurobi}, and {\sc Mosek} which implement complete BB algorithms for CGC.
 \end{itemize}

### 2016-09-08 QPLIB-1.0.tex
Add a table of applications

@@
 On the other hand, (MI)QP is, in some sense, a considerably more expressive class than (MI)LP.
-Quadratic expressions are found, either naturally or after appropriate reformulations, in very many practical problems (\cite{Kochenberger2014}, {\bf citations?}).
+Quadratic expressions are found, either naturally or after appropriate reformulations, in very many practical problems \cite{Kochenberger2014}.
+We survey several of these applications in Table \ref{tbl:application}.
 In general, any continuous function can be approximated with arbitrary accuracy (over a compact set) by a polynomial of arbitrary degree; in turn, every polynomial can be broken down to a system of quadratic expressions.
@@
 
-\ldots
+\begin{center}
+\begin{longtable}{ lcl }
+%
+\caption{Application Domains of (MI)QP}\label{tbl:application} \\
+%
+\toprule
+Problem & Discrete? & Contributions \\
+\midrule
+\endfirsthead
+%
+\multicolumn{3}{l}{\emph{Table \ref{tbl:application} (Application Domains of (MI)QP) continued}} \\
+\toprule
+Problem & Discrete? & Contributions \\
+\midrule
+\endhead
+%
+\midrule 
+\multicolumn{3}{l}{\multirow{1}{\textwidth}{$^{\ddagger}$Applications with many manuscripts cite reviews and recent works \\[-12pt] \phantom{hi} \hfill \emph{continued}}} \\ \\
+\endfoot
+%
+\bottomrule 
+\multicolumn{3}{l}{\multirow{1}{\textwidth}{$^{\ddagger}$Applications with many manuscripts cite reviews and recent works. $^{\star}$TSP $\equiv$ Traveling Salesman Problem}} \\ \\ \\
+\endlastfoot
+%
+\multicolumn{3}{l}{{\bf Classical Problems that can be formulated as MIQP}} \\[1pt]
+%
+\multirow{1}{.55\textwidth}{Quadratic Assignment Problem (QAP)$^{\ddagger}$} & \checkmark & \cite{anstreicher:2003,loiola-etal:2007} \\ \cmidrule(rl){2-3}
+%
+Max-Cut & \checkmark & \cite{Rendl2008} \\ \cmidrule(rl){2-3}
+%
+Maximum clique$^{\ddagger}$ & \checkmark & \cite{bomze-etal:1999} \\
+\midrule
+%
+\multicolumn{3}{l}{{\bf Computational chemistry \& Molecular biology}} \\[1pt]
+%
+Zeolites && \cite{gounaris-etal:2013} \\
+%
+\midrule
+%
+\multicolumn{3}{l}{{\bf Computational geometry}} \\[1pt]
+%
+\multirow{1}{.55\textwidth}{Layout design} & \checkmark & \cite{anjos-liers:2012,castillo-etal:2005,dorneich-sahinidis:1995} \\ \cmidrule(rl){2-3}
+%
+\multirow{1}{.55\textwidth}{Maximizing polygon dimensions} && \cite{audet-etal:2011,audet-etal:2007,Audet-etal:2009,audet-etal:2002,Audet-Ninin:2013} \\ \cmidrule(rl){2-3}
+%
+Packing circles$^{\ddagger}$ && \cite{hifi-mhallah:2009,szabo-etal:2005} \\ \cmidrule(rl){2-3}
+%
+Nesting polygons && \cite{kallrath:2009,rebennack-etal:2009} \\ \cmidrule(rl){2-3}
+%
+Cutting ellipses && \cite{kallrath-rebennack:2013} \\ %\cmidrule(rl){2-3}
+%Nonconvex shapes \\
+\midrule
+%
+\multirow{1}{.55\textwidth}{{\bf Finance:} Portfolio selection optimization} && \multirow{1}{.25\textwidth}{\cite{deng-etal:2013,kallrath:2003,lin-etal:2005,maranas-etal:1997,parpas-rustem:2006,rios-sahinidis:2010}} \\
+%
+\midrule
+%
+\multicolumn{3}{l}{{\bf Process Networks}} \\[1pt]
+%
+%\multicolumn{3}{l}{Process networks optimisation problems typically involve two} \\
+%
+\multirow{1}{.55\textwidth}{Crude oil scheduling} & \checkmark & \cite{li-etal:2007,li-etal:2011,li-etal:2012,mouret-etal:2009,mouret-etal:2011} \\ \cmidrule(rl){2-3}
+%
+\multirow{1}{.55\textwidth}{Data reconciliation} & \checkmark & \cite{ruiz-grossmann:2011} \\ \cmidrule(rl){2-3}
+%
+\multirow{1}{.55\textwidth}{Multi-commodity flow} & \checkmark & \cite{tadayon-smith:2013} \\ \cmidrule(rl){2-3}
+%
+\multirow{1}{.55\textwidth}{Multi-period blending} & \checkmark & \cite{kolodziej-etal:2013:jogo,kolodziej-etal:2013} \\ \cmidrule(rl){2-3}
+%
+\multirow{1}{.55\textwidth}{Natural gas networks} & \checkmark & \cite{hasan-etal:2011,li-etal:2011-aiche_journal,li-etal:2011-jogo} \\ \cmidrule(rl){2-3}
+%
+Pooling$^{\ddagger}$ & \checkmark & \multirow{1}{.25\textwidth}{\cite{Alfaki-Haugland:2013,Castillo-etal:2013,dambrosio-etal:2011pooling,Faria-Bagajewicz:2012,misener-floudas:2009,misener-floudas:2010-genpooling,Papageorgiou-etal:2012,pham-etal:2009,ruiz-etal:2013,visweswaran:2009}} \\ \\ \cmidrule(rl){2-3}
+%
+\multirow{1}{.55\textwidth}{Open-pit mine scheduling} & \checkmark & \cite{bley-etal:2012} \\ \cmidrule(rl){2-3}
+%
+Reverse osmosis & \checkmark & \cite{saif-etal:2008} \\ \cmidrule(rl){2-3}
+%
+Supply chain & \checkmark & \cite{nyberg-etal:2012} \\ \cmidrule(rl){2-3}
+%
+Water networks$^{\ddagger}$ & \checkmark & \multirow{1}{.25\textwidth}{\cite{ahmetovic-grossmann:2010,bagajewicz:2000,bragalli-etal:2011,castro-teles:2013,geissler-etal:2013,gleixner-etal:2012,jezowski-2010,khor-etal:2014,ponceortega-etal:2010,teles-etal:2012}} \\ \\
+%
+\midrule
+%
+\multicolumn{3}{l}{{\bf Robotics}} \\[1pt]
+%
+\multirow{1}{.55\textwidth}{TSP$^{\star}$ with Neighborhoods} & \checkmark & \cite{gentilini-etal:2013} \\
+%
+\end{longtable}
+\end{center}
 

### 2016-11-26 QPLIB-3.tex
added CUTEst
added a few citations, corrected a few typoes
Update QPLIB-3.tex

@@
 \begin{itemize}
- \item {\tt POLIP} \url{http://polip.zib.de/pipformat.php}
+ \item {\tt BARON Library} \url{http://www.minlp.com/nlp-and-minlp-test-problems}
  %
- \item {\tt MINLP} \url{http://www.gamsworld.org/minlp/minlplib.htm}
+ \item {\tt CUTEst Library} \url{https://ccpforge.cse.rl.ac.uk/gf/project/cutest/wiki}
+ %
+ \item {\tt GAMS Library} \url{http://www.gamsworld.org/performance/performlib.htm}
  %
@@
  %
- \item {\tt BARON Library} \url{http://www.minlp.com/nlp-and-minlp-test-problems}
+ \item {\tt MINLP} \url{http://www.gamsworld.org/minlp/minlplib.htm}
  %
- \item {\tt GAMS Library} \url{http://www.gamsworld.org/performance/performlib.htm}
+ \item {\tt POLIP} \url{http://polip.zib.de/pipformat.php}
 \end{itemize}
@@
 %
-\multirow{7}*{Linear}
-          & Binary  & Quadratic &  71 \\
+\multirow{5}*{Linear}
+          & \multirow{1}*{Binary}
+%                    & None      &   \\
+%          &         & Linear    &  \\
+                    & Quadratic &   9 \\[1.2 ex]
           & \multirow{2}*{Mixed}
-                    & Convex    &   5 \\
-          &         & Quadratic & 277 \\
-%          & \multirow{2}*{Integer}
-%                    & Convex    &  15 \\
-%          &         & Quadratic &   5 \\
-          & \multirow{2}*{General}
-                    & Convex    &  22 \\
-          &         & Quadratic &   8 \\
+                    & Convex    &   2\\[1.2 ex]
+          &         & Quadratic &    118\\[1.2 ex]
+          & \multirow{1}*{Integer}
+%                    & Linear    &    \\
+                   & Quadratic &    2\\[1.2 ex]
+          & \multirow{1}*{General}
+%                    & Linear    &    \\
+                   & Quadratic &    3\\[1.2 ex]
+%\multirow{5}*{Linear}
+%          & Binary  & Quadratic &   9\\
+%          & \multirow{2}*{Mixed}
+%                    & Convex    &   2\\
+%          &         & Quadratic &  118\\
+%%          & \multirow{2}*{Integer}
+%%                    & Convex    &  15 \\
+%%          &         & Quadratic &   5 \\
+%          & {General} & Quadratic &   3 \\
 %\hline
 \\
-\multirow{4}*{Convex}
-          & Binary  & Linear    &  12 \\
+\multirow{3}*{Convex}
+          & Binary  & Linear    &  2 \\[1.2 ex]
           & \multirow{2}*{Mixed}
-                    & Linear    &  37 \\
-          &         & Quadratic &   3 \\
-          & General & Linear    &   1 \\
+                    & Linear    &   13\\[1.2 ex]
+          &         & Quadratic &    1\\[1.2 ex]
+%          & General & Linear    &    \\
 %\hline
@@
           & \multirow{3}*{Binary}
-                    & None      &  29 \\
-          &         & Linear    & 180 \\
-%          &         & Quadratic &   8 \\
+                    & None      &   23\\[1.2 ex]
+          &         & Linear    &  59\\[1.2 ex]
+          &         & Quadratic &   5 \\[1.2 ex]
           & \multirow{2}*{Mixed}
-                    & Linear    &  12 \\
-          &         & Quadratic &   9 \\
-          & Integer & Linear    &   2 \\
-          & \multirow{2}*{General}
-                    & Linear    &   1 \\
-          &         & Quadratic &   2 \\
+                    & Linear    &   10\\[1.2 ex]
+          &         & Quadratic &    1\\[1.2 ex]
+%          & Integer & Linear    &    \\
+          & \multirow{1}*{Integer}
+                    & Linear    &    2\\[1.2 ex]
+          & \multirow{1}*{General}
+                    & Quadratic    &    1\\[1.2 ex]
+%          &         & Quadratic &    \\
 \hline
-Total     &         &           & 600 \\
+Total     &         &           & 251\\
 %
@@
 %
-\multirow{2}*{Linear}    & Convex    &  14 \\
-                         & Quadratic &  71 \\
-\multirow{3}*{Convex}    & None      &   6 \\
-                         & Box       &   6 \\
-                         & Linear    &  50 \\
-                         & Convex    &   3 \\
-                         & Quadratic &   6 \\
-\multirow{3}*{Quadratic} & Box       &   8 \\
-                         & Linear    &  19 \\
-                         & Convex    &  27 \\
-                         & Quadratic &  40 \\
+\multirow{2}*{Linear}    & Convex    &   11\\[1.2 ex]
+                         & Quadratic &   50\\[1.2 ex]
+\multirow{3}*{Convex}    
+                         & Box       &   3 \\[1.2 ex]
+                         & Linear    &   12\\[1.2 ex]
+                         & Convex    &    2\\[1.2 ex]
+                         & Quadratic &    5\\[1.2 ex]
+\multirow{3}*{Quadratic} 
+                         & Linear    &   5\\[1.2 ex]
+                         & Convex    &   11\\[1.2 ex]
+                         & Quadratic &   17\\[1.2 ex]
 \hline
-Total                    &           & 250 \\
+Total                    &           & 116 \\
 %
@@
  In a preliminary phase, all the instances received were divided according to their format and subsequently translated.
-In Sect.
+In \S~\ref{subsec:tools}
 %~\ref{subsec:CP_convex}
@@
 %to translate an instance from and to a given format to the \texttt{.gms} format
-are described more in detail.\\
+are described in more detail.\\
 

### 2016-12-07 QPLIB-3.tex
Updated tables

@@
                     & Quadratic &   9 \\[1.2 ex]
+\cmidrule(lr){2-4}
           & \multirow{2}*{Mixed}
@@
           &         & Quadratic &    118\\[1.2 ex]
+\cmidrule(lr){2-4}
           & \multirow{1}*{Integer}
@@
                    & Quadratic &    2\\[1.2 ex]
+\cmidrule(lr){2-4}
           & \multirow{1}*{General}
@@
                    & Quadratic &    3\\[1.2 ex]
+\cmidrule(lr){1-4}
 %\multirow{5}*{Linear}
@@
 %\hline
-\\
 \multirow{3}*{Convex}
           & Binary  & Linear    &  2 \\[1.2 ex]
+\cmidrule(lr){2-4}
           & \multirow{2}*{Mixed}
@@
 %\hline
-\\
+\cmidrule(lr){1-4}
 \multirow{7}*{Quadratic}
@@
           &         & Quadratic &   5 \\[1.2 ex]
+\cmidrule(lr){2-4}
           & \multirow{2}*{Mixed}
@@
 %          & Integer & Linear    &    \\
+\cmidrule(lr){2-4}
           & \multirow{1}*{Integer}
                     & Linear    &    2\\[1.2 ex]
+\cmidrule(lr){2-4}
           & \multirow{1}*{General}
@@
                          & Quadratic &   50\\[1.2 ex]
-\multirow{3}*{Convex}    
+\cmidrule(lr){1-3}
+\multirow{4}*{Convex}    
                          & Box       &   3 \\[1.2 ex]
@@
                          & Quadratic &    5\\[1.2 ex]
+\cmidrule(lr){1-3}
 \multirow{3}*{Quadratic} 

### 2016-12-07 QPLIB-3.tex
Added pics

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}\centering
+  \includegraphics[width=0.85\textwidth]{./pics/pic_overview.png}
+  \caption{...\label{fig:1}}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}\centering
+  \includegraphics[width=0.85\textwidth]{./pics/pic_var_small.png}
+  \caption{...\label{fig:4}}
+\end{figure}
+
+\begin{figure}\centering
+  \includegraphics[width=0.85\textwidth]{./pics/pic_var_medium.png}
+  \caption{...\label{fig:3}}
+\end{figure}
+
+\begin{figure}\centering
+  \includegraphics[width=0.85\textwidth]{./pics/pic_var_big.png}
+  \caption{...\label{fig:2}}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}\centering
+  \includegraphics[width=0.85\textwidth]{./pics/pic_constr_small.png}
+  \caption{...\label{fig:5}}
+\end{figure}
+
+\begin{figure}\centering
+  \includegraphics[width=0.85\textwidth]{./pics/pic_constr_medium.png}
+  \caption{...\label{fig:6}}
+\end{figure}
+
+\begin{figure}\centering
+  \includegraphics[width=0.85\textwidth]{./pics/pic_constr_big.png}
+  \caption{...\label{fig:7}}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}\centering
+  \includegraphics[width=0.85\textwidth]{./pics/pic_quad_vs_convex_constr.png}
+  \caption{...\label{fig:8}}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}\centering
+  \includegraphics[width=0.85\textwidth]{./pics/pic_fuffa.png}
+  \caption{...\label{fig:9}}
+\end{figure}
+
+\begin{figure}\centering
+  \includegraphics[width=0.85\textwidth]{./pics/pic_density.png}
+  \caption{...\label{fig:10}}
+\end{figure}
+
+\begin{figure}\centering
+  \includegraphics[width=0.85\textwidth]{./pics/pic_neg_eig.png}
+  \caption{...\label{fig:10}}
+\end{figure}
+
 %- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

### 2016-12-13 conclusion.tex
first draft of the conlcusions

\section{Conclusions}\label{sec:conclusions}

 This manuscript describes the first comprehensive library of instances for Quadratic Programming (QP). Since QP comprises  different and `varied'' categories of problems, we propose a classification and then we describe the principal solution methods representing the state-of-the-art algorithms for QP.

We then describe the steps of the adopted process  to filter the gathered instances  in order to build the new library. The goals we purposed are of different natures and we tried to build a library which is computationally challenging and as broad as possible, i.e., it represents the largest possible categories of QP problems.

We intentionally avoid to perform a computational comparison of the performances of the different solvers. Our goal is instead to provide a common test-bed of instances for practitioners and researchers in the field. This new library will hopefully serve as a point of reference to test new ideas and algorithms for QP problems.

### 2016-12-13 QPLIB-3.tex
Update QPLIB-3.tex
updated section 3

--- 2016-12-07 QPLIB-3.tex
@@
 
-In this section we present all the steps performed in order to build the library.
-In particular we describe the first set of gathered instances (Section \ref{subsec:instColl}), we discuss the issues concerning the format of the instances (Section \ref{subsec:format}), we present the feature used to classified the instances (Section \ref{subsec:feature}), and finally we then describe the selection process that we have used to filter the instances in order to construct the final library (Section \ref{subsec:selection}).
+In this section we present all the steps performed in order to build
+the library. In particular we describe the first set of gathered
+instances (Section \ref{subsec:instColl}),
+we present the feature used to classified the instances (Section
+\ref{subsec:feature}),
+we discuss the issues concerning the format of the instances (Section
+\ref{subsec:format}),  and finally we describe the selection process
+used to filter the instances in order to construct the final library
+(Section \ref{subsec:selection}).
 
@@
 
-In this section we describe the procedure we adopted to gather the instances.
-In January $2014$, we issued an online call for instances to the main international mailing lists of the mathematical optimization and numerical analysis community, in order to reach the largest possible set of interested researchers, both in academia and in industry.
-The call remained open for 10 months, during which we received a large number of contributions with different characteristics.
-The instances we received  are both artificial and coming from real-world applications.
+In this section we describe the procedure we adopted to gather the
+instances. In January $2014$, we issued an online call for instances
+using the principal international mailing lists.
+In particular, we addressed the mailing lists of the mathematical
+optimization and numerical analysis communities, reaching in this way
+the largest possible set of interested researchers and practitioners.
+The call remained open for 10 months, during which we received a large
+number of contributions with different characteristics. The instances
+we received  comes from theoretical studies and from real-world
+applications.
 
-In addition to spontaneous contribution we scanned the other known libraries of instances and we selected all the QP ones.
-In particular we cite here the libraries of ``generic'' QP instances from which we draw material:
+In addition to spontaneous contribution we analysed the other generic
+libraries of instances available  on internet and containing quadratic
+instances. Namely, we cite here the libraries from which we selected
+some quadratic instances:
 
 \begin{itemize}
- \item {\tt BARON Library} \url{http://www.minlp.com/nlp-and-minlp-test-problems}
- %
- \item {\tt CUTEst Library} \url{https://ccpforge.cse.rl.ac.uk/gf/project/cutest/wiki}
- %
- \item {\tt GAMS Library} \url{http://www.gamsworld.org/performance/performlib.htm}
- %
+ \item {\tt BARON Library}
+\url{http://www.minlp.com/nlp-and-minlp-test-problems}
+ \item {\tt CUTEst Library}
+\url{https://ccpforge.cse.rl.ac.uk/gf/project/cutest/wiki}
+ \item {\tt GAMS Library}
+\url{http://www.gamsworld.org/performance/performlib.htm}
  \item {\tt MacMINLP} \url{https://wiki.mcs.anl.gov/leyffer/index.php/MacMINLP}
- %
  \item {\tt Meszaros} \url{http://www.doc.ic.ac.uk/~im/00README.QP}
- %
  \item {\tt MINLP} \url{http://www.gamsworld.org/minlp/minlplib.htm}
- %
  \item {\tt POLIP} \url{http://polip.zib.de/pipformat.php}
 \end{itemize}
-%
-Other QP instances were found in libraries devoted to specific optimization problems that can be modeled as QP, such as Max-Cut, the Quadratic Assignment Problem, Portfolio Optimization problems and several others; for reason of space we refrain from listing all these sources.
 
-At the end of this process we had gathered more than eight thousand instances. Three fourths of them contained discrete variables, while the remaining ones contained only continuous variables.
+Other quadratic instances were found in online libraries devoted to
+specific optimization problems that can be modelled as quadratic ones:
+Max-Cut, Quadratic Assignment, Portfolio Optimization and several
+others.
+
+At the end of this process we gathered more than eight thousand
+instances. Three fourths of them contained discrete variables, while
+the remaining ones contained only continuous variables.
+
+More in details, we gathered a large number of Quadratic Binary Linear
+(QBL), Quadratic Continuous Quadratic (QCQ) and Quadratic General
+Quadratic (QGQ) instances, i.e., $\approx 1800$ instances, $\approx
+2000$ instances and $\approx 2500$ instances respectively.
+We gathered $\approx 1000$ Convex General Convex (CGC) instances. We
+gathered relatively few Quadratic Binary Quadratic (QBQ), Convex
+Continuous Convex (CCC) and Convex Mixed Convex (CMC) instances, i.e.,
+$\approx 150$ instances, $\approx 200$ instances and $\approx 200$
+instances respectively. Finally we gathered only $17$ Quadratic Mixed
+Linear (QML) instances.
+
+In the call for instances no specific formats were imposed
+%we did not impose a specific format
+for the submissions.
+%In this phase we also eliminated multiples copies of the same instance.
+To evaluate the instances we decided, for practical reasons, to use
+Gams as common platform for all the experiments involving commercial
+solvers.
+For this reason, we decided to translate all instances into the
+\texttt{.gms} Gams format and \texttt{.lp} format.
+ In a preliminary phase, all the instances received were divided
+according to their format and subsequently translated.
+%In \S~\ref{subsec:tools}
+%%~\ref{subsec:CP_convex}
+%the tools used
+%to translate an instance from a given format to the \texttt{.gms} format
+%%to translate an instance from and to a given format to the
+\texttt{.gms} format
+%are described in more detail.\\
+
+In addition, we have introduced
+a specific format \texttt{.qplib}. This new format is capable of describing
+all the instances of the library in a sparse form.
+In comparison to a more \emph{high level} format like \texttt{.gms}, the new
+format presents two advantages:
+it is easier to read by a self-made parser and it produces smaller files.
+See Appendix~\label{sec:format} for details.
+%\textcolor{red}{ToDo: describe \texttt{.qplib} format. Maybe in an appendix?}
+
+% is lighter in terms of sie
+%  significantly
+%less memory
+%in comparison and
+
+% for this reason we used the \texttt{.gms} format as one of the
+format of the library. %checked if one instance was present copies of
+the same instance, potentially written in different formats
+
+\subsection{Instance Features}\label{subsec:feature}
+
+For each instance of the starting set, we collected the following
+characteristics which allow to classify the instances into the
+categories described in Section~\ref{sec:QPbasic}.
+
+As far as the variable type is concerned, we gather the following information:
+\begin{itemize}
+\item number of binary variables (\emph{\# bin}),
+\item number of integer variables (\emph{\# int}),
+\item number of continuous variables (\emph{\# cont}).
+\end{itemize}
+In case at least one binary or integer variable is present, then the
+instance is categorized as \emph{discrete} instance, on the other
+hand, in case only continuous variables are presents, the instance is
+categorized as \emph{continuous}.
+
+As fas as the objective function is concerned, we gathered the
+following information:
+\begin{itemize}
+\item Percentage of negative eigenvalues (\emph{\% neg eig}),
+\item Density of the Hessian of the objective function (\emph{\% dens}).
+\end{itemize}
+ The number of negative eigenvalues of the Hessian matrix allows to
+identify the specific objective function type, i.e. in presence of at
+least one negative eigenvalue, the objective function is non convex.
+The density of the Hessian of the objective function is defined as the
+number of nonzero entries     over the total possible figure.
+
+As far as the constraint type is concerned, we gather the following information:
+\begin{itemize}
+\item number of linear constraints (\emph{\# lin}),
+\item number of quadratic constraints (\emph{\# quad}),
+\item number of convex constraints (\emph{\# conv}),
+\item number of box constraints (\emph{\# box}).
+\end{itemize}
+A constraint is considered quadratic if it contain a quadratic term.
+Among the quadratic constraints, the one with only non negative
+eigenvalues are classified as convex constraints.
+
+This information allows to have a static picture of the main features
+of the instances.
+
+\subsection{Instance Selection}\label{subsec:selection}
+
+The first goal of the library is to gather ``challenging'' instances,
+i.e., instances which can not be easily solved by the state of the art
+solvers.
+The second goal of the library is to represent the largest majority of
+the different categories of quadratic problems.
+The third goal is to include in each of the categories a set of well
+diversified instances.
+Finally, the fourth goal is to obtain a library whose set of instances
+is neither too small either too large, to preserve statistical
+relevance while at the same time being computationally manageable.
+
+To achieve the aforementioned goals, we performed the following two
+filters applied in cascade.
+
+\begin{itemize}
+\item \emph{First Instances Filter.}\\
+The first filter is design to drastically reduce the number of
+instances by eliminating the easy instances. An empirical measure way
+of testing the hardness of an instance is the CPU time needed by a
+complete solver (cf.~\S \ref{sec:algo}) to solve the instance to
+global optimality.
+In this step, we decided to use the broadest set of solvers able to
+tackle the different categories of quadratic problems.\\
+Accordingly, for each of the gathered instance, we run the solvers in
+{\tt Gams} (see Table \ref{}) able to solve it to global optimality.
+The number of solvers depends on the category of the instances under
+consideration.
+We the filtered according to a relative measure of computational
+difficulty, i.e., we discarded all instances that are solved by at
+least 30\% of the complete solvers within a time limit of 30 seconds.
+\item \emph{Second Instances Filter.}\\
+The goal of the second filter is to eliminate instances that are similar.
+We used the instance features presented in
+Section~\ref{subsec:feature} and the CPU solving time obtained in the
+first filter to guide this second step. \\
+We carefully analysed the instances and we grouped them accordingly to
+similar features and CPU time. To reduce the size of the test-bed
+while keeping the representativeness of the library,     we kept for
+each cluster of similar instances one representative.\\
+In order to have a deeper insight on the computational hardness of the
+instances we performed a second round of tests on the resulting set of
+instances.  We used a general purpose solverand we set the time limit
+to 120 seconds. Finally, all the instances which have been solved to
+proven optimality have been discarded.
+\end{itemize}
+In Table~\ref{tab:filters} we summarize the two filter steps, which
+allows us to identify the final set of $251$ discrete instances and
+$116$ continuous instances.
 
@@
 & $\approx$ 6000 Discr. Inst.  & $\approx$ 2500 Cont. inst. & \\
-First Filter  & $\Downarrow$  &  & \\
- & $\approx$ 3000 Discr. Inst.  & $\Downarrow$ & \\
-Second Filter & $\Downarrow$  &  & \\
- & 600 Discr. Inst.  & 250  Cont. inst. & \\
+First Filter  & $\Downarrow$  & $\Downarrow$ & \\
+ & $\approx$ 3000 Discr. Inst.  & $\approx$ 1000 Cont. Inst. & \\
+Second Filter & $\Downarrow$  & $\Downarrow$  & \\
+% & 600 Discr. Inst.  & 250  Cont. inst. & \\
+  & 251 Discr. Inst.  & 116  Cont. inst. & \\
 \end{tabular}
@@
 \end{center}
-
-\paragraph{First Instances Filter.} As already mentioned an important characteristic of an instances is its computational difficulty, i.e., the CPU time needed by a complete solver (cf.~\S \ref{sec:algo}) to solve the instance to global optimality.
-Accordingly, for each of the gathered instance we run the solvers in {\tt Gams} (see Table \ref{}) able to solve it to global optimality.
-The number of solvers depends on the category of the instances under consideration.
-Then in order to reduce the number of instances we performed a first filter based on a relative measure of computational difficulty, i.e., we discarded all ``easy'' instances which are solved by at least 30\% of the complete solvers within a time limit of 30 seconds.
-Thanks to this first filter we obtained what we call the {\it filtered test bed}.
-
-%In Table \ref{tab:initialSet}, we report the number of instances in the filtered set subdivided according to the classification proposed in Section \ref{sec:classification}.
-We present in the table two different options for the field {\it Objective Functions}, i.e., Quadratic and Convex; three different options for the field {\it Variables}, i.e., Binary, Continuous and General and three different options for the field {\it Constraints}, i.e., Linear, Quadratic and Convex.
-The other categories of instances were not represented in the initial test bed.
-In particular the gathered a large number of Quadratic Binary Linear (QBL), Quadratic Continuous Quadratic (QCQ) and Quadratic General Quadratic (QGQ) instances, i.e., 1791 instances, 1957 instances and 2524 instances respectively.
-We gathered 901 Convex General Convex (CGC) instances.
-We gathered relatively few Quadratic Binary Quadratic (QBQ), Convex Continuous Convex (CCC) and Convex Mixed Convex (CMC) instances, i.e., 175 instances, 195 instances and 201 instances respectively.
-Finally we gathered only 17 Quadratic Mixed Linear (QML) instances.
-
-\bigskip
-The characteristics of the instances in the final library are presented in Table \ref{tab:FinalSet-D} for \emph{discrete} instances (*\{B,M,I,G\}*) and in Table \ref{tab:FinalSet-C} for \emph{continuous} ones (*C*).
 
@@
 
-%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-\subsection{Instance Format}\label{subsec:format}
-\framebox{TASK X : write }
-In the call for instances no specific formats were imposed
-%we did not impose a specific format
-for the submissions.
-%In this phase we also eliminated multiples copies of the same instance.
-To evaluate the instances we decided, for practical reasons, to use Gams as common platform for all the experiments involving commercial solvers.
-For this reason, we decided to translate all instances into the \texttt{.gms} Gams format.
- In a preliminary phase, all the instances received were divided according to their format and subsequently translated.
-In \S~\ref{subsec:tools}
-%~\ref{subsec:CP_convex}
-the tools used
-to translate an instance from a given format to the \texttt{.gms} format
-%to translate an instance from and to a given format to the \texttt{.gms} format
-are described in more detail.\\
+\subsection{Analysis of the final set of instances}\label{subsec:final set}
+In this section we analyse the feature of the instances selected to be
+part of the final set.
+The characteristics of the instances in the final library are
+presented in Table \ref{tab:FinalSet-D} for \emph{discrete} instances
+(*\{B,M,I,G\}*) and in Table \ref{tab:FinalSet-C} for
+\emph{continuous} ones (*C*).
+For each category, the Tables report in column $\#$ the corresponding
+number of instances.
+The final set well respects the original distribution of the gathered
+instances among the different categories.
+It is worth noticing that the classical discrete categories (such as
+(L,M,Q) or (Q,B,L)) are well represented by $118$ and $59$  instances
+respectively. On the other side, the classical continuous categories
+(such as (L,C,Q) and (QCQ)) are also well represented by $50$ and $17$
+ instances respectively.
+Moreover, the library covers a large majority of all possible
+categories of instances.
 
-In addition, we have introduced
-a specific format \texttt{.qplib}. This new format is capable of describing
-all the instances of the library in a sparse form.
-In comparison to a more \emph{high level} format like \texttt{.gms}, the new
-format presents two advantages:
-it is easier to read by a self-made parser and it produces smaller files.
-See Appendix~\label{sec:format} for details.
-%\textcolor{red}{ToDo: describe \texttt{.qplib} format. Maybe in an appendix?}
+To graphically represent the final set of instances we present some
+graphical representation of their main features.
 
-% is lighter in terms of sie
-%  significantly
-%less memory
-%in comparison and
+In Figure~\ref{fig:distribution} we represent in a logarithmic scale
+on the horizontal (resp. vertical) axis the number of variables (resp.
+constrains).
+Each  ``$+$'' point represents a continuous instances, while each
+``$\times$'' point represents a discrete one. The figure shows that
+the library contains instances of different size in terms of variables
+and constraints. The number of constraints goes from few units up to
+one hundred thousands constraints, while the number of variables goes
+up to almost forty-five thousands.
 
-% for this reason we used the \texttt{.gms} format as one of the format of the library. %checked if one instance was present copies of the same instance, potentially written in different formats
+In Figures~\ref{fig:pic_var_small}~\ref{fig:pic_var_medium}~\ref{fig:pic_var_large},
+for each instances we report the total number of variable with a
+``$+$'' and  the total number of discrete variables with a
+``$\times$'' (i.e. binary and integer variables). The instances are
+sorted accordingly to the total number of variables and, for reason of
+readability, we splitted the set into three pictures, firs one hundred
+instances are presented in Figures~\ref{fig:pic_var_small},
+then the second two hundred instances are presented in
+Figure~\ref{fig:pic_var_medium} and finally, the remaining $67$
+instances are presented in Figure~\ref{fig:pic_var_large}.
+From the picture we evince that for all instances dimention it is
+possible to find discrete instances with different percentages of
+continuous and purely continuous ones.
 
-\subsection{Instance Features}\label{subsec:feature}
-\framebox{TASK X : write }
+In Figures~\ref{fig:pic_constr_small}~\ref{fig:pic_constr_medium}~\ref{fig:pic_constr_big},
+for each instances we report the total number of constraints with a
+``$+$'' and the total number of quadratic (either convex of non
+convex) constraints with a ``$\times$''. The instances are sorted
+accordingly to the total number of constraints  and divided in the
+same manner as Figures~\ref{fig:pic_var_small}~\ref{fig:pic_var_medium}~\ref{fig:pic_var_large}.
+Also in this case, we can see that for all sizes it is possible to
+find instances with only linear constraints and instances with
+different percentage of linear and quadratic constraints.
 
-%Characteristics that should help to analyze the instances
-%
-%General characteristics:
-
-%Each QP instance can be classified accordingly to several features.
-For each instance of the starting set, the following features have been collected:
-\begin{itemize}
-\item Objective function characteristics:
-\begin{itemize}
-\item Type of objective function:  Linear, Convex or Quadratic.
-\item Density of the objective function, i.e. the percentage of nonzero entries of $Q_0$.
-\item Percentage of negative eigenvalues of $Q_0$.
-\end{itemize}
-\item Variables characteristics:
-\begin{itemize}
-\item Number of Continuous, Binary and Integer variables.
-%\item number of binary variables
-%\item number of integer variables
-\end{itemize}
-\item Constraints characteristics:
-\begin{itemize}
-\item Number of Linear, Convex and Quadratic constraints.
-\item Density of the constraints, i.e. the percentage of nonzero entries of the coefficients of the Linear, Continuous and Quadratic constraints.
-\end{itemize}
-\end{itemize}
-
-The mentioned features
-\begin{itemize}
-\item constraint-types: big-M constraints? box-constraints? combinatorial constraints? linear network?
-\item nonconvex: whether they can be turned into cones through appropriate decomposition
-\item quadratic: are discrete variables contained in the quadratic part or in the linear part only?
-\item note: some remark like "max-cut", "quadratic knapsack", etc.
-\end{itemize}
-
-Solver-dependent characteristics:
-\begin{itemize}
-\item time (?) [how measure time? what architecture will be used for tests? how to deal with parallel machines/algorithms?]
-\item number of nodes in case of branch-and-bound algorithms
-\item memory (?)
-\item matrix-free access (?)
-\item lower bound at root node
-\item time of finding first feasible solution
-\end{itemize}
-
-\begin{itemize}
-\item The static analysis is not enough to identify the hardness of the instances.
-\item \textcolor{red}{An empirical way for testing the hardness of one instance is the time needed to solve it}.
-\item We decided to use a \textcolor{red}{broad set of solvers} to test the computational hardness of the instances.
-\end{itemize}
+In Figures~\ref{fig:pic_density} and~\ref{fig:pic_neg_eig},  we
+focused on the instances with a quadratic objective function.
+In Figures~\ref{fig:pic_density} we plot the density of the Hessian
+matrix and we sort the instance according to it. We observe that the
+majority of the instances have either a low or high density. However,
+also intermediate values are present.
+In Figures~\ref{fig:pic_neg_eig} we present  the percentage of
+negative eigenvalues and we sort the instance according to it. As
+expected, a vast majority of the instances have $50\% $ of negative
+eigenvalues. In addition, around $40$ instances are convex (i.e. $0\%$
+of negative eigenvalues) but also instances of high or low percentage
+of negative eigenvalues are present.
 
 %- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-\subsection{Instance Selection}\label{subsec:selection}
-\framebox{TASK X : write }
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}\centering
+  \includegraphics[width=0.85\textwidth]{pic_overview.png}
+  \caption{Distribution of variables and constrains  of the qplib
+instances \label{fig:1}}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}\centering
+  \includegraphics[width=0.85\textwidth]{pic_var_small.png}
+  \caption{Number of variables (a) \label{fig:pic_var_small}}
+\end{figure}
+
+\begin{figure}\centering
+  \includegraphics[width=0.85\textwidth]{pic_var_medium.png}
+  \caption{Number of variables (b) \label{fig:pic_var_medium}}
+\end{figure}
+
+\begin{figure}\centering
+  \includegraphics[width=0.85\textwidth]{pic_var_big.png}
+  \caption{Number of variables (c) \label{fig:pic_var_large}}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}\centering
+  \includegraphics[width=0.85\textwidth]{pic_constr_small.png}
+  \caption{Number of constraints (a) \label{fig:pic_constr_small}}
+\end{figure}
+
+\begin{figure}\centering
+  \includegraphics[width=0.85\textwidth]{pic_constr_medium.png}
+  \caption{Number of constraints (b) \label{fig:pic_constr_medium}}
+\end{figure}
+
+\begin{figure}\centering
+  \includegraphics[width=0.85\textwidth]{pic_constr_big.png}
+  \caption{Number of constraints (c) \label{fig:pic_constr_big}}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\begin{figure}\centering
+%  \includegraphics[width=0.85\textwidth]{pic_quad_vs_convex_constr.png}
+%  \caption{Number of quadratic constraints (convex and non-convex)
+%\label{fig:pic_quad_vs_convex_constr}}
+%\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\begin{figure}\centering
+%  \includegraphics[width=0.85\textwidth]{pic_fuffa.png}
+%  \caption{...\label{fig:9}}
+%\end{figure}
+
+\begin{figure}\centering
+  \includegraphics[width=0.85\textwidth]{pic_density.png}
+  \caption{Density \label{fig:pic_density}}
+\end{figure}
+
+\begin{figure}\centering
+  \includegraphics[width=0.85\textwidth]{pic_neg_eig.png}
+  \caption{Negative eigenvalues \label{fig:pic_neg_eig}}
+\end{figure}
+
+%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
 

### 2016-12-17 conclusion.tex
Update conclusion.tex

@@
 
-We then describe the steps of the adopted process  to filter the gathered instances  in order to build the new library.
+We then describe the steps of the adopted process used to filter the gathered instances  in order to build the new library.
 The goals we purposed are of different natures and we tried to build a library which is computationally challenging and as broad as possible, i.e., it represents the largest possible categories of QP problems.
 
-We intentionally avoid to perform a computational comparison of the performances of the different solvers.
+We want to stress once again that we intentionally avoid to perform a computational comparison of the performances of the different solvers.
 Our goal is instead to provide a common test-bed of instances for practitioners and researchers in the field.

### 2016-12-17 QPLIB-3.tex
restored and updated

@@
 In this section we present all the steps performed in order to build
-the library. In particular we describe the first set of gathered
-instances (Section \ref{subsec:instColl}),
-we present the feature used to classified the instances (Section
-\ref{subsec:feature}),
-we discuss the issues concerning the format of the instances (Section
-\ref{subsec:format}),  and finally we describe the selection process
-used to filter the instances in order to construct the final library
-(Section \ref{subsec:selection}).
+the new library. In \S \ref{subsec:instColl}, we describe the set of gathered
+instances. In \S \ref{subsec:selection}
+we present the features used to classified the instances and 
+we discuss the issues concerning the format of the instances. Finally in \S \ref{subsec:final set}, we describe the selection process
+used to filter the instances and we graphically present the main features of selected instances.
 
@@
 The call remained open for 10 months, during which we received a large
-number of contributions with different characteristics. The instances
-we received  comes from theoretical studies and from real-world
+number of contributions of different natures. The instances
+we gathered come from theoretical studies and from real-world
 applications.
@@
 In addition to spontaneous contribution we analysed the other generic
-libraries of instances available  on internet and containing quadratic
-instances. Namely, we cite here the libraries from which we selected
-some quadratic instances:
+libraries of instances available  on internet and containing QP
+instances. Namely, we cite here the libraries from which we gathered instances:
 
@@
 Other quadratic instances were found in online libraries devoted to
-specific optimization problems that can be modelled as quadratic ones:
+specific QP problems as
 Max-Cut, Quadratic Assignment, Portfolio Optimization and several
@@
 the remaining ones contained only continuous variables.
-
+%
 More in details, we gathered a large number of Quadratic Binary Linear
@@
 To evaluate the instances we decided, for practical reasons, to use
-Gams as common platform for all the experiments involving commercial
+Gams as common platform for all the experiments involving commercial and non-commercial
 solvers.
 For this reason, we decided to translate all instances into the
-\texttt{.gms} Gams format and \texttt{.lp} format.
- In a preliminary phase, all the instances received were divided
-according to their format and subsequently translated.
+Gams format (\texttt{.gms}). 
+% In a preliminary phase, all the instances received were divided
+%according to their format and subsequently translated.
 %In \S~\ref{subsec:tools}
@@
 %%to translate an instance from and to a given format to the
-\texttt{.gms} format
+%\texttt{.gms} format
 %are described in more detail.\\
@@
 In addition, we have introduced
-a specific format \texttt{.qplib}. This new format is capable of describing
+a specific format called \texttt{.qplib}. This new format is capable of describing
 all the instances of the library in a sparse form.
@@
 it is easier to read by a self-made parser and it produces smaller files.
-See Appendix~\label{sec:format} for details.
+See the Appendix~\label{sec:format} for more details.
 %\textcolor{red}{ToDo: describe \texttt{.qplib} format. Maybe in an appendix?}
@@
 
-% for this reason we used the \texttt{.gms} format as one of the
-format of the library. %checked if one instance was present copies of
-the same instance, potentially written in different formats
+% for this reason we used the \texttt{.gms} format as one of the format of the library. %checked if one instance was present copies of the same instance, potentially written in different formats
 
-\subsection{Instance Features}\label{subsec:feature}
+%\subsection{Instance Features}\label{subsec:feature}
 
-For each instance of the starting set, we collected the following
-characteristics which allow to classify the instances into the
-categories described in Section~\ref{sec:QPbasic}.
+For each instance of the starting set, we collected important 
+characteristics which allowed us to classify the instances into the
+QP categories described in Section~\ref{sec:QPbasic}.
 
-As far as the variable type is concerned, we gather the following information:
+As far as the variable types are concerned, we collected the following information:
 \begin{itemize}
-\item number of binary variables (\emph{\# bin}),
-\item number of integer variables (\emph{\# int}),
-\item number of continuous variables (\emph{\# cont}).
+\item number of binary variables, % (\emph{\# bin}),
+\item number of integer variables, % (\emph{\# int}),
+\item number of continuous variables. % (\emph{\# cont}).
 \end{itemize}
+
 In case at least one binary or integer variable is present, then the
-instance is categorized as \emph{discrete} instance, on the other
+instance is categorized as a \emph{discrete} instance, on the other
 hand, in case only continuous variables are presents, the instance is
-categorized as \emph{continuous}.
+categorized as a \emph{continuous} instance.
 
@@
 \begin{itemize}
-\item Percentage of negative eigenvalues (\emph{\% neg eig}),
-\item Density of the Hessian of the objective function (\emph{\% dens}).
+\item percentage of negative eigenvalues of the $Q^0$ matrix,  % (\emph{\% neg eig}),
+\item density of the $Q^0$ matrix . %(\emph{\% dens}).
 \end{itemize}
- The number of negative eigenvalues of the Hessian matrix allows to
-identify the specific objective function type, i.e. in presence of at
-least one negative eigenvalue, the objective function is non convex.
-The density of the Hessian of the objective function is defined as the
-number of nonzero entries     over the total possible figure.
+ The number of negative eigenvalues of $Q^0$ allows us to
+identify the specific objective function type, i.e., in presence of at
+least one negative eigenvalue the objective function is non convex.
+The density of $Q^0$ is defined as the
+number of nonzero entries  over the total number.
 
-As far as the constraint type is concerned, we gather the following information:
+As far as the constraint types are concerned, we collected the following information:
 \begin{itemize}
-\item number of linear constraints (\emph{\# lin}),
-\item number of quadratic constraints (\emph{\# quad}),
-\item number of convex constraints (\emph{\# conv}),
-\item number of box constraints (\emph{\# box}).
+\item number of linear constraints, % (\emph{\# lin}),
+\item number of quadratic constraints, % (\emph{\# quad}),
+\item number of convex constraints, % (\emph{\# conv}),
+\item number of box constraints. %(\emph{\# box}).
 \end{itemize}
@@
 
-This information allows to have a static picture of the main features
-of the instances.
+This information allowed us to analyze the gathered instances and to perform the filters described in the the next paragraph.
 
@@
 
-The first goal of the library is to gather ``challenging'' instances,
-i.e., instances which can not be easily solved by the state of the art
+The first goal of the library is to represent the largest possible majority of
+the different categories of QP problems.
+The second goal of the library is to gather ``challenging'' instances,
+i.e., instances which can not be easily solved by  state-of-the-art
 solvers.
-The second goal of the library is to represent the largest majority of
-the different categories of quadratic problems.
 The third goal is to include in each of the categories a set of well
@@
 \begin{itemize}
+
 \item \emph{First Instances Filter.}\\
 The first filter is design to drastically reduce the number of
-instances by eliminating the easy instances. An empirical measure way
+instances by eliminating the ``easy'' ones. An empirical measure way
 of testing the hardness of an instance is the CPU time needed by a
@@
 global optimality.
-In this step, we decided to use the broadest set of solvers able to
-tackle the different categories of quadratic problems.\\
+%n this step, we decided to use the broadest set of solvers able to tackle the different categories of quadratic problems.\\
 Accordingly, for each of the gathered instance, we run the solvers in
-{\tt Gams} (see Table \ref{}) able to solve it to global optimality.
+{\tt Gams} (see Table \ref{t:solvers}) able to solve it to global optimality.
 The number of solvers depends on the category of the instances under
 consideration.
-We the filtered according to a relative measure of computational
+We then filtered according to a relative measure of computational
 difficulty, i.e., we discarded all instances that are solved by at
 least 30\% of the complete solvers within a time limit of 30 seconds.
+
 \item \emph{Second Instances Filter.}\\
 The goal of the second filter is to eliminate instances that are similar.
-We used the instance features presented in
-Section~\ref{subsec:feature} and the CPU solving time obtained in the
-first filter to guide this second step. \\
+%We used the instance features presented in
+%Section~\ref{subsec:feature} and the CPU solving time obtained in the
+%first filter to guide this second step. \\
 We carefully analysed the instances and we grouped them accordingly to
-similar features and CPU time. To reduce the size of the test-bed
+similar features. To reduce the size of the test-bed
 while keeping the representativeness of the library,   we kept for
-each cluster of similar instances one representative.\\
-In order to have a deeper insight on the computational hardness of the
-instances we performed a second round of tests on the resulting set of
-instances.  We used a general purpose solverand we set the time limit
-to 120 seconds. Finally, all the instances which have been solved to
-proven optimality have been discarded.
+each only few instances.
+Finally in order to have a deeper insight on the computational hardness of the
+instances, we performed a second round of tests on the filtered set of
+instances.  We used a general purpose solver able to tackle all QP categories and we set a time limit of 120 seconds. 
+All the instances which have been solved to
+proven optimality within this time limit have been discarded.
 \end{itemize}
@@
 
+\begin{center}
 \begin{table}
@@
 \end{tabular}
-\label{tab:FinalSet-D}
-\caption{Classification of the final set of discrete instances}
+\caption{Classification of the final set of discrete instances}\label{tab:DD}
 \end{table}
+\end{center}
 
+\begin{center}
 \begin{table}
@@
 \end{tabular}
-\label{tab:FinalSet-C}
-\caption{Classification of the final set of continuous instances}
+\caption{Classification of the final set of continuous instances}\label{tab:CC}
 \end{table}
+\end{center}
 
@@
 In this section we analyse the feature of the instances selected to be
-part of the final set.
+part of library.
 The characteristics of the instances in the final library are
-presented in Table \ref{tab:FinalSet-D} for \emph{discrete} instances
-(*\{B,M,I,G\}*) and in Table \ref{tab:FinalSet-C} for
+presented in Table \ref{tab:DD} for \emph{discrete} instances
+(*\{B,M,I,G\}*) and in Table \ref{tab:CC} for
 \emph{continuous} ones (*C*).
@@
 
-To graphically represent the final set of instances we present some
-graphical representation of their main features.
+To graphically represent the library we report some
+graphical representations of the instance main features.
 
-In Figure~\ref{fig:distribution} we represent in a logarithmic scale
+In Figure~\ref{fig:distribution} we represent 
 on the horizontal (resp. vertical) axis the number of variables (resp.
-constrains).
-Each  ``$+$'' point represents a continuous instances, while each
-``$\times$'' point represents a discrete one. The figure shows that
+constrains). Both axis are in logarithmic scale.
+The ``$+$'' points represent  the continuous instances, while the 
+``$\times$'' points represent the discrete instances. The figure shows that
 the library contains instances of different size in terms of variables
@@
 In Figures~\ref{fig:pic_var_small}~\ref{fig:pic_var_medium}~\ref{fig:pic_var_large},
-for each instances we report the total number of variable with a
-``$+$'' and  the total number of discrete variables with a
-``$\times$'' (i.e. binary and integer variables). The instances are
+for each instances we report the total number of variables with a
+``$+$'' point and  the total number of discrete variables with a
+``$\times$'' point (i.e. binary and integer variables). The instances are
 sorted accordingly to the total number of variables and, for reason of
-readability, we splitted the set into three pictures, firs one hundred
+readability, we splitted the figure into three sub figures. The first one hundred
 instances are presented in Figures~\ref{fig:pic_var_small},
@@
 instances are presented in Figure~\ref{fig:pic_var_large}.
-From the picture we evince that for all instances dimention it is
+From the picture we evince that for all instance dimensions it is
 possible to find discrete instances with different percentages of
@@
 for each instances we report the total number of constraints with a
-``$+$'' and the total number of quadratic (either convex of non
-convex) constraints with a ``$\times$''. The instances are sorted
+``$+$'' point and the total number of quadratic (either convex of non
+convex) constraints with a ``$\times$'' point. The instances are sorted
 accordingly to the total number of constraints  and divided in the
-same manner as Figures~\ref{fig:pic_var_small}~\ref{fig:pic_var_medium}~\ref{fig:pic_var_large}.
+same manner as for Figures~\ref{fig:pic_var_small}~\ref{fig:pic_var_medium}~\ref{fig:pic_var_large}.
 Also in this case, we can see that for all sizes it is possible to
@@
 focused on the instances with a quadratic objective function.
-In Figures~\ref{fig:pic_density} we plot the density of the Hessian
+In Figures~\ref{fig:pic_density} we plot the density of the $Q^0$
 matrix and we sort the instance according to it. We observe that the
 majority of the instances have either a low or high density. However,
-also intermediate values are present.
+also intermediate values are present. It is worth noticing that around 60 instances have a 
+density less that 5\% (see the Appendix for further details). 
 In Figures~\ref{fig:pic_neg_eig} we present  the percentage of
@@
 eigenvalues. In addition, around $40$ instances are convex (i.e. $0\%$
-of negative eigenvalues) but also instances of high or low percentage
-of negative eigenvalues are present.
+of negative eigenvalues) but also instances of high or low percentages
+of negative eigenvalues are present in the library.
 

### 2016-12-18 QPLIB-3.tex
checked up to §3.1, will keep checking the rest

@@
 
-In this section we present all the steps performed in order to build
-the new library. In \S \ref{subsec:instColl}, we describe the set of gathered
-instances. In \S \ref{subsec:selection}
-we present the features used to classified the instances and 
-we discuss the issues concerning the format of the instances. Finally in \S \ref{subsec:final set}, we describe the selection process
-used to filter the instances and we graphically present the main features of selected instances.
+In this section we present all the steps performed in order to build the
+new library. In \S \ref{subsec:instColl}, we describe the set of gathered
+instances. In \S \ref{subsec:selection} we present the features used to
+classify the instances and we discuss the issues concerning the format of
+the instances. Finally, in \S \ref{subsec:final set}, we describe the
+selection process used to filter the instances and we graphically present
+the main features of the selected instances.
 
@@
 instances. In January $2014$, we issued an online call for instances
-using the principal international mailing lists.
-In particular, we addressed the mailing lists of the mathematical
+using the main international mailing lists of the mathematical
 optimization and numerical analysis communities, reaching in this way
@@
 The call remained open for 10 months, during which we received a large
-number of contributions of different natures. The instances
-we gathered come from theoretical studies and from real-world
+number of contributions of different nature. The instances we gathered
+come both from theoretical studies as well as from real-world
 applications.
@@
 libraries of instances available  on internet and containing QP
-instances. Namely, we cite here the libraries from which we gathered instances:
-
+instances. Namely, the libraries from which we gathered instances are
+%
 \begin{itemize}
- \item {\tt BARON Library}
-\url{http://www.minlp.com/nlp-and-minlp-test-problems}
- \item {\tt CUTEst Library}
-\url{https://ccpforge.cse.rl.ac.uk/gf/project/cutest/wiki}
- \item {\tt GAMS Library}
-\url{http://www.gamsworld.org/performance/performlib.htm}
- \item {\tt MacMINLP} \url{https://wiki.mcs.anl.gov/leyffer/index.php/MacMINLP}
- \item {\tt Meszaros} \url{http://www.doc.ic.ac.uk/~im/00README.QP}
- \item {\tt MINLP} \url{http://www.gamsworld.org/minlp/minlplib.htm}
- \item {\tt POLIP} \url{http://polip.zib.de/pipformat.php}
+ \item the {\tt BARON} library
+ \url{http://www.minlp.com/nlp-and-minlp-test-problems};
+%
+\item the {\tt CUTEst} library
+ \url{https://ccpforge.cse.rl.ac.uk/gf/project/cutest/wiki};
+%
+\item the {\tt GAMS} library
+ \url{http://www.gamsworld.org/performance/performlib.htm};
+%
+\item the {\tt MacMINLP} library
+ \url{https://wiki.mcs.anl.gov/leyffer/index.php/MacMINLP};
+%
+\item the {\tt Meszaros} library
+ \url{http://www.doc.ic.ac.uk/~im/00README.QP};
+%
+\item the {\tt MINLP} library
+ \url{http://www.gamsworld.org/minlp/minlplib.htm};
+%
+\item the {\tt POLIP} library
+ \url{http://polip.zib.de/pipformat.php}.
 \end{itemize}
@@
 Other quadratic instances were found in online libraries devoted to
-specific QP problems as
-Max-Cut, Quadratic Assignment, Portfolio Optimization and several
-others.
+specific QP problems as Max-Cut, Quadratic Assignment, Portfolio
+Optimization, and several others.
 
@@
 instances. Three fourths of them contained discrete variables, while
-the remaining ones contained only continuous variables.
-%
-More in details, we gathered a large number of Quadratic Binary Linear
-(QBL), Quadratic Continuous Quadratic (QCQ) and Quadratic General
-Quadratic (QGQ) instances, i.e., $\approx 1800$ instances, $\approx
-2000$ instances and $\approx 2500$ instances respectively.
-We gathered $\approx 1000$ Convex General Convex (CGC) instances. We
-gathered relatively few Quadratic Binary Quadratic (QBQ), Convex
+the remaining ones contained only continuous variables. More in details,
+we gathered $\approx 1800$ Quadratic Binary Linear (QBL) instances,
+$\approx 2000$ Quadratic Continuous Quadratic (QCQ) instances, and
+and $\approx 2500$ Quadratic General Quadratic (QGQ) instances. W
+also gathered $\approx 1000$ Convex General Convex (CGC) instances. We
+gathered relatively fewer Quadratic Binary Quadratic (QBQ), Convex
 Continuous Convex (CCC) and Convex Mixed Convex (CMC) instances, i.e.,
 $\approx 150$ instances, $\approx 200$ instances and $\approx 200$
-instances respectively. Finally we gathered only $17$ Quadratic Mixed
-Linear (QML) instances.
+instances respectively. Finally, we gathered only $17$ Quadratic Mixed
+Linear (QML) instances. In the call for instances, no specific formats
+requirements were imposed for the submissions.
 
-In the call for instances no specific formats were imposed
-%we did not impose a specific format
-for the submissions.
-%In this phase we also eliminated multiples copies of the same instance.
 To evaluate the instances we decided, for practical reasons, to use
-Gams as common platform for all the experiments involving commercial and non-commercial
-solvers.
-For this reason, we decided to translate all instances into the
-Gams format (\texttt{.gms}). 
+Gams as common platform for all the experiments involving commercial and
+non-commercial solvers. For this reason, we decided to translate all
+instances into the Gams format (\texttt{.gms}). 
 % In a preliminary phase, all the instances received were divided
@@
 %are described in more detail.\\
-
-In addition, we have introduced
-a specific format called \texttt{.qplib}. This new format is capable of describing
-all the instances of the library in a sparse form.
+%
+In addition, we have introduced a new, specific \texttt{.qplib} format. This
+new is capable of describing all the instances of the library in a sparse form.
 In comparison to a more \emph{high level} format like \texttt{.gms}, the new
-format presents two advantages:
-it is easier to read by a self-made parser and it produces smaller files.
-See the Appendix~\label{sec:format} for more details.
-%\textcolor{red}{ToDo: describe \texttt{.qplib} format. Maybe in an appendix?}
-
-% is lighter in terms of sie
-%  significantly
-%less memory
-%in comparison and
-
-% for this reason we used the \texttt{.gms} format as one of the format of the library. %checked if one instance was present copies of the same instance, potentially written in different formats
+format presents two advantages: it is easier to read by a self-made parser and
+it produces smaller files. See the Appendix~\label{sec:format} for more details.
 
@@
 
-For each instance of the starting set, we collected important 
-characteristics which allowed us to classify the instances into the
-QP categories described in Section~\ref{sec:QPbasic}.
-
-As far as the variable types are concerned, we collected the following information:
+For each instance of the starting set, we collected important characteristics
+which allowed us to classify the instances into the QP categories described in
+Section~\ref{sec:QPbasic}. As far as the variable types are concerned, we
+collected the following information:
+%
 \begin{itemize}
-\item number of binary variables, % (\emph{\# bin}),
-\item number of integer variables, % (\emph{\# int}),
+ \item number of binary variables; % (\emph{\# bin}),
+ %
+ \item number of integer variables; % (\emph{\# int}),
+ %
  \item number of continuous variables. % (\emph{\# cont}).
 \end{itemize}
-
-In case at least one binary or integer variable is present, then the
-instance is categorized as a \emph{discrete} instance, on the other
-hand, in case only continuous variables are presents, the instance is
-categorized as a \emph{continuous} instance.
-
-As fas as the objective function is concerned, we gathered the
-following information:
+%
+In case at least one binary or integer variable is present, then the instance is
+categorized as \emph{discrete}, otherwise it is categorized as \emph{continuous}.
+As fas as the objective function is concerned, we gathered the following
+information:
+%
 \begin{itemize}
-\item percentage of negative eigenvalues of the $Q^0$ matrix,  % (\emph{\% neg eig}),
-\item density of the $Q^0$ matrix . %(\emph{\% dens}).
+ \item percentage of negative eigenvalues of the $Q^0$ matrix;
+       % (\emph{\% neg eig}),
+ %
+ \item density of the $Q^0$ matrix (number of nonzero entries over the total
+       number). %(\emph{\% dens}).
 \end{itemize}
- The number of negative eigenvalues of $Q^0$ allows us to
-identify the specific objective function type, i.e., in presence of at
-least one negative eigenvalue the objective function is non convex.
-The density of $Q^0$ is defined as the
-number of nonzero entries  over the total number.
-
-As far as the constraint types are concerned, we collected the following information:
+%
+The number of negative eigenvalues of $Q^0$ allows us to identify the
+objective function type, as in presence of at least one negative eigenvalue
+the objective function is non convex. Finally, as far as the constraint types
+are concerned, we collected the following information:
+%
 \begin{itemize}
  \item number of linear constraints, % (\emph{\# lin}),
+ %
  \item number of quadratic constraints, % (\emph{\# quad}),
+ %
  \item number of convex constraints, % (\emph{\# conv}),
+ %
  \item number of box constraints. %(\emph{\# box}).
 \end{itemize}
-A constraint is considered quadratic if it contain a quadratic term.
-Among the quadratic constraints, the one with only non negative
-eigenvalues are classified as convex constraints.
+%
+A constraint is considered quadratic if it contains at least one nonzero in
+the quadratic term. Among the quadratic constraints, the ones with only
+non-negative eigenvalues are classified as convex constraints. All this
+information allowed us to analyze the gathered instances and to perform the
+filters described in the the next paragraph.
 
-This information allowed us to analyze the gathered instances and to perform the filters described in the the next paragraph.
-
+%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
 \subsection{Instance Selection}\label{subsec:selection}
@@
 
+%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
 \subsection{Analysis of the final set of instances}\label{subsec:final set}
+
 In this section we analyse the feature of the instances selected to be

### 2016-12-18 QPLIB-3.tex
Update QPLIB-3.tex
checked up to §3.2, will continue later

@@
 
-The first goal of the library is to represent the largest possible majority of
-the different categories of QP problems.
-The second goal of the library is to gather ``challenging'' instances,
-i.e., instances which can not be easily solved by  state-of-the-art
-solvers.
-The third goal is to include in each of the categories a set of well
-diversified instances.
-Finally, the fourth goal is to obtain a library whose set of instances
-is neither too small either too large, to preserve statistical
-relevance while at the same time being computationally manageable.
-
+During the development of the library, a discussion ensued about what
+the expected gols that we wanted to achieve. The following four goals
+where finally identified:
+%
+\begin{enumerate}
+ \item represent as much as possible all the different categories of QP
+       problems;
+ %
+ \item gather ``challenging'' instances, i.e., ones which can not be easily
+       solved by  state-of-the-art solvers;
+ %
+ \item include for each of the categories a set of well-diversified
+       instances;
+ %
+ \item obtain a set of instances which is neither too small, so as to
+       preserve statistical relevance, nor too large so as to being
+       computationally manageable.
+\end{enumerate
+%
 To achieve the aforementioned goals, we performed the following two
-filters applied in cascade.
-
+filters, applied in cascade.
+%
 \begin{itemize}
-
  \item \emph{First Instances Filter.}\\
-The first filter is design to drastically reduce the number of
-instances by eliminating the ``easy'' ones. An empirical measure way
-of testing the hardness of an instance is the CPU time needed by a
-complete solver (cf.~\S \ref{sec:algo}) to solve the instance to
-global optimality.
-%n this step, we decided to use the broadest set of solvers able to tackle the different categories of quadratic problems.\\
-Accordingly, for each of the gathered instance, we run the solvers in
-{\tt Gams} (see Table \ref{t:solvers}) able to solve it to global optimality.
-The number of solvers depends on the category of the instances under
-consideration.
-We then filtered according to a relative measure of computational
-difficulty, i.e., we discarded all instances that are solved by at
-least 30\% of the complete solvers within a time limit of 30 seconds.
-
+       The first filter was designed to drastically reduce the number of
+       instances by eliminating the ``easy'' ones. An empirical measure
+       for the hardness of an instance is the CPU time needed by a
+       complete solver (cf.~\S \ref{sec:algo}) to solve it to
+       global optimality. Accordingly, for each of the gathered instance we
+       ran the complete solvers in {\tt Gams} (cf.~Table \ref{t:solvers})
+       capable of solving it, whose number depends on the category of the
+       instance under consideration. We then filtered according to a relative
+       measure of computational difficulty, i.e., we discarded all instances
+       that are solved by at least 30\% of the complete solvers within a time
+       limit of 30 seconds.
+ %
  \item \emph{Second Instances Filter.}\\
-The goal of the second filter is to eliminate instances that are similar.
-%We used the instance features presented in
-%Section~\ref{subsec:feature} and the CPU solving time obtained in the
-%first filter to guide this second step. \\
-We carefully analysed the instances and we grouped them accordingly to
-similar features. To reduce the size of the test-bed
-while keeping the representativeness of the library,   we kept for
-each only few instances.
-Finally in order to have a deeper insight on the computational hardness of the
-instances, we performed a second round of tests on the filtered set of
-instances.  We used a general purpose solver able to tackle all QP categories and we set a time limit of 120 seconds. 
-All the instances which have been solved to
-proven optimality within this time limit have been discarded.
+       The goal of the second filter was to eliminate ``similar'' instances.
+       We carefully analysed the instances one by one, and we clustered them
+       according to their features; for each cluster we kept only a few
+       representatives. Finally, in order to only keep computationally
+       challenging instances we ran a general purpose solver able to tackle
+       all QP categories with a time limit of 120 seconds; all the instances
+       which have been solved to proven optimality within this time limit were
+       discarded.
 \end{itemize}
+%
 In Table~\ref{tab:filters} we summarize the two filter steps, which
-allows us to identify the final set of $251$ discrete instances and
+allowed us to identify the final set of $251$ discrete instances and
 $116$ continuous instances.
@@
  \centering
-
  \setlength{\tabcolsep}{5pt}
@@
 \end{center}
+
+%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
+\subsection{Analysis of the final set of instances}\label{subsec:final set}
+
+In this section we analyse the feature of the instances selected to be
+part of library.
+The characteristics of the instances in the final library are
+presented in Table \ref{tab:DD} for \emph{discrete} instances
+(*\{B,M,I,G\}*) and in Table \ref{tab:CC} for
+\emph{continuous} ones (*C*).
+For each category, the Tables report in column $\#$ the corresponding
+number of instances.
+The final set well respects the original distribution of the gathered
+instances among the different categories.
+It is worth noticing that the classical discrete categories (such as
+(L,M,Q) or (Q,B,L)) are well represented by $118$ and $59$  instances
+respectively. On the other side, the classical continuous categories
+(such as (L,C,Q) and (QCQ)) are also well represented by $50$ and $17$
+ instances respectively.
+Moreover, the library covers a large majority of all possible
+categories of instances.
 
@@
 \end{center}
-
-%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-\subsection{Analysis of the final set of instances}\label{subsec:final set}
-
-In this section we analyse the feature of the instances selected to be
-part of library.
-The characteristics of the instances in the final library are
-presented in Table \ref{tab:DD} for \emph{discrete} instances
-(*\{B,M,I,G\}*) and in Table \ref{tab:CC} for
-\emph{continuous} ones (*C*).
-For each category, the Tables report in column $\#$ the corresponding
-number of instances.
-The final set well respects the original distribution of the gathered
-instances among the different categories.
-It is worth noticing that the classical discrete categories (such as
-(L,M,Q) or (Q,B,L)) are well represented by $118$ and $59$  instances
-respectively. On the other side, the classical continuous categories
-(such as (L,C,Q) and (QCQ)) are also well represented by $50$ and $17$
- instances respectively.
-Moreover, the library covers a large majority of all possible
-categories of instances.
 

### 2016-12-18 QPLIB-3.tex
typo updated -- missing }
other typo -- format missing
Update QPLIB-3.tex

@@
 In addition, we have introduced a new, specific \texttt{.qplib} format. This
-new is capable of describing all the instances of the library in a sparse form.
+new format is capable of describing all the instances of the library in a sparse form.
 In comparison to a more \emph{high level} format like \texttt{.gms}, the new
@@
        computationally manageable.
-\end{enumerate
+\end{enumerate}
 %
@@
 possible to find discrete instances with different percentages of
-continuous and purely continuous ones.
+continuous variables and purely continuous ones.
 

### 2016-12-18 QPLIB-3.tex
Update QPLIB-3.tex
Finished checking all of §3; mostly cosmetic changes, with a few comments. AF

@@
        according to their features; for each cluster we kept only a few
-       representatives. Finally, in order to only keep computationally
-       challenging instances we ran a general purpose solver able to tackle
-       all QP categories with a time limit of 120 seconds; all the instances
-       which have been solved to proven optimality within this time limit were
-       discarded.
+       representatives. {\bf Comment: we may want to be a tad more specific
+       about what ``features'' exactly meant there}. Finally, in order to only
+       keep computationally challenging instances we ran a general purpose solver
+       able to tackle all QP categories with a time limit of 120 seconds; all the
+       instances which have been solved to proven optimality within this time limit
+       were discarded. {\bf Comment: this might actually be best characterized as
+       a third filter.}
 \end{itemize}
@@
 
-In this section we analyse the feature of the instances selected to be
-part of library.
-The characteristics of the instances in the final library are
-presented in Table \ref{tab:DD} for \emph{discrete} instances
-(*\{B,M,I,G\}*) and in Table \ref{tab:CC} for
-\emph{continuous} ones (*C*).
-For each category, the Tables report in column $\#$ the corresponding
-number of instances.
-The final set well respects the original distribution of the gathered
-instances among the different categories.
-It is worth noticing that the classical discrete categories (such as
-(L,M,Q) or (Q,B,L)) are well represented by $118$ and $59$  instances
-respectively. On the other side, the classical continuous categories
-(such as (L,C,Q) and (QCQ)) are also well represented by $50$ and $17$
- instances respectively.
-Moreover, the library covers a large majority of all possible
-categories of instances.
+We now analyse the features of the instances selected to be part of the
+library. The characteristics of the instances are presented in Table
+\ref{tab:DD} for \emph{discrete} instances (*\{B,M,I,G\}*) and in Table
+\ref{tab:CC} for \emph{continuous} ones (*C*). For each category, the Tables
+report in column $\#$ the corresponding number of instances. It can be seen
+that the final set well respects the original distribution of the gathered
+instances among the different categories. Indeed, the discrete categories
+(LMQ) or (QBL) are well represented by $118$ and $59$ instances,
+respectively. Similarly, the continuous categories (LCQ) and (QCQ) are well
+represented by $50$ and $17$ instances, respectively. Moreover, the library
+actually covers the large majority of all possible categories of instances.
 
@@
 \end{tabular}
-\caption{Classification of the final set of discrete instances}\label{tab:DD}
+\caption{Classification of the final set of discrete instances}
+\label{tab:DD}
 \end{table}
@@
 \end{tabular}
-\caption{Classification of the final set of continuous instances}\label{tab:CC}
+\caption{Classification of the final set of continuous instances}
+\label{tab:CC}
 \end{table}
@@
 
-To graphically represent the library we report some
-graphical representations of the instance main features.
-
-In Figure~\ref{fig:distribution} we represent 
-on the horizontal (resp. vertical) axis the number of variables (resp.
-constrains). Both axis are in logarithmic scale.
-The ``$+$'' points represent  the continuous instances, while the 
-``$\times$'' points represent the discrete instances. The figure shows that
-the library contains instances of different size in terms of variables
-and constraints. The number of constraints goes from few units up to
-one hundred thousands constraints, while the number of variables goes
-up to almost forty-five thousands.
-
-In Figures~\ref{fig:pic_var_small}~\ref{fig:pic_var_medium}~\ref{fig:pic_var_large},
-for each instances we report the total number of variables with a
-``$+$'' point and  the total number of discrete variables with a
-``$\times$'' point (i.e. binary and integer variables). The instances are
-sorted accordingly to the total number of variables and, for reason of
-readability, we splitted the figure into three sub figures. The first one hundred
-instances are presented in Figures~\ref{fig:pic_var_small},
-then the second two hundred instances are presented in
-Figure~\ref{fig:pic_var_medium} and finally, the remaining $67$
-instances are presented in Figure~\ref{fig:pic_var_large}.
-From the picture we evince that for all instance dimensions it is
-possible to find discrete instances with different percentages of
-continuous variables and purely continuous ones.
-
-In Figures~\ref{fig:pic_constr_small}~\ref{fig:pic_constr_medium}~\ref{fig:pic_constr_big},
-for each instances we report the total number of constraints with a
-``$+$'' point and the total number of quadratic (either convex of non
-convex) constraints with a ``$\times$'' point. The instances are sorted
-accordingly to the total number of constraints  and divided in the
-same manner as for Figures~\ref{fig:pic_var_small}~\ref{fig:pic_var_medium}~\ref{fig:pic_var_large}.
-Also in this case, we can see that for all sizes it is possible to
-find instances with only linear constraints and instances with
-different percentage of linear and quadratic constraints.
-
-In Figures~\ref{fig:pic_density} and~\ref{fig:pic_neg_eig},  we
-focused on the instances with a quadratic objective function.
-In Figures~\ref{fig:pic_density} we plot the density of the $Q^0$
-matrix and we sort the instance according to it. We observe that the
-majority of the instances have either a low or high density. However,
-also intermediate values are present. It is worth noticing that around 60 instances have a 
-density less that 5\% (see the Appendix for further details). 
-In Figures~\ref{fig:pic_neg_eig} we present  the percentage of
-negative eigenvalues and we sort the instance according to it. As
-expected, a vast majority of the instances have $50\% $ of negative
-eigenvalues. In addition, around $40$ instances are convex (i.e. $0\%$
-of negative eigenvalues) but also instances of high or low percentages
-of negative eigenvalues are present in the library.
-
-%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
+We now report some graphs that help in illustrating the main features
+of the instences. In Figure~\ref{fig:distribution} we plot the number
+of variables (horizontal axis) versus the number of constrains
+(vertical axis), both in logarithmic scale. Continuous instances are
+marked with ``$+$'', and discrete ones with ``$\times$''. The figure shows
+that the library contains a quite diverse set of instances of different in
+terms of number of variables and constraints. The number of constraints goes
+from few units up to one hundred thousands constraints, while the number of
+variables goes up to almost forty-five thousands.
 
@@
 \end{figure}
+
+Figure~\ref{fig:pic_var_small}, Figure~\ref{fig:pic_var_medium} and
+Figure~\ref{fig:pic_var_large} describe how discrete and continuous
+variables are distributed within the instances. The instances are
+sorted accordingly to the total number of variables and, for reason of
+readability, the graph is splitted in the three figures:
+Figure~\ref{fig:pic_var_small} describes the 100 smaller instances,
+Figure~\ref{fig:pic_var_medium} the 200 medium ones, and
+Figure~\ref{fig:pic_var_large} the 67 largest ones. For each instance
+we report the total number of variables with a ``$+$'', and the total
+number of discrete variables (binary or general integer) with a
+``$\times$''. The pictures clearly show that instances with different
+percentages of integer and continuous variables are present in the
+library, and well distributed across the whole spectrum of variable
+sizes.
 
@@
 
+Similarly, Figures~\ref{fig:pic_constr_small}, Figure~\ref{fig:pic_constr_medium}
+and Figure~\ref{fig:pic_constr_big} describe how linear and quadratic
+constraints are distributed within the instances. The instances are sorted
+accordingly to the total number of constraints and divided in the
+same manner as for Figures~\ref{fig:pic_var_small}--\ref{fig:pic_var_large}.
+For each instances we report the total number of constraints with a ``$+$''
+and the total number of (either convex of nonconvex) quadratic constraints
+with a ``$\times$''. Also in this case, different percentages of linear and
+quadratic constraints are present and well-distributed across the spectrum
+of constraint sizes, although both medium- and large-size instances show a
+prevalence of lower percentages of quadratic constraints. In particular,
+from Figure~\ref{fig:pic_constr_big} we learn that while the maximum number
+of linear constraints exceeds 100000, the maximum number of quadratic
+constraints tops up at around 10000. This is, however, reasonable, considering
+how quadratic constraints can, in general, be expected to be much more
+computationally challenging than linear ones, especially if nonconvex.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%\begin{figure}\centering
-%  \includegraphics[width=0.85\textwidth]{pic_quad_vs_convex_constr.png}
-%  \caption{Number of quadratic constraints (convex and non-convex)
-%\label{fig:pic_quad_vs_convex_constr}}
-%\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Speaking of nonconvexity, Figure~\ref{fig:pic_neg_eig} focuses on
+the instances with a quadratic objective function, ordered by the
+the percentage of negative eigenvalues (vertical axis). Although
+almost all percentages are represented, the instances are mostly
+clustered around two values. About 25\% of the instances are convex,
+i.e., they have 0\% of negative eigenvalues. Among the others, a vast
+majority have around 50\% of negative eigenvalues. However, instances
+with high or low percentages of negative eigenvalues are present.
 
-%\begin{figure}\centering
-%  \includegraphics[width=0.85\textwidth]{pic_fuffa.png}
-%  \caption{...\label{fig:9}}
-%\end{figure}
+\begin{figure}\centering
+  \includegraphics[width=0.85\textwidth]{pic_neg_eig.png}
+  \caption{Negative eigenvalues \label{fig:pic_neg_eig}}
+\end{figure}
+
+Similarly, Figure~\ref{fig:pic_density} shows the instances with a
+quadratic objective function sorted according to the density of the
+$Q^0$ matrix (vertical axis). The majority of the instances have either
+a very low or a rather high density: indeed, about 30\% of the instances
+have density smaller that 5\%, and about 30\% of the instances have
+density larger that 50\%. However, also intermediate values are present.
 
@@
 
-\begin{figure}\centering
-  \includegraphics[width=0.85\textwidth]{pic_neg_eig.png}
-  \caption{Negative eigenvalues \label{fig:pic_neg_eig}}
-\end{figure}
-
-%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
+{\bf Nothing about convexity of constraints? A graph plotting the
+number of convex quadratic constraints vs the total number of quadratic
+constraints?}
 

### 2016-12-20 QPLIB-1.0.tex
updated application table

@@
 
-\framebox{TASK 2 : write motivation}
-
 Optimization problems with quadratic constraints and/or objective function (QP) have been the subject of a considerable amount of research over the last 60 years.
@@
 On the other hand, (MI)QP is, in some sense, a considerably more expressive class than (MI)LP.
-Quadratic expressions are found, either naturally or after appropriate reformulations, in very many practical problems \cite{Kochenberger2014}.
-We survey several of these applications in Table \ref{tbl:application}.
-In general, any continuous function can be approximated with arbitrary accuracy (over a compact set) by a polynomial of arbitrary degree; in turn, every polynomial can be broken down to a system of quadratic expressions.
+Quadratic expressions are found, either naturally or after appropriate reformulations, in very many optimization problems \cite{Kochenberger2014}.
+Table \ref{tbl:application} provides a certainly non-exhaustive collection of applications that lead to formulations with either quadratic constraints, ot quadratic objective function, or both.
+Also, in general any continuous function can be approximated with arbitrary accuracy (over a compact set) by a polynomial of arbitrary degree; in turn, every polynomial can be broken down to a system of quadratic expressions.
 Hence, (MI)QP is, in some sense, roughly as expressive as the whole of (MI)NLP.
@@
 \toprule
-Problem & Discrete? & Contributions \\
+Problem & Discrete & Contributions \\
 \midrule
@@
 %
-\multirow{1}{.55\textwidth}{{\bf Finance:} Portfolio selection optimization} && \multirow{1}{.25\textwidth}{\cite{deng-etal:2013,kallrath:2003,lin-etal:2005,maranas-etal:1997,parpas-rustem:2006,rios-sahinidis:2010}} \\
+\multirow{1}{.55\textwidth}{{\bf Finance:} Portfolio optimization} & checkmark & \multirow{1}{.25\textwidth}{\cite{deng-etal:2013,kallrath:2003,lin-etal:2005,maranas-etal:1997,parpas-rustem:2006,rios-sahinidis:2010,FrFG16,FrGe06a,FrGe07a,FrGe09a}} \\
 %
@@
 \multirow{1}{.55\textwidth}{Multi-commodity flow} & \checkmark & \cite{tadayon-smith:2013} \\ \cmidrule(rl){2-3}
+%
+\multirow{1}{.55\textwidth}{Quadratic network design} & \checkmark & \cite{FrFG16,FGGP11} \\ \cmidrule(rl){2-3}
 %
@@
 %
+\multicolumn{3}{l}{{\bf Telecommunications}} \\[1pt]
+%
+\multirow{1}{.55\textwidth}{Delay-constrained routing} & \checkmark & \cite{FrGaSc14,FrGaSt16a} \\
+%
+%
+\multicolumn{3}{l}{{\bf Energy}} \\[1pt]
+%
+\multirow{1}{.55\textwidth}{Unit-commitment problem} & \checkmark & \cite{Tetal15} \\
+%
 \end{longtable}

### 2016-12-27 QPLIB-3.tex
Update QPLIB-3.tex

@@
 
-Figure~\ref{fig:pic_var_small}, Figure~\ref{fig:pic_var_medium} and
-Figure~\ref{fig:pic_var_large} describe how discrete and continuous
+%Figure~\ref{fig:pic_var_small}, Figure~\ref{fig:pic_var_medium} and
+Figure~\ref{fig:pic_var} 
+describes how discrete and continuous
 variables are distributed within the instances. The instances are
-sorted accordingly to the total number of variables and, for reason of
-readability, the graph is splitted in the three figures:
-Figure~\ref{fig:pic_var_small} describes the 100 smaller instances,
-Figure~\ref{fig:pic_var_medium} the 200 medium ones, and
-Figure~\ref{fig:pic_var_large} the 67 largest ones. For each instance
-we report the total number of variables with a ``$+$'', and the total
-number of discrete variables (binary or general integer) with a
-``$\times$''. The pictures clearly show that instances with different
-percentages of integer and continuous variables are present in the
-library, and well distributed across the whole spectrum of variable
-sizes.
+sorted accordingly to the total number of variables.
+% and, for reason of
+%readability, the graph is splitted in the three figures:
+%Figure~\ref{fig:pic_var_small} describes the 100 smaller instances,
+%Figure~\ref{fig:pic_var_medium} the 200 medium ones, and
+%Figure~\ref{fig:pic_var_large} the 67 largest ones. 
+For each instance we report the total number of variables with a ``$+$'', and the total number of discrete variables (binary or general integer) with a ``$\times$''.
+The pictures clearly show that instances with different percentages of integer and continuous variables are present in the library, and well distributed across the whole spectrum of variable sizes.
 
@@
 \begin{figure}\centering
-  \includegraphics[width=0.85\textwidth]{pic_var_small.png}
-  \caption{Number of variables (a) \label{fig:pic_var_small}}
+  \includegraphics[width=0.85\textwidth]{pic_var.png}
+  \caption{Number of variables \label{fig:pic_var}}
 \end{figure}
 
-\begin{figure}\centering
-  \includegraphics[width=0.85\textwidth]{pic_var_medium.png}
-  \caption{Number of variables (b) \label{fig:pic_var_medium}}
-\end{figure}
+%\begin{figure}\centering
+%  \includegraphics[width=0.85\textwidth]{pic_var_small.png}
+%  \caption{Number of variables (a) \label{fig:pic_var_small}}
+%\end{figure}
+%
+%\begin{figure}\centering
+%  \includegraphics[width=0.85\textwidth]{pic_var_medium.png}
+%  \caption{Number of variables (b) \label{fig:pic_var_medium}}
+%\end{figure}
+%
+%\begin{figure}\centering
+%  \includegraphics[width=0.85\textwidth]{pic_var_big.png}
+%  \caption{Number of variables (c) \label{fig:pic_var_large}}
+%\end{figure}
 
-\begin{figure}\centering
-  \includegraphics[width=0.85\textwidth]{pic_var_big.png}
-  \caption{Number of variables (c) \label{fig:pic_var_large}}
-\end{figure}
-
-Similarly, Figures~\ref{fig:pic_constr_small}, Figure~\ref{fig:pic_constr_medium}
-and Figure~\ref{fig:pic_constr_big} describe how linear and quadratic
+Similarly, Figures~\ref{fig:pic_constr} 
+%Figure~\ref{fig:pic_constr_medium}
+%and Figure~\ref{fig:pic_constr_big} 
+describes how linear and quadratic
 constraints are distributed within the instances. The instances are sorted
-accordingly to the total number of constraints and divided in the
-same manner as for Figures~\ref{fig:pic_var_small}--\ref{fig:pic_var_large}.
+accordingly to the total number of constraints
+% and divided in the same manner as for Figures~\ref{fig:pic_var_small}--\ref{fig:pic_var_large}.
 For each instances we report the total number of constraints with a ``$+$''
@@
 prevalence of lower percentages of quadratic constraints. In particular,
-from Figure~\ref{fig:pic_constr_big} we learn that while the maximum number
+from Figure~\ref{fig:pic_constr} we learn that while the maximum number
 of linear constraints exceeds 100000, the maximum number of quadratic
@@
 \begin{figure}\centering
-  \includegraphics[width=0.85\textwidth]{pic_constr_small.png}
-  \caption{Number of constraints (a) \label{fig:pic_constr_small}}
+  \includegraphics[width=0.85\textwidth]{pic_constr.png}
+  \caption{Number of constraints \label{fig:pic_constr}}
 \end{figure}
 
-\begin{figure}\centering
-  \includegraphics[width=0.85\textwidth]{pic_constr_medium.png}
-  \caption{Number of constraints (b) \label{fig:pic_constr_medium}}
-\end{figure}
+%\begin{figure}\centering
+%  \includegraphics[width=0.85\textwidth]{pic_constr_small.png}
+%  \caption{Number of constraints (a) \label{fig:pic_constr_small}}
+%\end{figure}
+%
+%\begin{figure}\centering
+%  \includegraphics[width=0.85\textwidth]{pic_constr_medium.png}
+%  \caption{Number of constraints (b) \label{fig:pic_constr_medium}}
+%\end{figure}
+%
+%\begin{figure}\centering
+%  \includegraphics[width=0.85\textwidth]{pic_constr_big.png}
+%  \caption{Number of constraints (c) \label{fig:pic_constr_big}}
+%\end{figure}
+
+Figure~\ref{fig:pic_conv_constr} shows the instances with at least a quadratic constraints sorted according to the number of quadratic constraints (vertical axis).
+For each instances we report the total number of constraints with a ``$+$''
+and the total number of nonconvex quadratic constraints
+with a ``$\times$''.
 
 \begin{figure}\centering
-  \includegraphics[width=0.85\textwidth]{pic_constr_big.png}
-  \caption{Number of constraints (c) \label{fig:pic_constr_big}}
+  \includegraphics[width=0.85\textwidth]{pic_quad_conv_vs_nonconv.png}
+  \caption{Quadratic constraints \label{fig:pic_conv_constr}}
 \end{figure}
@@
 
-{\bf Nothing about convexity of constraints? A graph plotting the
-number of convex quadratic constraints vs the total number of quadratic
-constraints?}
-
+Additional details on the instance features can be found in the Appendix~\ref{sec:instance_details}.
 %- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

### 2016-12-27 format.tex
Update format.tex

@@
 
-The QPLIB format is designed to cope with a variety of
-linear and quadratic programs. We distinguish, in particular,
-linear programs (LP)
+Following the notation of Section~\ref{sec:QPbasic}, a generic Quadratic Program (QP) can be written as follows:
+%
 \begin{align*}	
- \mininx{} & l^T x + f \\
- \st & c\s{L} \leq A x \leq c\s{U} \tim{and} x\s{L} \leq x \leq x\s{U}, \\
-\intertext{linear programs with quadratic constraints (LPQC)}
- \mininx{} & l^T x + f \\
- \st & c\s{L} \leq A x + \half  [x,Q\s{C},x] \leq c\s{U}
-     \tim{and} x\s{L} \leq x \leq x\s{U}, \\
-\intertext{bound-constrained quadratic programs (BQP)}
- \mininx{} & \half x^T Q x + l^T x + f \\
- \st & x\s{L} \leq x \leq x\s{U}, \\
-\intertext{quadratic programs (QP)}
- \mininx{} & \half x^T Q x + l^T x + f \\
- \st & c\s{L} \leq A x \leq c\s{U} \tim{and} x\s{L} \leq x \leq x\s{U}, \\
-\intertext{and quadratic program with quadratic constraints (QPQC)}
- \mininx{} & \half x^T Q x + l^T x + f \\
- \st & c\s{L} \leq A x + \half [x,Q\s{C},x] \leq c\s{U}
-     \tim{and} x\s{L} \leq x \leq x\s{U},
+ \min \;\;
+ & x^\top Q^0 x + b^0 x + f \\
+ %
+ & c^i_l \leq x^\top Q^i x + b^i x \leq c^i_u & i \in \Mcm \\ 
+ %
+ & l_j \leq x_j \leq u_j & j \in \Mcn  \\
+ %
+ & x_j \in \mathbb{Z} & j \in \Mcz
 \end{align*}
-\todo[inline]{\small Stefan: Are LPQC and QPQC frequently used abbreviations?
+
+where:
+
+\begin{itemize}
+ \item $\Mcn = \{ 1, \ldots, n \}$ is the set of (indices) of variables, and $\Mcm = \{ 1, \ldots, m \}$ is the set of (indices) of constraints;
+ %
+ \item $x = [x_j]_{j = 1}^n \in \RR^n$ is a finite vector of real variables;
+ %
+ \item $Q^i$ for $i \in \{ 0 \} \cup \Mcm$ are symmetric $n \times n$ real matrices;
+ %
+ \item $b^i$, $c^i_l$, $c^i_u$ for $i \in \{ 0 \} \cup \Mcm$ and $f$ are, respectively, real $n$-vectors, real constants, real constants and real constant;
+ %
+ \item $-\infty \leq l_j \leq u_j \leq \infty$ are the (extended) real upper and lower bounds on each variable $x_j$ for $j \in \Mcn$;
+ %
+ \item $\Mcm = \Mcq \cup \Mcl$ where $Q^i = 0$ for all $i \in \Mcl$;
+ %
+ \item the variables in $\Mcz \subseteq \Mcm$ are restricted to only attain integer values.
+\end{itemize}
+
+%Accordingly, the QPLIB format is designed to cope with a variety of linear and quadratic programs. 
+%We distinguish, in particular,
+%linear programs (LP)
+%\begin{align*}
+% \mininx{} & l^T x + f \\
+% \st & c\s{L} \leq A x \leq c\s{U} \tim{and} x\s{L} \leq x \leq x\s{U}, \\
+%\intertext{linear programs with quadratic constraints (LPQC)}
+% \mininx{} & l^T x + f \\
+% \st & c\s{L} \leq A x + \half  [x,Q\s{C},x] \leq c\s{U}
+%     \tim{and} x\s{L} \leq x \leq x\s{U}, \\
+%\intertext{bound-constrained quadratic programs (BQP)}
+% \mininx{} & \half x^T Q x + l^T x + f \\
+% \st & x\s{L} \leq x \leq x\s{U}, \\
+%\intertext{quadratic programs (QP)}
+% \mininx{} & \half x^T Q x + l^T x + f \\
+% \st & c\s{L} \leq A x \leq c\s{U} \tim{and} x\s{L} \leq x \leq x\s{U}, \\
+%\intertext{and quadratic program with quadratic constraints (QPQC)}
+% \mininx{} & \half x^T Q x + l^T x + f \\
+% \st & c\s{L} \leq A x + \half [x,Q\s{C},x] \leq c\s{U}
+%     \tim{and} x\s{L} \leq x \leq x\s{U},
+%\end{align*}
+%\todo[inline]{\small Stefan: Are LPQC and QPQC frequently used abbreviations?
 If a linear program has quadratic constraints, how can it then still be a linear program?
@@
 I further think that one could skip QCP and consider them as part of QCQP.}
-where $Q\s{C}$ is the diagonal matrix with blocks $Q\s{C}_k$, $k=1,\ldots,m$,
-$Q$ and the $Q\s{C}_k$ are symmetric $n$ by $n$ Hessian matrices,
-$A$ is a rectangular $m$ by $n$ constraint matrix,
-$l$, $x\s{L}$, $x\s{U}$, $c\s{L}$, $c\s{U}$
-are vectors of appropriate sizes and $f$ is a scalar; the notation
-$[ x, Q\s{C},x]$ denotes the vector whose $k$-th component is
-$x^T Q\s{C}_k x$.
-In addition, all or a subset of the variables $x$ may be required to be
-integer, while further restrictions (such as a variable being binary)
-may be imposed via explicit constraints.
+%where $Q\s{C}$ is the diagonal matrix with blocks $Q\s{C}_k$, $k=1,\ldots,m$,
+%$Q$ and the $Q\s{C}_k$ are symmetric $n$ by $n$ Hessian matrices,
+%$A$ is a rectangular $m$ by $n$ constraint matrix,
+%$l$, $x\s{L}$, $x\s{U}$, $c\s{L}$, $c\s{U}$
+%are vectors of appropriate sizes and $f$ is a scalar; the notation
+%$[ x, Q\s{C},x]$ denotes the vector whose $k$-th component is
+%$x^T Q\s{C}_k x$.
+%In addition, all or a subset of the variables $x$ may be required to be
+%integer, while further restrictions (such as a variable being binary)
+%may be imposed via explicit constraints.
 
@@
 $n$  & number of variables (integer) & \\
-$m$  & number of general linear constraints (integer) & [2] \\
+$m$  & number of constraints (integer) & [2] \\
 \hline
-$n\s{Q}$ & number of nonzeros  (integer) in lower triangle of $Q$  & [3] \\
-$i$\; $j$\; $Q_{i,j}$ & \lbox{row and column indices (integers) and value (real)
-for each nonzero entry of $Q$, if $n\s{Q} > 0$, one triple on each line} & \\
+$n^{Q^0}$ & number of nonzeros  (integer) in lower triangle of $Q^0$  & [3] \\
+$h$\; $k$\; $Q_{hk}^0$ & \lbox{row and column indices (integers) and value (real)
+for each nonzero entry of $Q^0$, if $n^{Q^0} > 0$, one triple on each line} & \\
 \hline
-$l_d$ & default value (real) for entries in $l$ & \\
-$n\s{l}$ & number of non-default entries (integer) in $l$ & \\
-$i$\; $l_i$ & \lbox{index (integer) and value (real) for each non-default
-term in $l$, if $n\s{l} > 0$, one pair per line} & \\
+$b^0_d$ & default value (real) for entries in $b^0$ & \\
+$n^{b^0}$ & number of non-default entries (integer) in $b^0$ & \\
+$j$\; $b^0_j$ & \lbox{index (integer) and value (real) for each non-default
+term in $b^0$, if $n^{b^0} > 0$, one pair per line} & \\
 \hline
-$f$ & value (real) of f & \\ \hline
-$n\s{Q${}_c$}$ & number of nonzeros (integer) in lower triangle of
-$Q\s{C}$  & [2,4] \\
-$k$\; $i$\; $j$\; $Q\s{C}_{k,i,j}$
+$f$ & constant part of the objective function & \\ \hline
+$n^{Q^i}$ & number of nonzeros (integer) in lower triangle of
+$Q^i$  & [2,4] \\
+$i$\; $h$\; $k$\; $Q^{i}_{hk}$
 & \lbox{constraint, row and column indices (integers) and value (real) for
- each entry of $Q\s{C}$, if $n\s{Q${}_c$} > 0$, one quadruple on each line} & \\
+ each entry of $Q^{i}$, if $n^{Q^i}$, one quadruple on each line} & \\
 \hline
-$n\s{A}$ & number of nonzeros  (integer) in $A$                     & [2] \\
-$i$\; $j$\; $A_{i,j}$ & \lbox{row and column indices (integers) and value (real)
+$n^{b}$ & number of nonzeros  (integer) in $b^i$, $i \in \Mcm$                     & [2] \\
+$i$\; $j$\; $b^i_{j}$ & \lbox{row and column indices (integers) and value (real)
 for each nonzero
-entry of $A$, if $n\s{A} > 0$, one triple on each line} & \\
+entry of $b^i$, if $n^{b} > 0$, one triple on each line} & \\
 \hline
@@
 \hline
-$c\s{L}_d$ & default value (real) for entries in $c\s{L}$            & [2] \\
-$n^{c\s{L}}$ & number of non-default entries (integer) in $c\s{L}$    & [2] \\
-$i$\;\ $c\s{L}_i$ & \lbox{index (integer) and value (real) for each non-default
+$c_{l,d}$ & default value (real) for entries in $c_{l}$            & [2] \\
+$n^{c_{l,d}}$ & number of non-default entries (integer) in $c_{l}$    & [2] \\
+$i$\;\ $c^i_{l}$ & \lbox{index (integer) and value (real) for each non-default
 term
-in $c\s{L}$, if $n^{c\s{L}} > 0$, one pair per line}        & [2] \\
+in $c_{l,d}$, if $n^{c_{l,d}} > 0$, one pair per line}        & [2] \\
 \hline
-$c\s{U}_d$ & default value (real) for entries in $c\s{U}$            & [2] \\
-$n^{c\s{U}}$ & number of non-default entries (integer) in $c\s{U}$    & [2] \\
-$i$\;\ $c\s{U}_i$ & \lbox{index (integer) and value (real) for each non-default
+$c_{u,d}$ & default value (real) for entries in $c_{u}$            & [2] \\
+$n^{c_{u,d}}$ & number of non-default entries (integer) in $c_{u}$    & [2] \\
+$i$\;\ $c^i_{u}$ & \lbox{index (integer) and value (real) for each non-default
 term
-in $c\s{U}$, if $n^{c\s{U}} > 0$, one pair per line}        & [2] \\
+in $c_{u,d}$, if $n^{c_{u,d}} > 0$, one pair per line}        & \\
 \hline
-$x\s{L}_d$ & default value (real) for entries in $x\s{L}$            &     \\
-$n^{x\s{L}}$ & number of non-default entries (integer) in $x\s{L}$    &     \\
-$i$\;\ $x\s{L}_i$ & \lbox{index (integer) and value (real) for each non-default
-term in $x\s{L}$, if $n^{x\s{L}} > 0$, one pair per line} & \\
+$l_{d}$ & default value (real) for entries in $l$            & \\
+$n^{l_{d}}$ & number of non-default entries (integer) in $l$    &  \\
+$i$\;\ $l_{i}$ & \lbox{index (integer) and value (real) for each non-default
+term
+in $l$, if $n^{l_{d}}> 0$, one pair per line}        &  \\
 \hline
-$x\s{U}_d$ & default value (real) for entries in $x\s{U}$            &     \\
-$n^{x\s{U}}$ & number of non-default entries (integer) in $x\s{U}$    &     \\
-$i$\;\ $x\s{U}_i$ & \lbox{index (integer) and value (real) for each non-default
-term in $x\s{U}$, if $n^{x\s{U}} > 0$, one pair per line} & \\
+$u_{d}$ & default value (real) for entries in $u$            & \\
+$n^{u_{d}}$ & number of non-default entries (integer) in $u$    &  \\
+$i$\;\ $u_{i}$ & \lbox{index (integer) and value (real) for each non-default
+term
+in $u$, if $n^{u_{d}}> 0$, one pair per line}        &  \\
 \hline
@@
 value $i$} &  \\
-$i$\; $x_i$ & \lbox{index (integer) and name (character string)
+$j$\; ${\tt var\_name}_j$ & \lbox{index (integer) and name (character string)
 for each non-default variable name, if $n^x_d > 0$,  one pair per line} & \\
@@
 the numerical value $i$} &  \\
-$i$\; $c_i$ & \lbox{index  (integer) and name (character string) for each
+$i$\; ${\tt cons\_name}_i$ & \lbox{index  (integer) and name (character string) for each
 non-default constraint name, if $n^c_d > 0$,  one pair per line} & \\
@@
 \item [{[1]}]
+\todo[inline]{\small Fabio and Emiliano: we think that, if possible, we should follow the classification proposed in Section 2.}
+
 The problem type is represented by a character string as one of the following:
@@
 As a simple example, consider the mixed-integer QP
-\[\arr{l}{\mininx{} x_1^2 + x_2^2 + x_3^2 - x_1 x_2 - x_2 x_3
+\[\arr{l}{\min_{x \in \smallRe^3} x_1^2 + x_2^2 + x_3^2 - x_1 x_2 - x_2 x_3
   - 0.2 x_1  -0.4 x_2 - 0.2 x_3 \\
@@
 2         # general linear constraints
-5         # nonzeros in lower triangle of Q
-1 1 2.0   5 lines of row & column index & value of nonzeros in lower triangle Q
+5         # nonzeros in lower triangle of Q^0
+1 1 2.0   5 lines of row & column index & value of nonzeros in lower triangle Q^0
 2 1 -1.0  |
@@
 3 3 2.0   |
--0.2      default value for entries in l
-1         # non default entries in l
-2 -0.4    1 lines of index & value of non-default values in l
+-0.2      default value for entries in b_0
+1         # non default entries in b_0
+2 -0.4    1 lines of index & value of non-default values in b_0
 0.0       value of f
-4         # nonzeros in A
-1 1 1.0   4 lines of row & column index & value of nonzeros in A
+4         # nonzeros in vectors b^i (i =0,...,m)
+1 1 1.0   4 lines of row & column index & value of nonzeros in b^i (i =0,...,m)
 1 2 1.0   |
@@
 1.0E+20   infinity
-1.0       default value for entries in c^L
-0         # non default entries in c^L
-1.0E+20   default value for entries in c^U
-0         # non default entries in c^U
-0.0       default value for entries in x^L
-0         # non default entries in x^L
-1.0       default value for entries in x^U
-1         # non default entries in x^U
-2 2.0     1 line of non-default indices and values in x^U
+1.0       default value for entries in c_l
+0         # non default entries in c_l
+1.0E+20   default value for entries in c_u
+0         # non default entries in c_u
+0.0       default value for entries in l
+0         # non default entries in l
+1.0       default value for entries in u
+1         # non default entries in u
+2 2.0     1 line of non-default indices and values in u
 0         default variable type is continuous

### 2016-12-27 QPLIB-1.0.tex
Update QPLIB-1.0.tex

@@
 %
-\multirow{1}{.55\textwidth}{{\bf Finance:} Portfolio optimization} & checkmark & \multirow{1}{.25\textwidth}{\cite{deng-etal:2013,kallrath:2003,lin-etal:2005,maranas-etal:1997,parpas-rustem:2006,rios-sahinidis:2010,FrFG16,FrGe06a,FrGe07a,FrGe09a}} \\
+\multirow{1}{.55\textwidth}{{\bf Finance:} Portfolio optimization} & \checkmark & \multirow{1}{.25\textwidth}{\cite{deng-etal:2013,kallrath:2003,lin-etal:2005,maranas-etal:1997,parpas-rustem:2006,rios-sahinidis:2010,FrFG16,FrGe06a,FrGe07a,FrGe09a}} \\
 %

### 2017-01-03 format.tex
update qplib format to include optimization sense (minimize or maximize)

@@
 type & problem type (character string) & [1] \\
+sense & one of the words minimize or maximize (character string, case irrelevant) & \\
 \hline
@@
 MIQP      # problem is a mixed-integer quadratic program
+Minimize  # minimize the objective function
 3         # variables

### 2017-01-20 QPLIB.tex
added a few citations, corrected a few typoes
update affiliation

@@
 \and       Stefan Vigerske \at
-           GAMS Software GmbH, P.O. Box 40 59,50216 Frechen, Germany.
-           \email{stefan@gams.com}
+           GAMS Software GmbH, c/o Zuse Institute Berlin, Takustr. 7, 14195 Berlin, Germany.
+           \email{svigerske@gams.com}
 %
 \and       Angelika Wiegele \at
-Institut für Mathematik, Alpen-Adria-Universität Klagenfurt, Austria.  \email{angelika.wiegele@aau.at}
+Institut f{\"u}r Mathematik, Alpen-Adria-Universit{\"a}t Klagenfurt, 9020 Klagenfurt am W{\"o}rthersee, Austria.  \email{angelika.wiegele@aau.at}
            }

### 2017-01-20 QPLIB-1.0.tex
update/add todo

@@
 Hence, in particular for the convex case, tools and techniques that have been honed during decades of research for Linear Programming (LP), typically with integrality constraints (MILP), can often be extended to the quadratic case with at least less effort than what would be required for tackling general Non Linear Programming (NLP) problems, without or with integrality constraints (MINLP).
-This has indeed happened over and over again with different algorithmic techniques, such as Interior Point methods, active-set methods (of which the simplex method is a prototypical example), enumeration methods, cut-generation techniques, reformulation techniques, and many others ({\bf citations?}).
+This has indeed happened over and over again with different algorithmic techniques, such as Interior Point methods, active-set methods (of which the simplex method is a prototypical example), enumeration methods, cut-generation techniques, reformulation techniques, and many others \todo[inline]{citations?
+http://www.optimization-online.org/DB\_HTML/2009/07/2338.html ?}.
 Similarly, nonconvex continuous QP are perhaps the ``simplest'' class of problems that require features like spatial enumeration techniques to be solved.
-Hence, they are both the natural basis for the development of general techniques for nonconvex NLP, and a very specific class so that specialized approaches can be developed \cite{Dur2010,Burer2012}.
+\todo[inline]{If only the objective is quadratic, i.e., what one usually would denote by a ``nonconvex QP'', then one can ``just'' enumerate local optima.
+Nitpicking?} Hence, they are both the natural basis for the development of general techniques for nonconvex NLP, and a very specific class so that specialized approaches can be developed \cite{Dur2010,Burer2012}.
 
@@
 Quadratic expressions are found, either naturally or after appropriate reformulations, in very many optimization problems \cite{Kochenberger2014}.
-Table \ref{tbl:application} provides a certainly non-exhaustive collection of applications that lead to formulations with either quadratic constraints, ot quadratic objective function, or both.
+Table \ref{tbl:application} provides a certainly non-exhaustive collection of applications that lead to formulations with either quadratic constraints, or quadratic objective function, or both.
 Also, in general any continuous function can be approximated with arbitrary accuracy (over a compact set) by a polynomial of arbitrary degree; in turn, every polynomial can be broken down to a system of quadratic expressions.

### 2017-01-20 QPLIB.tex
update/add todo

@@
 \begin{abstract}
- This paper describes a new instance library for Quadratic Programming (QP).
+ This paper describes a new instance library for Quadratic Programming (QP).\todo[inline]{As we don't use the usual definition of QP, we should already say in the abstract what we understand under a QP.}
  QP is a very ``varied'' class, comprising sub-classes of problems ranging from

### 2017-01-20 QPLIB-3.tex
say gams performance libraries, fix capitalization of GAMS
- GAMS has a number of libraries, so "GAMS library" is very unspecific

@@
 %
-\item the {\tt GAMS} library
+\item the {\tt GAMS Performance} libraries
  \url{http://www.gamsworld.org/performance/performlib.htm};
@@
 To evaluate the instances we decided, for practical reasons, to use
-Gams as common platform for all the experiments involving commercial and
+GAMS as common platform for all the experiments involving commercial and
 non-commercial solvers. For this reason, we decided to translate all
-instances into the Gams format (\texttt{.gms}). 
+instances into the GAMS format (\texttt{.gms}). 
 % In a preliminary phase, all the instances received were divided
@@
        global optimality. Accordingly, for each of the gathered instance we
-       ran the complete solvers in {\tt Gams} (cf.~Table \ref{t:solvers})
+       ran the complete solvers in {\tt GAMS} (cf.~Table \ref{t:solvers})
        capable of solving it, whose number depends on the category of the

### 2017-01-20 QPLIB-2.3.tex
start revising solution methods

@@
 Asymptotically complete methods can find a globally optimal solution with probability one in infinite time, but again they cannot prove that a given instance is infeasible.
-Complete methods find an approximate globally optimal solution to within a prescribed optimality tolerance within finite time, or prove that none such exists (but see Sect.~\ref{s:complete} below); they are often referred to as \emph{exact} methods in the computational optimization community.
-According to \cite{neumaier}, {\it rigorous} methods also exist which find global within given tolerances even in the presence of rounding errors, except in ``near-degenerate cases''.
-Since it is debatable whether any of the solver we are using can be classified as rigorous, we have elected to limit ourselves to declaring solvers complete.
+Complete methods find an approximate globally optimal solution within a prescribed optimality tolerance within finite time, or prove that none such exists (but see Sect.~\ref{s:complete} below); they are often referred to as \emph{exact} methods in the computational optimization community.
+According to \cite{neumaier}, {\it rigorous} methods also exist, which find globally optimal solutions within given tolerances even in the presence of rounding errors, except for ``near-degenerate cases''.
+Since none of the solvers we are using can be classified as rigorous, we limit ourselves to declaring solvers complete.
 
@@
 Therefore, when we state that a given algorithm is incomplete or (asymptotically) complete we mean that it is so the largest problem class that the solver naturally targets, although it may be complete on specific sub-classes.
-For example, {\sc SnOpt} naturally targets continuous nonconvex NLPs and it is incomplete on this class, and therefore on QCQ, but becomes complete for CCC.
+For example, {\sc SnOpt} naturally targets continuous NLPs and it is incomplete on nonconvex NLPs, and therefore on QCQ, but becomes complete for CCC.
 In general, all complete methods for a problem $P$ must be complete for any problem $Q \subseteq P$, while a complete method for $P$ might be incomplete for $R \supset P$.
@@
 In general, local search methods are incomplete.
-As remarked above, however, local search methods deployed on a convex problem behave like complete methods, possibly via devices that that allow one to find a feasible starting solution if there is one (like what was used to be called ``phase -1'' in the simplex method).
+As remarked above, however, local search methods deployed on a convex problem behave like complete methods, possibly via devices that allow one to find a feasible starting solution if there is one (like what was used to be called ``phase -1'' in the simplex method).
 
-Most local search methods for *C* are iterative in nature, and need a feasible starting point $x^0$ as input.
+Most local search methods for *C* are iterative in nature, and need a starting point $x^0$ as input.
 The ($k+1$)-st iterate is obtained as $x^{k+1} = x^k + \alpha_k d^k$, where $\alpha_k$ (a scalar) is the {\it step length}, and $d^k$ (a vector) is the {\it search direction}, e.g.~belonging to a tangent manifold and having a negative directional derivative at $x^k$, in order to improving optimality while reducing infeasibility.
@@
 
-The solvers in Table \ref{t:solvers} which implement incomplete methods are {\sc ConOpt}, {\sc IpOpt}, {\sc Minos}, {\sc SnOpt} and {\sc KNitro}.
-The former four solvers implement local search algorithms for continuous nonconvex NLPs (a problem class containing QCQ).
-Note that traditionally {\sc KNitro} used to be a local (nonconvex) NLP solver.
-All of these solvers implement essentially three types of local search methods for NLP:
-%
+The solvers in Table \ref{t:solvers} which implement incomplete methods for continuous NLPs (a problem class containing QCQ) are {\sc ConOpt}, {\sc IpOpt}, {\sc Minos}, {\sc SnOpt} and {\sc KNitro}\todo{check capitalization of solver names}.
+The methods are local search algorithms and are essentially of the following types: %Note that traditionally {\sc KNitro} used to be a local (nonconvex) NLP solver.
+%All of these solvers implement essentially three types of local search methods for continuous NLPs:
+\todo{sequential linear programming, sequential quadratic programming, trust-region}
 \begin{itemize}
@@
  \item projected gradient methods \cite{CalaMore87:mp,ChenGui13:coap}.
-\end{itemize}
+\end{itemize} \todo{doesn't Knitro additionally implement an active set or SQP algorithm?}
 %

### 2017-01-20 QPLIB-2.3.tex
knitro has an active set solver and multistart for nlps

@@
 \begin{itemize}
- \item active set methods ({\sc ConOpt}, {\sc Minos}, {\sc SnOpt}) \cite{Dost09};
+ \item active set methods ({\sc ConOpt}, {\sc Minos}, {\sc SnOpt}, {\sc KNitro}) \cite{Dost09};
  %
@@
  \item projected gradient methods \cite{CalaMore87:mp,ChenGui13:coap}.
-\end{itemize} \todo{doesn't Knitro additionally implement an active set or SQP algorithm?}
+\end{itemize}
 %
@@
 
-The solvers in Table \ref{t:solvers} which implement asymptotically complete methods are {\sc oqnlp}, {\sc msnlp} and certain sub-solvers of {\sc lgo}. Specifically, we consider the following methods:
+The solvers in Table \ref{t:solvers} which implement asymptotically complete methods are {\sc oqnlp}, {\sc msnlp}, {\sc knitro}, and certain sub-solvers of {\sc lgo}.
+Specifically, we consider the following methods:
 %
@@
  %
- \item multi-start ({\sc lgo\_ms}, {\sc msnlp}, {\sc oqnlp}); specifically, the former two apply to QCQ whereas the latter to QGQ.
+ \item multi-start ({\sc knitro}, {\sc lgo\_ms}, {\sc msnlp}, {\sc oqnlp}); specifically, the former three apply to QCQ whereas the latter to QGQ.
 \end{itemize}

### 2017-01-20 QPLIB-2.3.tex
lindo = lindoglobal = lindo api, lgo_bb is not a complete solver
- deja vu

@@
 \begin{itemize}
- \item {\sc Antigone}, {\sc Baron}, {\sc Couenne}, {\sc Lindo}, {\sc LindoGlobal}, {\sc SCIP}, {\sc lgo\_bb}, which implement complete BB algorithms for QGQ;
+ \item {\sc Antigone}, {\sc Baron}, {\sc Couenne}, {\sc Lindo API}, {\sc SCIP}, which implement complete BB algorithms for QGQ;
  %
@@
 We also remark that CPLEX can currently only target problems with linear constraints when the objective function is nonconvex \cite{bliek}.
+
+\todo[inline]{{\sc lgo\_bb} was in the listing above, but is it a complete solver for any class?}
 

### 2017-01-23 QPLIB-2.3.tex
clear up listing of complete b&b solvers

--- 2017-01-20 QPLIB-2.3.tex
@@
 
-The solvers in Table \ref{t:solvers} BB type methods are:
+The following solvers from Table \ref{t:solvers} implement complete BB algorithms for QGQ or some subclasses:
 %
 \begin{itemize}
- \item {\sc Antigone}, {\sc Baron}, {\sc Couenne}, {\sc Lindo}, {\sc LindoGlobal}, {\sc SCIP}, {\sc lgo\_bb}, which implement complete BB algorithms for QGQ;
+ \item {\sc Antigone}, {\sc Baron}, {\sc Couenne}, {\sc Lindo API}, {\sc SCIP} for QGQ;
  %
-\item {\sc CPLEX}, which implements a complete BB algorithm for QGL;
+\item {\sc CPLEX} for QGL and CGC;
 %
- \item {\sc Knitro\_bb}, {\sc BonMin}, {\sc sBB}, {\sc Xpress}, {\sc Gurobi}, and {\sc Mosek} which implement complete BB algorithms for CGC.
+ \item {\sc Knitro}, {\sc BonMin}, {\sc sBB}, {\sc Xpress}, {\sc Gurobi}, and {\sc Mosek} for CGC.
 \end{itemize}
@@
 We remark that the latter category can be used as incomplete solvers for QGQ.
-We also remark that CPLEX can currently only target problems with linear constraints when the objective function is nonconvex \cite{bliek}.
+%We also remark that CPLEX can currently only target problems with linear constraints when the objective function is nonconvex \cite{bliek}.
+We also remark that {\sc lgo} implements an incomplete BB algorithm for QCQ by using bounds obtained from sampling.
 

### 2017-01-23 QPLIB-2.3.tex
reorganize cutting plane methods
- first the method, then the solvers
- make it clearer that the milp is a relaxation (for CGC)

@@
 
-\paragraph{Cutting plane approaches}.
+\paragraph{Cutting plane approaches.}
 %
-The two remaining solvers in Table \ref{t:solvers} are {\sc AlphaECP} \cite{westerlund} and {\sc DICOpt} \cite{quesada}, which are complete cutting plane methods based on different principles.
-Characteristic of both solvers is that they need to employ a complete method for solving MILP (LIL) sub-problems at each iteration; in turn, this is typically based on BB, which is therefore a crucial technique also for this class of approaches.
+Cutting plane approaches construct and iteratively improve a MILP (LIL) relaxation of the problem \cite{quesada,westerlund}.
+The cutting planes for the MILP are generated by linearization (first-order Taylor approximation) of the nonlinearities.
+If the latter are convex, the MILP provides a valid lower bound for the problem.
 Additionally, incomplete methods can be used to provide local solutions.
-Other solvers, like {\sc BonMin}, also offer this kind of approach among their algorithmic options.
+Therefor, these methods are complete on CGC if a complete method is used to solve the MILP.
+The latter is typically based on BB, which is therefore a crucial technique also for this class of approaches.
+
+Solvers in Table \ref{t:solvers} that implement complete cutting plane methods for CGC are {\sc AlphaECP} \cite{westerlund}, {\sc BonMin} (in the algorithmic mode B-OA) \cite{BBCCGLLLMSW} and {\sc DICOpt} \cite{quesada}.
+\todo{here we give references to the solvers, above we don't}
 

### 2017-01-23 QPLIB-2.4.tex
comments and minor corrections from first proofreading
knitro has an active set solver and multistart for nlps
resolve ? in table, add todo regarding xpress-slp

@@
 \subsection{Solvers}\label{subsec:solver}
+\todo{\small Ambros: merge with Section~\ref{s:complete} or put list of solvers in Section~\ref{s:complete} here}
 
@@
 In Table \ref{t:solvers}, we mark with ``I'' a pair (solver, problem) if the solver accepts the problem in input but it is an incomplete solver for the problem, with ``A'' if it is asymptotically complete, with ``C'' if it is complete, and leave it blank if the solver won't accept the problem in input.
-
-{\bf The table has to be checked, as I've extrapolated from the text but I'm not 100\% sure. Also, ``?''s have to be removed.}
 
@@
 {\sc IpOpt}       &     &     &     &     &  C  &  I  \\
-{\sc lgo}         &     &     &     &     &  C? &  A  \\  %try it out
+{\sc lgo}         &     &     &     &     &  A  &  A  \\
 {\sc Minos}       &     &     &     &     &  C  &  I  \\
@@
 {\sc SnOpt}       &     &     &     &     &  C  &  I  \\
-{\sc XPress}      &  C  &     &  C  &     &  C  &  ?  \\ %try it out
-{\sc Gurobi}      &  C  &     &  C  &     &  C  &  ?  \\ %try it out
-{\sc CPLEX}       &  C  &  C  &  C  &     &  C  &  I? \\ %try it out
+{\sc XPress}      &  C  &     &  C  &     &  C  &     \\
+{\sc Gurobi}      &  C  &     &  C  &     &  C  &     \\
+{\sc CPLEX}       &  C  &  C  &  C  &     &  C  &     \\
 {\sc mosek}       &  C  &     &  C  &     &  C  &     \\
@@
 \end{tabular}                 
-\caption{Families of QP problems that can be tackled each solver} \label{t:solvers}             
+\caption{Families of QP problems that can be tackled by each solver} \label{t:solvers}
 }
+
+\todo[inline]{Xpress has a solver for convex MIQCPs called Xpress-SLP.
+However, it is not available within GAMS.
+Should Table 2 take Xpress-SLP into account?}
 

### 2017-01-23 solvers.tex
rename QPLIB-2.3 and QPLIB-2.4
- QPLIB-2.4 now contains Section 2.3
- QPLIB-2.3 now contains Section 2.4
sort solvers alphabetically
- xpress is officially called fico xpress

--- 2017-01-23 QPLIB-2.4.tex
@@
 \subsection{Solvers}\label{subsec:solver}
-\todo{\small Ambros: merge with Section~\ref{s:complete} or put list of solvers in Section~\ref{s:complete} here}
 
-We now provide a succinct list of the solvers we have tested, using the approaches described in \S \ref{sec:algo}.
+We now provide a succinct list of the solvers we have worked with.
 In Table \ref{t:solvers}, we mark with ``I'' a pair (solver, problem) if the solver accepts the problem in input but it is an incomplete solver for the problem, with ``A'' if it is asymptotically complete, with ``C'' if it is complete, and leave it blank if the solver won't accept the problem in input.
@@
 \hline
+{\sc AlphaECP}    &  C  &  I  &  C  &  I  &  C  &  I  \\
 {\sc Antigone}    &  C  &  C  &  C  &  C  &  C  &  C  \\
 {\sc Baron}       &  C  &  C  &  C  &  C  &  C  &  C  \\
+{\sc BonMin}      &  C  &  I  &  C  &  I  &  C  &  I  \\
+{\sc ConOpt}      &     &     &     &     &  C  &  I  \\
 {\sc Couenne}     &  C  &  C  &  C  &  C  &  C  &  C  \\
+{\sc CPLEX}       &  C  &  C  &  C  &     &  C  &     \\
+{\sc DICOpt}      &  C  &  I  &  C  &  I  &  C  &  I  \\
+{\sc FICO XPress} &  C  &     &  C  &     &  C  &     \\
+{\sc Gurobi}      &  C  &     &  C  &     &  C  &     \\
+{\sc IpOpt}       &     &     &     &     &  C  &  I  \\
 {\sc KNitro}      &  C  &  I  &  C  &  I  &  C  &  A  \\
 {\sc lindo api}   &  C  &  C  &  C  &  C  &  C  &  C  \\
-{\sc SCIP}        &  C  &  C  &  C  &  C  &  C  &  C  \\
-{\sc oqnlp}       &  A  &  A  &  A  &  A  &  C  &  A  \\
-{\sc AlphaECP}    &  C  &  I  &  C  &  I  &  C  &  I  \\
-{\sc BonMin}      &  C  &  I  &  C  &  I  &  C  &  I  \\
-{\sc DICOpt}      &  C  &  I  &  C  &  I  &  C  &  I  \\
-{\sc sBB}         &  C  &  I  &  C  &  I  &  C  &  I  \\
-{\sc ConOpt}      &     &     &     &     &  C  &  I  \\
-{\sc IpOpt}       &     &     &     &     &  C  &  I  \\
 {\sc lgo}         &     &     &     &     &  A  &  A  \\
 {\sc Minos}       &     &     &     &     &  C  &  I  \\
+{\sc mosek}       &  C  &     &  C  &     &  C  &     \\
 {\sc msnlp}       &     &     &     &     &  C  &  A  \\
+{\sc oqnlp}       &  A  &  A  &  A  &  A  &  C  &  A  \\
+{\sc sBB}         &  C  &  I  &  C  &  I  &  C  &  I  \\
+{\sc SCIP}        &  C  &  C  &  C  &  C  &  C  &  C  \\
 {\sc SnOpt}       &     &     &     &     &  C  &  I  \\
-{\sc XPress}      &  C  &     &  C  &     &  C  &     \\
-{\sc Gurobi}      &  C  &     &  C  &     &  C  &     \\
-{\sc CPLEX}       &  C  &  C  &  C  &     &  C  &     \\
-{\sc mosek}       &  C  &     &  C  &     &  C  &     \\
 \hline

### 2017-01-23 methods.tex
add latex defines for solver names
- to make it easier to change their writing throughout the paper

--- 2017-01-23 QPLIB-2.3.tex
@@
 Therefore, when we state that a given algorithm is incomplete or (asymptotically) complete we mean that it is so the largest problem class that the solver naturally targets, although it may be complete on specific sub-classes.
-For example, {\sc SnOpt} naturally targets continuous NLPs and it is incomplete on nonconvex NLPs, and therefore on QCQ, but becomes complete for CCC.
+For example, \snopt naturally targets continuous NLPs and it is incomplete on nonconvex NLPs, and therefore on QCQ, but becomes complete for CCC.
 In general, all complete methods for a problem $P$ must be complete for any problem $Q \subseteq P$, while a complete method for $P$ might be incomplete for $R \supset P$.
@@
 
-The solvers in Table \ref{t:solvers} which implement incomplete methods for continuous NLPs (a problem class containing QCQ) are {\sc ConOpt}, {\sc IpOpt}, {\sc Minos}, {\sc SnOpt} and {\sc KNitro}\todo{check capitalization of solver names}.
-The methods are local search algorithms and are essentially of the following types: %Note that traditionally {\sc KNitro} used to be a local (nonconvex) NLP solver.
+The solvers in Table \ref{t:solvers} which implement incomplete methods for continuous NLPs (a problem class containing QCQ) are \conopt, \ipopt, \minos, \snopt, and \knitro.
+The methods are local search algorithms and are essentially of the following types: %Note that traditionally \knitro used to be a local (nonconvex) NLP solver.
 %All of these solvers implement essentially three types of local search methods for continuous NLPs:
@@
 \begin{itemize}
- \item active set methods ({\sc ConOpt}, {\sc Minos}, {\sc SnOpt}, {\sc KNitro}) \cite{Dost09};
+ \item active set methods (\conopt, \minos, \snopt, \knitro) \cite{Dost09};
  %
- \item interior point methods ({\sc IpOpt}, {\sc KNitro}) \cite{Wright97};
+ \item interior point methods (\ipopt, \knitro) \cite{Wright97};
  %
@@
 
-The solvers in Table \ref{t:solvers} which implement asymptotically complete methods are {\sc oqnlp}, {\sc msnlp}, {\sc knitro}, and certain sub-solvers of {\sc lgo}.
-Specifically, we consider the following methods:
+The solvers in Table \ref{t:solvers} which implement asymptotically complete methods are \oqnlp, \msnlp, \knitro, and certain sub-solvers of \lgo. Specifically, we consider the following methods:
 %
 \begin{itemize}
- \item global adaptive random search ({\sc lgo\_gars});
+ \item global adaptive random search (\lgo GARS);
  %
- \item multi-start ({\sc knitro}, {\sc lgo\_ms}, {\sc msnlp}, {\sc oqnlp}); specifically, the former three apply to QCQ whereas the latter to QGQ.
+ \item multi-start (\knitro, \lgo MS, \msnlp, \oqnlp); specifically, the former three apply to QCQ whereas the latter to QGQ.
 \end{itemize}
 
-\paragraph{Global Adaptive Random Search.}
+\paragraph{Global Adaptive Random Search (GARS).}
 %
@@
 Since this method only depends on sampling and function evaluation, it is usually fast.
-In the {\sc lgo\_gars} solver, it provides a useful starting point for a subsequent local search procedure.
+In the GARS solver of \lgo, it provides a useful starting point for a subsequent local search procedure.
 Asymptotic global convergence is attained by restarting the random search from different initial points $x'$.
@@
 \begin{itemize}
- \item {\sc Antigone}, {\sc Baron}, {\sc Couenne}, {\sc Lindo API}, {\sc SCIP} for QGQ;
+ \item \antigone, \baron, \couenne, \lindo, \scip for QGQ;
  %
-\item {\sc CPLEX} for QGL and CGC;
+\item \cplex for QGL and CGC;
 %
- \item {\sc Knitro}, {\sc BonMin}, {\sc sBB}, {\sc Xpress}, {\sc Gurobi}, and {\sc Mosek} for CGC.
+ \item \knitro, \bonmin, \sbb, \xpress, \gurobi, and \mosek for CGC.
 \end{itemize}
@@
 %We also remark that CPLEX can currently only target problems with linear constraints when the objective function is nonconvex \cite{bliek}.
-We also remark that {\sc lgo} implements an incomplete BB algorithm for QCQ by using bounds obtained from sampling.
+We also remark that \lgo implements an incomplete BB algorithm for QCQ by using bounds obtained from sampling.
 
@@
 
-Solvers in Table \ref{t:solvers} that implement complete cutting plane methods for CGC are {\sc AlphaECP} \cite{westerlund}, {\sc BonMin} (in the algorithmic mode B-OA) \cite{BBCCGLLLMSW} and {\sc DICOpt} \cite{quesada}.
+Solvers in Table \ref{t:solvers} that implement complete cutting plane methods for CGC are \alphaecp \cite{westerlund}, \bonmin (in the algorithmic mode B-OA) \cite{BBCCGLLLMSW} and \dicopt \cite{quesada}.
 \todo{here we give references to the solvers, above we don't}

### 2017-01-23 solvers.tex
add latex defines for solver names
- to make it easier to change their writing throughout the paper

@@
 \hline
-{\sc AlphaECP}    &  C  &  I  &  C  &  I  &  C  &  I  \\
-{\sc Antigone}    &  C  &  C  &  C  &  C  &  C  &  C  \\
-{\sc Baron}       &  C  &  C  &  C  &  C  &  C  &  C  \\
-{\sc BonMin}      &  C  &  I  &  C  &  I  &  C  &  I  \\
-{\sc ConOpt}      &     &     &     &     &  C  &  I  \\
-{\sc Couenne}     &  C  &  C  &  C  &  C  &  C  &  C  \\
-{\sc CPLEX}       &  C  &  C  &  C  &     &  C  &     \\
-{\sc DICOpt}      &  C  &  I  &  C  &  I  &  C  &  I  \\
-{\sc FICO XPress} &  C  &     &  C  &     &  C  &     \\
-{\sc Gurobi}      &  C  &     &  C  &     &  C  &     \\
-{\sc IpOpt}       &     &     &     &     &  C  &  I  \\
-{\sc KNitro}      &  C  &  I  &  C  &  I  &  C  &  A  \\
-{\sc lindo api}   &  C  &  C  &  C  &  C  &  C  &  C  \\
-{\sc lgo}         &     &     &     &     &  A  &  A  \\
-{\sc Minos}       &     &     &     &     &  C  &  I  \\
-{\sc mosek}       &  C  &     &  C  &     &  C  &     \\
-{\sc msnlp}       &     &     &     &     &  C  &  A  \\
-{\sc oqnlp}       &  A  &  A  &  A  &  A  &  C  &  A  \\
-{\sc sBB}         &  C  &  I  &  C  &  I  &  C  &  I  \\
-{\sc SCIP}        &  C  &  C  &  C  &  C  &  C  &  C  \\
-{\sc SnOpt}       &     &     &     &     &  C  &  I  \\
+\alphaecp    &  C  &  I  &  C  &  I  &  C  &  I  \\
+\antigone    &  C  &  C  &  C  &  C  &  C  &  C  \\
+\baron       &  C  &  C  &  C  &  C  &  C  &  C  \\
+\bonmin      &  C  &  I  &  C  &  I  &  C  &  I  \\
+\conopt      &     &     &     &     &  C  &  I  \\
+\couenne     &  C  &  C  &  C  &  C  &  C  &  C  \\
+\cplex       &  C  &  C  &  C  &     &  C  &     \\
+\dicopt      &  C  &  I  &  C  &  I  &  C  &  I  \\
+\gurobi      &  C  &     &  C  &     &  C  &     \\
+\ipopt       &     &     &     &     &  C  &  I  \\
+\knitro      &  C  &  I  &  C  &  I  &  C  &  A  \\
+\lindo       &  C  &  C  &  C  &  C  &  C  &  C  \\
+\lgo         &     &     &     &     &  A  &  A  \\
+\minos       &     &     &     &     &  C  &  I  \\
+\mosek       &  C  &     &  C  &     &  C  &     \\
+\msnlp       &     &     &     &     &  C  &  A  \\
+\oqnlp       &  A  &  A  &  A  &  A  &  C  &  A  \\
+\sbb         &  C  &  I  &  C  &  I  &  C  &  I  \\
+\scip        &  C  &  C  &  C  &  C  &  C  &  C  \\
+\snopt       &     &     &     &     &  C  &  I  \\
+\xpress      &  C  &     &  C  &     &  C  &     \\
 \hline
@@
 }
-
-\todo[inline]{Xpress has a solver for convex MIQCPs called Xpress-SLP.
-However, it is not available within GAMS.
-Should Table 2 take Xpress-SLP into account?}
 

### 2017-01-23 solvers.tex
solvers was supposed to go between methods classification and methods

@@
-%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-%  QPLIB-2.4.tex
-%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-\subsection{Solvers}\label{subsec:solver}
+\subsubsection{Solvers}\label{subsec:solver}
 
-We now provide a succinct list of the solvers we have worked with.
-In Table \ref{t:solvers}, we mark with ``I'' a pair (solver, problem) if the solver accepts the problem in input but it is an incomplete solver for the problem, with ``A'' if it is asymptotically complete, with ``C'' if it is complete, and leave it blank if the solver won't accept the problem in input.
+When compiling QPLIB, we have worked with QP solvers that come with the GAMS distribution\footnote{\url{https://www.gams.com}}.
+A list of these solvers is provided by Table \ref{t:solvers}.
+
+We mark a pair (solver, problem) with ``I'' if the solver accepts the problem in input but it is an incomplete solver for the problem, with ``A'' if it is asymptotically complete, with ``C'' if it is complete, and leave it blank if the solver won't accept the problem in input.
+When a solver implements several algorithms, we have chosen, for each problem class, the algorithm that potentially provides the ``strongest'' results (``C'' $>$ ``A'' $>$ ``I'' $>$ blank).
 
@@
 \end{table}
-
-%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-%  End QPLIB-2.4.tex
-%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

### 2017-01-23 methods.tex
solvers was supposed to go between methods classification and methods
make intro part to solution methods independent of set of solvers

@@
 
-\subsection{Solution Methods}\label{sec:algo}
+\subsection{Solution Methods and Solvers}\label{sec:algo}
 
-In this section we provide a quick overview of existing solution methods for QP, restricting ourselves to these implemented by the set of solvers considered in this paper (Table \ref{t:solvers}).
-For each approach we briefly describe the main algorithmic ideas and point out the formulation they address according to the classification set out in Sect.~\ref{sec:classification}.
+In this section we provide a quick overview of existing solution methods for QP, restricting ourselves to these implemented by the set of solvers considered in this paper (\S~\ref{subsec:solver}).
+For each approach we briefly describe the main algorithmic ideas and point out the formulation they address according to the classification set out in \S~\ref{sec:classification}.
 We remark that many solvers implement more than one algorithm, among which the user can choose at runtime.
@@
 
-% antigone/glomiqo: sBB
-% baron: sBB
-% couenne: sBB
-% knitro: IPM with CG, SLQP, convexBB + outerapproxBB
-% lindo/lindoglobal: sBB
-% scip: sBB
-% msnlp/oqnlp: multistart
-% alphaecp: cutting plane for pseudoconvex MINLP
-% bonmin: convexBB
-% dicopt: alternating projection NLP/MILP
-% sbb: convexBB
-% conopt: GRG (active set -- like SQP)
-% ipopt: IPM w/ line search filter
-% lgo: Lipschitz BB, random search, multistart (default) [cassato?]
-% minos: active set method
-% snopt: active set SQP
-% xpress: convexBB
-% gurobi: convexBB
-% cplex: sBB
-
-The methods implemented by the solvers in Table~\ref{t:solvers} can, following \cite{neumaier}, be broadly organized in three categories: {\it incomplete}, {\it asymptotically complete}, and {\it complete}.
+Solution methods for QP can, following \cite{neumaier}, be broadly organized in three categories: {\it incomplete}, {\it asymptotically complete}, {\it complete}, and {\it rigorous}.
 Incomplete methods are only able to identify solutions, often locally optimal according to a suitable notion, and may even fail to find one even when one exists; in particular, they are typically not able to determine that an instance is empty.
@@
 Complete methods find an approximate globally optimal solution within a prescribed optimality tolerance within finite time, or prove that none such exists (but see Sect.~\ref{s:complete} below); they are often referred to as \emph{exact} methods in the computational optimization community.
-According to \cite{neumaier}, {\it rigorous} methods also exist, which find globally optimal solutions within given tolerances even in the presence of rounding errors, except for ``near-degenerate cases''.
+Finally, {\it rigorous} methods find globally optimal solutions within given tolerances even in the presence of rounding errors, except for ``near-degenerate cases''.
 Since none of the solvers we are using can be classified as rigorous, we limit ourselves to declaring solvers complete.
@@
 Therefore, when we state that a given algorithm is incomplete or (asymptotically) complete we mean that it is so the largest problem class that the solver naturally targets, although it may be complete on specific sub-classes.
-For example, \snopt naturally targets continuous NLPs and it is incomplete on nonconvex NLPs, and therefore on QCQ, but becomes complete for CCC.
+For example, interior point algorithms naturally target continuous NLPs and are incomplete on nonconvex NLPs, and therefore on QCQ, but become complete for CCC.
 In general, all complete methods for a problem $P$ must be complete for any problem $Q \subseteq P$, while a complete method for $P$ might be incomplete for $R \supset P$.
+
+\input{solvers.tex}
 

### 2017-01-23 methods.tex
start to add references for solvers

@@
 
-Typically, bounds in the optimization direction are computed by means of convex relaxations \cite{mccormick,couenne}, which replace nonconvex terms $t(x)$ with linearization variables $\hat{t}$, and then replace the corresponding defining constraints $\hat{t} = t(x)$ by means of lower and upper (respectively, convex and concave) bounding functions $\hat{t} \ge \underline{t}(x)$ and $\hat{t} \le\bar{t}(x)$.
+Typically, bounds in the optimization direction are computed by means of convex relaxations \cite{mccormick,BeLeLiMaWa08}, which replace nonconvex terms $t(x)$ with linearization variables $\hat{t}$, and then replace the corresponding defining constraints $\hat{t} = t(x)$ by means of lower and upper (respectively, convex and concave) bounding functions $\hat{t} \ge \underline{t}(x)$ and $\hat{t} \le\bar{t}(x)$.
 This is actually where finite (and tight) bounds on the variables are crucial, which differentiate also in practice the bounded case from the unbounded one.
-Different strategies are used when the nonconvexities are only quadratic \cite{glomiqo,bliek}.
+Different strategies are used when the nonconvexities are only quadratic \cite{MiFl13,bliek}.
 
@@
 %
-Cutting plane approaches construct and iteratively improve a MILP (LIL) relaxation of the problem \cite{quesada,westerlund}.
+Cutting plane approaches construct and iteratively improve a MILP (LIL) relaxation of the problem \cite{DuGr86,WePoe02}.
 The cutting planes for the MILP are generated by linearization (first-order Taylor approximation) of the nonlinearities.
@@
 
-Solvers in Table \ref{t:solvers} that implement complete cutting plane methods for CGC are \alphaecp \cite{westerlund}, \bonmin (in the algorithmic mode B-OA) \cite{BBCCGLLLMSW} and \dicopt \cite{quesada}.
-\todo{here we give references to the solvers, above we don't}
+Solvers in Table \ref{t:solvers} that implement complete cutting plane methods for CGC are \alphaecp, \bonmin (in the algorithmic mode B-OA), and \dicopt.
 

### 2017-01-23 solvers.tex
make intro part to solution methods independent of set of solvers
start to add references for solvers

@@
 
-When compiling QPLIB, we have worked with QP solvers that come with the GAMS distribution\footnote{\url{https://www.gams.com}}.
-A list of these solvers is provided by Table \ref{t:solvers}.
-
-We mark a pair (solver, problem) with ``I'' if the solver accepts the problem in input but it is an incomplete solver for the problem, with ``A'' if it is asymptotically complete, with ``C'' if it is complete, and leave it blank if the solver won't accept the problem in input.
+When compiling QPLIB, we have worked with the QP solvers that come with the GAMS distribution\footnote{\url{https://www.gams.com}}. 
+Table \ref{t:solvers} provides a list of these solvers, together with a classification of their algorithm, and references.
+For more details on the solvers, we refer to the given references, solver manuals, and the survey \cite{BuVi10}.
+In the table, we mark a pair (solver, problem) with ``I'' if the solver accepts the problem in input but it is an incomplete solver for the problem, with ``A'' if it is asymptotically complete, with ``C'' if it is complete, and leave it blank if the solver won't accept the problem in input.
 When a solver implements several algorithms, we have chosen, for each problem class, the algorithm that potentially provides the ``strongest'' results (``C'' $>$ ``A'' $>$ ``I'' $>$ blank).
@@
 \begin{table}
-{
  \centering
@@
  \setlength{\tabcolsep}{3pt}
-% \renewcommand \arraystretch{1}
 
-\begin{tabular}{lccccccc}
+\begin{tabular}{llccccccc}
 \toprule
-             & CGL & QGL & CGC & QGQ & CCC & QCQ \\
-\hline
-\alphaecp    &  C  &  I  &  C  &  I  &  C  &  I  \\
-\antigone    &  C  &  C  &  C  &  C  &  C  &  C  \\
-\baron       &  C  &  C  &  C  &  C  &  C  &  C  \\
-\bonmin      &  C  &  I  &  C  &  I  &  C  &  I  \\
-\conopt      &     &     &     &     &  C  &  I  \\
-\couenne     &  C  &  C  &  C  &  C  &  C  &  C  \\
-\cplex       &  C  &  C  &  C  &     &  C  &     \\
-\dicopt      &  C  &  I  &  C  &  I  &  C  &  I  \\
-\gurobi      &  C  &     &  C  &     &  C  &     \\
-\ipopt       &     &     &     &     &  C  &  I  \\
-\knitro      &  C  &  I  &  C  &  I  &  C  &  A  \\
-\lindo       &  C  &  C  &  C  &  C  &  C  &  C  \\
-\lgo         &     &     &     &     &  A  &  A  \\
-\minos       &     &     &     &     &  C  &  I  \\
-\mosek       &  C  &     &  C  &     &  C  &     \\
-\msnlp       &     &     &     &     &  C  &  A  \\
-\oqnlp       &  A  &  A  &  A  &  A  &  C  &  A  \\
-\sbb         &  C  &  I  &  C  &  I  &  C  &  I  \\
-\scip        &  C  &  C  &  C  &  C  &  C  &  C  \\
-\snopt       &     &     &     &     &  C  &  I  \\
-\xpress      &  C  &     &  C  &     &  C  &     \\
-\hline
+             & & CGL & QGL & CGC & QGQ & CCC & QCQ \\
+\midrule
+\alphaecp    & \cite{WeLu03,WePoe02} &  C  &  I  &  C  &  I  &  C  &  I  \\
+\antigone    & \cite{MiFl13,MiFl14} &  C  &  C  &  C  &  C  &  C  &  C  \\
+\baron       & \cite{TaSa02,TaSa04,TaSa05} &  C  &  C  &  C  &  C  &  C  &  C  \\
+\bonmin      & \cite{BBCCGLLLMSW} &  C  &  I  &  C  &  I  &  C  &  I  \\
+\conopt      &  &     &     &     &     &  C  &  I  \\
+\couenne     & \cite{BeLeLiMaWa08} &  C  &  C  &  C  &  C  &  C  &  C  \\
+\cplex       & &  C  &  C  &  C  &     &  C  &     \\
+\dicopt      & \cite{DuGr86,KoGr89,ViGr90} &  C  &  I  &  C  &  I  &  C  &  I  \\
+\gurobi      & \cite{Ro12} &  C  &     &  C  &     &  C  &     \\
+\ipopt       & &     &     &     &     &  C  &  I  \\
+\knitro      & \cite{ByNoWa06} &  C  &  I  &  C  &  I  &  C  &  A  \\
+\lindo       & \cite{LiSc09} &  C  &  C  &  C  &  C  &  C  &  C  \\
+\lgo         & &     &     &     &     &  A  &  A  \\
+\minos       & &     &     &     &     &  C  &  I  \\
+\mosek       & &  C  &     &  C  &     &  C  &     \\
+\msnlp       & &     &     &     &     &  C  &  A  \\
+\oqnlp       & \cite{ULPGKM02} &  A  &  A  &  A  &  A  &  C  &  A  \\
+\sbb         &  &  C  &  I  &  C  &  I  &  C  &  I  \\
+\scip        & \cite{Ach09,VigerskeGleixner2016} &  C  &  C  &  C  &  C  &  C  &  C  \\
+\snopt       & &     &     &     &     &  C  &  I  \\
+\xpress      & &  C  &     &  C  &     &  C  &     \\
+\bottomrule
 \end{tabular}
 \caption{Families of QP problems that can be tackled by each solver} \label{t:solvers}
-}
 
 \end{table}
+
+% antigone/glomiqo: sBB
+% baron: sBB
+% couenne: sBB
+% knitro: IPM with CG, SLQP, convexBB + outerapproxBB
+% lindo/lindoglobal: sBB
+% scip: sBB
+% msnlp/oqnlp: multistart
+% alphaecp: cutting plane for pseudoconvex MINLP
+% bonmin: convexBB
+% dicopt: alternating projection NLP/MILP
+% sbb: convexBB
+% conopt: GRG (active set -- like SQP)
+% ipopt: IPM w/ line search filter
+% lgo: Lipschitz BB, random search, multistart (default) [cassato?]
+% minos: active set method
+% snopt: active set SQP
+% xpress: convexBB
+% gurobi: convexBB
+% cplex: sBB

### 2017-01-23 QPLIB-1.0.tex
make intro part to solution methods independent of set of solvers
fix some latex warnings

@@
 \bottomrule 
-\multicolumn{3}{l}{\multirow{1}{\textwidth}{$^{\ddagger}$Applications with many manuscripts cite reviews and recent works. $^{\star}$TSP $\equiv$ Traveling Salesman Problem}} \\ \\ \\
+\multicolumn{3}{l}{\multirow{1}{\textwidth}{$^{\ddagger}$Applications with many manuscripts cite reviews and recent works.}} \\ \\ \\
 \endlastfoot
@@
 %
-\multirow{1}{.55\textwidth}{{\bf Finance:} Portfolio optimization} & \checkmark & \multirow{1}{.25\textwidth}{\cite{deng-etal:2013,kallrath:2003,lin-etal:2005,maranas-etal:1997,parpas-rustem:2006,rios-sahinidis:2010,FrFG16,FrGe06a,FrGe07a,FrGe09a}} \\
+{\bf Finance} \\
+Portfolio optimization & \checkmark & \multirow{1}{.25\textwidth}{\cite{deng-etal:2013,kallrath:2003,lin-etal:2005,maranas-etal:1997,parpas-rustem:2006,rios-sahinidis:2010,FrFG16,FrGe06a,FrGe07a,FrGe09a}} \\ \\
 %
@@
 %
-\multirow{1}{.55\textwidth}{TSP$^{\star}$ with Neighborhoods} & \checkmark & \cite{gentilini-etal:2013} \\
+\multirow{1}{.55\textwidth}{Traveling Salesman Problem with Neighborhoods} & \checkmark & \cite{gentilini-etal:2013} \\ \\
 %
@@
 The structure of this paper is the following.
-In \S~\ref{sec:QPbasic} we review the basic notions about QP.
+\todo{citation of sections should be uniform: ``\S'' vs.
+``Sect.'' vs.
+``Section''}In \S~\ref{sec:QPbasic} we review the basic notions about QP.
 In particular, \S~\ref{subsec:notation} sets out the notation, \S~\ref{sec:classification} proposes a---to the best of our knowledge, new---taxonomy of QP that helps us in discussing the (very) different classes of QPs, \S~\ref{sec:algo} very briefly reviews the solution methods for QP upon which the solvers we have employed are based, and \S~\ref{subsec:solver} describes the solvers.

### 2017-01-23 QPLIB-1.0.tex
unify referencing of ((sub)sub)sections, add other todos

@@
 
-On the other hand, (MI)QP is, in some sense, a considerably more expressive class than (MI)LP.
+On the other hand, (MI)QP\todo{wasn't QP supposed to already include the mixed-integer case?
+so why now again MIQP?} is, in some sense, a considerably more expressive class than (MI)LP.
 Quadratic expressions are found, either naturally or after appropriate reformulations, in very many optimization problems \cite{Kochenberger2014}.
@@
 The structure of this paper is the following.
-\todo{citation of sections should be uniform: ``\S'' vs.
-``Sect.'' vs.
-``Section''}In \S~\ref{sec:QPbasic} we review the basic notions about QP.
+In \S~\ref{sec:QPbasic} we review the basic notions about QP.
 In particular, \S~\ref{subsec:notation} sets out the notation, \S~\ref{sec:classification} proposes a---to the best of our knowledge, new---taxonomy of QP that helps us in discussing the (very) different classes of QPs, \S~\ref{sec:algo} very briefly reviews the solution methods for QP upon which the solvers we have employed are based, and \S~\ref{subsec:solver} describes the solvers.
 Then, \S~\ref{sec:lib} describes the process used to obtain the library and its results; the software tools that we have used, and that are freely released together with the library, are discussed in \S~\ref{subsec:tools}.
-Some conclusions are drawn in \S~\ref{sec:conclusions}, after which in the Appendix a complete description of all the instances of the library is provided.
+Some conclusions are drawn in \S~\ref{sec:conclusions}, after which in \S~\ref{sec:instance_details} a complete description of all the instances of the library is provided\todo{and then there is \S~\ref{sec:format}}.
 

### 2017-01-23 QPLIB-3.tex
fix some latex warnings
unify referencing of ((sub)sub)sections, add other todos

@@
 In this section we present all the steps performed in order to build the
-new library. In \S \ref{subsec:instColl}, we describe the set of gathered
-instances. In \S \ref{subsec:selection} we present the features used to
+new library. In \S~\ref{subsec:instColl}, we describe the set of gathered
+instances. In \S~\ref{subsec:selection} we present the features used to
 classify the instances and we discuss the issues concerning the format of
-the instances. Finally, in \S \ref{subsec:final set}, we describe the
+the instances. Finally, in \S~\ref{subsec:final set}, we describe the
 selection process used to filter the instances and we graphically present
@@
 format presents two advantages: it is easier to read by a self-made parser and
-it produces smaller files. See the Appendix~\label{sec:format} for more details.
+it produces smaller files. See the \S~\ref{sec:format} for more details.
 
@@
 which allowed us to classify the instances into the QP categories described in
-Section~\ref{sec:QPbasic}. As far as the variable types are concerned, we
+\S~\ref{sec:QPbasic}. As far as the variable types are concerned, we
 collected the following information:
@@
        for the hardness of an instance is the CPU time needed by a
-       complete solver (cf.~\S \ref{sec:algo}) to solve it to
+       complete solver (cf.~\S~\ref{sec:algo}) to solve it to
        global optimality. Accordingly, for each of the gathered instance we
@@
   \caption{Distribution of variables and constrains  of the qplib
-instances \label{fig:1}}
+instances \label{fig:distribution}}
 \end{figure}
@@
 
-Additional details on the instance features can be found in the Appendix~\ref{sec:instance_details}.
+Additional details on the instance features can be found in \S~\ref{sec:instance_details}.
 %- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

### 2017-01-23 QPLIB-2.2.tex
replace \bf comments by \todo

@@
 \subsubsection{Classification}\label{ssec:taxonomy}
+\todo[inline]{subsubsection has some title as subsection}
 
@@
 Indeed, while C*L (and, a fortiori, C*\{C,Q\}) can always be brought to D*L (D*\{C,Q\}), using the same technique C*B becomes D*L, which is significantly different from D*B.
-In practice, a diagonal convex objective function under linear constraints is found in many applications ({\bf citations?}), so that D*L still makes sense to distinguish the case where the objective function is ``naturally'' separable from that where separability is artificially introduced, although this is in theory always possible.
+In practice, a diagonal convex objective function under linear constraints is found in many applications\todo{citations?}, so that D*L still makes sense to distinguish the case where the objective function is ``naturally'' separable from that where separability is artificially introduced, although this is in theory always possible.
 
@@
 Of course, these problems can still require considerable time, like LCL, when the size of the instance grows.
-In this sense, like in the linear case, a significant divide is from solvers that need all the data of QP to work, and those that are ``matrix-free'', i.e., only require the solution of simple operations (typically, matrix-vector products) with the data of the problem to work ({\bf citations?}).
+In this sense, like in the linear case, a significant divide is from solvers that need all the data of QP to work, and those that are ``matrix-free'', i.e., only require the solution of simple operations (typically, matrix-vector products) with the data of the problem to work\todo{citations?}.
 While in our library we have never exploited such a characteristic, which is not suitable to the use of standard modeling tools, this may be relevant for the solution of very-large-scale CIC.
@@
  \item \emph{Nonconvex Continuous Quadratic Programs} QCQ are instead in general $\mathcal{NP}$-hard, even if the constraints are very specific (QCB) and only one eigenvalue of $Q^0$ is negative \cite{Hemmecke2010}.
-They therefore require enumerative techniques like the spatial Branch\&Bound ({\bf citations?}), to be solved to optimality.
+They therefore require enumerative techniques like the spatial Branch\&Bound\todo{citations?}, to be solved to optimality.
 Of course, local approaches are available that are able to efficiently provide saddle points (hopefully, local optima) of the CQC, but providing global guarantees about the quality of the obtained solutions is challenging.
@@
 However, the general approaches for CBC are basically the same as for CGC, so there is seldom the need to distinguish between the two classes as far as solvability is concerned, although matters can be different regarding actual solution cost.
-Programs with only binary variables CBC can be easier than mixed-binary or integer ones C\{M,I\}C because some techniques are specifically known for the binary-only case, cf.~the next point ({\bf citations?}).
-Absence of continuous variables, even in the presence of integer ones CIC, can also lead to specific techniques ({\bf citations??}).
+Programs with only binary variables CBC can be easier than mixed-binary or integer ones C\{M,I\}C because some techniques are specifically known for the binary-only case, cf.~the next point\todo{citations?}.
+Absence of continuous variables, even in the presence of integer ones CIC, can also lead to specific techniques\todo{citations??}.
   %
@@
 However, the special nature of binary variables combined with quadratic forms allows for quite specific techniques to be developed, among which is the reformulation of the problem as a LBL.
-Also, many well-known combinatorial problems can be naturally reformulated as problems of this class, and therefore a considerable number of results have been obtained by exploiting specific properties of the set of constraints ( \cite{Rendl2008}, {\bf some citations?
-some specific problem to mention apart from Max-Cut?}).
+Also, many well-known combinatorial problems can be naturally reformulated as problems of this class, and therefore a considerable number of results have been obtained by exploiting specific properties of the set of constraints \cite{Rendl2008}\todo{some citations?
+some specific problem to mention apart from Max-Cut?}.
   %
@@
 Due to the lack of convexity even when integrality requirements are removed, solution methods must typically combine several algorithmic ideas, such as enumeration (distinguishing the role of integral variables from that of continuous ones involved into nonconvex terms) and techniques (e.g., outer approximation, SDP relaxation, \ldots) that allow to efficiently compute bounds.
-As in the convex case, QBQ, QMQ, and QIQ can benefit from more specific properties of the variables (\cite{Buchheim2013,Dong2016}, {\bf citations?}).
+As in the convex case, QBQ, QMQ, and QIQ can benefit from more specific properties of the variables \cite{Buchheim2013,Dong2016}\todo{citations?}.
 \end{itemize}

### 2017-01-23 QPLIB-3.tex
replace \bf comments by \todo

@@
        according to their features; for each cluster we kept only a few
-       representatives. {\bf Comment: we may want to be a tad more specific
+       representatives\todo{we may want to be a tad more specific
        about what ``features'' exactly meant there}. Finally, in order to only
@@
        instances which have been solved to proven optimality within this time limit
-       were discarded. {\bf Comment: this might actually be best characterized as
-       a third filter.}
+       were discarded\todo{this might actually be best characterized as
+       a third filter}.
 \end{itemize}

### 2017-01-23 methods.tex
replace \bf comments by \todo

@@
 addresses that problem or not. 
-{\bf Antonio: the concept of ``addresses that problem'' is not very clear to me, this sentence may benefit from rephrasing (or the paragraph from deleting if it's not deemed to be utterly necessary).} Depending on the local search employed, Multi-start methods can address MINLP of the most general class.
+\todo[inline]{Antonio: the concept of ``addresses that problem'' is not very clear to me, this sentence may benefit from rephrasing (or the paragraph from deleting if it's not deemed to be utterly necessary).} Depending on the local search employed, Multi-start methods can address MINLP of the most general class.
 

### 2017-01-23 QPLIB-1.0.tex
replace \bf comments by \todo
replace {\tt by \texttt{ and similar for \it and \bf
- i think \tt, \it, \bf, etc., are deprecated

@@
 Hence, (MI)QP is, in some sense, roughly as expressive as the whole of (MI)NLP.
-Of course this is, in principle, true for (MI)LP as well, but at the cost of much larger and much more complex formulations ({\bf citations?}).
+Of course this is, in principle, true for (MI)LP as well, but at the cost of much larger and much more complex formulations\todo{citations?}.
 Hence, for many applications QP may represent the ``sweet spot'' between the effectiveness, but lower expressive power, of (MI)LP and the higher expressive power, but much higher computational cost, of (MI)NLP.
@@
 %
-\multicolumn{3}{l}{{\bf Classical Problems that can be formulated as MIQP}} \\[1pt]
+\multicolumn{3}{l}{\textbf{Classical Problems that can be formulated as MIQP}} \\[1pt]
 %
@@
 %
-\multicolumn{3}{l}{{\bf Computational chemistry \& Molecular biology}} \\[1pt]
+\multicolumn{3}{l}{\textbf{Computational chemistry \& Molecular biology}} \\[1pt]
 %
@@
 %
-\multicolumn{3}{l}{{\bf Computational geometry}} \\[1pt]
+\multicolumn{3}{l}{\textbf{Computational geometry}} \\[1pt]
 %
@@
 %
-{\bf Finance} \\
+\textbf{Finance} \\
 Portfolio optimization & \checkmark & \multirow{1}{.25\textwidth}{\cite{deng-etal:2013,kallrath:2003,lin-etal:2005,maranas-etal:1997,parpas-rustem:2006,rios-sahinidis:2010,FrFG16,FrGe06a,FrGe07a,FrGe09a}} \\ \\
@@
 %
-\multicolumn{3}{l}{{\bf Process Networks}} \\[1pt]
+\multicolumn{3}{l}{\textbf{Process Networks}} \\[1pt]
 %
@@
 %
-\multicolumn{3}{l}{{\bf Robotics}} \\[1pt]
+\multicolumn{3}{l}{\textbf{Robotics}} \\[1pt]
 %
@@
 %
-\multicolumn{3}{l}{{\bf Telecommunications}} \\[1pt]
+\multicolumn{3}{l}{\textbf{Telecommunications}} \\[1pt]
 %
@@
 %
-\multicolumn{3}{l}{{\bf Energy}} \\[1pt]
+\multicolumn{3}{l}{\textbf{Energy}} \\[1pt]
 %

### 2017-01-23 QPLIB-3.tex
replace {\tt by \texttt{ and similar for \it and \bf
- i think \tt, \it, \bf, etc., are deprecated

@@
 \begin{itemize}
- \item the {\tt BARON} library
+ \item the \texttt{BARON} library
  \url{http://www.minlp.com/nlp-and-minlp-test-problems};
 %
-\item the {\tt CUTEst} library
+\item the \texttt{CUTEst} library
  \url{https://ccpforge.cse.rl.ac.uk/gf/project/cutest/wiki};
 %
-\item the {\tt GAMS Performance} libraries
+\item the \texttt{GAMS Performance} libraries
  \url{http://www.gamsworld.org/performance/performlib.htm};
 %
-\item the {\tt MacMINLP} library
+\item the \texttt{MacMINLP} library
  \url{https://wiki.mcs.anl.gov/leyffer/index.php/MacMINLP};
 %
-\item the {\tt Meszaros} library
+\item the \texttt{Meszaros} library
  \url{http://www.doc.ic.ac.uk/~im/00README.QP};
 %
-\item the {\tt MINLP} library
+\item the \texttt{MINLP} library
  \url{http://www.gamsworld.org/minlp/minlplib.htm};
 %
-\item the {\tt POLIP} library
+\item the \texttt{POLIP} library
  \url{http://polip.zib.de/pipformat.php}.
@@
        global optimality. Accordingly, for each of the gathered instance we
-       ran the complete solvers in {\tt GAMS} (cf.~Table \ref{t:solvers})
+       ran the complete solvers in \texttt{GAMS} (cf.~Table \ref{t:solvers})
        capable of solving it, whose number depends on the category of the

### 2017-01-23 format.tex
unify referencing of ((sub)sub)sections, add other todos
replace {\tt by \texttt{ and similar for \it and \bf
- i think \tt, \it, \bf, etc., are deprecated

@@
 
-Following the notation of Section~\ref{sec:QPbasic}, a generic Quadratic Program (QP) can be written as follows:
+Following the notation of \S~\ref{sec:QPbasic}, a generic Quadratic Program (QP) can be written as follows:
 %
@@
 Real values may either by in decimal or exponential formats; for the latter,
-the exponent may be preceded by either the character {\tt D} or {\tt E},
-e.g. {\tt 12.56D+2} or {\tt 12.56E+2}.
+the exponent may be preceded by either the character \texttt{D} or \texttt{E},
+e.g. \texttt{12.56D+2} or \texttt{12.56E+2}.
 
@@
 value $i$} &  \\
-$j$\; ${\tt var\_name}_j$ & \lbox{index (integer) and name (character string)
+$j$\; $\texttt{var\_name}_j$ & \lbox{index (integer) and name (character string)
 for each non-default variable name, if $n^x_d > 0$,  one pair per line} & \\
@@
 the numerical value $i$} &  \\
-$i$\; ${\tt cons\_name}_i$ & \lbox{index  (integer) and name (character string) for each
+$i$\; $\texttt{cons\_name}_i$ & \lbox{index  (integer) and name (character string) for each
 non-default constraint name, if $n^c_d > 0$,  one pair per line} & \\
@@
 \item For continuous problems (i.e., all variables are continuous):
-\begin{description}[leftmargin=!,labelwidth=\widthof{\ttfamily SMILQPC}]
+\begin{description}[leftmargin=!,labelwidth=\widthof{\texttt{SMILQPC}}]
 \item [LP]      a linear program,
@@
 \item For integer problems (i.e., all variables are integer valued):
-\begin{description}[leftmargin=!,labelwidth=\widthof{\ttfamily SMILQPC}]
+\begin{description}[leftmargin=!,labelwidth=\widthof{\texttt{SMILQPC}}]
 \item [ILP]     an integer linear program,
@@
 integer variables):
-\begin{description}[leftmargin=!,labelwidth=\widthof{\ttfamily SMILQPC}]
+\begin{description}[leftmargin=!,labelwidth=\widthof{\texttt{SMILQPC}}]
 \item [MILP]    a mixed-integer linear program,

### 2017-01-23 methods.tex
replace {\tt by \texttt{ and similar for \it and \bf
- i think \tt, \it, \bf, etc., are deprecated

@@
 
-Solution methods for QP can, following \cite{neumaier}, be broadly organized in three categories: {\it incomplete}, {\it asymptotically complete}, {\it complete}, and {\it rigorous}.
+Solution methods for QP can, following \cite{neumaier}, be broadly organized in three categories: \emph{incomplete}, \emph{asymptotically complete}, \emph{complete}, and \emph{rigorous}.
 Incomplete methods are only able to identify solutions, often locally optimal according to a suitable notion, and may even fail to find one even when one exists; in particular, they are typically not able to determine that an instance is empty.
@@
 Complete methods find an approximate globally optimal solution within a prescribed optimality tolerance within finite time, or prove that none such exists (but see Sect.~\ref{s:complete} below); they are often referred to as \emph{exact} methods in the computational optimization community.
-Finally, {\it rigorous} methods find globally optimal solutions within given tolerances even in the presence of rounding errors, except for ``near-degenerate cases''.
+Finally, rigorous methods find globally optimal solutions within given tolerances even in the presence of rounding errors, except for ``near-degenerate cases''.
 Since none of the solvers we are using can be classified as rigorous, we limit ourselves to declaring solvers complete.
@@
 Most local search methods for *C* are iterative in nature, and need a starting point $x^0$ as input.
-The ($k+1$)-st iterate is obtained as $x^{k+1} = x^k + \alpha_k d^k$, where $\alpha_k$ (a scalar) is the {\it step length}, and $d^k$ (a vector) is the {\it search direction}, e.g.~belonging to a tangent manifold and having a negative directional derivative at $x^k$, in order to improving optimality while reducing infeasibility.
+The ($k+1$)-st iterate is obtained as $x^{k+1} = x^k + \alpha_k d^k$, where $\alpha_k$ (a scalar) is the \emph{step length}, and $d^k$ (a vector) is the \emph{search direction}, e.g.~belonging to a tangent manifold and having a negative directional derivative at $x^k$, in order to improving optimality while reducing infeasibility.
 Alternatively, $x^{k+1} = x^k + d^k(\alpha_k)$ where $d^k(\cdot)$ is an arc satisfying $d^k(0) = 0$.
@@
 %
-At each iteration $k$ the algorithm forms the {\it active set} $\mathcal{A}^k$ containing the (indices of the) {\it active constraints}, i.e., all of the equality constraints as well as the inequality constraints that are satisfied at equality at the current iterate $x^k$.
+At each iteration $k$ the algorithm forms the \emph{active set} $\mathcal{A}^k$ containing the (indices of the) \emph{active constraints}, i.e., all of the equality constraints as well as the inequality constraints that are satisfied at equality at the current iterate $x^k$.
 A subproblem, consisting of a minimization of a certain auxiliary objective function and including only active constraints, is then solved to identify a good search direction $d^k$.
@@
 %
-This is a modification of an algorithm called {\it pure random search}, which consists in sampling a random point $x'$ from a given compact set known to contain a global optimum, and then sampling a new candidate solution $y$ in a neighborhood of $x'$, setting $x' \leftarrow y$ if $y$ improves $x'$, and repeating as long as a termination condition is not satisfied.
+This is a modification of an algorithm called \emph{pure random search}, which consists in sampling a random point $x'$ from a given compact set known to contain a global optimum, and then sampling a new candidate solution $y$ in a neighborhood of $x'$, setting $x' \leftarrow y$ if $y$ improves $x'$, and repeating as long as a termination condition is not satisfied.
 The adaptivity stems from changing the distribution for sampling $y$ at run-time, depending on the quality of the solutions identified by the method.
@@
 
-When the BB algorithm is allowed to select coordinate directions corresponding to continuous variables, it is called {\it spatial} BB (sBB).
+When the BB algorithm is allowed to select coordinate directions corresponding to continuous variables, it is called \emph{spatial} BB (sBB).
 Branching on continuous (rather than integer or binary) variables becomes necessary in the presence of nonconvex nonlinearities, as it happens e.g.~in QCQ, since the quality of the bounds improves as the feasible set in the current node gets smaller.

### 2017-01-24 QPLIB-3.tex
revise instance selection section
- add some todos
- fix spelling and grammar

@@
 $\approx 2000$ Quadratic Continuous Quadratic (QCQ) instances, and
-and $\approx 2500$ Quadratic General Quadratic (QGQ) instances. W
+and $\approx 2500$ Quadratic General Quadratic (QGQ) instances. We
 also gathered $\approx 1000$ Convex General Convex (CGC) instances. We
@@
 %
-In case at least one binary or integer variable is present, then the instance is
+If at least one binary or integer variable is present, then the instance is
 categorized as \emph{discrete}, otherwise it is categorized as \emph{continuous}.
-As fas as the objective function is concerned, we gathered the following
+As far as the objective function is concerned, we gathered the following
 information:
@@
  %
- \item density of the $Q^0$ matrix (number of nonzero entries over the total
-       number). %(\emph{\% dens}).
+ \item density of the $Q^0$ matrix (number of nonzero entries divided by the total
+       number of entries). %(\emph{\% dens}).
 \end{itemize}
@@
 non-negative eigenvalues are classified as convex constraints. All this
-information allowed us to analyze the gathered instances and to perform the
+information allowed us to analyse the gathered instances and to perform the
 filters described in the the next paragraph.
@@
 
-During the development of the library, a discussion ensued about what
-the expected gols that we wanted to achieve. The following four goals
+During the development of the library, a discussion ensued about
+the expected goals that we wanted to achieve. The following four goals
 where finally identified:
@@
        global optimality. Accordingly, for each of the gathered instance we
-       ran the complete solvers in \texttt{GAMS} (cf.~Table \ref{t:solvers})
-       capable of solving it, whose number depends on the category of the
-       instance under consideration. We then filtered according to a relative
+       ran the complete solvers in GAMS, which number depends on the category
+       of the instance under consideration, cf.~Table \ref{t:solvers}.
+       We then filtered according to a relative
        measure of computational difficulty, i.e., we discarded all instances
@@
        keep computationally challenging instances we ran a general purpose solver
-       able to tackle all QP categories with a time limit of 120 seconds; all the
+       able to tackle all QP categories\todo{do you mean a complete solver for QGQ? if so, maybe say so}\xspace with a time limit of 120 seconds; all the
        instances which have been solved to proven optimality within this time limit
@@
 & \multicolumn{ 2}{c}{$\Downarrow$}& \\
-& $\approx$ 6000 Discr. Inst.  & $\approx$ 2500 Cont. inst. & \\
+& $\approx$ 6000 Discr. Inst.  & $\approx$ 2500 Cont. Inst. & \\
 First Filter  & $\Downarrow$  & $\Downarrow$ & \\
@@
 \ref{tab:DD} for \emph{discrete} instances (*\{B,M,I,G\}*) and in Table
-\ref{tab:CC} for \emph{continuous} ones (*C*). For each category, the Tables
-report in column $\#$ the corresponding number of instances. It can be seen
+\ref{tab:CC} for \emph{continuous} ones (*C*). For each category, the tables
+report in column ``$\#$'' the corresponding number of instances. It can be seen
 that the final set well respects the original distribution of the gathered
-instances among the different categories. Indeed, the discrete categories
+instances\todo{Maybe one could put the numbers for the original set of instances
+into Tables \ref{tab:DD} and \ref{tab:CC}, too?}\xspace among the different categories. Indeed, the discrete categories
 (LMQ) or (QBL) are well represented by $118$ and $59$ instances,
@@
 We now report some graphs that help in illustrating the main features
-of the instences. In Figure~\ref{fig:distribution} we plot the number
+of the instances. In Figure~\ref{fig:distribution} we plot the number
 of variables (horizontal axis) versus the number of constrains
@@
 marked with ``$+$'', and discrete ones with ``$\times$''. The figure shows
-that the library contains a quite diverse set of instances of different in
-terms of number of variables and constraints. The number of constraints goes
-from few units up to one hundred thousands constraints, while the number of
-variables goes up to almost forty-five thousands.
+that the library contains a quite diverse set of instances in
+terms of number of variables and constraints. The maximal number of constraints
+is 100000, while the maximal number of
+variables is almost 40000.
 
@@
 For each instance we report the total number of variables with a ``$+$'', and the total number of discrete variables (binary or general integer) with a ``$\times$''.
-The pictures clearly show that instances with different percentages of integer and continuous variables are present in the library, and well distributed across the whole spectrum of variable sizes.
+The pictures clearly show that instances with different percentages of integer and continuous variables are present in the library, and that these differences are well distributed across the whole spectrum of variable sizes.
 
@@
 %and Figure~\ref{fig:pic_constr_big} 
-describes how linear and quadratic
+describes how the number of linear and quadratic
 constraints are distributed within the instances. The instances are sorted
-accordingly to the total number of constraints
+accordingly to the total number of constraints.
 % and divided in the same manner as for Figures~\ref{fig:pic_var_small}--\ref{fig:pic_var_large}.
-For each instances we report the total number of constraints with a ``$+$''
-and the total number of (either convex of nonconvex) quadratic constraints
-with a ``$\times$''. Also in this case, different percentages of linear and
+For each instance we report the total number of constraints with a ``$+$''
+and the total number of quadratic constraints
+with a ``$\times$''. Also, in this case, different percentages of linear and
 quadratic constraints are present and well-distributed across the spectrum
@@
 
-Figure~\ref{fig:pic_conv_constr} shows the instances with at least a quadratic constraints sorted according to the number of quadratic constraints (vertical axis).
-For each instances we report the total number of constraints with a ``$+$''
+Figure~\ref{fig:pic_conv_constr} shows the instances with at least one quadratic constraint sorted according to the number of quadratic constraints (vertical axis).
+For each instance we report the total number of constraints with a ``$+$''
 and the total number of nonconvex quadratic constraints
@@
   \includegraphics[width=0.85\textwidth]{pic_quad_conv_vs_nonconv.png}
+  \todo[inline]{horizontal axis label could say ``Instances with at least one quadratic constraint''}
   \caption{Quadratic constraints \label{fig:pic_conv_constr}}
@@
 the instances with a quadratic objective function, ordered by the
-the percentage of negative eigenvalues (vertical axis). Although
-almost all percentages are represented, the instances are mostly
-clustered around two values. About 25\% of the instances are convex,
+the percentage of negative eigenvalues\todo{Was there some tolerance applied on when an eigenvalue is considered negative?
+$-10^{-9}$, $-10^{-6}$, etc.?} in the $Q^0$ matrix (vertical axis).
+The instances are mostly
+clustered around two values. About 25\% of the instances are convex\todo{so one also restrict to linear constraints, or may only the objective be convex?},
 i.e., they have 0\% of negative eigenvalues. Among the others, a vast
-majority have around 50\% of negative eigenvalues. However, instances
-with high or low percentages of negative eigenvalues are present.
+majority has around 50\% of negative eigenvalues. However, instances
+with high or low percentages of negative eigenvalues are present, too.
 
@@
   \includegraphics[width=0.85\textwidth]{pic_neg_eig.png}
+  \todo[inline]{horizontal axis label should say ``Instances with quadratic objective function''}
   \caption{Negative eigenvalues \label{fig:pic_neg_eig}}
@@
 a very low or a rather high density: indeed, about 30\% of the instances
-have density smaller that 5\%, and about 30\% of the instances have
-density larger that 50\%. However, also intermediate values are present.
+have density smaller than 5\%, and about 30\% of the instances have
+density larger than 50\%. However, also intermediate values are present.
 
@@
   \includegraphics[width=0.85\textwidth]{pic_density.png}
+  \todo[inline]{horizontal axis label should say ``Instances with quadratic objective function''}
   \caption{Density \label{fig:pic_density}}

### 2017-01-24 QPLIB-2.2.tex
revise instance selection section
- add some todos
- fix spelling and grammar
revise conclusions and appendix

@@
  %
- \item variables: (C)ontinuous only, (B)inary only, (M)ixed binary and continuous, (I)nteger only, (G)eneral (all three types);
+ \item variables: (C)ontinuous only, (B)inary only, (M)ixed binary and continuous, (I)nteger only, (G)eneral (all three types\todo{or mixed integer and continuous?});
  %
@@
   \item \emph{Nonconvex Integer Quadratic Programs} QGQ is the most general, and therefore is the most difficul, class.
-Due to the lack of convexity even when integrality requirements are removed, solution methods must typically combine several algorithmic ideas, such as enumeration (distinguishing the role of integral variables from that of continuous ones involved into nonconvex terms) and techniques (e.g., outer approximation, SDP relaxation, \ldots) that allow to efficiently compute bounds.
+Due to the lack of convexity even when integrality requirements are removed, solution methods must typically combine several algorithmic ideas, such as enumeration (distinguishing the role of integral variables from that of continuous ones involved into nonconvex terms) and techniques (e.g., outer approximation, semidefinite programming relaxation, \ldots) that allow to efficiently compute bounds.
 As in the convex case, QBQ, QMQ, and QIQ can benefit from more specific properties of the variables \cite{Buchheim2013,Dong2016}\todo{citations?}.

### 2017-01-24 conclusion.tex
revise conclusions and appendix

@@
  This manuscript describes the first comprehensive library of instances for Quadratic Programming (QP).
-Since QP comprises  different and `varied'' categories of problems, we propose a classification and then we describe the principal solution methods representing the state-of-the-art algorithms for QP.
+Since QP comprises  different and ``varied'' categories of problems, we propose a classification and then we describe the principal solution methods representing the state-of-the-art algorithms for QP.
 

### 2017-01-27 conclusion.tex
Add files via upload

--- 2016-12-17 conclusion.tex
@@
  This manuscript describes the first comprehensive library of instances for Quadratic Programming (QP).
-Since QP comprises  different and `varied'' categories of problems, we propose a classification and then we describe the principal solution methods representing the state-of-the-art algorithms for QP.
+Since QP comprises  different and ``varied'' categories of problems, we propose a classification and we briefly discuss the main classes of solution methods for QP.
 
 We then describe the steps of the adopted process used to filter the gathered instances  in order to build the new library.
-The goals we purposed are of different natures and we tried to build a library which is computationally challenging and as broad as possible, i.e., it represents the largest possible categories of QP problems.
+The goals we purposed are of different natures and we tried to build a library which is computationally challenging and as broad as possible, i.e., it represents the largest possible categories of QP problems, while remaining of manageable size.
 

### 2017-01-27 format.tex
Add files via upload

@@
 \newcommand{\lbox}[1]{\parbox[t]{105mm}{#1}\vspace*{1mm}}
-
-Following the notation of \S~\ref{sec:QPbasic}, a generic Quadratic Program (QP) can be written as follows:
-%
-\begin{align*}
- \min \;\;
- & x^\top Q^0 x + b^0 x + f \\
- %
- & c^i_l \leq x^\top Q^i x + b^i x \leq c^i_u & i \in \Mcm \\
- %
- & l_j \leq x_j \leq u_j & j \in \Mcn  \\
- %
- & x_j \in \mathbb{Z} & j \in \Mcz
-\end{align*}
-
-where:
-
-\begin{itemize}
- \item $\Mcn = \{ 1, \ldots, n \}$ is the set of (indices) of variables, and $\Mcm = \{ 1, \ldots, m \}$ is the set of (indices) of constraints;
- %
- \item $x = [x_j]_{j = 1}^n \in \RR^n$ is a finite vector of real variables;
- %
- \item $Q^i$ for $i \in \{ 0 \} \cup \Mcm$ are symmetric $n \times n$ real matrices;
- %
- \item $b^i$, $c^i_l$, $c^i_u$ for $i \in \{ 0 \} \cup \Mcm$ and $f$ are, respectively, real $n$-vectors, real constants, real constants and real constant;
- %
- \item $-\infty \leq l_j \leq u_j \leq \infty$ are the (extended) real upper and lower bounds on each variable $x_j$ for $j \in \Mcn$;
- %
- \item $\Mcm = \Mcq \cup \Mcl$ where $Q^i = 0$ for all $i \in \Mcl$;
- %
- \item the variables in $\Mcz \subseteq \Mcm$ are restricted to only attain integer values.
-\end{itemize}
 
@@
 
-The QPLIB format is defined in Table~\ref{tab-qplib-format}.
+The QPLIB format is defined in Table~\ref{tab-qplib-format}, with the notation of \S~\ref{sec:QPbasic}.
+
 The data is in free format (blanks separate values), but must occur in
@@
 \hline
-$f$ & constant part of the objective function & \\ \hline
+$q^0$ & constant part of the objective function & \\ \hline
 $n^{Q^i}$ & number of nonzeros (integer) in lower triangle of
@@
 \hline
-$n^{b}$ & number of nonzeros  (integer) in $b^i$, $i \in \Mcm$                     & [2] \\
+$n^{b^i}$ & number of nonzeros  (integer) in $b^i$, $i \in \Mcm$                     & [2] \\
 $i$\; $j$\; $b^i_{j}$ & \lbox{row and column indices (integers) and value (real)
@@
 \hline
-$\infty$ & \lbox{value (real) for infinity for constraint or variable
+$c_{\infty}$ & \lbox{value (real) for infinity for constraint or variable
 bounds---any bound greater than or equal to this in, absolute value,
@@
 %}
-
-\todo[inline]{\small Stefan: I would suggest to make it simpler, even though that
-would reduce flexibility. Always fix the default value for coefficients to 0
-(otherwise, how you parse this? First initialize a n x n matrix with the non-zero
-default and then overwrite with the given entries, which may then be zero?)
-and have no default bounds for variables. If you then also adapt the suggestion
-to use a constraint sense ($\leq$,$\geq$,$=$) and a right-hand-side instead of
-a left- and right-hand-side, then there is also no need to specify a value for infinity.
-This would leave out ranged constraints - but are they really used? One could
-do it like CPLEX and have an extra constraint sense for ranges and an extra
-section to specify the range.}
+%
+%\todo[inline]{\small Stefan: I would suggest to make it simpler, even though that
+%would reduce flexibility. Always fix the default value for coefficients to 0
+%(otherwise, how you parse this? First initialize a n x n matrix with the non-zero
+%default and then overwrite with the given entries, which may then be zero?)
+%and have no default bounds for variables. If you then also adapt the suggestion
+%to use a constraint sense ($\leq$,$\geq$,$=$) and a right-hand-side instead of
+%a left- and right-hand-side, then there is also no need to specify a value for infinity.
+%This would leave out ranged constraints - but are they really used? One could
+%do it like CPLEX and have an extra constraint sense for ranges and an extra
+%section to specify the range.}
+%
 
@@
 \item [{[1]}]
-\todo[inline]{\small Fabio and Emiliano: we think that, if possible, we should follow the classification proposed in Section 2.}
+%\todo[inline]{\small Fabio and Emiliano: we think that, if possible, we should follow the classification proposed in Section 2.}
 
@@
 \end{description}
-\todo[inline]{\small Stefan: I agree that leaving out information on
-numerical properties of the matrices and vectors (convexity, etc.) is good here.
-Such properties may also depend on tolerances, etc.
-As above, I would suggest to eliminate LPQC and rename QPQC into QCP.
-Having a string for ``unspecified'' may be nice, too, or one just uses MIQCQP in this case.}
+%\todo[inline]{\small Stefan: I agree that leaving out information on
+%numerical properties of the matrices and vectors (convexity, etc.) is good here.
+%Such properties may also depend on tolerances, etc.
+%As above, I would suggest to eliminate LPQC and rename QPQC into QCP.
+%Having a string for ``unspecified'' may be nice, too, or one just uses MIQCQP in this case.}
 \end{description}
@@
 2 -0.4    1 lines of index & value of non-default values in b_0
-0.0       value of f
+0.0       value of q^0
 4         # nonzeros in vectors b^i (i =0,...,m)

### 2017-01-27 methods.tex
show list of todos up front, minor changes
Add files via upload

@@
 The solvers in Table \ref{t:solvers} which implement incomplete methods for continuous NLPs (a problem class containing QCQ) are \conopt, \ipopt, \minos, \snopt, and \knitro.
-The methods are local search algorithms and are essentially of the following types: %Note that traditionally \knitro used to be a local (nonconvex) NLP solver.
-%All of these solvers implement essentially three types of local search methods for continuous NLPs:
-\todo{sequential linear programming, sequential quadratic programming, trust-region}
+Note that all these solvers tackle the more general class of continuous NLP, while we use them only for the considerably more restricted class of QP.
+While there are many different approaches for continuous NLPs in general, we therefore describe here on those that are particularly well-suited for QP, i.e.:
+%
 \begin{itemize}
@@
 \begin{itemize}
- \item global adaptive random search (\lgo GARS);
+ \item global adaptive random search (\lgo);
  %
- \item multi-start (\knitro, \lgo MS, \msnlp, \oqnlp); specifically, the former three apply to QCQ whereas the latter to QGQ.
+ \item multi-start (\knitro, \lgo, \msnlp, \oqnlp); specifically, the former three apply to QCQ whereas the latter to QGQ.
 \end{itemize}
@@
 
-In general, the applicability of meta-heuristics to a given problem depends 
-on whether the local search they utilize
-%on what whether the local search they rest on 
-addresses that problem or not. 
-\todo[inline]{Antonio: the concept of ``addresses that problem'' is not very clear to me, this sentence may benefit from rephrasing (or the paragraph from deleting if it's not deemed to be utterly necessary).} Depending on the local search employed, Multi-start methods can address MINLP of the most general class.
+In general, the applicability of meta-heuristics to a given problem relies on availability of a local search that exploit the structural properties of the problem. 
+Depending on the local search employed, Multi-start methods can address MINLP of the most general class.
 

### 2017-01-27 QPLIB.tex
make references clickable in pdf
- include hyperref package
move solvers section before algorithms section
- as in the algorithms section, we refer which solver implements what
rename QPLIB-2.3 and QPLIB-2.4
- QPLIB-2.4 now contains Section 2.3
- QPLIB-2.3 now contains Section 2.4
add latex defines for solver names
- to make it easier to change their writing throughout the paper
change capitalization of solver names
- use the variant found in solver logs
- drop the FICO for xpress again (they insisted on it when we did the
  minlp software paper, but their own log is not filled up with FICO)
-

--- 2017-01-20 QPLIB.tex
@@
 \usepackage{amsmath,amsfonts,amssymb}
-\usepackage{amssymb}
-\usepackage{amsmath}
 \usepackage{multirow}
-\usepackage{todonotes}
-
 \usepackage{longtable}
@@
 \usepackage{calc}
+\usepackage[hidelinks]{hyperref}
+\usepackage{xspace}
+\usepackage[numbers,sort&compress]{natbib}
+\usepackage{doi}
+
+% TODO remove these before submission
+\usepackage{todonotes}
+%\usepackage{showlabels}
 
@@
 
+\def\alphaecp{\textsc{AlphaECP}\xspace}
+\def\antigone{\textsc{ANTIGONE}\xspace}
+\def\baron{\textsc{BARON}\xspace}
+\def\bonmin{\textsc{BONMIN}\xspace}
+\def\conopt{\textsc{CONOPT}\xspace}
+\def\couenne{\textsc{Couenne}\xspace}
+\def\cplex{\textsc{Cplex}\xspace}
+\def\dicopt{\textsc{DICOPT}\xspace}
+\def\xpress{\textsc{Xpress-Optimizer}\xspace}
+\def\gurobi{\textsc{Gurobi}\xspace}
+\def\ipopt{\textsc{Ipopt}\xspace}
+\def\knitro{\textsc{Knitro}\xspace}
+\def\lindo{\textsc{Lindo API}\xspace}
+\def\lgo{\textsc{LGO}\xspace}
+\def\minos{\textsc{MINOS}\xspace}
+\def\mosek{\textsc{MOSEK}\xspace}
+\def\msnlp{\textsc{MsNlp}\xspace}
+\def\oqnlp{\textsc{OQNLP}\xspace}
+\def\sbb{\textsc{SBB}\xspace}
+\def\scip{\textsc{SCIP}\xspace}
+\def\snopt{\textsc{SNOPT}\xspace}
+
 %- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
@@
 \newenvironment{proof}[1][Proof]{\textbf{#1.} }{\ \rule{0.5em}{0.5em}}
-%\renewenvironment{proof}{{\bfseries Proof}}{\qed}
 
-\def\remark{\stepcounter{theorem}\par{\bf Remark}.
+\def\remark{\stepcounter{theorem}\par\textbf{Remark}.
 \ignorespaces}
@@
 \newenvironment{keywords}{\vspace{0.4in}\begin{quote}\small \em
-{\bf Keywords\/}:}{\end{quote}}
+\textbf{Keywords\/}:}{\end{quote}}
 \fi
@@
 %- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
+
+%\listoftodos
+\newpage
 
@@
 \and       Ruth Misener \at
-           Department of Computing,Imperial College London, UK. \email{r.misener@imperial.ac.uk}
+           Department of Computing, Imperial College London, UK. \email{r.misener@imperial.ac.uk}
 %
@@
 \begin{abstract}
- This paper describes a new instance library for Quadratic Programming (QP).
+ This paper describes a new instance library for Quadratic Programming (QP), i.e., the family of continuous and (mixed)-integer optimization problems where either the objective function or the constrains (or both) are quadratic.
  QP is a very ``varied'' class, comprising sub-classes of problems ranging from
@@
 
-\input{QPLIB-2.3.tex}
+\input{methods.tex}
 
 %- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-
-\input{QPLIB-2.4.tex}
 
@@
 
-\input{QPLIB-4.tex}
+%\input{QPLIB-4.tex}
 
@@
 \section{Acknowledgements}
+
+We are grateful to all the donors who provided instances for the library.
+We gratefully acknowledge the financial support of the Gaspard Monge Program for Optimization and operations research (PGMO)  and the logistic support of GAMS for having provided us the licence of their software.
+Finally, we would like to acknowledge the financial and networking support by the COST Action TD1207.
 

### 2017-01-27 QPLIB-1.0.tex
Add files via upload

@@
 Hence, in particular for the convex case, tools and techniques that have been honed during decades of research for Linear Programming (LP), typically with integrality constraints (MILP), can often be extended to the quadratic case with at least less effort than what would be required for tackling general Non Linear Programming (NLP) problems, without or with integrality constraints (MINLP).
-This has indeed happened over and over again with different algorithmic techniques, such as Interior Point methods, active-set methods (of which the simplex method is a prototypical example), enumeration methods, cut-generation techniques, reformulation techniques, and many others \todo[inline]{citations?
-http://www.optimization-online.org/DB\_HTML/2009/07/2338.html ?}.
+This has indeed happened over and over again with different algorithmic techniques, such as Interior Point methods, active-set methods (of which the simplex method is a prototypical example), enumeration methods, cut-generation techniques, reformulation techniques, and many others \cite{BS09}.
 Similarly, nonconvex continuous QP are perhaps the ``simplest'' class of problems that require features like spatial enumeration techniques to be solved. 
-\todo[inline]{If only the objective is quadratic, i.e., what one usually would denote by a ``nonconvex QP'', then one can ``just'' enumerate local optima.
-Nitpicking?} Hence, they are both the natural basis for the development of general techniques for nonconvex NLP, and a very specific class so that specialized approaches can be developed \cite{Dur2010,Burer2012}.
+Hence, they are both the natural basis for the development of general techniques for nonconvex NLP, and a very specific class so that specialized approaches can be developed \cite{Dur2010,Burer2012}.
 
-On the other hand, (MI)QP\todo{wasn't QP supposed to already include the mixed-integer case?
-so why now again MIQP?} is, in some sense, a considerably more expressive class than (MI)LP.
+On the other hand, (MI)QP is, in some sense, a considerably more expressive class than (MI)LP.
 Quadratic expressions are found, either naturally or after appropriate reformulations, in very many optimization problems \cite{Kochenberger2014}.
@@
 Hence, (MI)QP is, in some sense, roughly as expressive as the whole of (MI)NLP.
-Of course this is, in principle, true for (MI)LP as well, but at the cost of much larger and much more complex formulations\todo{citations?}.
+Of course this is, in principle, true for (MI)LP as well, but at the cost of much larger and much more complex formulations.
 Hence, for many applications QP may represent the ``sweet spot'' between the effectiveness, but lower expressive power, of (MI)LP and the higher expressive power, but much higher computational cost, of (MI)NLP.
@@
 %
-Max-Cut & \checkmark & \cite{Rendl2008} \\ \cmidrule(rl){2-3}
+Max-Cut & \checkmark & \cite{Rendl2008,KrislockMalickRoupin2016} \\ \cmidrule(rl){2-3}
 %
@@
 %
-Packing circles$^{\ddagger}$ && \cite{hifi-mhallah:2009,szabo-etal:2005} \\ \cmidrule(rl){2-3}
+Packing circles$^{\ddagger}$ & \checkmark & \cite{FrFG16,FGGP11,hifi-mhallah:2009,szabo-etal:2005} \\ \cmidrule(rl){2-3}
 %
@@
 %
-Pooling$^{\ddagger}$ & \checkmark & \multirow{1}{.25\textwidth}{\cite{Alfaki-Haugland:2013,Castillo-etal:2013,dambrosio-etal:2011pooling,Faria-Bagajewicz:2012,misener-floudas:2009,misener-floudas:2010-genpooling,Papageorgiou-etal:2012,pham-etal:2009,ruiz-etal:2013,visweswaran:2009}} \\ \\ \cmidrule(rl){2-3}
+Pooling$^{\ddagger}$ & \checkmark & \multirow{1}{.25\textwidth}{\cite{Alfaki-Haugland:2013,Castillo-etal:2013,dambrosio-etal:2011pooling,Faria-Bagajewicz:2012,misener-floudas:2009,misener-floudas:2010-genpooling,Papageorgiou-etal:2012,pham-etal:2009,ruiz-etal:2013}} \\ \\ \cmidrule(rl){2-3}
 %
@@
 %
-\multirow{1}{.55\textwidth}{Unit-commitment problem} & \checkmark & \cite{Tetal15} \\
+\multirow{1}{.55\textwidth}{Unit-commitment} & \checkmark & \cite{FrFG16,FrGe06a,FrGe09a,Tetal15} \\
 %
+\multicolumn{3}{l}{\textbf{Data Confidentiality}} \\[1pt]
+%
+\multirow{1}{.55\textwidth}{Controlled Tabular Adjustment} & \checkmark & \cite{CaFG14} \\
 \end{longtable}
@@
 In particular, \S~\ref{subsec:notation} sets out the notation, \S~\ref{sec:classification} proposes a---to the best of our knowledge, new---taxonomy of QP that helps us in discussing the (very) different classes of QPs, \S~\ref{sec:algo} very briefly reviews the solution methods for QP upon which the solvers we have employed are based, and \S~\ref{subsec:solver} describes the solvers.
-Then, \S~\ref{sec:lib} describes the process used to obtain the library and its results; the software tools that we have used, and that are freely released together with the library, are discussed in \S~\ref{subsec:tools}.
-Some conclusions are drawn in \S~\ref{sec:conclusions}, after which in \S~\ref{sec:instance_details} a complete description of all the instances of the library is provided\todo{and then there is \S~\ref{sec:format}}.
+Then, \S~\ref{sec:lib} describes the process used to obtain the library and its results.
+%; the software tools that we have used, and that are freely released together with the library, are discussed in \S~\ref{subsec:tools}. 
+Some conclusions are drawn in \S~\ref{sec:conclusions}, after which Appendix A provides a complete description of all the instances of the library, while Appendix B describes the QPLIB file format.
 

### 2017-01-27 QPLIB-2.2.tex
Add files via upload

@@
 %- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
+%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
 
@@
 A taxonomy of QP instances should therefore strive to identify the set of properties that an instance should have in order to apply the most relevant computational methods.
-However, the sheer number of different existing approaches, and the fact that new ones are proposed, makes it hard to provide a taxonomy that is both simple and covers all possible special cases.
+However, the sheer number of different existing approaches, and the fact that new ones are continuously proposed, makes it hard to provide a taxonomy that is both simple and covers all possible special cases.
 This is why, in this paper, we propose an approach that aims at finding a good balance between simplicity and coverage of the main families of computational methods.
@@
 % -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -
-\subsubsection{Classification}\label{ssec:taxonomy}
-\todo[inline]{subsubsection has some title as subsection}
+\subsubsection{Taxonomy}\label{ssec:taxonomy}
 
@@
  %
- \item variables: (C)ontinuous only, (B)inary only, (M)ixed binary and continuous, (I)nteger only, (G)eneral (all three types\todo{or mixed integer and continuous?});
+ \item variables: (C)ontinuous only, (B)inary only, (M)ixed binary and continuous, (I)nteger only, (G)eneral (all other cases);
  %
- \item constraints: (N)one, (B)ox, (L)inear, (D)iagonal convex quadratic, (C)onvex quadratic, nonconvex (Q)ua-dratic.
+ \item constraints: (N)one, (B)ox, (L)inear, (D)iagonal convex quadratic, (C)onvex quadratic, nonconvex (Q)uadratic.
 \end{itemize}
@@
 \subsubsection{Examples and reformulations}\label{ssec:reform}
-\todo{\small Ambros: maybe switch this with the next section? the next one contains more familiar material to the reader and might be an easier first introduction to the notation.}
 
@@
  %
- & x^\top Q^0 x + b^0 x \leq x^0 & \\ 
+ & -\infty \leq x^\top Q^0 x + b^0 x \leq x^0 & \\ 
  %
- & x^\top Q^i x + b^i x \leq c^i & i \in \Mcm \\ 
+ & c^i_l \leq x^\top Q^i x + b^i x \leq c^i_u & i \in \Mcm \\ 
  %
@@
 When a $Q^i$ is positive semidefinite (PSD), i.e., the corresponding constraint/objective function is convex, general quadratic constraints are in fact equivalent to diagonal ones.
-In fact, every PSD matrix can be factorized as $Q^i = L^i (L^i)^\top$, e.g.~by the (incomplete) Cholesky factorization, $f^i(x) = x^\top Q^i x = \sum_{j \in \Mcn} z^2_jz$ where $z = x^\top L^i$.
+In fact, since every PSD matrix can be factorized as $Q^i = L^i (L^i)^\top$, e.g.~by the (incomplete) Cholesky factorization, $f^i(x) = x^\top Q^i x = \sum_{j \in \Mcn} z^2_jz$ where $z = x^\top L^i$.
 Hence, one could think that D** problems need not be distinguished from C** ones; however, this is true only for ``complicated'' constraints, but not for ``simple'' ones, because the above reformulation technique introduces linear constraints.
 Indeed, while C*L (and, a fortiori, C*\{C,Q\}) can always be brought to D*L (D*\{C,Q\}), using the same technique C*B becomes D*L, which is significantly different from D*B.
-In practice, a diagonal convex objective function under linear constraints is found in many applications\todo{citations?}, so that D*L still makes sense to distinguish the case where the objective function is ``naturally'' separable from that where separability is artificially introduced, although this is in theory always possible.
+In practice, a diagonal convex objective function under linear constraints is found in many applications (e.g., \cite{FrFG16,FrGe06a,FrGe09a,FGGP11}), so that D*L still makes sense to distinguish the case where the objective function is ``naturally'' separable from that where separability is artificially introduced.
 
@@
 Of course, these problems can still require considerable time, like LCL, when the size of the instance grows.
-In this sense, like in the linear case, a significant divide is from solvers that need all the data of QP to work, and those that are ``matrix-free'', i.e., only require the solution of simple operations (typically, matrix-vector products) with the data of the problem to work\todo{citations?}.
+In this sense, like in the linear case, a significant divide is from solvers that need all the data of QP to work, and those that are ``matrix-free'', i.e., only require the solution of simple operations (typically, matrix-vector products) with the data of the problem to work.
 While in our library we have never exploited such a characteristic, which is not suitable to the use of standard modeling tools, this may be relevant for the solution of very-large-scale CIC.
@@
  \item \emph{Nonconvex Continuous Quadratic Programs} QCQ are instead in general $\mathcal{NP}$-hard, even if the constraints are very specific (QCB) and only one eigenvalue of $Q^0$ is negative \cite{Hemmecke2010}.
-They therefore require enumerative techniques like the spatial Branch\&Bound\todo{citations?}, to be solved to optimality.
+They therefore require enumerative techniques like the spatial Branch\&Bound \cite{FV90,BLLMW09}, to be solved to optimality.
 Of course, local approaches are available that are able to efficiently provide saddle points (hopefully, local optima) of the CQC, but providing global guarantees about the quality of the obtained solutions is challenging.
@@
 However, the general approaches for CBC are basically the same as for CGC, so there is seldom the need to distinguish between the two classes as far as solvability is concerned, although matters can be different regarding actual solution cost.
-Programs with only binary variables CBC can be easier than mixed-binary or integer ones C\{M,I\}C because some techniques are specifically known for the binary-only case, cf.~the next point\todo{citations?}.
-Absence of continuous variables, even in the presence of integer ones CIC, can also lead to specific techniques\todo{citations??}.
+Programs with only binary variables CBC can be easier than mixed-binary or integer ones C\{M,I\}C because some techniques are specifically known for the binary-only case, cf.~the next point \cite{BEP09}.
+Absence of continuous variables, even in the presence of integer ones CIC, can also lead to specific techniques \cite{BEL13}.
   %
@@
 However, the special nature of binary variables combined with quadratic forms allows for quite specific techniques to be developed, among which is the reformulation of the problem as a LBL.
-Also, many well-known combinatorial problems can be naturally reformulated as problems of this class, and therefore a considerable number of results have been obtained by exploiting specific properties of the set of constraints \cite{Rendl2008}\todo{some citations?
-some specific problem to mention apart from Max-Cut?}.
+Also, many well-known combinatorial problems can be naturally reformulated as problems of this class, and therefore a considerable number of results have been obtained by exploiting specific properties of the set of constraints \cite{Rendl2008,Loiola2007657}.
   %
@@
 Due to the lack of convexity even when integrality requirements are removed, solution methods must typically combine several algorithmic ideas, such as enumeration (distinguishing the role of integral variables from that of continuous ones involved into nonconvex terms) and techniques (e.g., outer approximation, semidefinite programming relaxation, \ldots) that allow to efficiently compute bounds.
-As in the convex case, QBQ, QMQ, and QIQ can benefit from more specific properties of the variables \cite{Buchheim2013,Dong2016}\todo{citations?}.
+As in the convex case, QBQ, QMQ, and QIQ can benefit from more specific properties of the variables \cite{Buchheim2013,Dong2016}.
 \end{itemize}
@@
 These can have a significant algorithmic impact, and therefore can be of interest to researchers.
-Common structures are, e.g., network flow or knapsack-type linear constraints, semi-continuous variables, or the fact that a nonconvex quadratic objective function/constraint can be reformulated as a second-order cone (hence, convex) one.
+Common structures are, e.g., network flow \cite{FrFG16,FrGaSc14,FrGaSt16a,FGGP11,tadayon-smith:2013} or knapsack-type linear constraints \cite{FrFG16,FrGo13b,FGGP11}, semi-continuous variables \cite{FrFG16,FrGaSc14,FrGaSt16a,FrGe06a,FrGe07a,FrGe09a,FGGP11}, or the fact that a nonconvex quadratic objective function/constraint can be reformulated as a second-order cone (hence, convex) one \cite{FrFG16,FrGaSc14,FrGaSt16a,FrGe09a,FGGP11}.
 It would be rather hard to collect a comprehensive list of all types of structures that may be of interest to any individual researcher, since these are as varied as the different possible approaches for specialized sub-classes of QP.

### 2017-01-27 QPLIB-3.tex
Add files via upload

@@
 %
-The number of negative eigenvalues of $Q^0$ allows us to identify the
+
+The number of negative (i.e., smaller than $-10^{-16}$) eigenvalues of $Q^0$ allows us to identify the
 objective function type, as in presence of at least one negative eigenvalue
@@
 the quadratic term. Among the quadratic constraints, the ones with only
-non-negative eigenvalues are classified as convex constraints. All this
-information allowed us to analyse the gathered instances and to perform the
+non-negative eigenvalues are classified as convex constraints.
+Note that this might lead to classify some instances that have conic constraints as non-convex ones, although their feasible region is in fact convex.
+However, only some solvers are capable of properly exploiting this property.
+ All this information allowed us to analyse the gathered instances and to perform the
 filters described in the the next paragraph.
@@
        of the instance under consideration, cf.~Table \ref{t:solvers}.
-       We then filtered according to a relative
+       We then filtered according to a first
        measure of computational difficulty, i.e., we discarded all instances
@@
        according to their features; for each cluster we kept only a few
-       representatives\todo{we may want to be a tad more specific
-       about what ``features'' exactly meant there}. Finally, in order to only
-       keep computationally challenging instances we ran a general purpose solver
-       able to tackle all QP categories\todo{do you mean a complete solver for QGQ? if so, maybe say so}\xspace with a time limit of 120 seconds; all the
+       representatives (e.g., very similar size, same donor,\ldots). Finally, in order to only
+       keep computationally challenging instances we ran a a complete solver for QGQ with a time limit of 120 seconds; all the
        instances which have been solved to proven optimality within this time limit
-       were discarded\todo{this might actually be best characterized as
-       a third filter}.
+       were discarded.
 \end{itemize}
@@
 that the final set well respects the original distribution of the gathered
-instances\todo{Maybe one could put the numbers for the original set of instances
-into Tables \ref{tab:DD} and \ref{tab:CC}, too?}\xspace among the different categories. Indeed, the discrete categories
+instances among the different categories. Indeed, the discrete categories
 (LMQ) or (QBL) are well represented by $118$ and $59$ instances,
@@
 with a ``$\times$''.
+It can be seen that the majority of instances only have convex constraints; of those that have nonconvex ones, however, a significant fraction have no convex ones at all.
 
@@
   \includegraphics[width=0.85\textwidth]{pic_quad_conv_vs_nonconv.png}
-  \todo[inline]{horizontal axis label could say ``Instances with at least one quadratic constraint''}
-  \caption{Quadratic constraints \label{fig:pic_conv_constr}}
+  \caption{Quadratic constraints (instances with at least one) \label{fig:pic_conv_constr}}
 \end{figure}
@@
 the instances with a quadratic objective function, ordered by the
-the percentage of negative eigenvalues\todo{Was there some tolerance applied on when an eigenvalue is considered negative?
-$-10^{-9}$, $-10^{-6}$, etc.?} in the $Q^0$ matrix (vertical axis).
-The instances are mostly
-clustered around two values. About 25\% of the instances are convex\todo{so one also restrict to linear constraints, or may only the objective be convex?},
+the percentage of negative eigenvalues in the $Q^0$ matrix (vertical axis). The instances are mostly
+clustered around two values. About 25\% of the instances have convex objective function,
 i.e., they have 0\% of negative eigenvalues. Among the others, a vast
@@
   \includegraphics[width=0.85\textwidth]{pic_neg_eig.png}
-  \todo[inline]{horizontal axis label should say ``Instances with quadratic objective function''}
-  \caption{Negative eigenvalues \label{fig:pic_neg_eig}}
+  \caption{Negative eigenvalues (instances with quadratic objective function) \label{fig:pic_neg_eig}}
 \end{figure}
@@
   \includegraphics[width=0.85\textwidth]{pic_density.png}
-  \todo[inline]{horizontal axis label should say ``Instances with quadratic objective function''}
-  \caption{Density \label{fig:pic_density}}
+  \caption{Density of $Q^0$ (instances with quadratic objective function) \label{fig:pic_density}}
 \end{figure}
@@
 Additional details on the instance features can be found in \S~\ref{sec:instance_details}.
+
+%The all set of instances can be downloaded at \url{http://qplib.zib.de/html/index.html}.
+%Since, as described in Section \ref{subsec:final set} the library contains instances of different nature, the website allows to filter the instances according to the classification proposed in Section \ref{ssec:taxonomy}.
+
+\subsection{Website}
+%Select subset or categories of instances (EXTRACT FROM THE LIBRARY A SUBSET OF INSTANCES WITH SPECIFIC CHARACTERISTICS)
+%\todo[inline]{TASK X : write }
+
+The instances of QPLIB are publically accessible at the website \url{http://qplib.zib.de}.
+%
+Beyond links to the \texttt{gms} and \texttt{lp} format files, the website provides a rich set of metadata for each
+instance: the three letter problem classification (as described in Section \ref{subsec:final set}), basic properties such as the number of variables and constraints of
+different types, the sense and curvature of the objective function, and information on the nonzero structure of the
+problem.
+%
+In addition, we display a visualization of the sparsity patterns of the Jacobian and the Hessian matrix of the
+Lagrangian function.  In the plots of the Jacobian nonzero pattern, the linear and nonlinear entries are distinguished
+by color.  Figure... shows an example for instance QPLIB...
+
+The entire set of instances can be explored in a searchable and sortable table of selected instance features: problem
+classification, convexity of the continuous relaxation, and number of variables, constraints, and nonzeros.
+%
+Finally, a statistics page displays diagrams on the composition of the library according to different criteria: the
+number instances according to problem type, variable types, convexity, number of variables and constraints, and density.
+%
+A table containing the relevant metadata for each instance can be downloaded in \texttt{csv} format and as spreadsheets
+such that researchers can easily compile their own subset of instances according to these statistics.
+
+The complete library can be downloaded as one archive, which contains the website for offline browsing and exploration.  
+%
+In the future, we plan to extend the website by the addition of contributor information and references to the literature.
+
 %- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

### 2017-01-27 format.tex
put back todos that haven't been discussed yet

@@
 %\end{align*}
-%\todo[inline]{\small Stefan: Are LPQC and QPQC frequently used abbreviations?
-If a linear program has quadratic constraints, how can it then still be a linear program?
-I am more familiar with the abbreviations QCP instead of LPQC and QCQP instead of QPQC.
-I further think that one could skip QCP and consider them as part of QCQP.}
 %where $Q\s{C}$ is the diagonal matrix with blocks $Q\s{C}_k$, $k=1,\ldots,m$,
@@
 %
-%\todo[inline]{\small Stefan: I would suggest to make it simpler, even though that
-%would reduce flexibility. Always fix the default value for coefficients to 0
-%(otherwise, how you parse this? First initialize a n x n matrix with the non-zero
-%default and then overwrite with the given entries, which may then be zero?)
-%and have no default bounds for variables. If you then also adapt the suggestion
-%to use a constraint sense ($\leq$,$\geq$,$=$) and a right-hand-side instead of
-%a left- and right-hand-side, then there is also no need to specify a value for infinity.
-%This would leave out ranged constraints - but are they really used? One could
-%do it like CPLEX and have an extra constraint sense for ranges and an extra
-%section to specify the range.}
-%
+\todo[inline]{\small Stefan: I would suggest to make it simpler, even though that
+would reduce flexibility. Always fix the default value for coefficients to 0
+(otherwise, how you parse this? First initialize a n x n matrix with the non-zero
+default and then overwrite with the given entries, which may then be zero?)
+and have no default bounds for variables. If you then also adapt the suggestion
+to use a constraint sense ($\leq$,$\geq$,$=$) and a right-hand-side instead of
+a left- and right-hand-side, then there is also no need to specify a value for infinity.
+This would leave out ranged constraints - but are they really used? One could
+do it like CPLEX and have an extra constraint sense for ranges and an extra
+section to specify the range.}
 
@@
 \end{description}
-%\todo[inline]{\small Stefan: I agree that leaving out information on
-%numerical properties of the matrices and vectors (convexity, etc.) is good here.
-%Such properties may also depend on tolerances, etc.
-%As above, I would suggest to eliminate LPQC and rename QPQC into QCP.
-%Having a string for ``unspecified'' may be nice, too, or one just uses MIQCQP in this case.}
 \end{description}
+\todo[inline]{\small Stefan: Are LPQC and QPQC frequently used abbreviations?
+If a linear program has quadratic constraints, how can it then still be a linear program?
+I am more familiar with the abbreviations QCP instead of LPQC and QCQP instead of QPQC.
+I further think that one could skip QCP and consider them as part of QCQP.}
 

### 2017-01-27 QPLIB-1.0.tex
solvers have merged into solution methods subsection, no need for extra citation

@@
 In \S~\ref{sec:QPbasic} we review the basic notions about QP.
-In particular, \S~\ref{subsec:notation} sets out the notation, \S~\ref{sec:classification} proposes a---to the best of our knowledge, new---taxonomy of QP that helps us in discussing the (very) different classes of QPs, \S~\ref{sec:algo} very briefly reviews the solution methods for QP upon which the solvers we have employed are based, and \S~\ref{subsec:solver} describes the solvers.
+In particular, \S~\ref{subsec:notation} sets out the notation, \S~\ref{sec:classification} proposes a---to the best of our knowledge, new---taxonomy of QP that helps us in discussing the (very) different classes of QPs, \S~\ref{sec:algo} very briefly reviews the solution methods for QP and the solvers we have employed.
 Then, \S~\ref{sec:lib} describes the process used to obtain the library and its results.

### 2017-01-27 format.tex
moved todos into #2, #3, #4

@@
 %
-\todo[inline]{\small Stefan: I would suggest to make it simpler, even though that
-would reduce flexibility. Always fix the default value for coefficients to 0
-(otherwise, how you parse this? First initialize a n x n matrix with the non-zero
-default and then overwrite with the given entries, which may then be zero?)
-and have no default bounds for variables. If you then also adapt the suggestion
-to use a constraint sense ($\leq$,$\geq$,$=$) and a right-hand-side instead of
-a left- and right-hand-side, then there is also no need to specify a value for infinity.
-This would leave out ranged constraints - but are they really used? One could
-do it like CPLEX and have an extra constraint sense for ranges and an extra
-section to specify the range.}
 
@@
 \end{description}
-\todo[inline]{\small Stefan: Are LPQC and QPQC frequently used abbreviations?
-If a linear program has quadratic constraints, how can it then still be a linear program?
-I am more familiar with the abbreviations QCP instead of LPQC and QCQP instead of QPQC.
-I further think that one could skip QCP and consider them as part of QCQP.}
 

### 2017-01-28 QPLIB-3.tex
update pictures
- use most current numbers
- corrected count on nonconvex constraints
- use positive eigenvalues in Q^0 if maximization

@@
 \begin{figure}\centering
-  \includegraphics[width=0.85\textwidth]{pic_overview.png}
-  \caption{Distribution of variables and constrains  of the qplib
-instances \label{fig:distribution}}
+  \includegraphics[width=0.55\textwidth]{pic_overview.pdf}
+  \caption{Distribution of number of variables and constraints of QPLIB instances
+instances. \label{fig:distribution}}
 \end{figure}
@@
 \begin{figure}\centering
-  \includegraphics[width=0.85\textwidth]{pic_var.png}
-  \caption{Number of variables \label{fig:pic_var}}
+  \includegraphics[width=0.55\textwidth]{pic_var.pdf}
+  \caption{Number of variables of QPLIB instances. \label{fig:pic_var}}
 \end{figure}
 
-%\begin{figure}\centering
-%  \includegraphics[width=0.85\textwidth]{pic_var_small.png}
-%  \caption{Number of variables (a) \label{fig:pic_var_small}}
-%\end{figure}
-%
-%\begin{figure}\centering
-%  \includegraphics[width=0.85\textwidth]{pic_var_medium.png}
-%  \caption{Number of variables (b) \label{fig:pic_var_medium}}
-%\end{figure}
-%
-%\begin{figure}\centering
-%  \includegraphics[width=0.85\textwidth]{pic_var_big.png}
-%  \caption{Number of variables (c) \label{fig:pic_var_large}}
-%\end{figure}
-
-Similarly, Figures~\ref{fig:pic_constr} 
+Similarly, Figures~\ref{fig:pic_constr} (left)
 %Figure~\ref{fig:pic_constr_medium}
@@
 prevalence of lower percentages of quadratic constraints. In particular,
-from Figure~\ref{fig:pic_constr} we learn that while the maximum number
+from Figure~\ref{fig:pic_constr} (left) we learn that while the maximum number
 of linear constraints exceeds 100000, the maximum number of quadratic
@@
 \begin{figure}\centering
-  \includegraphics[width=0.85\textwidth]{pic_constr.png}
-  \caption{Number of constraints \label{fig:pic_constr}}
+  \includegraphics[width=0.49\textwidth]{pic_constr.pdf}
+  \includegraphics[width=0.49\textwidth]{pic_quad_conv_vs_nonconv.pdf}
+  \caption{Number of constraints, quadratic constraints, and nonconvex quadratic constraints of QPLIB instances. \label{fig:pic_constr}}
 \end{figure}
 
-%\begin{figure}\centering
-%  \includegraphics[width=0.85\textwidth]{pic_constr_small.png}
-%  \caption{Number of constraints (a) \label{fig:pic_constr_small}}
-%\end{figure}
-%
-%\begin{figure}\centering
-%  \includegraphics[width=0.85\textwidth]{pic_constr_medium.png}
-%  \caption{Number of constraints (b) \label{fig:pic_constr_medium}}
-%\end{figure}
-%
-%\begin{figure}\centering
-%  \includegraphics[width=0.85\textwidth]{pic_constr_big.png}
-%  \caption{Number of constraints (c) \label{fig:pic_constr_big}}
-%\end{figure}
-
-Figure~\ref{fig:pic_conv_constr} shows the instances with at least one quadratic constraint sorted according to the number of quadratic constraints (vertical axis).
+Figure~\ref{fig:pic_constr} (right) shows the instances with at least one quadratic constraint sorted according to the number of quadratic constraints (vertical axis).
 For each instance we report the total number of constraints with a ``$+$''
@@
 with a ``$\times$''.
-It can be seen that the majority of instances only have convex constraints; of those that have nonconvex ones, however, a significant fraction have no convex ones at all.
+It can be seen that the majority of instances only have nonconvex constraints. %; of those that have nonconvex ones, however, a significant fraction have no convex ones at all.
+
+Speaking of nonconvexity, Figure~\ref{fig:pic_neg_eig} (left) focuses on
+the instances with a quadratic objective function, ordered by the
+the percentage ``hard'' eigenvalues in the $Q^0$ matrix (vertical axis), by which
+we mean negative eigenvalues in case of a minimization problem and positive eigenvalues in case of a maximization problem.
+The instances are mostly clustered around two values. About 25\% of the instances have a convex (if minimization) or concave (if maximization) objective function,
+i.e., they have 0\% of ``hard'' eigenvalues. Among the others, a vast
+majority has around 50\% of ``hard'' eigenvalues. However, instances
+with high or low percentages of ``hard'' eigenvalues are present, too.
 
 \begin{figure}\centering
-  \includegraphics[width=0.85\textwidth]{pic_quad_conv_vs_nonconv.png}
-  \caption{Quadratic constraints (instances with at least one) \label{fig:pic_conv_constr}}
+  \includegraphics[width=0.49\textwidth]{pic_neg_eig.pdf}
+  \includegraphics[width=0.49\textwidth]{pic_density.pdf}
+  \caption{``Hard'' eigenvalues (left) and density (right) of $Q^0$ matrix for QPLIB instances with a quadratic objective function. \label{fig:pic_neg_eig}}
 \end{figure}
 
-Speaking of nonconvexity, Figure~\ref{fig:pic_neg_eig} focuses on
-the instances with a quadratic objective function, ordered by the
-the percentage of negative eigenvalues in the $Q^0$ matrix (vertical axis). The instances are mostly
-clustered around two values. About 25\% of the instances have convex objective function,
-i.e., they have 0\% of negative eigenvalues. Among the others, a vast
-majority has around 50\% of negative eigenvalues. However, instances
-with high or low percentages of negative eigenvalues are present, too.
-
-\begin{figure}\centering
-  \includegraphics[width=0.85\textwidth]{pic_neg_eig.png}
-  \caption{Negative eigenvalues (instances with quadratic objective function) \label{fig:pic_neg_eig}}
-\end{figure}
-
-Similarly, Figure~\ref{fig:pic_density} shows the instances with a
+Similarly, Figure~\ref{fig:pic_neg_eig} (right) shows the instances with a
 quadratic objective function sorted according to the density of the
@@
 density larger than 50\%. However, also intermediate values are present.
-
-\begin{figure}\centering
-  \includegraphics[width=0.85\textwidth]{pic_density.png}
-  \caption{Density of $Q^0$ (instances with quadratic objective function) \label{fig:pic_density}}
-\end{figure}
 

### 2017-01-28 QPLIB-2.2.tex
update classification count tables
- reflect current set of instances
- regard objective sense for D and C category

@@
 \begin{itemize}
- \item objective function: (L)inear, (D)iagonal convex quadratic, (C)onvex quadra-tic, nonconvex (Q)uadratic;
+ \item objective function: (L)inear, (D)iagonal convex (if minimization) or concave (if maximization) quadratic, (C)onvex (if minimization) or (C)oncave (if maximization) quadratic,  (Q)uadratic (all other cases);
  %
- \item variables: (C)ontinuous only, (B)inary only, (M)ixed binary and continuous, (I)nteger only, (G)eneral (all other cases);
+ \item variables: (C)ontinuous only, (B)inary only, (M)ixed binary and continuous, (I)nteger (incl.\ binary) only, (G)eneral (all other cases);
  %

### 2017-01-28 QPLIB-3.tex
update classification count tables
- reflect current set of instances
- regard objective sense for D and C category

@@
 
-\begin{center}
 \begin{table}
@@
 %\hline
-\multirow{3}*{Convex}
+\multirow{1}*{Convex (if min)}
           & Binary  & Linear    &  2 \\[1.2 ex]
 \cmidrule(lr){2-4}
+\multirow{1}*{or}
           & \multirow{2}*{Mixed}
-                    & Linear    &   13\\[1.2 ex]
+                    & Linear    &   14\\[1.2 ex]
+\multirow{2}*{Concave (if max)}
           &         & Quadratic &    1\\[1.2 ex]
 %          & General & Linear    &    \\
+\cmidrule(lr){2-4}
+          & \multirow{1}*{Integer}
+                    & Linear    &   2\\[1.2 ex]
 %\hline
 \cmidrule(lr){1-4}
-\multirow{7}*{Quadratic}
+\multirow{6}*{Quadratic}
           & \multirow{3}*{Binary}
-                    & None      &   23\\[1.2 ex]
+                    & None      &   23\\[1.2 ex]  % there are actually QBB due to the {0,1} bounds on the binary variables
           &         & Linear    &  59\\[1.2 ex]
@@
           & \multirow{2}*{Mixed}
-                    & Linear    &   10\\[1.2 ex]
+                    & Linear    &   9\\[1.2 ex]
           &         & Quadratic &    1\\[1.2 ex]
 %          & Integer & Linear    &    \\
-\cmidrule(lr){2-4}
-          & \multirow{1}*{Integer}
-                    & Linear    &    2\\[1.2 ex]
 \cmidrule(lr){2-4}
@@
 \end{table}
-\end{center}
 
-\begin{center}
 \begin{table}
@@
 %
-\multirow{2}*{Linear}    & Convex    &   11\\[1.2 ex]
-                         & Quadratic &   50\\[1.2 ex]
+\multirow{2}*{Linear}    & Convex    &   15\\[1.2 ex]
+                         & Quadratic &   46\\[1.2 ex]
 \cmidrule(lr){1-3}
-\multirow{4}*{Convex}
+\multirow{1}*{Convex (if min)}
                          & Box       &   3 \\[1.2 ex]
-                         & Linear    &   12\\[1.2 ex]
-                         & Convex    &    2\\[1.2 ex]
-                         & Quadratic &    5\\[1.2 ex]
+\multirow{1}*{or}
+                         & Linear    &   11\\[1.2 ex]
+                        % & Convex    &    2\\[1.2 ex]
+\multirow{1}*{Concave (if max)}
+                         & Quadratic &    9\\[1.2 ex]
 \cmidrule(lr){1-3}
 \multirow{3}*{Quadratic}
-                         & Linear    &   5\\[1.2 ex]
-                         & Convex    &   11\\[1.2 ex]
-                         & Quadratic &   17\\[1.2 ex]
+                         & Linear    &   6\\[1.2 ex]
+                         & Convex    &   1\\[1.2 ex]
+                         & Quadratic &   25\\[1.2 ex]
 \hline
@@
 \end{table}
-\end{center}
 

### 2017-01-28 QPLIB-3.tex
revise/update section on website

@@
 %Select subset or categories of instances (EXTRACT FROM THE LIBRARY A SUBSET OF INSTANCES WITH SPECIFIC CHARACTERISTICS)
-%\todo[inline]{TASK X : write }
 
-The instances of QPLIB are publically accessible at the website \url{http://qplib.zib.de}.
+The instances of QPLIB are publicly accessible at the website \url{http://qplib.zib.de}.
 %
-Beyond links to the \texttt{gms} and \texttt{lp} format files, the website provides a rich set of metadata for each
+Beyond all instances in GAMS and \texttt{.lp} format, the website provides a rich set of metadata for each
 instance: the three letter problem classification (as described in Section \ref{subsec:final set}), basic properties such as the number of variables and constraints of
-different types, the sense and curvature of the objective function, and information on the nonzero structure of the
+different types, the sense and convexity/concavity of the objective function, and information on the nonzero structure of the
 problem.
@@
 Lagrangian function.  In the plots of the Jacobian nonzero pattern, the linear and nonlinear entries are distinguished
-by color.  Figure... shows an example for instance QPLIB...
+by color.  Figure... shows an example for instance QPLIB... \todo{add figure}
 
 The entire set of instances can be explored in a searchable and sortable table of selected instance features: problem
-classification, convexity of the continuous relaxation, and number of variables, constraints, and nonzeros.
+classification, convexity of the continuous relaxation, number of (all, binary, integer) variables, (all, quadratic) constraints, nonzeros, hard eigenvalues in $Q^0$, and density of $Q^0$.
 %
 Finally, a statistics page displays diagrams on the composition of the library according to different criteria: the
-number instances according to problem type, variable types, convexity, number of variables and constraints, and density.
+number instances according to problem type, variable and constraint types, convexity, number of variables and constraints, and density.
 %
-A table containing the relevant metadata for each instance can be downloaded in \texttt{csv} format and as spreadsheets
-such that researchers can easily compile their own subset of instances according to these statistics.
+A file containing the relevant metadata for each instance can be downloaded in comma-separated-values (\texttt{csv}) format,
+so that researchers can easily compile their own subset of instances according to these statistics.
 

### 2017-01-29 QPLIB-2.1.tex
Add files via upload
emphasize early that optimization sense can be maximization

@@
 \begin{align*}
- \min \;\;
- & x^\top Q^0 x + b^0 x \\
- %
- & x^\top Q^i x + b^i x \leq c^i & i \in \Mcm \\ 
+ \min \textrm{or} \max \;\;
+ & x^\top Q^0 x + b^0 x + q^0 \\
+ \textrm{such that}\;\;
+ & c^i_l \leq x^\top Q^i x + b^i x \leq c^i_u & i \in \Mcm \\
  %
@@
 \end{align*}
+
+%
+%\begin{align*}	
+% \min \;\;
+% & x^\top Q^0 x + b^0 x \\
+% %
+% & x^\top Q^i x + b^i x \leq c^i & i \in \Mcm \\ 
+% %
+% & l_j \leq x_j \leq u_j & j \in \Mcn  \\
+% %
+% & x_j \in \mathbb{Z} & j \in \Mcz
+%\end{align*}
 %
@@
  %
- \item $b^i$ and $c^i$ for $i \in \{ 0 \} \cup \Mcm$ are, respectively, real $n$-vectors and real constants;
+ \item $b^i$, $c_u^i$, $c_l^i$ for $i \in \{ 0 \} \cup \Mcm$, and $q^0$ are, respectively, real $n$-vectors and real constants;
  %
@@
 %
-Due to the presence of integral requirements on the variables, this class of problems is often referred to as Mixed-Integer Quadratic Program (MIQP).
+Due to the presence of integral requirements on the variables and quadratic constraints, this class of problems is often referred to as Mixed-Integer Quadratically Constraint Quadratic Program (MIQCQP).
 It will be sometimes useful to refer to the (sub)set $\Mcb =  \{ \, i \in \Mcz : l_j = 0 , u_j = 1 \, \} \subseteq \Mcz$ of the binary variables, and to $\Mcr = \Mcn \setminus \Mcz$ as the set of continuous ones.
@@
 The relative flexibility offered by quadratic functions, as opposed e.g.~to linear ones, allows several reformulation techniques to be applicable to this family of problems in order to emphasize different properties of the various components.
-Some of these reformulation techniques will be commented later on; here we remark that, for instance, integrality requirements, in particular in the form of binary variables could always be ``hidden'' by introducing (non convex) quadratic constraints utilizing the celebrated relationship $x_j \in \{0, 1\} \iff x_j^2 = x_j$.
+Some of these reformulation techniques will be commented later on; here we remark that, for instance, integrality requirements, in particular in the form of binary variables could always be ``hidden'' by introducing (nonconvex) quadratic constraints utilizing the celebrated relationship $x_j \in \{0, 1\} \iff x_j^2 = x_j$.
 Therefore, when discussing these problems an effort has to be made to distinguish between features that come from the original model, and those that can be introduced by reformulation techniques in order to extract (and algorithmically exploit) specific properties.

### 2017-01-29 QPLIB-3.tex
add picture about nonzeros
- provides a bit more insight next to number of vars/cons

@@
 We now report some graphs that help in illustrating the main features
-of the instances. In Figure~\ref{fig:distribution} we plot the number
+of the instances. In Figure~\ref{fig:distribution}~(left) we plot the number
 of variables (horizontal axis) versus the number of constrains
@@
 variables is almost 40000.
+Figure~\ref{fig:distribution}~(right) plots the number of nonzero elements in the gradient of the objective function and the Jacobian and the number of these nonzeros corresponding to nonlinear variables, that is, it counts the appearances of variables in objectives and constraints and how often such an appearance is in a quadratic term.
 
@@
 \begin{figure}\centering
-  \includegraphics[width=0.55\textwidth]{pic_overview.pdf}
+  \includegraphics[width=0.50\textwidth]{pic_overview.pdf}
+  \includegraphics[width=0.49\textwidth]{pic_nz.pdf}
   \caption{Distribution of number of variables and constraints of QPLIB instances
-instances. \label{fig:distribution}}
+(left). Number of (nonlinear) nonzeros of QPLIB instances (right).\label{fig:distribution}}
 \end{figure}

### 2017-01-29 QPLIB-3.tex
add sparsity pattern for one instance

@@
 The instances of QPLIB are publicly accessible at the website \url{http://qplib.zib.de}.
+Scripts and tools developed for MINLPLib 2 \cite{Vigerske2014} were used and extended to create the website.
 %
@@
 Lagrangian function.  In the plots of the Jacobian nonzero pattern, the linear and nonlinear entries are distinguished
-by color.  Figure... shows an example for instance QPLIB... \todo{add figure}
+by color.  Figure \ref{fig:sparsitypattern} shows an example for instance QPLIB\_2967.
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.5]{{QPLIB_2967.jac}.png} \qquad
+\includegraphics{{QPLIB_2967.hess}.png}
+\caption{Sparsity pattern of Gradient of objective (first line) and Jacobian (left) and of Hessian of Lagrangian function (right) for instance QPLIB\_2967.}
+\label{fig:sparsitypattern}
+\end{figure}
 

### 2017-01-29 QPLIB-3.tex
make clear that variable bounds were not counted for binary variables, more about neg->hard ev.

@@
 \begin{itemize}
- \item percentage of negative eigenvalues of the $Q^0$ matrix;
+ \item percentage of positive and negative eigenvalues of the $Q^0$ matrix;
        % (\emph{\% neg eig}),
@@
 
-The number of negative (i.e., smaller than $-10^{-16}$) eigenvalues of $Q^0$ allows us to identify the
-objective function type, as in presence of at least one negative eigenvalue
-the objective function is non convex. Finally, as far as the constraint types
+The number of positive (i.e., larger than $10^{-16}$) and negative (i.e., smaller than $-10^{-16}$) eigenvalues of $Q^0$ allows us to identify the
+objective function type, as in presence of at least one negative (positive) eigenvalue
+the objective function is non convex (non concave). Finally, as far as the constraint types
 are concerned, we collected the following information:
@@
  %
- \item number of box constraints. %(\emph{\# box}).
+ \item number of variable bounds (for non-binary variables). %(\emph{\# box}).
 \end{itemize}
@@
           & \multirow{3}*{Binary}
-                    & None      &   23\\[1.2 ex]  % there are actually QBB due to the {0,1} bounds on the binary variables
+                    & None      &   23\\[1.2 ex] on the binary variables
           &         & Linear    &  59\\[1.2 ex]

### 2017-01-29 QPLIB-3.tex
non convex -> nonconvex

@@
 objective function type, as in presence of at least one negative (positive) eigenvalue
-the objective function is non convex (non concave). Finally, as far as the constraint types
+the objective function is nonconvex (nonconcave). Finally, as far as the constraint types
 are concerned, we collected the following information:

### 2017-01-29 QPLIB-1.0.tex
added Hans paragraph

@@
 
+While no performance issues of solvers for QP/QCP problems are considered in 
+this paper, we refer to the comprehensive benchmark site \url{http://plato.asu.edu/bench.html}. Of the benchmarks
+on this site three deal exlusively with QP/QCP problems: 
+the (1) Large SOCP, (2) MISOCP, and the (3) MIQ(C)P benchmarks. 
+The following contain partly results for such problems: 
+(4) Parallel Barrier Solvers on Large LP/QP problems, (5) AMPL-NLP, (6) MINLP.
+The benchmarks (1,2,4) have only convex instances, while the others include 
+nonconvex instances. Global optima are obtained by several of the solvers in
+the benchmarks (3,5).
+All solvers in the latest addition (6) compute global optima, they are ANTIGONE,
+BARON, COUENNE, LINDO, and SCIP. This benchmark is based on MINLPLIB2 \url{http://www.gamsworld.org/minlp/minlplib2/html/}, a
+collection of presently 1388 instances. In order to give a first representative
+impression of solver performance, care was taken to reduce the number and 
+allow all solvers to finish in reasonable time. More than half of the selected
+instances are QP or QCP. A time limit of two hours on a
+modern workstation with sufficent memory was used. Computations were done in
+GAMS with scripts derived from the MIPLIB scripts and available at Github.
+For details we refer to \url{http://plato.asu.edu/ftp/minlp.html}.
+
 %- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

### 2017-01-30 QPLIB-1.0.tex
minor revision on hans paragraph
- add references to solvers subsection and minlplib 2 citation
- the scripts were taken from scip, not miplib

@@
 All solvers in the latest addition (6) compute global optima, they are ANTIGONE,
-BARON, COUENNE, LINDO, and SCIP. This benchmark is based on MINLPLIB2 \url{http://www.gamsworld.org/minlp/minlplib2/html/}, a
+BARON, COUENNE, LINDO, and SCIP, c.f.~\S~\ref{subsec:solver}. This benchmark is based on MINLPLib~2 \cite{Vigerske2014}, a
 collection of presently 1388 instances. In order to give a first representative
-impression of solver performance, care was taken to reduce the number and 
+impression of solver performance, care was taken to reduce the number of instances and 
 allow all solvers to finish in reasonable time. More than half of the selected
 instances are QP or QCP. A time limit of two hours on a
-modern workstation with sufficent memory was used. Computations were done in
-GAMS with scripts derived from the MIPLIB scripts and available at Github.
+modern workstation with sufficient memory was used. Computations were done in
+GAMS with scripts derived from the SCIP check scripts and available at GitHub.
 For details we refer to \url{http://plato.asu.edu/ftp/minlp.html}.

### 2017-01-30 QPLIB-1.0.tex
fix some latex warnings

@@
 
-\begin{center}
-\begin{longtable}{ lcl }
+\begin{longtable}[c]{lcl}
 %
@@
 \midrule 
-\multicolumn{3}{l}{\multirow{1}{\textwidth}{$^{\ddagger}$Applications with many manuscripts cite reviews and recent works \\[-12pt] \phantom{hi} \hfill \emph{continued}}} \\ \\
+\multicolumn{3}{l}{\multirow{1}{.95\textwidth}{$^{\ddagger}$Applications with many manuscripts cite reviews and recent works \\[-12pt] \phantom{hi}  \hfill\emph{continued}}}
 \endfoot
@@
 \bottomrule 
-\multicolumn{3}{l}{\multirow{1}{\textwidth}{$^{\ddagger}$Applications with many manuscripts cite reviews and recent works.}} \\ \\ \\
+\multicolumn{3}{l}{\multirow{1}{.95\textwidth}{$^{\ddagger}$Applications with many manuscripts cite reviews and recent works.}}
 \endlastfoot
@@
 %
-\multirow{1}{.55\textwidth}{Quadratic Assignment Problem (QAP)$^{\ddagger}$} & \checkmark & \cite{anstreicher:2003,loiola-etal:2007} \\ \cmidrule(rl){2-3}
+Quadratic Assignment Problem$^{\ddagger}$ & \checkmark & \cite{anstreicher:2003,loiola-etal:2007} \\ \cmidrule(rl){2-3}
 %
@@
 Zeolites && \cite{gounaris-etal:2013} \\
-%
 \midrule
@@
 %
-\multirow{1}{.55\textwidth}{Layout design} & \checkmark & \cite{anjos-liers:2012,castillo-etal:2005,dorneich-sahinidis:1995} \\ \cmidrule(rl){2-3}
+Layout design & \checkmark & \cite{anjos-liers:2012,castillo-etal:2005,dorneich-sahinidis:1995} \\ \cmidrule(rl){2-3}
 %
-\multirow{1}{.55\textwidth}{Maximizing polygon dimensions} && \cite{audet-etal:2011,audet-etal:2007,Audet-etal:2009,audet-etal:2002,Audet-Ninin:2013} \\ \cmidrule(rl){2-3}
+Maximizing polygon dimensions && \cite{audet-etal:2011,audet-etal:2007,Audet-etal:2009,audet-etal:2002,Audet-Ninin:2013} \\ \cmidrule(rl){2-3}
 %
@@
 %
-\textbf{Finance} \\
-Portfolio optimization & \checkmark & \multirow{1}{.25\textwidth}{\cite{deng-etal:2013,kallrath:2003,lin-etal:2005,maranas-etal:1997,parpas-rustem:2006,rios-sahinidis:2010,FrFG16,FrGe06a,FrGe07a,FrGe09a}} \\ \\
+\multicolumn{3}{l}{\textbf{Finance}} \\[1pt]
+Portfolio optimization & \checkmark & \multirow{1}{.3\textwidth}{\cite{deng-etal:2013,kallrath:2003,lin-etal:2005,maranas-etal:1997,parpas-rustem:2006,rios-sahinidis:2010,FrFG16,FrGe06a,FrGe07a,FrGe09a}} \\ \\
 %
@@
 %
-\multirow{1}{.55\textwidth}{Crude oil scheduling} & \checkmark & \cite{li-etal:2007,li-etal:2011,li-etal:2012,mouret-etal:2009,mouret-etal:2011} \\ \cmidrule(rl){2-3}
+Crude oil scheduling & \checkmark & \cite{li-etal:2007,li-etal:2011,li-etal:2012,mouret-etal:2009,mouret-etal:2011} \\ \cmidrule(rl){2-3}
 %
-\multirow{1}{.55\textwidth}{Data reconciliation} & \checkmark & \cite{ruiz-grossmann:2011} \\ \cmidrule(rl){2-3}
+Data reconciliation & \checkmark & \cite{ruiz-grossmann:2011} \\ \cmidrule(rl){2-3}
 %
-\multirow{1}{.55\textwidth}{Multi-commodity flow} & \checkmark & \cite{tadayon-smith:2013} \\ \cmidrule(rl){2-3}
+Multi-commodity flow & \checkmark & \cite{tadayon-smith:2013} \\ \cmidrule(rl){2-3}
 %
-\multirow{1}{.55\textwidth}{Quadratic network design} & \checkmark & \cite{FrFG16,FGGP11} \\ \cmidrule(rl){2-3}
+Quadratic network design & \checkmark & \cite{FrFG16,FGGP11} \\ \cmidrule(rl){2-3}
 %
-\multirow{1}{.55\textwidth}{Multi-period blending} & \checkmark & \cite{kolodziej-etal:2013:jogo,kolodziej-etal:2013} \\ \cmidrule(rl){2-3}
+Multi-period blending & \checkmark & \cite{kolodziej-etal:2013:jogo,kolodziej-etal:2013} \\ \cmidrule(rl){2-3}
 %
-\multirow{1}{.55\textwidth}{Natural gas networks} & \checkmark & \cite{hasan-etal:2011,li-etal:2011-aiche_journal,li-etal:2011-jogo} \\ \cmidrule(rl){2-3}
+Natural gas networks & \checkmark & \cite{hasan-etal:2011,li-etal:2011-aiche_journal,li-etal:2011-jogo} \\ \cmidrule(rl){2-3}
 %
-Pooling$^{\ddagger}$ & \checkmark & \multirow{1}{.25\textwidth}{\cite{Alfaki-Haugland:2013,Castillo-etal:2013,dambrosio-etal:2011pooling,Faria-Bagajewicz:2012,misener-floudas:2009,misener-floudas:2010-genpooling,Papageorgiou-etal:2012,pham-etal:2009,ruiz-etal:2013}} \\ \\ \cmidrule(rl){2-3}
+Pooling$^{\ddagger}$ & \checkmark & \multirow{1}{.3\textwidth}{\cite{Alfaki-Haugland:2013,Castillo-etal:2013,dambrosio-etal:2011pooling,Faria-Bagajewicz:2012,misener-floudas:2009,misener-floudas:2010-genpooling,Papageorgiou-etal:2012,pham-etal:2009,ruiz-etal:2013}} \\ \\ \cmidrule(rl){2-3}
 %
-\multirow{1}{.55\textwidth}{Open-pit mine scheduling} & \checkmark & \cite{bley-etal:2012} \\ \cmidrule(rl){2-3}
+Open-pit mine scheduling & \checkmark & \cite{bley-etal:2012} \\ \cmidrule(rl){2-3}
 %
@@
 %
-Water networks$^{\ddagger}$ & \checkmark & \multirow{1}{.25\textwidth}{\cite{ahmetovic-grossmann:2010,bagajewicz:2000,bragalli-etal:2011,castro-teles:2013,geissler-etal:2013,gleixner-etal:2012,jezowski-2010,khor-etal:2014,ponceortega-etal:2010,teles-etal:2012}} \\ \\
+Water networks$^{\ddagger}$ & \checkmark & \multirow{1}{.3\textwidth}{\cite{ahmetovic-grossmann:2010,bagajewicz:2000,bragalli-etal:2011,castro-teles:2013,geissler-etal:2013,gleixner-etal:2012,jezowski-2010,khor-etal:2014,ponceortega-etal:2010,teles-etal:2012}} \\ \\
 %
@@
 %
-\multirow{1}{.55\textwidth}{Traveling Salesman Problem with Neighborhoods} & \checkmark & \cite{gentilini-etal:2013} \\ \\
+Traveling Salesman Problem \\
+with Neighborhoods & \checkmark & \cite{gentilini-etal:2013} \\
+\midrule
 %
@@
 %
-\multirow{1}{.55\textwidth}{Delay-constrained routing} & \checkmark & \cite{FrGaSc14,FrGaSt16a} \\
-%
+Delay-constrained routing & \checkmark & \cite{FrGaSc14,FrGaSt16a} \\
+\midrule
 %
@@
 %
-\multirow{1}{.55\textwidth}{Unit-commitment} & \checkmark & \cite{FrFG16,FrGe06a,FrGe09a,Tetal15} \\
+Unit-commitment & \checkmark & \cite{FrFG16,FrGe06a,FrGe09a,Tetal15} \\
+\midrule
 %
@@
 %
-\multirow{1}{.55\textwidth}{Controlled Tabular Adjustment} & \checkmark & \cite{CaFG14} \\
+Controlled Tabular Adjustment & \checkmark & \cite{CaFG14}
 \end{longtable}
-\end{center}
 

### 2017-01-30 QPLIB-2.2.tex
emphasize early that optimization sense can be maximization
fixes to bibliography
- remove duplicates, update references, etc

@@
 We will now provide a first general discussion about the different problem classes that our proposed taxonomy defines.
-Some of them are actually ``too simple'' to make sense in our context.
+For simplicity, we will assume minimization problems for the remaining of this section.
+Some problem classes are actually ``too simple'' to make sense in our context.
 For instance, D*B problems have only diagonal quadratic (hence separable) objective function and bound constraints; as such, they read
@@
  \item \emph{Nonconvex Continuous Quadratic Programs} QCQ are instead in general $\mathcal{NP}$-hard, even if the constraints are very specific (QCB) and only one eigenvalue of $Q^0$ is negative \cite{Hemmecke2010}.
-They therefore require enumerative techniques like the spatial Branch\&Bound \cite{FV90,BLLMW09}, to be solved to optimality.
+They therefore require enumerative techniques like the spatial Branch\&Bound \cite{FV90,BeLeLiMaWa08}, to be solved to optimality.
 Of course, local approaches are available that are able to efficiently provide saddle points (hopefully, local optima) of the CQC, but providing global guarantees about the quality of the obtained solutions is challenging.

### 2017-01-30 format.tex
fixup formatting of appendix tables

@@
 
-\vspace*{2mm}
-
-%\vspace*{-10mm}
-%{\tiny
-%\LTcapwidth=16.17cm
-\begin{longtable}{|llr|}
+\begin{longtable}{|lp{.7\textwidth}r|}
 \caption{\label{tab-qplib-format}{The QPLIB file format: refer to the notes
@@
 \endlastfoot
-%\begin{tabular}{|llr|}
-%\hline
 name & problem name (character string) & \\
@@
 $n^{Q^0}$ & number of nonzeros  (integer) in lower triangle of $Q^0$  & [3] \\
-$h$\; $k$\; $Q_{hk}^0$ & \lbox{row and column indices (integers) and value (real)
-for each nonzero entry of $Q^0$, if $n^{Q^0} > 0$, one triple on each line} & \\
+$h$\; $k$\; $Q_{hk}^0$ & row and column indices (integers) and value (real)
+for each nonzero entry of $Q^0$, if $n^{Q^0} > 0$, one triple on each line & \\
 \hline
@@
 $n^{b^0}$ & number of non-default entries (integer) in $b^0$ & \\
-$j$\; $b^0_j$ & \lbox{index (integer) and value (real) for each non-default
-term in $b^0$, if $n^{b^0} > 0$, one pair per line} & \\
+$j$\; $b^0_j$ & index (integer) and value (real) for each non-default
+term in $b^0$, if $n^{b^0} > 0$, one pair per line & \\
 \hline
@@
 $i$\; $h$\; $k$\; $Q^{i}_{hk}$
-& \lbox{constraint, row and column indices (integers) and value (real) for
- each entry of $Q^{i}$, if $n^{Q^i}$, one quadruple on each line} & \\
+& constraint, row and column indices (integers) and value (real) for
+ each entry of $Q^{i}$, if $n^{Q^i}$, one quadruple on each line & \\
 \hline
 $n^{b^i}$ & number of nonzeros  (integer) in $b^i$, $i \in \Mcm$                     & [2] \\
-$i$\; $j$\; $b^i_{j}$ & \lbox{row and column indices (integers) and value (real)
+$i$\; $j$\; $b^i_{j}$ & row and column indices (integers) and value (real)
 for each nonzero
-entry of $b^i$, if $n^{b} > 0$, one triple on each line} & \\
+entry of $b^i$, if $n^{b} > 0$, one triple on each line & \\
 \hline
-$c_{\infty}$ & \lbox{value (real) for infinity for constraint or variable
+$c_{\infty}$ & value (real) for infinity for constraint or variable
 bounds---any bound greater than or equal to this in, absolute value,
-is infinite} & \\
+is infinite & \\
 \hline
@@
 $n^{c_{l,d}}$ & number of non-default entries (integer) in $c_{l}$    & [2] \\
-$i$\;\ $c^i_{l}$ & \lbox{index (integer) and value (real) for each non-default
+$i$\;\ $c^i_{l}$ & index (integer) and value (real) for each non-default
 term
-in $c_{l,d}$, if $n^{c_{l,d}} > 0$, one pair per line}        & [2] \\
+in $c_{l,d}$, if $n^{c_{l,d}} > 0$, one pair per line        & [2] \\
 \hline
@@
 $n^{c_{u,d}}$ & number of non-default entries (integer) in $c_{u}$    & [2] \\
-$i$\;\ $c^i_{u}$ & \lbox{index (integer) and value (real) for each non-default
+$i$\;\ $c^i_{u}$ & index (integer) and value (real) for each non-default
 term
-in $c_{u,d}$, if $n^{c_{u,d}} > 0$, one pair per line}        & \\
+in $c_{u,d}$, if $n^{c_{u,d}} > 0$, one pair per line        & \\
 \hline
@@
 $n^{l_{d}}$ & number of non-default entries (integer) in $l$    &  \\
-$i$\;\ $l_{i}$ & \lbox{index (integer) and value (real) for each non-default
+$i$\;\ $l_{i}$ & index (integer) and value (real) for each non-default
 term
-in $l$, if $n^{l_{d}}> 0$, one pair per line}        &  \\
+in $l$, if $n^{l_{d}}> 0$, one pair per line        &  \\
 \hline
@@
 $n^{u_{d}}$ & number of non-default entries (integer) in $u$    &  \\
-$i$\;\ $u_{i}$ & \lbox{index (integer) and value (real) for each non-default
+$i$\;\ $u_{i}$ & index (integer) and value (real) for each non-default
 term
-in $u$, if $n^{u_{d}}> 0$, one pair per line}        &  \\
+in $u$, if $n^{u_{d}}> 0$, one pair per line        &  \\
 \hline
-$v_d$ & \lbox{default variable type  (integer, 0 for continuous variables,
- 1 for integer variables)} & [5] \\
+$v_d$ & default variable type  (integer, 0 for continuous variables,
+ 1 for integer variables) & [5] \\
 $n^v$ & number of non-default variables (integer)  & [5] \\
-$i$\; $v_i$ & \lbox{index  and type (integers) for each non-default
-variable type, if  $n^v > 0$, one pair per line} & [5] \\
+$i$\; $v_i$ & index  and type (integers) for each non-default
+variable type, if  $n^v > 0$, one pair per line & [5] \\
 \hline
-$x\s{0}_d$ & \lbox{default value (real) for the components of the starting
-point  $x^0$ for the  variables $x$} & \\
+$x\s{0}_d$ & default value (real) for the components of the starting
+point  $x^0$ for the  variables $x$ & \\
 $n^{x^0}$ & number of non-default starting entries (integer) in $x$ & \\
-$i$\; $x_i^0$ & \lbox{index (integer) and value (real) for each non-default
-starting value in $x^0$, if $n^{x^0} > 0$, one pair per line} &  \\
+$i$\; $x_i^0$ & index (integer) and value (real) for each non-default
+starting value in $x^0$, if $n^{x^0} > 0$, one pair per line &  \\
 \hline
-$y\s{0}_d$ & \lbox{default value (real) for the components of the starting
+$y\s{0}_d$ & default value (real) for the components of the starting
 point $y^0$ for the Lagrange multipliers $y$ for the general
-constraints} &  [2] \\
+constraints &  [2] \\
 $n^{y^0}$ & number of non-default starting entries  (integer) in $y$ &  [2] \\
-$i$\; $y_i^0$ & \lbox{index (integer) and value (real) for each non-default
-starting value in $y^0$, if $n^{y^0} > 0$, one pair per line} &  [2] \\
+$i$\; $y_i^0$ & index (integer) and value (real) for each non-default
+starting value in $y^0$, if $n^{y^0} > 0$, one pair per line &  [2] \\
 \hline
-$z\s{0}_d$ & \lbox{default value (real) for the components of the starting
-point $z^0$ for the dual variables $z$ for the simple bound constraints} & \\
+$z\s{0}_d$ & default value (real) for the components of the starting
+point $z^0$ for the dual variables $z$ for the simple bound constraints & \\
 $n^{z^0}$ & number of non-default starting entries (integer) in $z$ & \\
-$i$\; $z_i^0$ & \lbox{index (integer) and value (real) for each
-non-default starting value in $z^0$, if $n^{z^0} > 0$, one pair per line} &  \\
+$i$\; $z_i^0$ & index (integer) and value (real) for each
+non-default starting value in $z^0$, if $n^{z^0} > 0$, one pair per line &  \\
 \hline
-$n^x_d$ & \lbox{number of non-default names (integer) of variables---default
+$n^x_d$ & number of non-default names (integer) of variables---default
 for variable $i$ is the character string representing the numerical
-value $i$} &  \\
-$j$\; $\texttt{var\_name}_j$ & \lbox{index (integer) and name (character string)
-for each non-default variable name, if $n^x_d > 0$,  one pair per line} & \\
+value $i$ &  \\
+$j$\; $\texttt{var\_name}_j$ & index (integer) and name (character string)
+for each non-default variable name, if $n^x_d > 0$,  one pair per line & \\
 \hline
-$n^c_d$ & \lbox{number of non-default names (integer) of general
+$n^c_d$ & number of non-default names (integer) of general
 constraints---default for constraint $i$ is the character string representing
-the numerical value $i$} &  \\
-$i$\; $\texttt{cons\_name}_i$ & \lbox{index  (integer) and name (character string) for each
-non-default constraint name, if $n^c_d > 0$,  one pair per line} & \\
+the numerical value $i$ &  \\
+$i$\; $\texttt{cons\_name}_i$ & index  (integer) and name (character string) for each
+non-default constraint name, if $n^c_d > 0$,  one pair per line & \\
 \hline
 \end{longtable}
-%}
-%
-
-%\end{tabular}
 
 \renewcommand{\descriptionlabel}[1]{\hspace{\labelsep}\texttt{#1}}
-
 \begin{description}
@@
 5         # nonzeros in lower triangle of Q^0
-1 1 2.0   5 lines of row & column index & value of nonzeros in lower triangle Q^0
+1 1 2.0   5 lines row & column index & value of nz. in lower triangle Q^0
 2 1 -1.0  |
@@
 4         # nonzeros in vectors b^i (i =0,...,m)
-1 1 1.0   4 lines of row & column index & value of nonzeros in b^i (i =0,...,m)
+1 1 1.0   4 lines row & column index & value of nz. in b^i (i =0,...,m)
 1 2 1.0   |

### 2017-01-30 QPLIB-3.tex
fix some latex warnings
emphasis that variables/constraints order fixed in .qplib format

@@
 \item the \texttt{CUTEst} library
- \url{https://ccpforge.cse.rl.ac.uk/gf/project/cutest/wiki};
+ \url{https://ccpforge.cse.rl.ac.uk/gf/project/cutest};
 %
@@
 In comparison to a more \emph{high level} format like \texttt{.gms}, the new
-format presents two advantages: it is easier to read by a self-made parser and
-it produces smaller files. See the \S~\ref{sec:format} for more details.
+format presents three advantages: it is easier to read by a self-made parser, it produces smaller files, and the order of variables and constraints is fixed.
+See the \S~\ref{sec:format} for more details.
 
@@
 
-\begin{center}
-\begin{table}[]
+\begin{table}
  \centering
@@
 %
-Beyond all instances in GAMS and \texttt{.lp} format, the website provides a rich set of metadata for each
+Beyond all instances in GAMS, \texttt{.lp}, and \texttt{.qplib} format, the website provides a rich set of metadata for each
 instance: the three letter problem classification (as described in Section \ref{subsec:final set}), basic properties such as the number of variables and constraints of

### 2017-01-30 format.tex
rename problem types to something more familiar (issue #4)
I'll update the qplib instances on the up-to-be qplib website accordingly.

@@
 %\todo[inline]{\small Fabio and Emiliano: we think that, if possible, we should follow the classification proposed in Section 2.}
-
 The problem type is represented by a character string as one of the following:
@@
 \begin{description}[leftmargin=!,labelwidth=\widthof{\texttt{SMILQPC}}]
-\item [LP]      a linear program,
-\item [LPQC]    a linear program with quadratic constraints,
-\item [BQP]     a bound-constrained quadratic program,
-\item [QP]      a quadratic program, and
-\item [QPQC]    a quadratic program with quadratic constraints.
+\item [LP]   a problem with linear objective and linear constraints,
+\item [QCP]  a problem with linear objective and linear and quadratic constraints,
+\item [BQP]  a problem with quadratic objective and variable bounds,
+\item [QP]   a problem with quadratic objective and linear constraints, and
+\item [QCQP] a problem with quadratic objective and linear and quadratic constraints.
 \end{description}
+
 \item For integer problems (i.e., all variables are integer valued):
 \begin{description}[leftmargin=!,labelwidth=\widthof{\texttt{SMILQPC}}]
-\item [ILP]     an integer linear program,
-\item [ILPQC]   an integer linear program with quadratic constraints,
-\item [IBQP]    an integer bound-constrained quadratic program,
-\item [IQP]     an integer quadratic program, and
-\item [IQPQC]   an integer quadratic program with quadratic constraints.
+\item [ILP]   an LP plus integer requirement for all variables,
+\item [IQCP]  an QCP plus integer requirement for all variables,
+\item [IBQP]  an BQP plus integer requirement for all variables,
+\item [IQP]   an QP plus integer requirement for all variables, and
+\item [IQCQP] an QCQP plus integer requirement for all variables.
 \end{description}
+
 \item For mixed-integer problems (i.e., there is a mix of continuous and
@@
 \begin{description}[leftmargin=!,labelwidth=\widthof{\texttt{SMILQPC}}]
-\item [MILP]    a mixed-integer linear program,
-\item [MILPQC]  a mixed-integer linear program with quadratic constraints,
-\item [MIBQP]   a mixed-integer bound-constrained quadratic program,
-\item [MIQP]    a mixed-integer quadratic program, and
-\item [MIQPQC]  a mixed-integer quadratic program with quadratic constraints.
+\item [MILP]   an LP plus integer requirement for some variables,
+\item [MIQCP]  an QCP plus integer requirement for some variables,
+\item [MIBQP]  an BQP plus integer requirement for some variables,
+\item [MIQP]   an QP plus integer requirement for some variables, and
+\item [MIQCQP] an QCQP plus integer requirement for some variables.
 \end{description}
@@
 \item [{[2]}]
-for bound-constrained QPs, these sections are omitted.
+for BQPs, IBQPs, and MIBQPs, these sections are omitted.
 \item [{[3]}]
-for linear program with quadratic constraints, this section is omitted.
+for LPs, QCPs, ILPs, IQCPs, MILPs, MIQCPs, this section is omitted.
 \item [{[4]}]
-for problems without quadratic constraints, this section is omitted.
+for LPs, BQPs, QPs, ILPs, IBQPs, IBQPs, MILPs, MIBQPs, MIBQPs, this section is omitted.
 \item [{[5]}]

### 2017-01-30 QPLIB-2.2.tex
remove extra z, add top
- see issue #4, https://github.com/fabiofurini/QPLIB/issues/4#issuecomment-276062961

@@
 When a $Q^i$ is positive semidefinite (PSD), i.e., the corresponding constraint/objective function is convex, general quadratic constraints are in fact equivalent to diagonal ones.
-In fact, since every PSD matrix can be factorized as $Q^i = L^i (L^i)^\top$, e.g.~by the (incomplete) Cholesky factorization, $f^i(x) = x^\top Q^i x = \sum_{j \in \Mcn} z^2_jz$ where $z = x^\top L^i$.
+In fact, since every PSD matrix can be factorized as $Q^i = L^i (L^i)^\top$, e.g.~by the (incomplete) Cholesky factorization, $f^i(x) = x^\top Q^i x = \sum_{j \in \Mcn} z^2_j$ where $z^\top = x^\top L^i$.
 Hence, one could think that D** problems need not be distinguished from C** ones; however, this is true only for ``complicated'' constraints, but not for ``simple'' ones, because the above reformulation technique introduces linear constraints.

### 2017-02-01 QPLIB-1.0.tex
update to qplib format and that pesky 1/2

@@
 The structure of this paper is the following.
-In \S~\ref{sec:QPbasic} we review the basic notions about QP.
-In particular, \S~\ref{subsec:notation} sets out the notation, \S~\ref{sec:classification} proposes a---to the best of our knowledge, new---taxonomy of QP that helps us in discussing the (very) different classes of QPs, \S~\ref{sec:algo} very briefly reviews the solution methods for QP and the solvers we have employed.
-Then, \S~\ref{sec:lib} describes the process used to obtain the library and its results.
-%; the software tools that we have used, and that are freely released together with the library, are discussed in \S~\ref{subsec:tools}. 
-Some conclusions are drawn in \S~\ref{sec:conclusions}, after which Appendix A provides a complete description of all the instances of the library, while Appendix B describes the QPLIB file format.
+In \S\ref{sec:QPbasic} we review the basic notion of QP.
+In particular, \S\ref{subsec:notation} sets out the notation, \S\ref{sec:classification} proposes a---to the best of our knowledge, new---taxonomy of QP that helps us in discussing the (very) different classes of QPs, \S\ref{sec:algo} very briefly reviews the solution methods for QP and the solvers we have employed.
+Then, \S\ref{sec:lib} describes the process used to obtain the library and its results.
+%; the software tools that we have used, and that are freely released together with the library, are discussed in \S\ref{subsec:tools}.
+Some conclusions are drawn in \S\ref{sec:conclusions}, after which Appendix A provides a complete description of all the instances of the library, while Appendix B describes the QPLIB file format.
 
@@
 All solvers in the latest addition (6) compute global optima, they are ANTIGONE,
-BARON, COUENNE, LINDO, and SCIP, c.f.~\S~\ref{subsec:solver}. This benchmark is based on MINLPLib~2 \cite{Vigerske2014}, a
+BARON, COUENNE, LINDO, and SCIP, c.f.~\S\ref{subsec:solver}. This benchmark is based on MINLPLib~2 \cite{Vigerske2014}, a
 collection of presently 1388 instances. In order to give a first representative

### 2017-02-01 QPLIB-2.1.tex
update to qplib format and that pesky 1/2

@@
  \min \textrm{or} \max \;\;
- & x^\top Q^0 x + b^0 x + q^0 \\
+ & \half x^\top Q^0 x + b^0 x + q^0 \\
  \textrm{such that}\;\;
- & c^i_l \leq x^\top Q^i x + b^i x \leq c^i_u & i \in \Mcm \\
+ & c^i_l \leq \half x^\top Q^i x + b^i x \leq c^i_u & i \in \Mcm \\
  %
@@
  %
- \item $Q^i$ for $i \in \{ 0 \} \cup \Mcm$ are symmetric $n \times n$ real matrices: because one is always only interested in the value of quadratic functions of the type $x^\top Q^i x$, symmetry can be assumed without loss of generality by just replacing both $Q^i_{hk}$ and $Q^i_{kh}$ with their average $(Q^i_{hk} + Q^i_{kh}) / 2$;
+ \item $Q^i$ for $i \in \{ 0 \} \cup \Mcm$ are symmetric $n \times n$ real (Hessian) matrices: since one is only interested in the value of quadratic forms of the type $x^\top Q^i x$, symmetry can be assumed without loss of generality by just replacing off diagonal pairs $Q^i_{hk}$ and $Q^i_{kh}$ with their average $(Q^i_{hk} + Q^i_{kh}) / 2$;
  %

### 2017-02-01 QPLIB-2.2.tex
update to qplib format and that pesky 1/2

--- 2017-01-30 QPLIB-2.2.tex
@@
 
-Despite the apparent simplicity of the definition given in \S~\ref{subsec:notation}, Quadratic Programming instances can be of several rather different ``types'' in practice, depending on the fine details of the data.
+Despite the apparent simplicity of the definition given in \S\ref{subsec:notation}, Quadratic Programming instances can be of several rather different ``types'' in practice, depending on the fine details of the data.
 In particular, many algorithmic approaches can only be applied to QP when the data of the problem has specific properties.
@@
  \min \; \Big\{ \,
- \sum_{j \in \Mcn} \big( \, Q^0_j x_j^2 + b^0_j x_j \, \big) \;:\;
+ \sum_{j \in \Mcn} \big( \, \half Q^0_j x_j^2 + b^0_j x_j \, \big) \;:\;
  l_j \leq x_j \leq u_j \quad j \in \Mcn \;\;,\;\;
@@
  \min \; \Big\{ \, b^0 x \;:\;
- x^\top Q^i x + b^i x \leq c^i \quad i \in \Mcm \;\;,\;\;
+ \half x^\top Q^i x + b^i x \leq c^i \quad i \in \Mcm \;\;,\;\;
  x_j \in \mathbb{Z} \quad j \in \Mcn \, \Big\}
@@
  %
- & -\infty \leq x^\top Q^0 x + b^0 x \leq x^0 & \\ 
+ & -\infty \leq \half x^\top Q^0 x + b^0 x \leq x^0 & \\
  %
- & c^i_l \leq x^\top Q^i x + b^i x \leq c^i_u & i \in \Mcm \\ 
+ & c^i_l \leq \half x^\top Q^i x + b^i x \leq c^i_u & i \in \Mcm \\
  %
@@
 When a $Q^i$ is positive semidefinite (PSD), i.e., the corresponding constraint/objective function is convex, general quadratic constraints are in fact equivalent to diagonal ones.
-In fact, since every PSD matrix can be factorized as $Q^i = L^i (L^i)^\top$, e.g.~by the (incomplete) Cholesky factorization, $f^i(x) = x^\top Q^i x = \sum_{j \in \Mcn} z^2_jz$ where $z = x^\top L^i$.
+In fact, since every PSD matrix can be factorized as $Q^i = L^i (L^i)^\top$, e.g.~by the (incomplete) Cholesky factorization, the term $\half x^\top Q^i x \equiv \half \sum_{j \in \Mcn} z^2_j$ where $z^\top = x^\top L^i$.
 Hence, one could think that D** problems need not be distinguished from C** ones; however, this is true only for ``complicated'' constraints, but not for ``simple'' ones, because the above reformulation technique introduces linear constraints.
 Indeed, while C*L (and, a fortiori, C*\{C,Q\}) can always be brought to D*L (D*\{C,Q\}), using the same technique C*B becomes D*L, which is significantly different from D*B.
-In practice, a diagonal convex objective function under linear constraints is found in many applications (e.g., \cite{FrFG16,FrGe06a,FrGe09a,FGGP11}), so that D*L still makes sense to distinguish the case where the objective function is ``naturally'' separable from that where separability is artificially introduced.
+In practice, a diagonal convex objective function under linear constraints is found in many applications (e.g., \cite{FrFG16,FrGe06a,FrGe09a,FGGP11}), so that it still makes sense to distinguish the D*L case where the objective function is ``naturally'' separable from that where separability is artificially introduced.
 

### 2017-02-01 QPLIB-3.tex
update to qplib format and that pesky 1/2

@@
 In this section we present all the steps performed in order to build the
-new library. In \S~\ref{subsec:instColl}, we describe the set of gathered
-instances. In \S~\ref{subsec:selection} we present the features used to
+new library. In \S\ref{subsec:instColl}, we describe the set of gathered
+instances. In \S\ref{subsec:selection} we present the features used to
 classify the instances and we discuss the issues concerning the format of
-the instances. Finally, in \S~\ref{subsec:final set}, we describe the
+the instances. Finally, in \S\ref{subsec:final set}, we describe the
 selection process used to filter the instances and we graphically present
@@
 %according to their format and subsequently translated.
-%In \S~\ref{subsec:tools}
+%In \S\ref{subsec:tools}
 %%~\ref{subsec:CP_convex}
@@
 format presents three advantages: it is easier to read by a self-made parser, it produces smaller files, and the order of variables and constraints is fixed.
-See the \S~\ref{sec:format} for more details.
+See the \S\ref{sec:format} for more details.
 
@@
 which allowed us to classify the instances into the QP categories described in
-\S~\ref{sec:QPbasic}. As far as the variable types are concerned, we
+\S\ref{sec:QPbasic}. As far as the variable types are concerned, we
 collected the following information:
@@
 \begin{itemize}
- \item percentage of positive and negative eigenvalues of the $Q^0$ matrix;
+ \item percentage of positive and negative eigenvalues of the Hessian $Q^0$;
        % (\emph{\% neg eig}),
  %
- \item density of the $Q^0$ matrix (number of nonzero entries divided by the total
+ \item density of the Hessian $Q^0$ (number of nonzero entries divided by the total
        number of entries). %(\emph{\% dens}).
@@
        for the hardness of an instance is the CPU time needed by a
-       complete solver (cf.~\S~\ref{sec:algo}) to solve it to
+       complete solver (cf.~\S\ref{sec:algo}) to solve it to
        global optimality. Accordingly, for each of the gathered instance we
@@
 the instances with a quadratic objective function, ordered by the
-the percentage ``hard'' eigenvalues in the $Q^0$ matrix (vertical axis), by which
+the percentage ``hard'' eigenvalues in the Hessian $Q^0$ (vertical axis), by which
 we mean negative eigenvalues in case of a minimization problem and positive eigenvalues in case of a maximization problem.
@@
   \includegraphics[width=0.49\textwidth]{pic_density.pdf}
-  \caption{``Hard'' eigenvalues (left) and density (right) of $Q^0$ matrix for QPLIB instances with a quadratic objective function. \label{fig:pic_neg_eig}}
+  \caption{``Hard'' eigenvalues (left) and density (right) of the Hessian $Q^0$ for QPLIB instances with a quadratic objective function. \label{fig:pic_neg_eig}}
 \end{figure}
@@
 quadratic objective function sorted according to the density of the
-$Q^0$ matrix (vertical axis). The majority of the instances have either
+Hessian $Q^0$ (vertical axis). The majority of the instances have either
 a very low or a rather high density: indeed, about 30\% of the instances
@@
 
-Additional details on the instance features can be found in \S~\ref{sec:instance_details}.
+Additional details on the instance features can be found in \S\ref{sec:instance_details}.
 

### 2017-02-01 format.tex
add missing [2]
update to qplib format and that pesky 1/2

--- 2017-01-30 format.tex
@@
 \newcommand{\s}[1]{^{\mbox{\protect\tiny #1}}}
-\newcommand{\sfrac}[2]{{\scriptstyle \frac{#1}{#2}}}
-\newcommand{\half}{\sfrac{1}{2}}
 \newcommand{\smallRe}
@@
 
-The QPLIB format is defined in Table~\ref{tab-qplib-format}, with the notation of \S~\ref{sec:QPbasic}.
+The QPLIB format is defined in Table~\ref{tab-qplib-format}, with the notation of \S\ref{sec:QPbasic}.
 
@@
 $q^0$ & constant part of the objective function & \\ \hline
-$n^{Q^i}$ & number of nonzeros (integer) in lower triangle of
+$n^{Q^i}$ & total number of nonzeros (integer) in lower triangles of
 $Q^i$  & [2,4] \\
 $i$\; $h$\; $k$\; $Q^{i}_{hk}$
-& constraint, row and column indices (integers) and value (real) for
- each entry of $Q^{i}$, if $n^{Q^i}$, one quadruple on each line & \\
+& i, row and column indices (integers) and value (real) for
+ each entry of $Q^{i}$ for every $i \in \Mcm$,  if $n^{Q^i}$,
+ one quadruple on each line & \\
 \hline
-$n^{b^i}$ & number of nonzeros  (integer) in $b^i$, $i \in \Mcm$                     & [2] \\
-$i$\; $j$\; $b^i_{j}$ & row and column indices (integers) and value (real)
-for each nonzero
-entry of $b^i$, if $n^{b} > 0$, one triple on each line & \\
+$n^{b^i}$ & total number of nonzeros  (integer) in $b^i$ for all $i \in \Mcm$ &
+ [2] \\
+$i$\; $j$\; $b^i_{j}$ & i and index (integers) and value (real)
+for each nonzero entry of $b^i$ for every $i \in \Mcm$, if $n^{b} > 0$,
+one triple on each line & \\
 \hline
@@
 $i$\;\ $c^i_{u}$ & index (integer) and value (real) for each non-default
-term
-in $c_{u,d}$, if $n^{c_{u,d}} > 0$, one pair per line        & \\
+term in $c_{u,d}$, if $n^{c_{u,d}} > 0$, one pair per line        & [2] \\
 \hline
@@
 $v_d$ & default variable type  (integer, 0 for continuous variables,
- 1 for integer variables) & [5] \\
+ 1 for integer variables, 2 for binary variables) & [5] \\
 $n^v$ & number of non-default variables (integer)  & [5] \\
@@
 %\todo[inline]{\small Fabio and Emiliano: we think that, if possible, we should follow the classification proposed in Section 2.}
-
-The problem type is represented by a character string as one of the following:
-\begin{description}
-\item For continuous problems (i.e., all variables are continuous):
-\begin{description}[leftmargin=!,labelwidth=\widthof{\texttt{SMILQPC}}]
-\item [LP]      a linear program,
-\item [LPQC]    a linear program with quadratic constraints,
-\item [BQP]     a bound-constrained quadratic program,
-\item [QP]      a quadratic program, and
-\item [QPQC]    a quadratic program with quadratic constraints.
-\end{description}
-\item For integer problems (i.e., all variables are integer valued):
-\begin{description}[leftmargin=!,labelwidth=\widthof{\texttt{SMILQPC}}]
-\item [ILP]     an integer linear program,
-\item [ILPQC]   an integer linear program with quadratic constraints,
-\item [IBQP]    an integer bound-constrained quadratic program,
-\item [IQP]     an integer quadratic program, and
-\item [IQPQC]   an integer quadratic program with quadratic constraints.
-\end{description}
-\item For mixed-integer problems (i.e., there is a mix of continuous and
-integer variables):
-\begin{description}[leftmargin=!,labelwidth=\widthof{\texttt{SMILQPC}}]
-\item [MILP]    a mixed-integer linear program,
-\item [MILPQC]  a mixed-integer linear program with quadratic constraints,
-\item [MIBQP]   a mixed-integer bound-constrained quadratic program,
-\item [MIQP]    a mixed-integer quadratic program, and
-\item [MIQPQC]  a mixed-integer quadratic program with quadratic constraints.
-\end{description}
-\end{description}
+The problem type is represented by a three character string as given
+in \S\ref{ssec:taxonomy}
+%\begin{description}
+%\item For continuous problems (i.e., all variables are continuous):
+%\begin{description}[leftmargin=!,labelwidth=\widthof{\texttt{SMILQPC}}]
+%\item [LP]   a problem with linear objective and linear constraints,
+%\item [QCP]  a problem with linear objective and linear and quadratic constraints,
+%\item [BQP]  a problem with quadratic objective and variable bounds,
+%\item [QP]   a problem with quadratic objective and linear constraints, and
+%\item [QCQP] a problem with quadratic objective and linear and quadratic constraints.
+%\end{description}
+%\item For integer problems (i.e., all variables are integer valued):
+%\begin{description}[leftmargin=!,labelwidth=\widthof{\texttt{SMILQPC}}]
+%\item [ILP]   an LP plus integer requirement for all variables,
+%\item [IQCP]  an QCP plus integer requirement for all variables,
+%\item [IBQP]  an BQP plus integer requirement for all variables,
+%\item [IQP]   an QP plus integer requirement for all variables, and
+%\item [IQCQP] an QCQP plus integer requirement for all variables.
+%\end{description}
+%\item For mixed-integer problems (i.e., there is a mix of continuous and
+%integer variables):
+%\begin{description}[leftmargin=!,labelwidth=\widthof{\texttt{SMILQPC}}]
+%\item [MILP]   an LP plus integer requirement for some variables,
+%\item [MIQCP]  an QCP plus integer requirement for some variables,
+%\item [MIBQP]  an BQP plus integer requirement for some variables,
+%\item [MIQP]   an QP plus integer requirement for some variables, and
+%\item [MIQCQP] an QCQP plus integer requirement for some variables.
+%\end{description}
+%\end{description}
 
 \item [{[2]}]
-for bound-constrained QPs, these sections are omitted.
+for problems of type \textit{**N} or \textit{**B}, these lines/sections
+are omitted.
 \item [{[3]}]
-for linear program with quadratic constraints, this section is omitted.
+for problems of type \textit{L**}, this section is omitted.
 \item [{[4]}]
-for problems without quadratic constraints, this section is omitted.
+for problems of type \textit{**N}, \textit{**B} or \textit{**L},
+this section is omitted.
 \item [{[5]}]
-for purely-continuous or purely-integer problems, this section is omitted.
+for problems of type \textit{*C*}, \textit{*B*} or \textit{*I*},
+this section is omitted.
 \end{description}
@@
 \st 1 \leq x_1 + x_2, \; 1 \leq x_1 + x_3, \; 0 \leq x_1 \leq 1,
-\; 0 \leq x_2 \leq 2, \; \mbox{and binary} \; x_3.}
+\; 0 \leq x_2 \leq 2, \; \mbox{and binary} \; x_3,}
 \]
+for which the Hessian of the objective function is
+\[Q^0 = \mat{ccc}{2 & -1 & 0 \\ -1 & 2 & -1 \\ 0 & -1 & 2}.\]
 This may then be represent in QPLIB format as follows:
@@
 MIPBAND   # problem name
-MIQP      # problem is a mixed-integer quadratic program
+QML       # problem is a mixed-integer quadratic program
 Minimize  # minimize the objective function
@@
 5         # nonzeros in lower triangle of Q^0
-1 1 2.0   5 lines row & column index & value of nz. in lower triangle Q^0
+1 1 2.0   5 lines row & column index & value of nonzero in lower triangle Q^0
 2 1 -1.0  |
@@
 1         # non default entries in b_0
-2 -0.4    1 lines of index & value of non-default values in b_0
+2 -0.4    1 line of index & value of non-default values in b_0
 0.0       value of q^0
-4         # nonzeros in vectors b^i (i =0,...,m)
-1 1 1.0   4 lines row & column index & value of nz. in b^i (i =0,...,m)
+4         # nonzeros in vectors b^i (i=0,...,m)
+1 1 1.0   4 lines constraint, index & value of nonzero in b^i (i=0,...,m)
 1 2 1.0   |
@@
 1         # non default variable types
-3 1       variable 3 is integer
+3 2       variable 3 is binary
 1.0       default value for initial values for x

### 2017-02-01 methods.tex
update to qplib format and that pesky 1/2

@@
 
-In this section we provide a quick overview of existing solution methods for QP, restricting ourselves to these implemented by the set of solvers considered in this paper (\S~\ref{subsec:solver}).
-For each approach we briefly describe the main algorithmic ideas and point out the formulation they address according to the classification set out in \S~\ref{sec:classification}.
+In this section we provide a quick overview of existing solution methods for QP, restricting ourselves to these implemented by the set of solvers considered in this paper (\S\ref{subsec:solver}).
+For each approach we briefly describe the main algorithmic ideas and point out the formulation they address according to the classification set out in \S\ref{sec:classification}.
 We remark that many solvers implement more than one algorithm, among which the user can choose at runtime.

### 2017-02-01 format.tex
minor fixes
add comment that binary vars become integer when type is *I*

@@
 $h$\; $k$\; $Q_{hk}^0$ & row and column indices (integers) and value (real)
-for each nonzero entry of $Q^0$, if $n^{Q^0} > 0$, one triple on each line & \\
+for each nonzero entry of $Q^0$, if $n^{Q^0} > 0$, one triple on each line & [3]\\
 \hline
@@
 for each nonzero entry of $b^i$ for every $i \in \Mcm$, if $n^{b} > 0$,
-one triple on each line & \\
+one triple on each line & [2] \\
 \hline
@@
 \item [{[2]}]
-for problems of type \textit{**N} or \textit{**B}, these lines/sections
+for problems of type \texttt{**N} or \texttt{**B}, these lines/sections
 are omitted.
 \item [{[3]}]
-for problems of type \textit{L**}, this section is omitted.
+for problems of type \texttt{L**}, this section is omitted.
 \item [{[4]}]
-for problems of type \textit{**N}, \textit{**B} or \textit{**L},
+for problems of type \texttt{**N}, \texttt{**B} or \texttt{**L},
 this section is omitted.
 \item [{[5]}]
-for problems of type \textit{*C*}, \textit{*B*} or \textit{*I*},
-this section is omitted.
+for problems of type \texttt{*C*}, \texttt{*B*} or \texttt{*I*},
+this section is omitted. In case of \texttt{*I*}, binary variable need to be specified as general integer variables with bounds 0 and 1.
 \end{description}

### 2017-02-01 methods.tex
addded citations to commonly-used non-GAMS continuous QP solvers

@@
 Note that all these solvers tackle the more general class of continuous NLP, while we use them only for the considerably more restricted class of QP.
-While there are many different approaches for continuous NLPs in general, we therefore describe here on those that are particularly well-suited for QP, i.e.:
+Aside from solvers provided by GAMS, there are a number of other, specialized incomplete QP solvers, such as
+\cqp \cite{GoulOrbaRobi13:mpc},
+\dqp \cite{GoulRobi16:coap} and
+\ooqp \cite{GertWrig03:toms}
+for convex problems, and
+\bqpd \cite{Flet:2000},
+\qpa \cite{GoulToin02i} and
+\qpb \cite{ConnGoulOrbaToin:2000},
+\qpc \cite{GoulOrbaToin03:toms},
+\sqic \cite{Gill2015}
+for nonconvex ones
+While there are many different approaches for continuous NLPs in general, we mention here those that are particularly well-suited for QP, i.e.:
 %
 \begin{itemize}
- \item active set methods (\conopt, \minos, \snopt, \knitro) \cite{Dost09};
+ \item active set methods (\conopt, \minos, \snopt, \knitro, \bqpd, \qpa, \sqic) \cite{Dost09};
  %
- \item interior point methods (\ipopt, \knitro) \cite{Wright97};
+ \item interior point methods (\ipopt, \knitro, \cqp, \ooqp, \qpb, \qpc) \cite{Wright97};
  %
- \item projected gradient methods \cite{CalaMore87:mp,ChenGui13:coap}.
+ \item projected gradient methods (\dqp) \cite{CalaMore87:mp,ChenGui13:coap}.
 \end{itemize}
 %
-Active set and interior point methods have been defined for *C* but also for general (continuous) NLPs. In fact, all of the solvers named above target this larger problem class.
+Active set and interior point methods have been defined for *C* but also for general (continuous) NLPs, and many of the solvers named above target this
+larger problem class.
 
@@
 %
-These approaches are similar to active-set methods in that at each iteration they consider the gradient projected onto the set of active constraints, in order to (try to) guarantee feasibility of the next iterate.
-They differ from active-set methods in that the set of active constraints can change dramatically on each iteration, and they have stronger convergence properties, more akin to those of interior-point methods.
+These approaches are similar to active-set methods in that at each iteration they project the steepest descent direction onto the constraint set in order to (try to) predict the optimal active set; the predicted set may then be searched more exhaustively to improve the objective function.
+They differ from active-set methods in that the set of active constraints can change dramatically on each iteration, and they often exhibit stronger practical convergence behaviour, more akin to those of interior-point methods.
+However they rely heavily on projection, and thus are only suited to simple constraint geometries such as boxes and simplices.
 

### 2017-02-01 format.tex
add note that for binary variables, a different default for variable bounds may apply

@@
 \hline
-$l_{d}$ & default value (real) for entries in $l$            & \\
-$n^{l_{d}}$ & number of non-default entries (integer) in $l$    &  \\
+$l_{d}$ & default value (real) for entries in $l$            & [6] \\
+$n^{l_{d}}$ & number of non-default entries (integer) in $l$    & [6] \\
 $i$\;\ $l_{i}$ & index (integer) and value (real) for each non-default
 term
-in $l$, if $n^{l_{d}}> 0$, one pair per line        &  \\
+in $l$, if $n^{l_{d}}> 0$, one pair per line        &  [6] \\
 \hline
-$u_{d}$ & default value (real) for entries in $u$            & \\
-$n^{u_{d}}$ & number of non-default entries (integer) in $u$    &  \\
+$u_{d}$ & default value (real) for entries in $u$            & [6] \\
+$n^{u_{d}}$ & number of non-default entries (integer) in $u$    &  [6] \\
 $i$\;\ $u_{i}$ & index (integer) and value (real) for each non-default
 term
-in $u$, if $n^{u_{d}}> 0$, one pair per line        &  \\
+in $u$, if $n^{u_{d}}> 0$, one pair per line        & [6] \\
 \hline
@@
 \item [{[2]}]
-for problems of type \texttt{**N} or \texttt{**B}, these lines/sections
+For problems of type \texttt{**N} or \texttt{**B}, these lines/sections
 are omitted.
 \item [{[3]}]
-for problems of type \texttt{L**}, this section is omitted.
+For problems of type \texttt{L**}, this section is omitted.
 \item [{[4]}]
-for problems of type \texttt{**N}, \texttt{**B} or \texttt{**L},
+For problems of type \texttt{**N}, \texttt{**B} or \texttt{**L},
 this section is omitted.
 \item [{[5]}]
-for problems of type \texttt{*C*}, \texttt{*B*} or \texttt{*I*},
+For problems of type \texttt{*C*}, \texttt{*B*} or \texttt{*I*},
 this section is omitted. In case of \texttt{*I*}, binary variable need to be specified as general integer variables with bounds 0 and 1.
+\item [{[6]}]
+For problems of type \texttt{*B*} and for variables that are specified explicitly as binary variables, the default values for lower and upper bounds of a binary variables are assumed to be 0 and 1, respectively, independent of the specification of $l_d$ and $u_d$.
+%Note, that in case of \texttt{*I*}, binary variable need to be specified as general integer variables with bounds 0 and 1.
 \end{description}

### 2017-02-01 QPLIB-1.0.tex
cleanup problem notation
- LPs and NLPs are always continuous
- LP/NLP + integer variables is called MILP/MINLP
- QP can be continuous and discrete
- closes #1

@@
 However, this diversity also implies that QP means very different things to different researchers.
-This is illustrated by the simple fact that the class of problems that we simply refer to here as ``QP'' is called in different ways, among which Mixed-Integer Quadratically Constrained Quadratic Problems (MI-QCQP).
+This is illustrated by the simple fact that the class of problems that we simply refer to here as ``QP'' is called in different ways, among which Mixed-Integer Quadratically Constrained Quadratic Problems (MIQCQP).
 It is perhaps therefore not surprising that, unlike for ``simpler'' problems classes \cite{Koch2011}, so far there has never been a single library containing all different kinds of instances of QP.
@@
 At least some of the rationale for this interest is likely due to the fact that they are the ``least-nonlinear nonlinear problems''.
-Hence, in particular for the convex case, tools and techniques that have been honed during decades of research for Linear Programming (LP), typically with integrality constraints (MILP), can often be extended to the quadratic case with at least less effort than what would be required for tackling general Non Linear Programming (NLP) problems, without or with integrality constraints (MINLP).
+Hence, in particular for the convex case, tools and techniques that have been honed during decades of research for Linear Programming (LP), typically with integrality constraints (MILP), can often be extended to the quadratic case with at least less effort than what would be required for tackling general (Mixed-Integer) Nonlinear Programming ((MI)NLP) problems.
 This has indeed happened over and over again with different algorithmic techniques, such as Interior Point methods, active-set methods (of which the simplex method is a prototypical example), enumeration methods, cut-generation techniques, reformulation techniques, and many others \cite{BS09}.
@@
 
-On the other hand, (MI)QP is, in some sense, a considerably more expressive class than (MI)LP.
+On the other hand, QP, that is, quadratic problems with continuous or integer variables, is, in some sense, a considerably more expressive class than (MI)LP.
 Quadratic expressions are found, either naturally or after appropriate reformulations, in very many optimization problems \cite{Kochenberger2014}.
@@
 Also, in general any continuous function can be approximated with arbitrary accuracy (over a compact set) by a polynomial of arbitrary degree; in turn, every polynomial can be broken down to a system of quadratic expressions.
-Hence, (MI)QP is, in some sense, roughly as expressive as the whole of (MI)NLP.
-Of course this is, in principle, true for (MI)LP as well, but at the cost of much larger and much more complex formulations.
-Hence, for many applications QP may represent the ``sweet spot'' between the effectiveness, but lower expressive power, of (MI)LP and the higher expressive power, but much higher computational cost, of (MI)NLP.
+Hence, QP is, in some sense, roughly as expressive as the whole of MINLP.
+Of course this is, in principle, true for MILP as well, but at the cost of much larger and much more complex formulations.
+Hence, for many applications QP may represent the ``sweet spot'' between the effectiveness, but lower expressive power, of MILP and the higher expressive power, but much higher computational cost, of MINLP.
 
@@
 %
-\caption{Application Domains of (MI)QP}\label{tbl:application} \\
+\caption{Application Domains of QP}\label{tbl:application} \\
 %
@@
 %
-\multicolumn{3}{l}{\emph{Table \ref{tbl:application} (Application Domains of (MI)QP) continued}} \\
+\multicolumn{3}{l}{\emph{Table \ref{tbl:application} (Application Domains of QP) continued}} \\
 \toprule

### 2017-02-01 methods.tex
cleanup problem notation
- LPs and NLPs are always continuous
- LP/NLP + integer variables is called MILP/MINLP
- QP can be continuous and discrete
- closes #1

@@
 Therefore, when we state that a given algorithm is incomplete or (asymptotically) complete we mean that it is so the largest problem class that the solver naturally targets, although it may be complete on specific sub-classes.
-For example, interior point algorithms naturally target continuous NLPs and are incomplete on nonconvex NLPs, and therefore on QCQ, but become complete for CCC.
+For example, interior point algorithms naturally target NLPs and are incomplete on NLPs, and therefore on QCQ, but become complete for CCC.
 In general, all complete methods for a problem $P$ must be complete for any problem $Q \subseteq P$, while a complete method for $P$ might be incomplete for $R \supset P$.
@@
 
-The solvers in Table \ref{t:solvers} which implement incomplete methods for continuous NLPs (a problem class containing QCQ) are \conopt, \ipopt, \minos, \snopt, and \knitro.
-Note that all these solvers tackle the more general class of continuous NLP, while we use them only for the considerably more restricted class of QP.
+The solvers in Table \ref{t:solvers} which implement incomplete methods for NLPs (a problem class containing QCQ) are \conopt, \ipopt, \minos, \snopt, and \knitro.
+Note that all these solvers tackle the more general class of NLP, while we use them only for the considerably more restricted class of QP.
 Aside from solvers provided by GAMS, there are a number of other, specialized incomplete QP solvers, such as
@@
 for nonconvex ones
-While there are many different approaches for continuous NLPs in general, we mention here those that are particularly well-suited for QP, i.e.:
+While there are many different approaches for NLPs in general, we mention here those that are particularly well-suited for QP, i.e.:
 %
@@
 %
-Active set and interior point methods have been defined for *C* but also for general (continuous) NLPs, and many of the solvers named above target this
+Active set and interior point methods have been defined for *C* but also for general NLPs, and many of the solvers named above target this
 larger problem class.
@@
 In general, the applicability of meta-heuristics to a given problem relies on availability of a local search that exploit the structural properties of the problem.
-Depending on the local search employed, Multi-start methods can address MINLP of the most general class.
+Depending on the local search employed, multi-start methods can address MINLP of the most general class.
 

### 2017-02-02 format.tex
binary variables now always have lower bound 0 and upper bound 1
- thus, no bounds can be specified for binary variables
- thus, bounds sections can be ignored for *B* type instances
- see discussion in d4e031c1

@@
 \item [{[6]}]
-For problems of type \texttt{*B*} and for variables that are specified explicitly as binary variables, the default values for lower and upper bounds of a binary variables are assumed to be 0 and 1, respectively, independent of the specification of $l_d$ and $u_d$.
-%Note, that in case of \texttt{*I*}, binary variable need to be specified as general integer variables with bounds 0 and 1.
+For problems of type \texttt{*B*}, this section is omitted.
+Bounds for binary variables must not be specified in this section.
+They are assumed to have a lower bound of 0 and an upper bound of 1, disregarding any setting of $l_d$ and $u_d$.
+To fix a binary variable to 0 or 1, either a corresponding linear equation must be added or the variable type should be changed to continuous or general integer.
 \end{description}

### 2017-02-02 QPLIB-3.tex
proofread website section

@@
 
-The instances of QPLIB are publicly accessible at the website \url{http://qplib.zib.de}.
-Scripts and tools developed for MINLPLib 2 \cite{Vigerske2014} were used and extended to create the website.
+The instances of QPLIB are publicly accessible at the website \url{http://qplib.zib.de}, which was created by extending
+scripts and tools initially developed for MINLPLib\,2~\cite{Vigerske2014}.
 %
@@
 \includegraphics{{QPLIB_2967.hess}.png}
-\caption{Sparsity pattern of Gradient of objective (first line) and Jacobian (left) and of Hessian of Lagrangian function (right) for instance QPLIB\_2967.}
+\caption{Example for the sparsity pattern of the Jacobian of the constraint functions (left) and
+  of the Hessian of the Lagrangian function (right) for instance QPLIB\_2967.
+  The gradient of the objective function is displayed as the first row
+  of the Jacobian matrix.}
 \label{fig:sparsitypattern}
@@
 The entire set of instances can be explored in a searchable and sortable table of selected instance features: problem
-classification, convexity of the continuous relaxation, number of (all, binary, integer) variables, (all, quadratic) constraints, nonzeros, hard eigenvalues in $Q^0$, and density of $Q^0$.
+classification, convexity of the continuous relaxation, number of (all, binary, integer) variables, (all, quadratic) constraints, nonzeros, hard eigenvalues in~$Q^0$, and density of~$Q^0$.
 %
 Finally, a statistics page displays diagrams on the composition of the library according to different criteria: the
-number instances according to problem type, variable and constraint types, convexity, number of variables and constraints, and density.
+number of instances according to problem type, variable and constraint types, convexity, problem size, and density.
 %

### 2017-02-02 QPLIB.tex
fix some latex warnings
update to qplib format and that pesky 1/2
the min-1/2 looks bad, make it a proper 1/2
- also good to have this clearly visible
addded citations to commonly-used non-GAMS continuous QP solvers
add funding acknowledgement

@@
 \usepackage{doi}
+\usepackage[english]{babel}
+\usepackage{microtype}
 
@@
 \def\snopt{\textsc{SNOPT}\xspace}
+\def\qpa{\textsc{QPA}\xspace}
+\def\qpb{\textsc{QPB}\xspace}
+\def\qpc{\textsc{QPC}\xspace}
+\def\cqp{\textsc{CQP}\xspace}
+\def\dqp{\textsc{DQP}\xspace}
+\def\ooqp{\textsc{OOQP}\xspace}
+\def\sqic{\textsc{SQIC}\xspace}
+\def\bqpd{\textsc{BQPD}\xspace}
 
@@
 \fi
+
+%\newcommand{\half}{{\scriptstyle \frac{1}{2}}}
+%\newcommand{\sfrac}[2]{{\scriptstyle \frac{#1}{#2}}}
+\newcommand{\half}{\frac{1}{2}}
 
@@
 We are grateful to all the donors who provided instances for the library.
-We gratefully acknowledge the financial support of the Gaspard Monge Program for Optimization and operations research (PGMO)  and the logistic support of GAMS for having provided us the licence of their software.
+We gratefully acknowledge the financial support of the Gaspard Monge Program for Optimization and operations research (PGMO)  and the logistic support of GAMS for having provided us with the license of their software.
 Finally, we would like to acknowledge the financial and networking support by the COST Action TD1207.
+The fifth author was supported by the Research Campus MODAL \emph{Mathematical Optimization and Data Analysis Laboratories} funded by the Federal Ministry of Education and Research (BMBF Grant~05M14ZAM).
+All responsibilty for the content of this publication is assumed by the authors.
 

### 2017-02-02 QPLIB.tex
update acknowledgement

@@
 Finally, we would like to acknowledge the financial and networking support by the COST Action TD1207.
-The fifth author was supported by the Research Campus MODAL \emph{Mathematical Optimization and Data Analysis Laboratories} funded by the Federal Ministry of Education and Research (BMBF Grant~05M14ZAM).
+The work of the fifth and twelfth author was supported by the Research Campus MODAL \emph{Mathematical Optimization and Data Analysis Laboratories} funded by the Federal Ministry of Education and Research (BMBF Grant~05M14ZAM).
 All responsibilty for the content of this publication is assumed by the authors.

### 2017-02-02 format.tex
slight clarification of the format regarding binary variables

@@
 For problems of type \texttt{*C*}, \texttt{*B*} or \texttt{*I*},
-this section is omitted. In case of \texttt{*I*}, binary variable need to be specified as general integer variables with bounds 0 and 1.
+this section is omitted. For problems of \texttt{*I*}, binary variables
+should be specified with lower and upper bounds 0 and 1.
 \item [{[6]}]
 For problems of type \texttt{*B*}, this section is omitted.
-Bounds for binary variables must not be specified in this section.
-They are assumed to have a lower bound of 0 and an upper bound of 1, disregarding any setting of $l_d$ and $u_d$.
-To fix a binary variable to 0 or 1, either a corresponding linear equation must be added or the variable type should be changed to continuous or general integer.
+%Bounds for binary variables must not be specified in this section.
 \end{description}
+Binary variables defined either implicitly via the type \texttt{*B*}
+or explicitly in the variable type section will be assumed to
+have lower and upper bounds 0 and 1, and this will override any
+explicit bounds $l_d$, $u_d$ $l_i$ and $u_i$ set in the lower and upper
+bound sections.
+To fix a binary variable to 0 or 1, its variable type should be changed
+to continuous or general integer and the corresponding lower and upper
+bounds set accordingly in the lower and upper bound sections.
 

### 2017-02-02 format.tex
fix typos
- put back the word "integer" to make clear that type must be changed in *I*

@@
 For problems of type \texttt{*C*}, \texttt{*B*} or \texttt{*I*},
-this section is omitted. For problems of \texttt{*I*}, binary variables
-should be specified with lower and upper bounds 0 and 1.
+this section is omitted. For problems of type \texttt{*I*}, binary variables
+should be specified as integer variables with lower and upper bounds 0 and 1.
 \item [{[6]}]
@@
 have lower and upper bounds 0 and 1, and this will override any
-explicit bounds $l_d$, $u_d$ $l_i$ and $u_i$ set in the lower and upper
+explicit bounds $l_d$, $u_d$, $l_i$, and $u_i$ set in the lower and upper
 bound sections.

### 2017-02-03 QPLIB-1.0.tex
small adjustments to improve the English. NB, I have commented out some of the last paragraph as the level of detail seemed disproportionate, and as these details are available in the citations provided. Please re-insert if it is generally felt that this is an ommision.

@@
 
-Quadratic Programming (QP) problems, where either the objective function \cite{wiki:qp}, or the constraints \cite{wiki:qcqp}, or both contain (at most) expressions in the variables of degree $2$, include a surprisingly diverse set of rather different instances.
-This is not surprising, given the vast scope of practical applications of these problems, and of solution methods designed to solve them \cite{GoulToin00a}.
-According to the fine details of the formulation, solving a QP may require employing either fundamentally combinatorial techniques, or ideas rooted in nonlinear optimization principles, or a mix of the two.
-In this sense, QP is likely one of the classes of problems where the collaboration between the communities interested in combinatorial and nonlinear optimization is more necessary, and potentially fruitful.
+Quadratic Programming (QP) problems---mathematical optimization problems for which either the objective function \cite{wiki:qp} or the constraints \cite{wiki:qcqp} (or both) are polynomial function of the variables of degree two---include a notably diverse set of different instances.
+This is not surprising, given the vast scope of practical applications of such problems, and of solution methods designed to solve them \cite{GoulToin00a}.
+Depending on the fine details of the formulation, solving a QP may require either fundamentally combinatorial techniques, or ideas rooted in nonlinear optimization principles, or a mix of the two.
+In this sense, QP is arguably one of the classes of problems where collaboration between the communities interested in combinatorial and in nonlinear optimization is most needed, and potentially fruitful.
 
 However, this diversity also implies that QP means very different things to different researchers.
-This is illustrated by the simple fact that the class of problems that we simply refer to here as ``QP'' is called in different ways, among which Mixed-Integer Quadratically Constrained Quadratic Problems (MIQCQP).
-It is perhaps therefore not surprising that, unlike for ``simpler'' problems classes \cite{Koch2011}, so far there has never been a single library containing all different kinds of instances of QP.
-Several libraries devoted to special cases of QP are indeed available; however, each of them is either focussed on one application (a specific problem that can be modeled as QP), or on QPs with specific structural properties that make them suitable to be solved with some given class of algorithmic approaches.
-To the best of our knowledge, collecting a set of QP instances that is at the same time not overwhelmingly numerous and significant for the many different interested communities has not been attempted, yet.
+This is illustrated by the simple fact that the class of problems that we simply refer to here as ``QP'' might more accurately be called Mixed-Integer Quadratically-Constrained Quadratic Programming (MIQCQP) in the most general case.
+It is, therefore, perhaps not surprising that, unlike for ``simpler'' problems classes \cite{Koch2011},  there has been, to date, no single library devoted to all different kinds of instances of QP.
+While several specialised libraries devoted to particular cases of QP are available, each of them is either focussed on a particular application (a specific problem that can be modelled as a QP), or on QPs with specific structural properties that make them suitable for solution by some given class of algorithmic approaches.
+To the best of our knowledge, collecting a set of QP instances that is at the same time not overwhelmingly numerous and significant for the many different interested communities has not been attempted.
 This work constitutes a first step in this direction.
 
-In this paper we report the steps that have been done to collect a (hopefully) significant library of QP instances, filtering the large set of available (or specifically provided) instances in order to end up with a manageable set that still contains a meaningful sample of all possible QP types.
+In this paper we report the steps that have been done to collect (we believe) a significant library of QP instances, filtering the a much larger set of currently available (or specifically provided) instances in order to end up with a manageable set that still contains a meaningful sample of all possible QP types.
 A particularly thorny issue in this process is how to select instances that are ``interesting''.
@@
 Our filtering process has then been in part based on the idea that if a significant fraction of the solvers that can solve a QP instance do so in a ``short'' time, then the instance is not challenging enough to be included in the library.
-Yet, if very few (maybe one) solvers can solve it very efficiently by exploiting some specific structure, but most other approaches cannot, then the instance can still be deemed ``interesting''.
-Putting this approach in practice requires a nontrivial number of technical steps and decisions that are detailed in the paper.
+Conversely, if very few (maybe one) solvers can solve it very efficiently by exploiting some specific structure, but most other approaches cannot, then the instance should be deemed ``interesting''.
+Putting this approach into practice requires a nontrivial number of technical steps and decisions that are detailed in the paper.
 We hope that our work can provide useful guidelines for other interested researchers.
 
-A consequence of our focus is that this paper is \emph{not} concerned about the different performance of the very diverse QP solvers; we will \emph{not} report any data comparing them.
-The only reason why solvers are used (and, therefore, described) in this context is to ensure that the instances of the library are nontrivial at least for a significant fraction of the current solution methods; providing guidance about which solver is most suited to some specific class of QPs is entirely outside the scope of our work.
+A consequence of our focus is that this paper is \emph{not} concerned with the different performance of the very diverse available set of QP solvers; we will \emph{not} report any data comparing them.
+The only reason that solvers are used (and, therefore, described) in this context is to ensure that the instances of the library are nontrivial at least for a significant fraction of the current solution methods.
+Providing guidance about which solvers are most suited to some specific class of QPs is entirely outside the scope of our work.
 
@@
 
-Optimization problems with quadratic constraints and/or objective function (QP) have been the subject of a considerable amount of research over the last 60 years.
+Optimization problems with quadratic constraints and/or objective function (QP) have been the subject of a considerable amount of research over the last almost seventy years.
 At least some of the rationale for this interest is likely due to the fact that they are the ``least-nonlinear nonlinear problems''.
-Hence, in particular for the convex case, tools and techniques that have been honed during decades of research for Linear Programming (LP), typically with integrality constraints (MILP), can often be extended to the quadratic case with at least less effort than what would be required for tackling general (Mixed-Integer) Nonlinear Programming ((MI)NLP) problems.
-This has indeed happened over and over again with different algorithmic techniques, such as Interior Point methods, active-set methods (of which the simplex method is a prototypical example), enumeration methods, cut-generation techniques, reformulation techniques, and many others \cite{BS09}.
-Similarly, nonconvex continuous QP are perhaps the ``simplest'' class of problems that require features like spatial enumeration techniques to be solved.
+Hence, in particular for the convex case, tools and techniques that have been honed during decades of research for Linear Programming (LP), typically with integrality constraints (MILP), can often be extended to the quadratic case more easily than would be required to tackle general (Mixed-Integer) Nonlinear Programming ((MI)NLP) problems.
+This has indeed happened over-and-over again with different algorithmic techniques, such as interior-point methods, active-set methods (of which the simplex method is a prototypical example), enumeration methods, cut-generation techniques, reformulation techniques, and many others \cite{BS09}.
+Similarly, nonconvex continuous QP are perhaps the ``simplest'' class of problems that require features, such as spatial enumeration techniques, for their solution.
 Hence, they are both the natural basis for the development of general techniques for nonconvex NLP, and a very specific class so that specialized approaches can be developed \cite{Dur2010,Burer2012}.
 
-On the other hand, QP, that is, quadratic problems with continuous or integer variables, is, in some sense, a considerably more expressive class than (MI)LP.
+In addition, QP, that is, quadratic problems with continuous or integer variables, is arguably a considerably more expressive class than (MI)LP.
 Quadratic expressions are found, either naturally or after appropriate reformulations, in very many optimization problems \cite{Kochenberger2014}.
 Table \ref{tbl:application} provides a certainly non-exhaustive collection of applications that lead to formulations with either quadratic constraints, or quadratic objective function, or both.
-Also, in general any continuous function can be approximated with arbitrary accuracy (over a compact set) by a polynomial of arbitrary degree; in turn, every polynomial can be broken down to a system of quadratic expressions.
+Also, in general any continuous function can be approximated with arbitrary accuracy (over a compact set) by a polynomial of arbitrary degree.
+In turn, every polynomial can be broken down to a system of quadratic expressions.
 Hence, QP is, in some sense, roughly as expressive as the whole of MINLP.
-Of course this is, in principle, true for MILP as well, but at the cost of much larger and much more complex formulations.
+Of course this is, in principle, true for MILP as well, but at the cost of much larger and much more complicated formulations.
 Hence, for many applications QP may represent the ``sweet spot'' between the effectiveness, but lower expressive power, of MILP and the higher expressive power, but much higher computational cost, of MINLP.
@@
 %
-\multicolumn{3}{l}{\textbf{Classical Problems that can be formulated as MIQP}} \\[1pt]
+\multicolumn{3}{l}{\textbf{Fundamental problems that can be formulated as MIQP}} \\[1pt]
 %
@@
 
-The structure of this paper is the following.
+The structure of this paper is as follows.
 In \S\ref{sec:QPbasic} we review the basic notion of QP.
-In particular, \S\ref{subsec:notation} sets out the notation, \S\ref{sec:classification} proposes a---to the best of our knowledge, new---taxonomy of QP that helps us in discussing the (very) different classes of QPs, \S\ref{sec:algo} very briefly reviews the solution methods for QP and the solvers we have employed.
-Then, \S\ref{sec:lib} describes the process used to obtain the library and its results.
+In particular, \S\ref{subsec:notation} sets out the notation, \S\ref{sec:classification} proposes a (to the best of our knowledge, new) taxonomy of QP that helps us in discussing the (very) different classes of QPs, and \S\ref{sec:algo} very briefly reviews the solution methods for QP and the solvers we have employed.
+Next \S\ref{sec:lib} describes the process used to obtain the library and its results.
 %; the software tools that we have used, and that are freely released together with the library, are discussed in \S\ref{subsec:tools}.
-Some conclusions are drawn in \S\ref{sec:conclusions}, after which Appendix A provides a complete description of all the instances of the library, while Appendix B describes the QPLIB file format.
+Some conclusions are drawn in \S\ref{sec:conclusions}, after which Appendix A provides a complete description of all the instances of the library, while Appendix B describes a simple (QPLIB) file format that encodes all of our examples.
 
-While no performance issues of solvers for QP/QCP problems are considered in
-this paper, we refer to the comprehensive benchmark site \url{http://plato.asu.edu/bench.html}. Of the benchmarks
-on this site three deal exlusively with QP/QCP problems:
-the (1) Large SOCP, (2) MISOCP, and the (3) MIQ(C)P benchmarks.
-The following contain partly results for such problems:
-(4) Parallel Barrier Solvers on Large LP/QP problems, (5) AMPL-NLP, (6) MINLP.
-The benchmarks (1,2,4) have only convex instances, while the others include
-nonconvex instances. Global optima are obtained by several of the solvers in
-the benchmarks (3,5).
-All solvers in the latest addition (6) compute global optima, they are ANTIGONE,
-BARON, COUENNE, LINDO, and SCIP, c.f.~\S\ref{subsec:solver}. This benchmark is based on MINLPLib~2 \cite{Vigerske2014}, a
-collection of presently 1388 instances. In order to give a first representative
-impression of solver performance, care was taken to reduce the number of instances and
-allow all solvers to finish in reasonable time. More than half of the selected
-instances are QP or QCP. A time limit of two hours on a
-modern workstation with sufficient memory was used. Computations were done in
-GAMS with scripts derived from the SCIP check scripts and available at GitHub.
+While no performance issues of solvers for QP problems are considered in
+this paper, we refer to the comprehensive benchmark site \url{http://plato.asu.edu/bench.html}. Of the result on this site, three deal exclusively with
+QP problems, namely the (1) Large SOCP, (2) MISOCP, and the (3) MIQ(C)P benchmarks, while three others contain have partial results for such problems,
+namely those for (4) Parallel Barrier Solvers on Large LP/QP problems, (5) AMPL-NLP and (6) MINLP.
+Benchmarks (1, 2 \& 4) contain only convex instances, while the others
+include nonconvex ones. Global optima are obtained by several of the solvers in
+benchmarks (3 \& 5), while all solvers in the latest addition (6) compute
+global optima.
+%, they are ANTIGONE,
+%BARON, COUENNE, LINDO, and SCIP, c.f.~\S\ref{subsec:solver}.
+Benchmark (6) is based on MINLPLib~2 \cite{Vigerske2014}, a
+collection of currently 1388 instances. In order to give a first representative
+impression of solver performance, care was taken there to reduce the number of
+instances and allow all solvers to finish in a reasonable time.
+More than half of the selected instances are QP or QCP.
+%A time limit of two hours on a
+%modern workstation with sufficient memory was used. Computations were done in
+%GAMS with scripts derived from the SCIP check scripts and available at GitHub.
 For details we refer to \url{http://plato.asu.edu/ftp/minlp.html}.

### 2017-02-03 QPLIB-2.1.tex
minor fixes
small adjustments to improve the English

@@
  \textrm{such that}\;\;
- & c^i_l \leq \half x^\top Q^i x + b^i x \leq c^i_u & i \in \Mcm \\
+ & c^i_l \leq \half x^\top Q^i x + b^i x \leq c^i_u & i \in \Mcm, \\
  %
- & l_j \leq x_j \leq u_j & j \in \Mcn  \\
+ & l_j \leq x_j \leq u_j & j \in \Mcn,  \\
  %
- & x_j \in \mathbb{Z} & j \in \Mcz
+ \textrm{and}\;\;
+ & x_j \in \mathbb{Z} & j \in \Mcz,
 \end{align*}
-
 %
@@
 %
-where:
+where
 %
@@
  %
- \item $x = [x_j]_{j = 1}^n \in \RR^n$ is a finite vector of real variables;
+ \item $x = [x_j]_{j = 1}^n$ %\in \RR^n$
+is a finite vector of real variables;
  %
@@
  %
- \item $-\infty \leq l_j \leq u_j \leq \infty$ are the (extended) real upper and lower bounds on each variable $x_j$ for $j \in \Mcn$;
+ \item $-\infty \leq l_j \leq u_j \leq \infty$ are the (extended) real lower and upper bounds on each variable $x_j$ for $j \in \Mcn$;
  %
- \item $\Mcm = \Mcq \cup \Mcl$ where $Q^i = 0$ for all $i \in \Mcl$ (i.e., these are the linear constraints, as opposed to the truly quadratic ones);
+ \item $\Mcm = \Mcq \cup \Mcl$ where $Q^i = 0$ for all $i \in \Mcl$ (i.e., these are the linear constraints, as opposed to the truly quadratic ones); and
  %
@@
 %
-Due to the presence of integral requirements on the variables and quadratic constraints, this class of problems is often referred to as Mixed-Integer Quadratically Constraint Quadratic Program (MIQCQP).
-It will be sometimes useful to refer to the (sub)set $\Mcb =  \{ \, i \in \Mcz : l_j = 0 , u_j = 1 \, \} \subseteq \Mcz$ of the binary variables, and to $\Mcr = \Mcn \setminus \Mcz$ as the set of continuous ones.
+Due to the presence of integrality requirements on the variables and of quadratic constraints, this class of problems is often referred to as Mixed-Integer Quadratically Constraint Quadratic Program (MIQCQP).
+It will be sometimes useful to refer to the (sub)set $\Mcb =  \{ \, j \in \Mcz : l_j = 0 , u_j = 1 \, \} \subseteq \Mcz$ of the binary variables, and to $\Mcr = \Mcn \setminus \Mcz$ as the set of continuous ones.
 Similarly, it will be sometimes useful to distinguish the (sub)set $\Mcx = \{ \, j : l_j > -\infty \lor u_j < \infty \, \}$ of the box-constrained variables from $\Mcu = \Mcn \setminus \Mcx$ of the unconstrained ones (in the sense that finite bounds are not explicitly provided in the data of the problem, although they may be implied by the other constraints).
@@
 Some of these reformulation techniques will be commented later on; here we remark that, for instance, integrality requirements, in particular in the form of binary variables could always be ``hidden'' by introducing (nonconvex) quadratic constraints utilizing the celebrated relationship $x_j \in \{0, 1\} \iff x_j^2 = x_j$.
-Therefore, when discussing these problems an effort has to be made to distinguish between features that come from the original model, and those that can be introduced by reformulation techniques in order to extract (and algorithmically exploit) specific properties.
+Therefore, when discussing these problems some effort has to be made to distinguish between features that come from the original model, and those that can be introduced by reformulation techniques in order to extract (and algorithmically exploit) specific properties.
 

### 2017-02-03 QPLIB-2.2.tex
small adjustments to improve the English, etc

@@
 
-Despite the apparent simplicity of the definition given in \S\ref{subsec:notation}, Quadratic Programming instances can be of several rather different ``types'' in practice, depending on the fine details of the data.
+Despite the apparent simplicity of the definition given in \S\ref{subsec:notation}, Quadratic Programming instances can be of several rather different ``types'' in practice, depending on fine details of the data.
 In particular, many algorithmic approaches can only be applied to QP when the data of the problem has specific properties.
-A taxonomy of QP instances should therefore strive to identify the set of properties that an instance should have in order to apply the most relevant computational methods.
-However, the sheer number of different existing approaches, and the fact that new ones are continuously proposed, makes it hard to provide a taxonomy that is both simple and covers all possible special cases.
+A taxonomy of QP instances should thus strive to identify the set of properties that an instance should have in order to apply the most relevant computational methods.
+However, the sheer number of different existing approaches, and the fact that new ones are frequently proposed, makes it hard to provide a taxonomy that is both simple and covers all possible special cases.
 This is why, in this paper, we propose an approach that aims at finding a good balance between simplicity and coverage of the main families of computational methods.
@@
 
-Our taxonomy is based on a three-fields code of the form ``\textit{OVC}'', where \textit{O} indicates the objective function, \textit{V} the variables, and \textit{C} the constraints of the problem.
+Our taxonomy is based on a three-fields code of the form ``\textit{OVC}'', where \textit{O} indicates the type of objective function considered, \textit{V} records the types of variables, and \textit{C} designates the types of constraints imposed on the variables.
 The fields can be given the following values:
@@
 \begin{itemize}
- \item objective function: (L)inear, (D)iagonal convex (if minimization) or concave (if maximization) quadratic, (C)onvex (if minimization) or (C)oncave (if maximization) quadratic,  (Q)uadratic (all other cases);
+ \item objective function: (\textit{L})inear, (\textit{D})iagonal convex (if minimization) or concave (if maximization) quadratic, (\textit{C})onvex (if minimization) or (\textit{C})oncave (if maximization) quadratic,  (\textit{Q})uadratic (all other cases);
  %
- \item variables: (C)ontinuous only, (B)inary only, (M)ixed binary and continuous, (I)nteger (incl.\ binary) only, (G)eneral (all other cases);
+ \item variables: (\textit{C})ontinuous only, (\textit{B})inary only, (\textit{M})ixed binary and continuous, (\textit{I})nteger (including binary) only, (\textit{G})eneral (all other cases);
  %
- \item constraints: (N)one, (B)ox, (L)inear, (D)iagonal convex quadratic, (C)onvex quadratic, nonconvex (Q)uadratic.
+ \item constraints: (\textit{N})one, (\textit{B})ox, (\textit{L})inear, (\textit{D})iagonal convex quadratic, (\textit{C})onvex quadratic, nonconvex (\textit{Q})uadratic.
 \end{itemize}
 %
-The wildcard ``*'' will be used to indicate any possible choice, and lists of the form ``\{X, Y, Z\}'' will indicate that the value of the given field can freely attain any of the specified values.
+The wildcard ``*'' will be used below to indicate any possible choice, and lists of the form ``\{\textit{X}, \textit{Y}, \textit{Z}\}'' will indicate that the value of the given field can freely attain any of the specified values.
 
 The ordering of the values in the previous lists is not irrelevant; in general, problems become ``harder'' when going from left to right.
-More specifically, for the \textit{O} and \textit{C} fields the order is that of \emph{strict} containment between problem classes: for instance, Linear objective functions are strictly contained in Diagonal convex quadratic ones (by just allowing the diagonal elements to be all-zero), which are strictly contained into general Convex quadratic ones (by allowing the off-diagonal elements to be all-zero), which in turn are strictly contained into general nonconvex Quadratic ones (by allowing any symmetric $Q^0$, hence possibly positive semidefinite ones as well).
+More specifically, for the \textit{O} and \textit{C} fields the order is that of \emph{strict} containment between problem classes: for instance, linear objective functions are strictly a special case of diagonal convex quadratic ones (by allowing the diagonal elements all to be zero), the latter are a strict subset of general convex quadratic objectives (by allowing the off-diagonal elements all to be zero), and these are strictly subsets of general nonconvex quadratic ones (since these permit any symmetric Hessian including positive semidefinite ones).
 The only field for which the containment relationship is not a total order is \textit{V}, for which only the partial orderings
 \[
- \mbox{C} \subset \mbox{M} \subset \mbox{G}
- \qquad,\qquad
- \mbox{B} \subset \mbox{M} \subset \mbox{G}
- \qquad,\qquad
- \mbox{B} \subset \mbox{I} \subset \mbox{G}
+ \mbox{\textit{C}} \subset \mbox{\textit{M}} \subset \mbox{\textit{G}},
+ \qquad
+ \mbox{\textit{B}} \subset \mbox{\textit{M}} \subset \mbox{\textit{G}},
+ \qquad
+ \mbox{and}
+ \qquad
+ \mbox{\textit{B}} \subset \mbox{\textit{I}} \subset \mbox{\textit{G}}
 \]
@@
 
-We will now provide a first general discussion about the different problem classes that our proposed taxonomy defines.
+We now give a general discussion about the different problem classes that our proposed taxonomy defines.
 For simplicity, we will assume minimization problems for the remaining of this section.
@@
 \]
-Hence, their solution only requires the independent minimization of a convex quadratic univariate function in each single variable $x_j$ over a box constraint and possibly integrality requirements, which can be attained trivially in $O(1)$ (per variable) by closed-form formul{\ae}, projection and rounding arguments.
-\emph{A fortiori}, the even simpler cases L*B, D*N and L*N (the latter obviously unbounded unless $b^0 = 0$) will not be discussed here.
-Similarly, CCN are immediately solved by linear algebra techniques, and therefore are of no interest in this context.
-On the other end of the spectrum, in general QP is a hard problem.
-Actually, LIQ---linear objective function and quadratic constraints in integer variables with no finite bounds, i.e.
+Hence, their solution only requires the independent minimization of a convex quadratic univariate function in each single variable $x_j$ over a box constraint and possibly integrality requirements, which can be attained trivially in $O(1)$ operations (per variable) by closed-form formul{\ae}, projection and rounding arguments.
+\emph{A fortiori}, the even simpler cases \textit{L*B}, \textit{D*N} and \textit{L*N} (the latter obviously unbounded unless $b^0 = 0$) will not be discussed here.
+Similarly, \textit{CCN} are immediately solved by linear algebra techniques, and therefore are of no interest in this context.
+At the other end of the spectrum, in general QP is a hard problem.
+Actually, \textit{LIQ}---linear objective function and quadratic constraints in integer variables with no finite bounds, i.e.
 \[
@@
 \]
-is not only $\mathcal{NP}$-hard, but downright undecidable \cite{jeroslow}. Hence so are the ``harder'' \{C,Q\}IQ.
+is not only $\mathcal{NP}$-hard, but downright undecidable \cite{jeroslow}. Hence so are the ``harder'' \{\textit{C},\textit{Q}\}\textit{IQ}.
 
 \smallskip
-It is important to note that the relationships between the different classes can be somehow blurred because reformulation techniques may allow to move one instance from one class to the other.
-The already recalled fact that $x^2 = x \iff x \in \{0, 1\}$, for instance, says that *M*---instances with only binary and continuous variables---can be recast as *CQ: nonconvex quadratic constraints can always take the place of binary variables.
-Actually, this is also true for *G* as long as $\Mcu = \emptyset$, as bounded general integer variables can be represented by binary ones.
+It is important to note that the relationships between the different classes can be somehow blurred because reformulation techniques may allow one to move an instance from one class to another.
+We already mentioned that $x^2 = x \iff x \in \{0, 1\}$, and in general \textit{*M*}---instances with only binary and continuous variables---can be recast as \textit{*CQ}; here nonconvex quadratic constraints take the place of binary variables.
+More generally, this is also true for \textit{*G*} as long as $\Mcu = \emptyset$, as bounded general integer variables can be represented by binary ones.
 
-Another relevant reformulation trick concerns the fact that, as soon as quadratic constraints are allowed, then a linear objective function can be assumed w.l.o.g..
-Indeed, any Q** (C*C) problem can always be rewritten as
+Another relevant reformulation trick concerns the fact that, as soon as quadratic constraints are allowed, then there is no loss of generality in assuming a linear objective function.
+Indeed, any \textit{Q**} (\textit{C*C}) problem can always be rewritten as
 %
@@
  %
- & -\infty \leq \half x^\top Q^0 x + b^0 x \leq x^0 & \\
+ & \hspace*{-5mm} -\infty \leq \half x^\top Q^0 x + b^0 x \leq x^0 & \\
  %
@@
 %
-i.e., a L*Q (L*C). Hence, it is clear that quadratic constraints are, in a well-defined sense, the most general situation (cf.~also the result above about hardness of LIQ).
+i.e., a \textit{L*Q} (\textit{L*C}) pronlem.
+Hence, it is clear that quadratic constraints are, in a well-defined sense, the most general situation (cf.~also the result above about hardness of \textit{LIQ}).
 
-When a $Q^i$ is positive semidefinite (PSD), i.e., the corresponding constraint/objective function is convex, general quadratic constraints are in fact equivalent to diagonal ones.
-In fact, since every PSD matrix can be factorized as $Q^i = L^i (L^i)^\top$, e.g.~by the (incomplete) Cholesky factorization, the term $\half x^\top Q^i x \equiv \half \sum_{j \in \Mcn} z^2_j$ where $z^\top = x^\top L^i$.
-Hence, one could think that D** problems need not be distinguished from C** ones; however, this is true only for ``complicated'' constraints, but not for ``simple'' ones, because the above reformulation technique introduces linear constraints.
-Indeed, while C*L (and, a fortiori, C*\{C,Q\}) can always be brought to D*L (D*\{C,Q\}), using the same technique C*B becomes D*L, which is significantly different from D*B.
-In practice, a diagonal convex objective function under linear constraints is found in many applications (e.g., \cite{FrFG16,FrGe06a,FrGe09a,FGGP11}), so that it still makes sense to distinguish the D*L case where the objective function is ``naturally'' separable from that where separability is artificially introduced.
+When a $Q^i$ is positive semidefinite (PSD), i.e., the corresponding constraint/objective function is convex, general Hessians are in fact equivalent to diagonal ones.
+In particular, since every PSD matrix can be factorized as $Q^i = L^i (L^i)^T$, e.g.~by the (incomplete) Cholesky factorization, the term $\half x^T Q^i x \equiv \half \sum_{j \in \Mcn} z^{i\;2}_j$ where $z^{i\;T} = x^T L^i$.
+Hence, one might maintain that D** problems need not be distinguished from C** ones.
+However in reality, this is only true for ``complicated'' constraints but not for ``simple'' ones, because the above reformulation technique introduces additional linear constraints, $L^{i\;T} x - z^i = 0$.
+Indeed, while \textit{C*L} (and, a fortiori, \textit{C*}\{\textit{C},\textit{Q}\}) can always be brought to \textit{D*L} (\textit{D*}\{\textit{C},\textit{Q}\}), using the above technique \textit{C*B} becomes \textit{D*L}, which may be significantly different from \textit{D*B}.
+In practice, a diagonal convex objective function under linear constraints is found in many applications (e.g., \cite{FrFG16,FrGe06a,FrGe09a,FGGP11}), so that it still makes sense to distinguish the \textit{D*L} case where the objective function is ``naturally'' separable from that where separability is artificially introduced.
 
@@
 \begin{itemize}
- \item \emph{Linear Programs} LCL and \emph{Mixed-Integer Linear Programs} LGL have been subject of an enormous amount of research and have their well-established instance libraries \cite{Koch2011}, so they won't be explicitly addressed here.
+ \item \emph{Linear Programs} \textit{LCL} and \emph{Mixed-Integer Linear Programs} \textit{LGL} have been subject of an enormous amount of research and have their well-established instance libraries \cite{Koch2011}, so they will not be explicitly addressed here.
  %
- \item \emph{Convex Continuous Quadratic Programs} CCC can be solved in polynomial time by Interior-Point techniques \cite{Wright97}; the simpler CCL can also be solved by means of ``simplex-like'' techniques, usually referred to as active-set methods \cite{Dost09}.
+ \item \emph{Convex Continuous Quadratic Programs} \textit{CCC} can be solved in polynomial time by Interior-Point techniques \cite{Wright97}; the simpler \textit{CCL} can also be solved by means of ``simplex-like'' techniques, usually referred to as active-set methods \cite{Dost09}.
 Actually, a slightly larger class of problems can be solved with Interior-Point methods: those that can be represented by Second-Order Cone Programs.
-When written as QPs the corresponding $Q^i$ may not be positive semidefinite, but still the problems can be efficiently solved.
-Of course, these problems can still require considerable time, like LCL, when the size of the instance grows.
-In this sense, like in the linear case, a significant divide is from solvers that need all the data of QP to work, and those that are ``matrix-free'', i.e., only require the solution of simple operations (typically, matrix-vector products) with the data of the problem to work.
-While in our library we have never exploited such a characteristic, which is not suitable to the use of standard modeling tools, this may be relevant for the solution of very-large-scale CIC.
+When written as QPs the corresponding $Q^i$ may not be positive semidefinite, but nonetheless such problems can be efficiently solved.
+Of course just as for \textit{LCL}, these problems may still require considerable computational effort when the size of the instance grows.
+In this sense, like in the linear case, there is a significant distinction between solvers that need all the data of QP to work, and those that are ``matrix-free'', i.e., only require the application of simple operations (typically, matrix-vector products) with the problem data.
+While when building our instance library we never exploited such characteristics, since they are not amenable to standard modelling tools, but this may be relevant for the solution of very-large-scale \textit{CIC}.
  %
- \item \emph{Nonconvex Continuous Quadratic Programs} QCQ are instead in general $\mathcal{NP}$-hard, even if the constraints are very specific (QCB) and only one eigenvalue of $Q^0$ is negative \cite{Hemmecke2010}.
-They therefore require enumerative techniques like the spatial Branch\&Bound \cite{FV90,BeLeLiMaWa08}, to be solved to optimality.
-Of course, local approaches are available that are able to efficiently provide saddle points (hopefully, local optima) of the CQC, but providing global guarantees about the quality of the obtained solutions is challenging.
-In our library we have specifically focussed on  \emph{exact} solution of the instances.
+ \item \emph{Nonconvex Continuous Quadratic Programs} \textit{QCQ} are generally $\mathcal{NP}$-hard, even if the constraints are very specific (\textit{QCB}) and only a single eigenvalue of $Q^0$ is negative \cite{Hemmecke2010}.
+They therefore require enumerative techniques, such as spatial Branch\&Bound \cite{FV90,BeLeLiMaWa08}, to be solved to optimality.
+Of course, local approaches are available that are able to efficiently provide saddle points (hopefully, local optima) of the \textit{CQC}, but providing global guarantees about the quality of the obtained solutions is challenging.
+In our library we have specifically focused on  \emph{exact} solution of the instances.
   %
- \item \emph{Convex Integer Quadratic Programs} CGC are in general $\mathcal{NP}$-hard, and therefore require enumerative techniques to be solved.
-However, convexity of the objective function and constraints implies that efficient techniques (see CCC) can be used at least to solve the continuous relaxation.
-The general view is that CGC are not, all other things being equal,  substantially more difficult than LGL to solve, especially if the objective function and/or the constraints have specific properties (e.g., DGL, CGL).
-Often integer variables are in fact binary ones, so several CCC models are C\{B,M\}C ones.
+ \item \emph{Convex Integer Quadratic Programs} \textit{CGC} are, in general, $\mathcal{NP}$-hard, and therefore require enumerative techniques to be solved.
+However, convexity of the objective function and constraints implies that efficient techniques (see \textit{CCC}) can be used at least to solve continuous relaxations.
+The general view is that \textit{CGC} are not, all other things being equal,  substantially more difficult than \textit{LGL} to solve, especially if the objective function and/or the constraints have specific properties (e.g., \textit{DGL}, \textit{CGL}).
+Often integer variables are in fact binary ones, so several \textit{CCC} models are \textit{C}\{\textit{B},\textit{M}\}\textit{C} ones.
 In practice binary variables are considered to lead to somewhat easier problems than general integer ones (cf.~the cited result about hardness of unbounded integer quadratic programs), and several algorithmic techniques have been specifically developed for this special case.
-However, the general approaches for CBC are basically the same as for CGC, so there is seldom the need to distinguish between the two classes as far as solvability is concerned, although matters can be different regarding actual solution cost.
-Programs with only binary variables CBC can be easier than mixed-binary or integer ones C\{M,I\}C because some techniques are specifically known for the binary-only case, cf.~the next point \cite{BEP09}.
-Absence of continuous variables, even in the presence of integer ones CIC, can also lead to specific techniques \cite{BEL13}.
+However, the general approaches for \textit{CBC} are basically the same as for \textit{CGC}, so there is seldom the need to distinguish between the two classes as far as solvability is concerned, although matters can be different regarding actual solution cost.
+Programs with only binary variables \textit{CBC} can be easier than mixed-binary or integer ones \textit{C}\{\textit{M},\textit{I}\}\textit{C} because some techniques are specifically known for the binary-only case, cf.~the next point \cite{BEP09}.
+Absence of continuous variables, even in the presence of integer ones \textit{CIC}, can also lead to specific techniques \cite{BEL13}.
   %
-  \item \emph{Nonconvex Binary Quadratic Programs} QB\{B, N, L\} obviously are $\mathcal{NP}$-hard.
-However, the special nature of binary variables combined with quadratic forms allows for quite specific techniques to be developed, among which is the reformulation of the problem as a LBL.
+  \item \emph{Nonconvex Binary Quadratic Programs} \textit{QB}\{\textit{B}, \textit{N}, \textit{L}\} obviously are $\mathcal{NP}$-hard.
+However, the special nature of binary variables combined with quadratic forms allows for quite specific techniques to be developed, one of which is the reformulation of the problem as a \textit{LBL}.
 Also, many well-known combinatorial problems can be naturally reformulated as problems of this class, and therefore a considerable number of results have been obtained by exploiting specific properties of the set of constraints \cite{Rendl2008,Loiola2007657}.
   %
-  \item \emph{Nonconvex Integer Quadratic Programs} QGQ is the most general, and therefore is the most difficul, class.
-Due to the lack of convexity even when integrality requirements are removed, solution methods must typically combine several algorithmic ideas, such as enumeration (distinguishing the role of integral variables from that of continuous ones involved into nonconvex terms) and techniques (e.g., outer approximation, semidefinite programming relaxation, \ldots) that allow to efficiently compute bounds.
-As in the convex case, QBQ, QMQ, and QIQ can benefit from more specific properties of the variables \cite{Buchheim2013,Dong2016}.
+  \item \emph{Nonconvex Integer Quadratic Programs} \textit{QGQ} is the most general, and therefore is the most difficult, class.
+Due to the lack of convexity even when integrality requirements are removed, solution methods must typically combine several algorithmic ideas, such as enumeration (distinguishing the role of integral variables from that of continuous ones involved into nonconvex terms) and techniques (e.g., outer approximation, semidefinite programming relaxation, \ldots) that allow the efficient computation of bounds.
+As in the convex case, \textit{QBQ}, \textit{QMQ}, and \textit{QIQ} can benefit from more specific properties of the variables \cite{Buchheim2013,Dong2016}.
 \end{itemize}
 %
-This description is purposely coarse; each of these classes can be subdivided into several others on the grounds of more detailed information about structures present in their constraints/objective function.
+This description is deliberately coarse; each of these classes can be subdivided into several others on the grounds of more detailed information about structures present in their constraints/objective function.
 These can have a significant algorithmic impact, and therefore can be of interest to researchers.
-Common structures are, e.g., network flow \cite{FrFG16,FrGaSc14,FrGaSt16a,FGGP11,tadayon-smith:2013} or knapsack-type linear constraints \cite{FrFG16,FrGo13b,FGGP11}, semi-continuous variables \cite{FrFG16,FrGaSc14,FrGaSt16a,FrGe06a,FrGe07a,FrGe09a,FGGP11}, or the fact that a nonconvex quadratic objective function/constraint can be reformulated as a second-order cone (hence, convex) one \cite{FrFG16,FrGaSc14,FrGaSt16a,FrGe09a,FGGP11}.
-It would be rather hard to collect a comprehensive list of all types of structures that may be of interest to any individual researcher, since these are as varied as the different possible approaches for specialized sub-classes of QP.
-For this reason we do not attempt such a more refined classification, and limit ourselves to the coarser one described in this paragraph.
+Common structures are, e.g., network flows \cite{FrFG16,FrGaSc14,FrGaSt16a,FGGP11,tadayon-smith:2013} or knapsack-type linear constraints \cite{FrFG16,FrGo13b,FGGP11}, and semi-continuous variables \cite{FrFG16,FrGaSc14,FrGaSt16a,FrGe06a,FrGe07a,FrGe09a,FGGP11}, or the fact that a nonconvex quadratic objective function/constraint can be reformulated as a second-order cone (hence, convex) one \cite{FrFG16,FrGaSc14,FrGaSt16a,FrGe09a,FGGP11}.
+It would be very hard to collect a comprehensive list of all types of structures that might be of interest to any individual researcher, since these are as varied as the different possible approaches for specialized sub-classes of QP.
+For this reason we do not attempt such a more refined classification, and limit ourselves to the coarser one described in this section.
 

### 2017-02-03 methods.tex
very minor adjustments to improve the English, etc.

@@
 
-In this section we provide a quick overview of existing solution methods for QP, restricting ourselves to these implemented by the set of solvers considered in this paper (\S\ref{subsec:solver}).
+In this section we provide a quick overview of existing solution methods for QP, restricting ourselves to these implemented by the set of solvers considered in this paper (see \S\ref{subsec:solver}).
 For each approach we briefly describe the main algorithmic ideas and point out the formulation they address according to the classification set out in \S\ref{sec:classification}.
-We remark that many solvers implement more than one algorithm, among which the user can choose at runtime.
-Moreover, algorithms are typically implemented by different solvers in different ways, so that the same conceptual algorithm can sometimes yield wildly different results or performance measures on the same instances.
+We remark that many solvers implement more than one algorithm, from which the user can choose at runtime.
+Moreover, algorithms are typically implemented in different ways within different solvers, so that the same conceptual algorithm can sometimes yield wildly different results or performance measures on the same instances.
 
-Solution methods for QP can, following \cite{neumaier}, be broadly organized in three categories: \emph{incomplete}, \emph{asymptotically complete}, \emph{complete}, and \emph{rigorous}.
-Incomplete methods are only able to identify solutions, often locally optimal according to a suitable notion, and may even fail to find one even when one exists; in particular, they are typically not able to determine that an instance is empty.
+Solution methods for QP can, following \cite{neumaier}, be broadly organized in four categories: \emph{incomplete}, \emph{asymptotically complete}, \emph{complete}, and \emph{rigorous}.
+Incomplete methods are only able to identify solutions, often locally optimal according to a suitable notion, and may even fail to find one even when one exists; in particular, they are typically not able to determine that an instance
+%is empty.
+has no solution.
 Asymptotically complete methods can find a globally optimal solution with probability one in infinite time, but again they cannot prove that a given instance is infeasible.
-Complete methods find an approximate globally optimal solution within a prescribed optimality tolerance within finite time, or prove that none such exists (but see Sect.~\ref{s:complete} below); they are often referred to as \emph{exact} methods in the computational optimization community.
+Complete methods find an approximate globally optimal solution within a prescribed optimality tolerance within finite time, or prove that none such exists (but see \S\ref{s:complete} below); they are often referred to as \emph{exact} methods in the computational optimization community.
 Finally, rigorous methods find globally optimal solutions within given tolerances even in the presence of rounding errors, except for ``near-degenerate cases''.
@@
 However, these three categories may exhibit some overlap.
-For example, any deterministic method for solving QCQ locally is incomplete in general, but becomes complete for CCC, since any local optimum of a convex QP is also global.
+For example, any deterministic method for solving \textit{QCQ} locally is incomplete in general, but becomes complete for \textit{CCC}, since any local optimum of a convex QP is also global.
 Therefore, when we state that a given algorithm is incomplete or (asymptotically) complete we mean that it is so the largest problem class that the solver naturally targets, although it may be complete on specific sub-classes.
-For example, interior point algorithms naturally target NLPs and are incomplete on NLPs, and therefore on QCQ, but become complete for CCC.
-In general, all complete methods for a problem $P$ must be complete for any problem $Q \subseteq P$, while a complete method for $P$ might be incomplete for $R \supset P$.
+For example, interior point algorithms naturally target NLPs and are incomplete on NLPs, and therefore on \textit{QCQ}, but become complete for \textit{CCC}.
+In general, all complete methods for a problem class $P$ must be complete for any problem class $Q \subseteq P$, while a complete method for $P$ might be incomplete for a class $R \supset P$.
 
@@
 
-Local search methods typically require as an input a solution $x'$ and attempt to improve it---either towards feasibility, or towards optimality, or both---using only information that is available in a neighborhood of $x'$.
+Local search methods typically require as an input an estimate $x'$ of the required  solution, and attempt to improve it---either towards feasibility, or towards optimality, or both---using only information that is available in a neighborhood of $x'$.
 In general, local search methods are incomplete.
-As remarked above, however, local search methods deployed on a convex problem behave like complete methods, possibly via devices that allow one to find a feasible starting solution if there is one (like what was used to be called ``phase -1'' in the simplex method).
+As remarked above, however, local search methods deployed on a convex problem behave like complete methods, possibly via devices that allow one to find a feasible starting estimate if there is one (like what used to be called ``phase-1'' in the simplex method).
 
-Most local search methods for *C* are iterative in nature, and need a starting point $x^0$ as input.
+Most local search methods for \textit{*C*} are iterative in nature, and need a starting point $x^0$ as input.
 The ($k+1$)-st iterate is obtained as $x^{k+1} = x^k + \alpha_k d^k$, where $\alpha_k$ (a scalar) is the \emph{step length}, and $d^k$ (a vector) is the \emph{search direction}, e.g.~belonging to a tangent manifold and having a negative directional derivative at $x^k$, in order to improving optimality while reducing infeasibility.
@@
 
-The solvers in Table \ref{t:solvers} which implement incomplete methods for NLPs (a problem class containing QCQ) are \conopt, \ipopt, \minos, \snopt, and \knitro.
+The solvers in Table \ref{t:solvers} which implement incomplete methods for NLPs (a problem class containing \textit{QCQ}) are \conopt, \ipopt, \minos, \snopt, and \knitro.
 Note that all these solvers tackle the more general class of NLP, while we use them only for the considerably more restricted class of QP.
-Aside from solvers provided by GAMS, there are a number of other, specialized incomplete QP solvers, such as
+Aside from solvers provided by GAMS, there are a number of other, specialized, incomplete QP solvers, such as
 \cqp \cite{GoulOrbaRobi13:mpc},
@@
 %
-Active set and interior point methods have been defined for *C* but also for general NLPs, and many of the solvers named above target this
+Active set and interior point methods have been defined for \textit{*C*} but also for general NLPs, and many of the solvers named above target this
 larger problem class.
@@
 %
-These approaches can be seen as recasting the original constrained problem QP as a parametrized family of unconstrained ones QP$_\mu$, where the constraints are moved in the objective function via a \emph{barrier term}---that goes to $+\infty$ as the boundary of the feasible region is approached---weighted with the \emph{barrier parameter} $\mu \in (0, \infty)$.
-In the convex case, the \emph{central path}---the continuous line formed from the optimal solutions of QP$_\mu$ for all varying $\mu$, typically unique e.g.~if the classical barrier based on the logarithmic function is employed---leads to a (central) optimal solution of QP when $\mu \to 0$.
+These approaches can be seen as recasting the original constrained problem QP as a parametrized family QP$_\mu$ of unconstrained ones, where the constraints are moved into the objective function via a \emph{barrier term}---that goes to $+\infty$ as the boundary of the feasible region is approached---weighted with the \emph{barrier parameter} $\mu \in (0, \infty)$.
+In the convex case, the \emph{central path}---the continuous line formed from the optimal solutions of QP$_\mu$ for all varying $\mu$, which is typically unique e.g.~if the classical barrier based on the logarithmic function is employed---leads to a (central) optimal solution of QP when $\mu \to 0$.
 Starting from an appropriately constructed ``central'' point (close to the solution of QP$_\mu$ for ``large'' $\mu$), these algorithms strive to follow the central path by performing $O(1)$ Newton steps before (substantially) reducing $\mu$.
-The algorithms can be shown to converge in a small number of iterations, each of which can be costly due to the need of solving an appropriately modified version of the KKT conditions for QP (``slackened'' with $\mu$), a (possibly) large-scale linear system.
-Actually, nowadays the algorithm is most often implemented in the primal-dual version, where the nonlinear KKT system is iteratively solved with Newton-like iterations; this has the extra advantage of allowing to remove the need of a feasible (central) starting point.
+The algorithms can be shown to converge in a small number of iterations, each of which can be costly due to the need of solving an appropriately perturbed version of the KKT conditions for QP (``perturbed'' with $\mu$), a (possibly) large-scale linear system.
+Nowadays the algorithm is most often implemented in the primal-dual version, where the nonlinear KKT system is iteratively solved with Newton-like iterations; this has the extra advantage of allowing to remove the need of a feasible (central) starting point.
 
@@
 However they rely heavily on projection, and thus are only suited to simple constraint geometries such as boxes and simplices.
+This is particularly useful for strictly-convex \textit{CCL} examples,
+since their duals may be reformulated exclusively to involve box constraints.
 
@@
  %
- \item multi-start (\knitro, \lgo, \msnlp, \oqnlp); specifically, the former three apply to QCQ whereas the latter to QGQ.
+ \item multi-start (\knitro, \lgo, \msnlp, \oqnlp); specifically, the former three apply to \textit{QCQ} whereas the latter to \textit{QGQ}.
 \end{itemize}
@@
 The method is asymptotically complete if the termination condition in Step \ref{step:ms2} is a certificate of global optimality for $x^\ast$, which is usually hard to obtain.
-However, typically some bound on the total CPU time, or number of function evaluations, or any other criteria that makes sense for the application at hand, is needed, which renders the method incomplete.
+In practice, typically some bound on the permitted total CPU time or number of function evaluations, or any other criteria that makes sense for the application at hand, is needed, which renders the method incomplete.
 
@@
 
-Complete methods are often referred to as \emph{exact} in a large part of the mathematical optimization community.
+%Complete methods are often referred to as \emph{exact} in a large part of the mathematical optimization community.
 This nomenclature has to be used with care, as it implicitly makes assumptions on the underlying computational model that may not be acceptable in all cases.
-To see that, consider that, as already mentioned, QPs (more precisely, LIQ) are generally undecidable \cite{jeroslow}; and yet, there exists a general decision method for deciding feasibility of systems of polynomial equations and inequalities \cite{tarski-reals}, including the solution of LCQ with zero objective function.
+To see that, consider that, as already mentioned, QPs (more precisely, \textit{LIQ}) are generally undecidable \cite{jeroslow}; and yet, there exists a general decision method for deciding feasibility of systems of polynomial equations and inequalities \cite{tarski-reals}, including the solution of \textit{LCQ} with zero objective function.
 This apparent contradiction is due to the fact that the two statements refer to different computational models: the former is based on the Turing Machine (TM), whereas the latter is based on the Real RAM (RRAM) machine \cite{blum}.
@@
 %
-Nearly all of the complete solvers in Table \ref{t:solvers} that address $\mathcal{NP}$-hard problems (i.e.~those in QGQ$\smallsetminus$CCC) are based on Branch-and-Bound (BB).
+Nearly all of the complete solvers in Table \ref{t:solvers} that address $\mathcal{NP}$-hard problems (i.e.~those in \textit{QGQ}$\smallsetminus$\textit{CCC}) are based on Branch-and-Bound (BB).
 This is an implicit but exhaustive search process based on exploring a \emph{branching tree} of the problem, where each node in the tree represents a subset of the feasible region.
@@
 When the BB algorithm is allowed to select coordinate directions corresponding to continuous variables, it is called \emph{spatial} BB (sBB).
-Branching on continuous (rather than integer or binary) variables becomes necessary in the presence of nonconvex nonlinearities, as it happens e.g.~in QCQ, since the quality of the bounds improves as the feasible set in the current node gets smaller.
+Branching on continuous (rather than integer or binary) variables becomes necessary in the presence of nonconvex nonlinearities, as it happens e.g.~in \textit{QCQ}, since the quality of the bounds improves as the feasible set in the current node gets smaller.
 
@@
 
-The following solvers from Table \ref{t:solvers} implement complete BB algorithms for QGQ or some subclasses:
+The following solvers from Table \ref{t:solvers} implement complete BB algorithms for \textit{QGQ} or some subclasses:
 %
 \begin{itemize}
- \item \antigone, \baron, \couenne, \lindo, \scip for QGQ;
+ \item \antigone, \baron, \couenne, \lindo, \scip for \textit{QGQ};
  %
-\item \cplex for QGL and CGC;
+\item \cplex for \textit{QGL} and \textit{CGC};
 %
- \item \knitro, \bonmin, \sbb, \xpress, \gurobi, and \mosek for CGC.
+ \item \knitro, \bonmin, \sbb, \xpress, \gurobi, and \mosek for \textit{CGC}.
 \end{itemize}
 %
-We remark that the latter category can be used as incomplete solvers for QGQ.
+We remark that the latter category can be used as incomplete solvers for \textit{QGQ}.
 %We also remark that CPLEX can currently only target problems with linear constraints when the objective function is nonconvex \cite{bliek}.
-We also remark that \lgo implements an incomplete BB algorithm for QCQ by using bounds obtained from sampling.
+We also note that \lgo implements an incomplete BB algorithm for \textit{QCQ} by using bounds obtained from sampling.
 
@@
 %
-Cutting plane approaches construct and iteratively improve a MILP (LIL) relaxation of the problem \cite{DuGr86,WePoe02}.
+Cutting plane approaches construct and iteratively improve a MILP (\textit{LIL}) relaxation of the problem \cite{DuGr86,WePoe02}.
 The cutting planes for the MILP are generated by linearization (first-order Taylor approximation) of the nonlinearities.
@@
 Additionally, incomplete methods can be used to provide local solutions.
-Therefor, these methods are complete on CGC if a complete method is used to solve the MILP.
+Therefor, these methods are complete on \textit{CGC} if a complete method is used to solve the MILP.
 The latter is typically based on BB, which is therefore a crucial technique also for this class of approaches.
 
-Solvers in Table \ref{t:solvers} that implement complete cutting plane methods for CGC are \alphaecp, \bonmin (in the algorithmic mode B-OA), and \dicopt.
+Solvers in Table \ref{t:solvers} that implement complete cutting plane methods for \textit{CGC} are \alphaecp, \bonmin (in the algorithmic mode B-OA), and \dicopt.
 

### 2017-02-03 solvers.tex
add some more references to solver papers
Add files via upload
put back todos that haven't been discussed yet
moved todos into #2, #3, #4
very minor adjustments to improve the English, etc.

@@
 For more details on the solvers, we refer to the given references, solver manuals, and the survey \cite{BuVi10}.
-In the table, we mark a pair (solver, problem) with ``I'' if the solver accepts the problem in input but it is an incomplete solver for the problem, with ``A'' if it is asymptotically complete, with ``C'' if it is complete, and leave it blank if the solver won't accept the problem in input.
+In the table, we mark a pair (solver, problem) with ``I'' if the solver accepts the problem as input but it is an incomplete solver for the problem, with ``A'' if it is asymptotically complete, with ``C'' if it is complete, and leave it blank if the solver won't accept the provided problem.
 When a solver implements several algorithms, we have chosen, for each problem class, the algorithm that potentially provides the ``strongest'' results (``C'' $>$ ``A'' $>$ ``I'' $>$ blank).
@@
 \bonmin      & \cite{BBCCGLLLMSW} &  C  &  I  &  C  &  I  &  C  &  I  \\
-\conopt      &  &     &     &     &     &  C  &  I  \\
+\conopt      & \cite{Drud1985,Drud1994} &     &     &     &     &  C  &  I  \\
 \couenne     & \cite{BeLeLiMaWa08} &  C  &  C  &  C  &  C  &  C  &  C  \\
-\cplex       & &  C  &  C  &  C  &     &  C  &     \\
+\cplex       & \cite{Bixby2000,cplex} &  C  &  C  &  C  &     &  C  &     \\
 \dicopt      & \cite{DuGr86,KoGr89,ViGr90} &  C  &  I  &  C  &  I  &  C  &  I  \\
 \gurobi      & \cite{Ro12} &  C  &     &  C  &     &  C  &     \\
-\ipopt       & &     &     &     &     &  C  &  I  \\
+\ipopt       & \cite{WaBi06} &     &     &     &     &  C  &  I  \\
 \knitro      & \cite{ByNoWa06} &  C  &  I  &  C  &  I  &  C  &  A  \\
 \lindo       & \cite{LiSc09} &  C  &  C  &  C  &  C  &  C  &  C  \\
-\lgo         & &     &     &     &     &  A  &  A  \\
-\minos       & &     &     &     &     &  C  &  I  \\
-\mosek       & &  C  &     &  C  &     &  C  &     \\
-\msnlp       & &     &     &     &     &  C  &  A  \\
-\oqnlp       & \cite{ULPGKM02} &  A  &  A  &  A  &  A  &  C  &  A  \\
-\sbb         &  &  C  &  I  &  C  &  I  &  C  &  I  \\
+\lgo         & \cite{Pinter1997,Pinter1998} &     &     &     &     &  A  &  A  \\
+\minos       & \cite{murtagh78,murtagh82} &     &     &     &     &  C  &  I  \\
+\mosek       & \cite{Andersen2000,Andersen2003} &  C  &     &  C  &     &  C  &     \\
+\msnlp       & \cite{ULPGKM02,LasdonPlummerUgrayBussieck2006} &     &     &     &     &  C  &  A  \\
+\oqnlp       & \cite{ULPGKM02,LasdonPlummerUgrayBussieck2006} &  A  &  A  &  A  &  A  &  C  &  A  \\
+\sbb         & \cite{sbb} &  C  &  I  &  C  &  I  &  C  &  I  \\
 \scip        & \cite{Ach09,VigerskeGleixner2016} &  C  &  C  &  C  &  C  &  C  &  C  \\
-\snopt       & &     &     &     &     &  C  &  I  \\
-\xpress      & &  C  &     &  C  &     &  C  &     \\
+\snopt       & \cite{GMS02,GMS05} &     &     &     &     &  C  &  I  \\
+\xpress      & \cite{xpress} &  C  &     &  C  &     &  C  &     \\
 \bottomrule

### 2017-02-03 QPLIB.tex
Update QPLIB.tex
Changed my name and affiliation; fixed the language in a couple of places in the abstract.

@@
         Hans Mittelmann \and
-        Nick Sahinidis \and
+        Nikolaos V Sahinidis \and
         Stefan Vigerske \and
@@
 %
-\and       Nick Sahinidis \at
-Chemical Engineering, Carnegie Mellon University, U.S.A. \email{sahinidis@cmu.edu}
+\and       Nikolaos V Sahinidis \at
+Departmet of Chemical Engineering, Carnegie Mellon University, Pittsburgh, PA 15213, U.S.A. \email{sahinidis@cmu.edu}
 %
@@
 \begin{abstract}
- This paper describes a new instance library for Quadratic Programming (QP), i.e., the family of continuous and (mixed)-integer optimization problems where either the objective function or the constrains (or both) are quadratic.
- QP is a very ``varied'' class, comprising sub-classes of problems ranging from
+ This paper describes a new instance library for Quadratic Programming (QP), i.e., the family of continuous and (mixed)-integer optimization problems where the objective function, the constrains, or both are quadratic.
+ QP is a very ``varied'' class of problems, comprising sub-classes of problems ranging from
  trivial to undecidable. Solution methods for QP are very diverse, ranging from
@@
  which both aspects are fundamental. Selecting a set of instances of QP that is
- at the same time not overwhelmingly numerous and sufficiently challenging for the
+ at the same time not overwhelmingly onerous but sufficiently challenging for the
  many different interested communities is therefore important. We propose a simple
@@
  We then briefly survey the field of QP, giving an overview of theory, methods and
- solvers. Finally we describe how the library was put together, and detail its final
+ solvers. Finally, we describe how the library was put together, and detail its final
  contents.

### 2017-02-03 QPLIB-1.0.tex
Update QPLIB-1.0.tex
Improved the english

@@
 
-Quadratic Programming (QP) problems---mathematical optimization problems for which either the objective function \cite{wiki:qp} or the constraints \cite{wiki:qcqp} (or both) are polynomial function of the variables of degree two---include a notably diverse set of different instances.
+Quadratic Programming (QP) problems---mathematical optimization problems for which the objective function \cite{wiki:qp}, the constraints \cite{wiki:qcqp}, or both are polynomial function of the variables of degree two---include a notably diverse set of different instances.
 This is not surprising, given the vast scope of practical applications of such problems, and of solution methods designed to solve them \cite{GoulToin00a}.
-Depending on the fine details of the formulation, solving a QP may require either fundamentally combinatorial techniques, or ideas rooted in nonlinear optimization principles, or a mix of the two.
+Depending on the specifics of the formulation, solving a QP may require primarily combinatorial techniques, ideas rooted in nonlinear optimization principles, or a mix of the two.
 In this sense, QP is arguably one of the classes of problems where collaboration between the communities interested in combinatorial and in nonlinear optimization is most needed, and potentially fruitful.
@@
 However, this diversity also implies that QP means very different things to different researchers.
-This is illustrated by the simple fact that the class of problems that we simply refer to here as ``QP'' might more accurately be called Mixed-Integer Quadratically-Constrained Quadratic Programming (MIQCQP) in the most general case.
+This is illustrated by the fact that the class of problems that we simply refer to here as ``QP'' might more accurately be called Mixed-Integer Quadratically-Constrained Quadratic Programming (MIQCQP) in the most general case.
 It is, therefore, perhaps not surprising that, unlike for ``simpler'' problems classes \cite{Koch2011},  there has been, to date, no single library devoted to all different kinds of instances of QP.
 While several specialised libraries devoted to particular cases of QP are available, each of them is either focussed on a particular application (a specific problem that can be modelled as a QP), or on QPs with specific structural properties that make them suitable for solution by some given class of algorithmic approaches.
-To the best of our knowledge, collecting a set of QP instances that is at the same time not overwhelmingly numerous and significant for the many different interested communities has not been attempted.
+To the best of our knowledge, collecting a set of QP instances that is at the same time not overwhelmingly onerous but sufficiently challenging for the many different interested communities has not been attempted.
 This work constitutes a first step in this direction.
 
-In this paper we report the steps that have been done to collect (we believe) a significant library of QP instances, filtering the a much larger set of currently available (or specifically provided) instances in order to end up with a manageable set that still contains a meaningful sample of all possible QP types.
-A particularly thorny issue in this process is how to select instances that are ``interesting''.
+In this paper, we report the steps that have been done to collect what we consider to be a quality library of QP instances, filtering a much larger set of currently available (or specifically provided) instances in order to end up with a manageable set that still contains a meaningful sample of possible QP types.
+A particularly thorny issue in this process was how to select instances that are ``interesting''.
 Our choice has been to take this to mean ``challenging for a significant set of solution methods''.
-Our filtering process has then been in part based on the idea that if a significant fraction of the solvers that can solve a QP instance do so in a ``short'' time, then the instance is not challenging enough to be included in the library.
-Conversely, if very few (maybe one) solvers can solve it very efficiently by exploiting some specific structure, but most other approaches cannot, then the instance should be deemed ``interesting''.
+Our filtering process has then been in part based on the idea that, if a significant fraction of the solvers that can solve a QP instance do so in a ``short'' time, then the instance is not challenging enough to be included in the library.
+Conversely, if very few (maybe one) of the solvers can solve it very efficiently by exploiting some specific structure, but most other approaches cannot, then the instance should be deemed ``interesting''.
 Putting this approach into practice requires a nontrivial number of technical steps and decisions that are detailed in the paper.
-We hope that our work can provide useful guidelines for other interested researchers.
+We hope that our work can provide useful guidelines for other researchers interested in the constructions of benchmarks for mathematical optimization problems.
 
-A consequence of our focus is that this paper is \emph{not} concerned with the different performance of the very diverse available set of QP solvers; we will \emph{not} report any data comparing them.
-The only reason that solvers are used (and, therefore, described) in this context is to ensure that the instances of the library are nontrivial at least for a significant fraction of the current solution methods.
+A consequence of our focus is that this paper is \emph{not} concerned with the performance of the very diverse available set of QP solvers; we will \emph{not} report any data comparing them.
+The only reason that solvers are used (and, therefore, described) in this context is to ensure that the instances of the library are nontrivial---at least for a significant fraction of the current solution methods.
 Providing guidance about which solvers are most suited to some specific class of QPs is entirely outside the scope of our work.
@@
 Optimization problems with quadratic constraints and/or objective function (QP) have been the subject of a considerable amount of research over the last almost seventy years.
-At least some of the rationale for this interest is likely due to the fact that they are the ``least-nonlinear nonlinear problems''.
+At least some of the rationale for this interest is likely due to the fact that QPs are the ``least-nonlinear nonlinear problems''.
 Hence, in particular for the convex case, tools and techniques that have been honed during decades of research for Linear Programming (LP), typically with integrality constraints (MILP), can often be extended to the quadratic case more easily than would be required to tackle general (Mixed-Integer) Nonlinear Programming ((MI)NLP) problems.
 This has indeed happened over-and-over again with different algorithmic techniques, such as interior-point methods, active-set methods (of which the simplex method is a prototypical example), enumeration methods, cut-generation techniques, reformulation techniques, and many others \cite{BS09}.
-Similarly, nonconvex continuous QP are perhaps the ``simplest'' class of problems that require features, such as spatial enumeration techniques, for their solution.
-Hence, they are both the natural basis for the development of general techniques for nonconvex NLP, and a very specific class so that specialized approaches can be developed \cite{Dur2010,Burer2012}.
+Similarly, nonconvex continuous QP is perhaps the ``simplest'' class of problems that require features such as spatial enumeration techniques for their solution.
+Hence, QPs are both a natural basis for the development of general techniques for nonconvex NLP, and a very specific class so that specialized approaches can be developed \cite{Dur2010,Burer2012}.
 
-In addition, QP, that is, quadratic problems with continuous or integer variables, is arguably a considerably more expressive class than (MI)LP.
+In addition, QP, with continuous or integer variables, is arguably a considerably more expressive class than (MI)LP.
 Quadratic expressions are found, either naturally or after appropriate reformulations, in very many optimization problems \cite{Kochenberger2014}.
-Table \ref{tbl:application} provides a certainly non-exhaustive collection of applications that lead to formulations with either quadratic constraints, or quadratic objective function, or both.
-Also, in general any continuous function can be approximated with arbitrary accuracy (over a compact set) by a polynomial of arbitrary degree.
+Table \ref{tbl:application} provides a certainly non-exhaustive collection of applications that lead to formulations with quadratic constraints, quadratic objective function, or both.
+In general, any continuous function can be approximated with arbitrary accuracy (over a compact set) by a polynomial of arbitrary degree.
 In turn, every polynomial can be broken down to a system of quadratic expressions.
-Hence, QP is, in some sense, roughly as expressive as the whole of MINLP.
-Of course this is, in principle, true for MILP as well, but at the cost of much larger and much more complicated formulations.
-Hence, for many applications QP may represent the ``sweet spot'' between the effectiveness, but lower expressive power, of MILP and the higher expressive power, but much higher computational cost, of MINLP.
+Hence, QP is, in some sense, roughly as expressive as MINLP.
+This is, in principle, true for MILP as well, but at the cost of much larger and much more complicated formulations.
+Hence, for many applications QP may represent the ``sweet spot'' between the effectiveness, but lower expressive power, of MILP and the higher expressive power, but much higher computational cost of MINLP.
 
@@
 %
-Quadratic Assignment Problem$^{\ddagger}$ & \checkmark & \cite{anstreicher:2003,loiola-etal:2007} \\ \cmidrule(rl){2-3}
+Quadratic assignment problem$^{\ddagger}$ & \checkmark & \cite{anstreicher:2003,loiola-etal:2007} \\ \cmidrule(rl){2-3}
 %
-Max-Cut & \checkmark & \cite{Rendl2008,KrislockMalickRoupin2016} \\ \cmidrule(rl){2-3}
+Max-cut & \checkmark & \cite{Rendl2008,KrislockMalickRoupin2016} \\ \cmidrule(rl){2-3}
 %
@@
 %
-\multicolumn{3}{l}{\textbf{Process Networks}} \\[1pt]
+\multicolumn{3}{l}{\textbf{Process networks}} \\[1pt]
 %
@@
 %
-Traveling Salesman Problem \\
-with Neighborhoods & \checkmark & \cite{gentilini-etal:2013} \\
+Traveling salesman problem \\
+with neighborhoods & \checkmark & \cite{gentilini-etal:2013} \\
 \midrule
@@
 %
-\multicolumn{3}{l}{\textbf{Data Confidentiality}} \\[1pt]
+\multicolumn{3}{l}{\textbf{Data confidentiality}} \\[1pt]
 %
@@
 In \S\ref{sec:QPbasic} we review the basic notion of QP.
-In particular, \S\ref{subsec:notation} sets out the notation, \S\ref{sec:classification} proposes a (to the best of our knowledge, new) taxonomy of QP that helps us in discussing the (very) different classes of QPs, and \S\ref{sec:algo} very briefly reviews the solution methods for QP and the solvers we have employed.
+In particular, \S\ref{subsec:notation} sets out the notation, \S\ref{sec:classification} proposes a new taxonomy of QP that helps us in discussing the (very) different classes of QPs, and \S\ref{sec:algo} very briefly reviews the solution methods for QP and the solvers we have employed.
 Next \S\ref{sec:lib} describes the process used to obtain the library and its results.
@@
 this paper, we refer to the comprehensive benchmark site \url{http://plato.asu.edu/bench.html}. Of the result on this site, three deal exclusively with
-QP problems, namely the (1) Large SOCP, (2) MISOCP, and the (3) MIQ(C)P benchmarks, while three others contain have partial results for such problems,
-namely those for (4) Parallel Barrier Solvers on Large LP/QP problems, (5) AMPL-NLP and (6) MINLP.
+QP problems, namely the (1) large SOCP, (2) MISOCP, and the (3) MIQ(C)P benchmarks, while three others contain have partial results for such problems,
+namely those for (4) parallel barrier solvers on large LP/QP problems, (5) AMPL-NLP and (6) MINLP.
 Benchmarks (1, 2 \& 4) contain only convex instances, while the others

### 2017-02-07 QPLIB-3.tex
corrected tiny typo
mention that there is now also ampl .mod available

@@
 new format is capable of describing all the instances of the library in a sparse form.
-In comparison to a more \emph{high level} format like \texttt{.gms}, the new
+In comparison to a more \emph{high level} and human readable format like \texttt{.gms}, the new
 format presents three advantages: it is easier to read by a self-made parser, it produces smaller files, and the order of variables and constraints is fixed.
@@
 
-Similarly, Figures~\ref{fig:pic_constr} (left)
+Similarly, Figure~\ref{fig:pic_constr} (left)
 %Figure~\ref{fig:pic_constr_medium}
@@
 %
-Beyond all instances in GAMS, \texttt{.lp}, and \texttt{.qplib} format, the website provides a rich set of metadata for each
+Beyond all instances in GAMS, \texttt{.lp}, AMPL (\texttt{.mod}), and \texttt{.qplib} format, the website provides a rich set of metadata for each
 instance: the three letter problem classification (as described in Section \ref{subsec:final set}), basic properties such as the number of variables and constraints of

### 2017-02-07 QPLIB-2.2.tex
fix one overfull hbox

@@
  %
- \item constraints: (\textit{N})one, (\textit{B})ox, (\textit{L})inear, (\textit{D})iagonal convex quadratic, (\textit{C})onvex quadratic, nonconvex (\textit{Q})uadratic.
+ \item constraints: (\textit{N})one, (\textit{B})ox, (\textit{L})inear, (\textit{D})iagonal convex quadratic, (\textit{C})on\-vex quadratic, nonconvex (\textit{Q})uadratic.
 \end{itemize}

### 2017-02-07 QPLIB-3.tex
more accurate description of the advantages of .qplib provided

@@
 
-In this section we present all the steps performed in order to build the
-new library. In \S\ref{subsec:instColl}, we describe the set of gathered
-instances. In \S\ref{subsec:selection} we present the features used to
-classify the instances and we discuss the issues concerning the format of
-the instances. Finally, in \S\ref{subsec:final set}, we describe the
-selection process used to filter the instances and we graphically present
-the main features of the selected instances.
+In this section we present all the steps we performed in order to build
+the new instance library. In \S\ref{subsec:instColl}, we describe the set
+of gathered instances, and
+in \S\ref{subsec:selection} we present the features used to
+classify the instances.
+%and discuss issues concerning the format of the instances.
+We describe the selection process used to filter the instances, and
+graphically present the main features of the selected instances in
+\S\ref{subsec:final set}, while in \S\ref{subsec:website}
+we provide information on how to access the test collection.
 
@@
 
-At the end of this process we gathered more than eight thousand
-instances. Three fourths of them contained discrete variables, while
-the remaining ones contained only continuous variables. More in details,
+At the end of this process we had gathered more than eight thousand
+instances. Three quarters of them contained discrete variables, while
+the remainder contained only continuous variables. In more detail,
 we gathered $\approx 1800$ Quadratic Binary Linear (QBL) instances,
@@
 and $\approx 2500$ Quadratic General Quadratic (QGQ) instances. We
-also gathered $\approx 1000$ Convex General Convex (CGC) instances. We
-gathered relatively fewer Quadratic Binary Quadratic (QBQ), Convex
-Continuous Convex (CCC) and Convex Mixed Convex (CMC) instances, i.e.,
-$\approx 150$ instances, $\approx 200$ instances and $\approx 200$
-instances respectively. Finally, we gathered only $17$ Quadratic Mixed
+also received $\approx 1000$ Convex General Convex (CGC) instances. We
+obtained relatively fewer Quadratic Binary Quadratic (QBQ), Convex
+Continuous Convex (CCC) and Convex Mixed Convex (CMC) instances,
+($\approx 150$, $\approx 200$ and $\approx 200$ instances respectively).
+Finally, we found only $17$ Quadratic Mixed
 Linear (QML) instances. In the call for instances, no specific formats
@@
 To evaluate the instances we decided, for practical reasons, to use
-GAMS as common platform for all the experiments involving commercial and
-non-commercial solvers. For this reason, we decided to translate all
-instances into the GAMS format (\texttt{.gms}).
+GAMS as common platform for all our final selection computations.
+%involving commercial and non-commercial solvers.
+For this reason, we translated all the
+instances we received into the GAMS format (\texttt{.gms}).
 % In a preliminary phase, all the instances received were divided
@@
 %
-In addition, we have introduced a new, specific \texttt{.qplib} format. This
-new format is capable of describing all the instances of the library in a sparse form.
-In comparison to a more \emph{high level} and human readable format like \texttt{.gms}, the new
-format presents three advantages: it is easier to read by a self-made parser, it produces smaller files, and the order of variables and constraints is fixed.
-See the \S\ref{sec:format} for more details.
 
@@
 
-For each instance of the starting set, we collected important characteristics
+For each instance in this large starting set,
+we collected important characteristics
 which allowed us to classify the instances into the QP categories described in
@@
 \begin{itemize}
- \item number of binary variables; % (\emph{\# bin}),
+ \item the number of binary variables; % (\emph{\# bin}),
  %
- \item number of integer variables; % (\emph{\# int}),
+ \item the number of integer variables; and % (\emph{\# int}),
  %
- \item number of continuous variables. % (\emph{\# cont}).
+ \item the number of continuous variables. % (\emph{\# cont}).
 \end{itemize}
@@
 \begin{itemize}
- \item percentage of positive and negative eigenvalues of the Hessian $Q^0$;
-       % (\emph{\% neg eig}),
+ \item the percentage of positive and negative eigenvalues of the Hessian
+      $Q^0$; and
+       % (\emph{\% neg eig}), and
  %
- \item density of the Hessian $Q^0$ (number of nonzero entries divided by the total
+ \item the density of the Hessian $Q^0$ (number of nonzero entries divided by the total
        number of entries). %(\emph{\% dens}).
@@
 
-The number of positive (i.e., larger than $10^{-16}$) and negative (i.e., smaller than $-10^{-16}$) eigenvalues of $Q^0$ allows us to identify the
+The number of positive (i.e., larger than $10^{-16}$) and negative (i.e., smaller than $-10^{-16}$) eigenvalues of $Q^0$ allowed us to identify the
 objective function type, as in presence of at least one negative (positive) eigenvalue
@@
 \begin{itemize}
- \item number of linear constraints, % (\emph{\# lin}),
+ \item the number of linear constraints, % (\emph{\# lin}),
  %
- \item number of quadratic constraints, % (\emph{\# quad}),
+ \item the number of quadratic constraints, % (\emph{\# quad}),
  %
- \item number of convex constraints, % (\emph{\# conv}),
+ \item the number of convex constraints, and % (\emph{\# conv}),
  %
- \item number of variable bounds (for non-binary variables). %(\emph{\# box}).
+ \item the number of variable bounds (for non-binary variables). %(\emph{\# box}).
 \end{itemize}
@@
 A constraint is considered quadratic if it contains at least one nonzero in
-the quadratic term. Among the quadratic constraints, the ones with only
-non-negative eigenvalues are classified as convex constraints.
-Note that this might lead to classify some instances that have conic constraints as non-convex ones, although their feasible region is in fact convex.
-However, only some solvers are capable of properly exploiting this property.
+a quadratic term (if present). Among the quadratic constraints, the ones whose
+Hessians have only non-negative eigenvalues (when $c_u^i < \infty$) and
+and non-positive eigenvalues (when $c_l^i > - \infty$)
+are classified as convex constraints; thus, a quadratic constraint with
+two sided, finite bounds is non-convex.
+Note that this might might occasionally lead us to classify some instances that have conic constraints as non-convex ones, although their feasible region is in fact convex---fortunately, only some solvers are capable of properly exploiting this property.
  All this information allowed us to analyse the gathered instances and to perform the
@@
 During the development of the library, a discussion ensued about
-the expected goals that we wanted to achieve. The following four goals
+the expected goals that we wished to achieve. The following four goals
 where finally identified:
@@
 \begin{enumerate}
- \item represent as much as possible all the different categories of QP
+ \item to represent as far as possible all the different categories of QP
        problems;
  %
- \item gather ``challenging'' instances, i.e., ones which can not be easily
+ \item to gather ``challenging'' instances, i.e., ones which can not be easily
        solved by  state-of-the-art solvers;
  %
- \item include for each of the categories a set of well-diversified
-       instances;
+ \item to include, for each of the categories, a set of well-diversified
+       instances; and
  %
- \item obtain a set of instances which is neither too small, so as to
+ \item to obtain a set of instances which is neither too small, so as to
        preserve statistical relevance, nor too large so as to being
@@
 %
-To achieve the aforementioned goals, we performed the following two
-filters, applied in cascade.
+To achieve such goals, we performed the following two filters, applied in
+a cascade.
 %
@@
 We now analyse the features of the instances selected to be part of the
-library. The characteristics of the instances are presented in Table
-\ref{tab:DD} for \emph{discrete} instances (*\{B,M,I,G\}*) and in Table
-\ref{tab:CC} for \emph{continuous} ones (*C*). For each category, the tables
+instance library. The characteristics of the instances are presented in Table
+\ref{tab:DD} for \emph{discrete} instances
+(\textit{*}\{\textit{B,M,I,G}\}\textit{*}) and in Table
+\ref{tab:CC} for \emph{continuous} ones (\textit{*C*}).
+For each category, the tables
 report in column ``$\#$'' the corresponding number of instances. It can be seen
@@
 instances among the different categories. Indeed, the discrete categories
-(LMQ) or (QBL) are well represented by $118$ and $59$ instances,
-respectively. Similarly, the continuous categories (LCQ) and (QCQ) are well
+(\textit{LMQ}) or (\textit{QBL}) are well represented by $118$ and $59$
+instances, respectively. Similarly, the continuous categories
+(\textit{LCQ}) and (\textit{QCQ}) are well
 represented by $50$ and $17$ instances, respectively. Moreover, the library
@@
 
-Speaking of nonconvexity, Figure~\ref{fig:pic_neg_eig} (left) focuses on
+On the theme of nonconvexity, Figure~\ref{fig:pic_neg_eig} (left) focuses on
 the instances with a quadratic objective function, ordered by the
@@
 
-Additional details on the instance features can be found in \S\ref{sec:instance_details}.
+Additional details on the instance features can be found in Appendix~\ref{sec:instance_details}.
 
@@
 
-\subsection{Website}
+\subsection{Website}\label{subsec:website}
 %Select subset or categories of instances (EXTRACT FROM THE LIBRARY A SUBSET OF INSTANCES WITH SPECIFIC CHARACTERISTICS)
 
-The instances of QPLIB are publicly accessible at the website \url{http://qplib.zib.de}, which was created by extending
+The QPLIB instances are publicly accessible at the website \url{http://qplib.zib.de}, which was created by extending
 scripts and tools initially developed for MINLPLib\,2~\cite{Vigerske2014}.
 %
-Beyond all instances in GAMS, \texttt{.lp}, AMPL (\texttt{.mod}), and \texttt{.qplib} format, the website provides a rich set of metadata for each
-instance: the three letter problem classification (as described in Section \ref{subsec:final set}), basic properties such as the number of variables and constraints of
+We provide all instances in GAMS (\texttt{.gms}), AMPL (\texttt{.mod}), CPLEX (\texttt{.lp})
+\cite{,cplex}, and  QPLIB (\texttt{.qplib}) formats. The latter is a new format
+specifically for QP instances. In comparison to more \emph{high level}
+formats such as \texttt{.gms} and \texttt{.lp}, the new
+format offers three main advantages: it is easier to read by a stand-alone
+parser (provided), it typically produces smaller files, and it permits the
+inclusion of two-sided inequalities without needless repetition of data.
+%the order of variables and constraints is fixed.
+See Appendix~\ref{sec:format} for more details.
+
+Beyond the example instances, the website provides a rich set of metadata for
+each instance: the three letter problem classification (as described in \S\ref{subsec:final set}), basic properties such as the number of variables and constraints of
 different types, the sense and convexity/concavity of the objective function, and information on the nonzero structure of the

### 2017-02-07 conclusion.tex
more accurate description of the advantages of .qplib provided

@@
  This manuscript describes the first comprehensive library of instances for Quadratic Programming (QP).
-Since QP comprises  different and ``varied'' categories of problems, we propose a classification and we briefly discuss the main classes of solution methods for QP.
+Since QP comprises  different and ``varied'' categories of problems, we proposed a classification and we briefly discuss the main classes of solution methods for QP.
 
 We then describe the steps of the adopted process used to filter the gathered instances  in order to build the new library.
-The goals we purposed are of different natures and we tried to build a library which is computationally challenging and as broad as possible, i.e., it represents the largest possible categories of QP problems, while remaining of manageable size.
+Our design goals were to build a library which is computationally challenging and as broad as possible, i.e., it represents the largest possible categories of QP problems, while remaining of manageable size.
+We have also proposed a stand-alone QP format that is intended
+for the convenient exchange and use of our QP instances.
 

### 2017-02-07 methods.tex
Update methods.tex
Added citations to BARON-related algorithmic papers
more accurate description of the advantages of .qplib provided

@@
  %
- \item multi-start (\knitro, \lgo, \msnlp, \oqnlp); specifically, the former three apply to \textit{QCQ} whereas the latter to \textit{QGQ}.
+ \item multi-start (\knitro, \lgo, \msnlp, \oqnlp); specifically, the first three apply to \textit{QCQ} and the last to \textit{QGQ}.
 \end{itemize}
@@
 This is actually where finite (and tight) bounds on the variables are crucial, which differentiate also in practice the bounded case from the unbounded one.
-Different strategies are used when the nonconvexities are only quadratic \cite{MiFl13,bliek}.
+Different strategies are used when the nonconvexities are only quadratic \cite{bst:09:oms,MiFl13,zs:13:oms,zs:13:iecr,bliek}.
 

### 2017-02-07 QPLIB-3.tex
fix duplicate word
- the qplib instances are not examples, they are *the* instances

@@
 two sided, finite bounds is non-convex.
-Note that this might might occasionally lead us to classify some instances that have conic constraints as non-convex ones, although their feasible region is in fact convex---fortunately, only some solvers are capable of properly exploiting this property.
+Note that this might occasionally lead us to classify some instances that have conic constraints as non-convex ones, although their feasible region is in fact convex---fortunately, only some solvers are capable of properly exploiting this property.
  All this information allowed us to analyse the gathered instances and to perform the
@@
 
-Beyond the example instances, the website provides a rich set of metadata for
+Beyond the instances, the website provides a rich set of metadata for
 each instance: the three letter problem classification (as described in \S\ref{subsec:final set}), basic properties such as the number of variables and constraints of

### 2017-02-07 QPLIB-2.2.tex
additional comment on the taxonomy

@@
  \item constraints: (\textit{N})one, (\textit{B})ox, (\textit{L})inear, (\textit{D})iagonal convex quadratic, (\textit{C})on\-vex quadratic, nonconvex (\textit{Q})uadratic.
+  Note that (\textit{D}) and (\textit{C}) are intended to mean that the feasible set of the constraint is convex, i.e., that either $Q^i$ is positive semidefinite and $c_l^i = -\infty$, or $Q^i$ is negative semidefinite and $c_u^i = \infty$.
+Quadratic constraints with both finite bounds cannot ever be convex (unless $Q^i = 0$, i.e., they are not ``truly'' quadratic constraints).
 \end{itemize}

### 2017-02-08 format.tex
make it more clear that indices are 1-based and only summed up number of quadratic/linear terms in constraints is given

@@
 the exponent may be preceded by either the character \texttt{D} or \texttt{E},
-e.g. \texttt{12.56D+2} or \texttt{12.56E+2}.
+e.g. \texttt{12.56D+2} or \texttt{12.56E+2}. Variable and constraint indices
+are \textbf{1-based}.
 
@@
 $q^0$ & constant part of the objective function & \\ \hline
-$n^{Q^i}$ & total number of nonzeros (integer) in lower triangles of
-$Q^i$  & [2,4] \\
+$\sum_{i\in\Mcm} n^{Q^i}$ & number of nonzeros (integer) in lower triangles of
+$Q^i$, sum up over all $i\in\Mcm$  & [2,4] \\
 $i$\; $h$\; $k$\; $Q^{i}_{hk}$
-& i, row and column indices (integers) and value (real) for
- each entry of $Q^{i}$ for every $i \in \Mcm$,  if $n^{Q^i}$,
+& $i$, row and column indices (integers) and value (real) for
+ each entry of $Q^{i}$ for every $i \in \Mcm$,  if $n^{Q^i}>0$,
  one quadruple on each line & \\
 \hline
-$n^{b^i}$ & total number of nonzeros  (integer) in $b^i$ for all $i \in \Mcm$ &
+$\sum_{i\in\Mcm} n^{b^i}$ & number of nonzeros (integer) in $b^i$, sum up over all $i \in \Mcm$ &
  [2] \\
-$i$\; $j$\; $b^i_{j}$ & i and index (integers) and value (real)
-for each nonzero entry of $b^i$ for every $i \in \Mcm$, if $n^{b} > 0$,
+$i$\; $j$\; $b^i_{j}$ & $i$ and index (integers) and value (real)
+for each nonzero entry of $b^i$ for every $i \in \Mcm$, if $n^{b^i} > 0$,
 one triple on each line & [2] \\

### 2017-02-08 format.tex
further tiny clarifications and corrections

@@
 the exponent may be preceded by either the character \texttt{D} or \texttt{E},
-e.g. \texttt{12.56D+2} or \texttt{12.56E+2}. Variable and constraint indices
-are \textbf{1-based}.
+e.g. \texttt{12.56D+2} or \texttt{12.56E+2}. Variable indices, $j$, must lie in
+the range $1 \leq j \leq n$, while constraint indices, $i$, must satisfy
+$1 \leq i \leq m$, that is they are both \textbf{one-based}. The case for
+character strings is irrelevant.
 
@@
 type & problem type (character string) & [1] \\
-sense & one of the words minimize or maximize (character string, case irrelevant) & \\
+sense & one of the words minimize or maximize (character string) & \\
 \hline
@@
 $q^0$ & constant part of the objective function & \\ \hline
-$\sum_{i\in\Mcm} n^{Q^i}$ & number of nonzeros (integer) in lower triangles of
-$Q^i$, sum up over all $i\in\Mcm$  & [2,4] \\
+$\displaystyle \sum_{i\in\Mcm} n^{Q^i}$ & number of nonzeros (integer) in lower triangles of
+$Q^i$, summed over all $i\in\Mcm$  & [2,4] \\
 $i$\; $h$\; $k$\; $Q^{i}_{hk}$
@@
 \hline
-$\sum_{i\in\Mcm} n^{b^i}$ & number of nonzeros (integer) in $b^i$, sum up over all $i \in \Mcm$ &
+$\displaystyle \sum_{i\in\Mcm} n^{b^i}$ & number of nonzeros (integer) in $b^i$, summed over all $i \in \Mcm$ &
  [2] \\
@@
 0.0       value of q^0
-4         # nonzeros in vectors b^i (i=0,...,m)
-1 1 1.0   4 lines constraint, index & value of nonzero in b^i (i=0,...,m)
+4         # nonzeros in vectors b^i (i=1,...,m)
+1 1 1.0   4 lines constraint, index & value of nonzero in b^i (i=1,...,m)
 1 2 1.0   |

### 2017-05-04 QPLIB-3.tex
remove remainings of old comment
update tables 4 and 5
- after fixing convexity check

@@
           & \multirow{2}*{Mixed}
-                    & Linear    &   14\\[1.2 ex]
+                    & Linear    &   12\\[1.2 ex]
 \multirow{2}*{Concave (if max)}
@@
           & \multirow{1}*{Integer}
-                    & Linear    &   2\\[1.2 ex]
+                    & Linear    &   1\\[1.2 ex]
 %\hline
@@
           & \multirow{3}*{Binary}
-                    & None      &   23\\[1.2 ex] on the binary variables
+                    & None      &   23\\[1.2 ex]
           &         & Linear    &  59\\[1.2 ex]
@@
           & \multirow{2}*{Mixed}
-                    & Linear    &   9\\[1.2 ex]
+                    & Linear    &   11\\[1.2 ex]
           &         & Quadratic &    1\\[1.2 ex]
-%          & Integer & Linear    &    \\
+\cmidrule(lr){2-4}
+          & Integer & Linear    &    1\\
 \cmidrule(lr){2-4}
@@
 \multirow{1}*{Concave (if max)}
-                         & Quadratic &    9\\[1.2 ex]
+                         & Quadratic &    6\\[1.2 ex]
 \cmidrule(lr){1-3}
@@
                          & Linear    &   6\\[1.2 ex]
-                         & Convex    &   1\\[1.2 ex]
-                         & Quadratic &   25\\[1.2 ex]
+                         & Convex    &   3\\[1.2 ex]
+                         & Quadratic &   26\\[1.2 ex]
 \hline

### 2017-05-05 QPLIB-3.tex
update instance classification after tightening tolerance on zero eigvals
- and put the correct tolerance into the paper

@@
 
-The number of positive (i.e., larger than $10^{-16}$) and negative (i.e., smaller than $-10^{-16}$) eigenvalues of $Q^0$ allowed us to identify the
+The number of positive (i.e., larger than $10^{-12}$) and negative (i.e., smaller than $-10^{-12}$) eigenvalues of $Q^0$ allowed us to identify the
 objective function type, as in presence of at least one negative (positive) eigenvalue
@@
                     & Linear    &   12\\[1.2 ex]
-\multirow{2}*{Concave (if max)}
+\multirow{1}*{Concave (if max)}
           &         & Quadratic &    1\\[1.2 ex]
 %          & General & Linear    &    \\
-\cmidrule(lr){2-4}
-          & \multirow{1}*{Integer}
-                    & Linear    &   1\\[1.2 ex]
+%\cmidrule(lr){2-4}
+%          & \multirow{1}*{Integer}
+%                    & Linear    &   1\\[1.2 ex]
 %\hline
@@
 \cmidrule(lr){2-4}
-          & Integer & Linear    &    1\\
+          & Integer & Linear    &    2\\[1.2 ex]
 \cmidrule(lr){2-4}
@@
 %
-\multirow{2}*{Linear}    & Convex    &   15\\[1.2 ex]
-                         & Quadratic &   46\\[1.2 ex]
+\multirow{2}*{Linear}    & Convex    &   13\\[1.2 ex]
+                         & Quadratic &   48\\[1.2 ex]
 \cmidrule(lr){1-3}
@@
                          & Linear    &   6\\[1.2 ex]
-                         & Convex    &   3\\[1.2 ex]
-                         & Quadratic &   26\\[1.2 ex]
+                         & Convex    &   1\\[1.2 ex]
+                         & Quadratic &   28\\[1.2 ex]
 \hline

### 2017-08-16 AnswerToTheReferees.tex
comments of the Referees

\documentclass[11pt]{article}
\usepackage[american]{babel}
%\usepackage{a4wide,paralist,url,xcolor,xspace}
\usepackage{a4wide,url,xcolor,xspace}
\definecolor{acblue}{rgb}{0,0,1}
\definecolor{acgreen}{rgb}{0,1,0}
\definecolor{acred}{rgb}{1,0,0}
\newcommand{\rep}[1]{{\textcolor{acblue}{#1}}}
%\newcommand{\comm}[1]{{\textcolor{acgreen}{#1}}}
\newcommand{\comm}[1]{}
\newcommand{\todo}[1]{{\textcolor{acred}{#1}}}

\usepackage{amsmath}

\newcommand{\eg}{e.g.,\xspace}
\newcommand{\ie}{i.e.,\xspace}

\parindent0pt
\parskip.8ex

\begin{document}

\section*{QPLIB: a Library of a Quadratic Programming Instances (MPC-022017-00003)-- Answer to the Referees}

\rep{
 Following the Area Editor suggestions, we 
 performed a deep revision of our manuscript. We carefully took into account all the comments made by the Referees and the Associate Editor.
 [\dots]
 The new submitted document highlights in red the text that has been
added or changed.
In the following, the comments of the Associate Editor and of the Referees are reported, together with our point-by-point replies.
}

\subsection*{Reply to the Associate Editor}

{\it
I have received two reviews for this manuscript, and while there is consensus that the library makes an important contribution, one of the reviewers in particular had significant concerns with the library and the manuscript describing it. The most significant concern, which I share, is that the paper is significantly longer than it needs to be. In particular, I did not see a great need to review the solution methods as is done in Section 2.3. Mentioning the current solvers, and which classes of problems they can solve according to your taxonomy is perhaps useful for a reader's reference. But attempting to describe the algorithms themselves is somewhat futile, as it takes significant space and yet is still an inadequate description. For that, it would be better to simply provide some references for the benefit of readers who are interested in understanding the algorithms in the solvers. I also agree with the reviewer's concern that some of the instance classes have only a single instance in them. If I am a researcher designing a method for solving instances in that class, I will be pretty disappointed to find that QPLIB only has a single relevant test instance. 
}

\rep{[\dots] }

{\it
I recommend the authors be given a chance to revise the manuscript (and library) to address the concerns raised by the reviewers, or otherwise provide justification for comments they choose not to address in their revision. I think that if the library and manuscript can be improved along these lines, this cab be a substantial and important contribution to the computational optimization field. 
}

\rep{We would like to thank the Editor-in-Chief for having given us the possibility to perform a revision of our manuscript. In the following we present the improvements we made in accordance with the comments of the referees. [\dots] }

\subsection*{Point-by-point reply to the Referee 1}

{\it
This paper presents a very thoughtfully constructed library of Quadratic Programming problems. I am not sure if there is a clear MPC policy on library
papers. I think it my merit a discussion at some time. Personally, I think libraries are very important so thoughtfully constructed ones such as this should
be encouraged and supported by MPC. However, I worry a bit about the over-
proliferation of such libraries, hence the reason for a possible discussion. Fortunately, I do not believe this is an issue with the library proposed in this
manuscript. The closest library to it I can think of is the CBLIB library I
mention bellow and the conceptual overlap is rather minimal. Furthermore, as
I believe the collection and selection of instances was very thorough and well
reasoned. Finally, I believe the brief survey and taxonomy of QP problems has
its own value beyond the library. I do have some comments bellow, but I believe
they can be easily addressed without the need for an additional detailed review.
Hence, my recommendation is the paper be accepted after a minor revision.
}

\rep{[\dots] }

{\it
Minor Comments:
\begin{enumerate}

\item I think it would be interesting to mention the Conic library CBLIB \url{http://cblib.zib.de} and its relation to the library described in the manuscript.

\rep{[\dots] }

\item Page 7, line -4. You say that convex means that the feasible region is
convex, but in page 19 you discuss the fact that SOCP problems that have
a convex feasible region could be considered nonconvex. Hence, maybe it
is more accurate to say that what you mean is that the constraint can be
converted to $f (x) \le c$ with $f$ convex by rearranging terms.

\rep{[\dots] }

\item Page 17. Doesn't CPLEX support some binary non-convex quadratic
problems.

\rep{[\dots] }

\item Page 19. I think it would be interesting to give a bit more details about the conic constraints and relate it with the issue of what is a convex
constraint I discuss in point 2. I think the argument is that problems with
non-convex quadratic constraints (i.e. $f (x) \le 0$ with $f$ non-convex) may
have a convex feasible region. I imagine that detecting this in general may
be NP-hard even though the conic case (1 negative eigenvalue) is relatively
easy (e.g. if the problem includes an inequality of the form $a \cdot x \ge$ where
a is the eigenvector associated to the negative eigenvalue). It would then
be interesting to mark the problems with non-convex constraints, but for
which it is known how to show the convexity of the complete feasible
region (because they are conic or another ad-hoc reason).

\rep{[\dots] }

\item Page 20. ``second similar instance filter''. It would be interesting to provide a reference to the similar instances that were filtered out in case
someone wants to test on them. For that matter I am not sure I saw if
the instances have a link/reference to their source.

\rep{[\dots] }

\item Page 36. I am not sure if the reasoning for the name of the instances was mentioned. Is the non-consecutive order because of the filtering? (e.g.
was there a ``0017'' instance that was excluded?)

\rep{[\dots] }

\end{enumerate}
}

\subsection*{Point-by-point reply to the Referee 2}

{\it
The paper presents a library of quadratic programming (QP) instances, where QP is broadly interpreted
to include quadratically constrained, and integer programs too. While I appreciate the overall goal of
the project, the paper is poorly written. It is often long-winded, and contains numerous factual errors,
and some omissions that must be corrected. Overall, the paper can easily be cut by 5-10 pages.
}

\rep{[\dots] }

{\it
\begin{enumerate}

\item There are two classes of important missing application domains of QPs in Table 1: (1) trust-region subproblems, and (2) PDE-constrained optimization problems. In fact, there exist a number of PDE-constrained instances in CUTEr, which would be useful additions to the library. The authors should add at least a few such examples. the lack of these problems is worrying, because
it (implicitly) assumes that direct solvers are used, which means that this library is less relevant for matrix-free solvers.

\rep{[\dots] }

\item The taxonomy in Section 2.2.1 seems to be unaware of the fact that an indefinite Hessian does not mean that a problem is difficult. For example, saddle-point problems such as

$$ \begin{bmatrix}
G & A^T\\
A & 0
\end{bmatrix}
\begin{pmatrix}
x \\
z 
\end{pmatrix}
=
\begin{pmatrix}
-g \\
b 
\end{pmatrix}
$$

can be solved to global optimality as long as the matrix is second-order sufficient. This means
that even some QPs with indefinite Hessians can easily be solved to global optimality.

\rep{[\dots] }

\item The discussion on p.9 recounts some textbook facts, and is largely superfluous. Delete it.

\rep{[\dots] }

\item What do you mean by ``exact solution of the instances'' (p.10, l.-18)? I don’t believe we can even
solve $Ax = b$ ``exactly'', except in some special cases. Please clarify this statement.

\rep{[\dots] }

\item The description of NLP solvers in Section 2.3.2 is at somewhat misleading. For example, KNI-
TRO is also a primal-dual interior-point method, not just a barrier code (in fact primal-dual
methods are preferred). The description of the “active-set” methods is wrong, unless you refer
to the inner method, namely the active-set QP solves. Otherwise, SNOPT and KNITRO are
SQP methods, while CONOPT is a gradient-projection method, and MINOS is an augmented
Lagrangian method. You must correct this section

\rep{[\dots] }

\item Some classes of problems in Table 4 do not seem to have many instances. Please create more.

\rep{[\dots] }

\item The y-axis labels are missing on most figures.

\rep{[\dots] }

\item Reword the sentence ``percentage of ``hard'' eigenvalues'', p.23, l.-6. The notion of hard has a
special meaning in trust-region problems. Also, this is not a measure of hardness, as explained
above. Instead, you should consider the number of negative eigenvalues of the reduced Hessian
(wrt equality constraints). Fig. 4 is meaningless, unless you replace the definition of hardness.

\rep{[\dots] }

\item Improve Fig.5: (1) What do the colors mean? (2) The plot on the right is not a sparsity pattern
of any Hessian, but at best the upper triangular sparsity pattern.

\rep{[\dots] }

\item I am concerned that problems are only available in GAMS, and tested on GAMS solvers. It would make the paper stronger if the authors also released AMPL model, for example, or made use of
JuMP.

\rep{[\dots] }

\end{enumerate}

}

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2017-09-07 AnswerToTheReferees.tex
Fabio is forcing me to use this antiquated windows-driven system instead of a more user-friendly command line

@@
 \usepackage{amsmath}
+\usepackage{amsfonts}
 
@@
 \parskip.8ex
+
+\newcommand{\leo}[1]{{\color{red}#1}}
+\newcommand{\fabio}[1]{{\color{cyan}#1}}
+\newcommand{\emiliano}[1]{{\color{blue}#1}}   
+\newcommand{\antonio}[1]{{\color{ForestGreen}#1}}     
+\newcommand{\andrea}[1]{{\color{blue}{\bf #1}}}
 
@@
 For that, it would be better to simply provide some references for the benefit of readers who are interested in understanding the algorithms in the solvers.
+}
+
+\rep{\leo{We removed 3 whole pages of algorithmic explanations.}}
+
+{\it
 I also agree with the reviewer's concern that some of the instance classes have only a single instance in them.
@@
 
-\rep{[\dots] }
+\rep{\leo{Sono d'accordo, ma come/dove troviamo le istanze supplementari?
+Sono le categorie con  $<10$ istanze nelle tavole 4 e 5, giusto?
+Dire che ce ne sono parecchie.
+15 categorie su 23 hanno meno di 10 istanze (il numero 10 e` magico perche' le nostre mani hanno dieci dita, ovviamente).} }
 
@@
 
-\rep{[\dots] }
+\rep{\leo{sono d'accordo, menzioniamo la libreria conica} }
 
@@
 
-\rep{[\dots] }
+\rep{ \leo{secondo me il secondo commento si riferisce a pagina 19...
+a pagina 7 riga -4 non trovo niente su SOCP.
+Pero` sono d'accordo che a pagina 19 l'ultimo paragrafo della sezione 3.1 potrebbe essere migliorato.
+Cosi` e` un po' criptico.
+Perche' non inserire un bell'esempio di una forma bilineare nonconvessa (scrivendola esplicitamente come xQx e dando esplicitamente lo spettro di Q con la parte definita negativa) inserita in un vincolo che definisce una regione convessa?
+E poi io direi che facciamo la scelta di mettere queste istanze come "nonconvesse" perche' dimostrare convessita` in generale e` difficile (a questo proposito, io conosco i risultati del PhD di Amir Ali Ahmadi sulla difficolta` di riconoscere la convessita` in polinomi di grado $>= 4$, ma non ho mai visto una dimostrazione di difficolta` di riconoscere la convessita` della feasible region in vincoli quadratici; spero che qualcuno dei nostri co-autori ne sia al corrente)...
+Il referee ricorda che riconoscere i casi di forme quadratiche con una negative eigenvalue inserite in un vincolo convesso e` facile e  da` un metodo, di cui sarebbe bene ritrovare il primo caso in cui viene menzionato, oppure darne una breve dimostrazione...
+io non ne ero a conoscenza (e nemmeno mi viene in mente sui due piedi come dimostrarlo).} }
 
@@
 
-\rep{[\dots] }
+\rep{\leo{rispondiamo al referee che CPLEX e` classificato 'C' (complete) in tavola 2 rispetto a QGC e ricordiamogli che pero` CPLEX non risolve (a mia conoscenza e per il momento) QGQ.
+Tra l'altro, c'e` a p.
+17 una menzione di "Table 2 below", tranne che la table 2 appare above.
+Diciamo solo "Table 2" senza il below e l'above, dato che la posizione di tavole e figure puo` comunque essere modificata sia da LaTeX sia dal typesetting finale.} }
 
@@
 
-\rep{[\dots] }
+\rep{ \leo{sono d'accordo col referee (ovvero: massi`, facciamolo contento)} \fabio{va qui? o e' un commento al punto di sotto?}}
 
@@
 
-\rep{[\dots] }
+\rep{\leo{ basta dargli una risposta secondo me} }
 
@@
 
-\rep{[\dots] }
+\rep{ \leo{sono d'accordo, dovremmo aggiungere delle istanze da trust-region subproblem e delle istanze da problemi PDE-constrained.
+Chiediamo a Gould per le istanze trust-region e a Gould, Mittelmann e Sahinidis per PDE-constrained.
+Specifichiamo che vorremmo istanze NON gia` in CUTE.} }
 
@@
 
-\rep{[\dots] }
+\rep{\leo{It is well known that every hard problem, seen as an infinite class of instances, contains efficiently solvable subclasses.
+We feel it is outside of the scope of this paper to characterize all known polytime solvable subclasses, and we do not really see how we can convincingly discuss the specific one you point out without further (long) digressions.}}
 
@@
 
-\rep{[\dots] }
+\rep{\leo{cancelliamo tutto!} }
 
-\item What do you mean by ``exact solution of the instances'' (p.10, l.-18)? I don’t believe we can even
+\item What do you mean by ``exact solution of the instances'' (p.10, l.-18)? I don't believe we can even
 solve $Ax = b$ ``exactly'', except in some special cases. Please clarify this statement.
 
-\rep{[\dots] }
+\rep{ \leo{We mean exactly the text given in the first paragraph of Sect.~2.2.4 in the submitted manuscript, which has now been revised (and hopefully made clearer) as shown in the first two paragraphs of Sect.~2.3.3 ``Complete Methods'': in short, the difference between computational models such as TM and RRAM.} }
 
@@
 
-\rep{[\dots] }
+\rep{\leo{correggiamo gli errori e ringraziamo il referee} }
 
@@
 
-\rep{[\dots] }
+\rep{\leo{stesso commento dell'AE -- dobbiamo mettere piu` istanze} }
 
@@
 
-\rep{[\dots] }
+\rep{\leo{mettiamo la label al y axis  (duh)} }
 
@@
 
-\rep{[\dots] }
+\rep{ \leo{non so nulla di "hard eigenvalues" nel senso del trust region; chiediamo agli esperti} }
 
@@
 
-\rep{[\dots] }
+\rep{\leo{Possiamo spiegare che proporre NOI le istanze in formati multipli darebbe la responsabilita` del traduttore a noi, che invece non vogliamo accettarla.
+D'altro canto penso che, sebbene sia stata una scelta discussa e effettuata democraticamente, non sia necessariamente la scelta giusta per ogni criterio.
+Per esempio, se il nostro criterio fosse la visibilita` della QPLIB (il numero di gente che la cita), dovremmo ASSOLUTAMENTE tradurre le istanze nel maggior numero possibile di formati.
+E se qualcuno ci fara` una correzione in futuro chi se ne frega, ringrazieremo e correggeremo.
+Francamente, se io fossi il dittatore della QPLIB, sceglierei di tradurre le istanze in quanti piu` formati sia possibile (almeno AMPL e SIF), e poi distribuirle con un warning.} \fabio{Adesso abbiamo anche aggiunto il formato AMPL, secondo me con questo il referee si mette tranquillo!} }
 

### 2017-09-08 QPLIB-2.2.tex
Additional comment on QCR
others {\textstyle \half}
leo's changes

@@
 
+\Leo{\sout{Following the remark of one of our anonymous referees,} We note that, although the left-to-right progression in the above text marks ``harder problems'', this does \textit{not} mean that \textit{every} instance of a hard problem is itself ``difficult to solve''.
+As is well known, problems as infinite collections of instances; and it is always possible to find subclasses that can be solved efficiently.}
+
 % -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -
@@
  \min \; \Big\{ \,
- \sum_{j \in \Mcn} \big( \, \half Q^0_j x_j^2 + b^0_j x_j \, \big) \;:\;
+ \sum_{j \in \Mcn} \big( \, {\textstyle \half} Q^0_j x_j^2 + b^0_j x_j \, \big) \;:\;
  l_j \leq x_j \leq u_j \quad j \in \Mcn \;\;,\;\;
@@
  \min \; \Big\{ \, b^0 x \;:\;
- \half x^\top Q^i x + b^i x \leq c^i \quad i \in \Mcm \;\;,\;\;
+ {\textstyle \half} x^\top Q^i x + b^i x \leq c^i \quad i \in \Mcm \;\;,\;\;
  x_j \in \mathbb{Z} \quad j \in \Mcn \, \Big\}
@@
 More generally, this is also true for \textit{*G*} as long as $\Mcu = \emptyset$, as bounded general integer variables can be represented by binary ones.
+Hence, the nonconvexity due to binary variables can always be expressed by means of (nonconvex) quadratic constraints.
+The converse is also true: when only binary variables are present, all quadratic constraints can be converted into convex ones \cite{BEP09,BEL13}.
 
@@
  %
- & \hspace*{-5mm} -\infty \leq \half x^\top Q^0 x + b^0 x \leq x^0 & \\
+ & \hspace*{-5mm} -\infty \leq {\textstyle \half} x^\top Q^0 x + b^0 x \leq x^0 & \\
  %
- & c^i_l \leq \half x^\top Q^i x + b^i x \leq c^i_u & i \in \Mcm \\
+ & c^i_l \leq {\textstyle \half} x^\top Q^i x + b^i x \leq c^i_u & i \in \Mcm \\
  %

### 2017-09-08 methods.tex
leo's changes

--- 2017-02-03 methods.tex
@@
 In this section we provide a quick overview of existing solution methods for QP, restricting ourselves to these implemented by the set of solvers considered in this paper (see \S\ref{subsec:solver}).
-For each approach we briefly describe the main algorithmic ideas and point out the formulation they address according to the classification set out in \S\ref{sec:classification}.
-We remark that many solvers implement more than one algorithm, from which the user can choose at runtime.
-Moreover, algorithms are typically implemented in different ways within different solvers, so that the same conceptual algorithm can sometimes yield wildly different results or performance measures on the same instances.
+For each approach we briefly describe  the formulation they address according to the classification set out in \S\ref{sec:classification}.
+We remark that many solvers implement more than one algorithm, which the user can choose at runtime.
+Moreover, algorithms are typically implemented in different ways within different solvers, so that the same conceptual algorithm can sometimes yield different results or performance measures on the same instances.
 
-Solution methods for QP can, following \cite{neumaier}, be broadly organized in four categories: \emph{incomplete}, \emph{asymptotically complete}, \emph{complete}, and \emph{rigorous}.
-Incomplete methods are only able to identify solutions, often locally optimal according to a suitable notion, and may even fail to find one even when one exists; in particular, they are typically not able to determine that an instance
-%is empty.
-has no solution.
-Asymptotically complete methods can find a globally optimal solution with probability one in infinite time, but again they cannot prove that a given instance is infeasible.
-Complete methods find an approximate globally optimal solution within a prescribed optimality tolerance within finite time, or prove that none such exists (but see \S\ref{s:complete} below); they are often referred to as \emph{exact} methods in the computational optimization community.
-Finally, rigorous methods find globally optimal solutions within given tolerances even in the presence of rounding errors, except for ``near-degenerate cases''.
+  Solution methods for QP can be broadly organized in four categories \cite{neumaier}: \emph{incomplete}, \emph{asymptotically complete}, \emph{complete}, and \emph{rigorous}.
+  \begin{itemize}
+  \item \textit{Incomplete} methods are only able to identify solutions, often locally optimal according to a suitable notion, and may even fail to find one even when one exists; in particular, they are typically unable to determine that an instance has no solution.
+  \item \textit{Asymptotically complete} methods can find a globally optimal solution with probability one in infinite time, but again they cannot prove that a given instance is infeasible.
+  \item \textit{Complete} methods find an approximate globally optimal solution within a prescribed optimality tolerance within finite time, or prove that none such exists (but see \S\ref{s:complete} below); they are often referred to as \emph{exact} methods in the computational optimization community.
+  \item \textit{Rigorous} methods find globally optimal solutions within given tolerances even in the presence of rounding errors, except for ``near-degenerate cases''.
 Since none of the solvers we are using can be classified as rigorous, we limit ourselves to declaring solvers complete.
+  \end{itemize}
+
+\input{solvers.tex}
+
+\subsubsection{Incomplete methods}
 
@@
 In general, all complete methods for a problem class $P$ must be complete for any problem class $Q \subseteq P$, while a complete method for $P$ might be incomplete for a class $R \supset P$.
-
-\input{solvers.tex}
-
-% -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -
-\subsubsection{Incomplete methods}\label{s:incomplete}
-
-Local search methods typically require as an input an estimate $x'$ of the required  solution, and attempt to improve it---either towards feasibility, or towards optimality, or both---using only information that is available in a neighborhood of $x'$.
-In general, local search methods are incomplete.
-As remarked above, however, local search methods deployed on a convex problem behave like complete methods, possibly via devices that allow one to find a feasible starting estimate if there is one (like what used to be called ``phase-1'' in the simplex method).
-
-Most local search methods for \textit{*C*} are iterative in nature, and need a starting point $x^0$ as input.
-The ($k+1$)-st iterate is obtained as $x^{k+1} = x^k + \alpha_k d^k$, where $\alpha_k$ (a scalar) is the \emph{step length}, and $d^k$ (a vector) is the \emph{search direction}, e.g.~belonging to a tangent manifold and having a negative directional derivative at $x^k$, in order to improving optimality while reducing infeasibility.
-Alternatively, $x^{k+1} = x^k + d^k(\alpha_k)$ where $d^k(\cdot)$ is an arc satisfying $d^k(0) = 0$.
-Local search methods for *I* are also iterative, but since defining a useful notion of descent direction is harder in presence of integer variables, $d^k$ and $\alpha_k$ are usually computed in different ways, depending on the application at hand.
 
@@
 \sqic \cite{Gill2015}
-for nonconvex ones
-While there are many different approaches for NLPs in general, we mention here those that are particularly well-suited for QP, i.e.:
-%
-\begin{itemize}
- \item active set methods (\conopt, \minos, \snopt, \knitro, \bqpd, \qpa, \sqic) \cite{Dost09};
- %
- \item interior point methods (\ipopt, \knitro, \cqp, \ooqp, \qpb, \qpc) \cite{Wright97};
- %
- \item projected gradient methods (\dqp) \cite{CalaMore87:mp,ChenGui13:coap}.
-\end{itemize}
-%
-Active set and interior point methods have been defined for \textit{*C*} but also for general NLPs, and many of the solvers named above target this
-larger problem class.
-
-\paragraph{Active set methods.}
-%
-At each iteration $k$ the algorithm forms the \emph{active set} $\mathcal{A}^k$ containing the (indices of the) \emph{active constraints}, i.e., all of the equality constraints as well as the inequality constraints that are satisfied at equality at the current iterate $x^k$.
-A subproblem, consisting of a minimization of a certain auxiliary objective function and including only active constraints, is then solved to identify a good search direction $d^k$.
-An appropriate step length $\alpha_k$ is then found by checking for the inactive constraints (since these must be satisfied too).
-If at $x^{k+1}$ some of the previously inactive constraints have become active, or vice versa, $\mathcal{A}^k$ is updated accordingly.
-This general scheme works because the step is chosen so that the objective decreases, and the active set cannot repeat.
-
-\paragraph{Interior point methods.}
-%
-These approaches can be seen as recasting the original constrained problem QP as a parametrized family QP$_\mu$ of unconstrained ones, where the constraints are moved into the objective function via a \emph{barrier term}---that goes to $+\infty$ as the boundary of the feasible region is approached---weighted with the \emph{barrier parameter} $\mu \in (0, \infty)$.
-In the convex case, the \emph{central path}---the continuous line formed from the optimal solutions of QP$_\mu$ for all varying $\mu$, which is typically unique e.g.~if the classical barrier based on the logarithmic function is employed---leads to a (central) optimal solution of QP when $\mu \to 0$.
-Starting from an appropriately constructed ``central'' point (close to the solution of QP$_\mu$ for ``large'' $\mu$), these algorithms strive to follow the central path by performing $O(1)$ Newton steps before (substantially) reducing $\mu$.
-The algorithms can be shown to converge in a small number of iterations, each of which can be costly due to the need of solving an appropriately perturbed version of the KKT conditions for QP (``perturbed'' with $\mu$), a (possibly) large-scale linear system.
-Nowadays the algorithm is most often implemented in the primal-dual version, where the nonlinear KKT system is iteratively solved with Newton-like iterations; this has the extra advantage of allowing to remove the need of a feasible (central) starting point.
-
-\paragraph{Projected gradient methods.}
-%
-These approaches are similar to active-set methods in that at each iteration they project the steepest descent direction onto the constraint set in order to (try to) predict the optimal active set; the predicted set may then be searched more exhaustively to improve the objective function.
-They differ from active-set methods in that the set of active constraints can change dramatically on each iteration, and they often exhibit stronger practical convergence behaviour, more akin to those of interior-point methods.
-However they rely heavily on projection, and thus are only suited to simple constraint geometries such as boxes and simplices.
-This is particularly useful for strictly-convex \textit{CCL} examples,
-since their duals may be reformulated exclusively to involve box constraints.
+for nonconvex ones.
 
@@
 
-The solvers in Table \ref{t:solvers} which implement asymptotically complete methods are \oqnlp, \msnlp, \knitro, and certain sub-solvers of \lgo. Specifically, we consider the following methods:
-%
-\begin{itemize}
- \item global adaptive random search (\lgo);
- %
- \item multi-start (\knitro, \lgo, \msnlp, \oqnlp); specifically, the former three apply to \textit{QCQ} whereas the latter to \textit{QGQ}.
-\end{itemize}
-
-\paragraph{Global Adaptive Random Search (GARS).}
-%
-This is a modification of an algorithm called \emph{pure random search}, which consists in sampling a random point $x'$ from a given compact set known to contain a global optimum, and then sampling a new candidate solution $y$ in a neighborhood of $x'$, setting $x' \leftarrow y$ if $y$ improves $x'$, and repeating as long as a termination condition is not satisfied.
-The adaptivity stems from changing the distribution for sampling $y$ at run-time, depending on the quality of the solutions identified by the method.
-Since this method only depends on sampling and function evaluation, it is usually fast.
-In the GARS solver of \lgo, it provides a useful starting point for a subsequent local search procedure.
-Asymptotic global convergence is attained by restarting the random search from different initial points $x'$.
-
-\paragraph{Multi-start.}
-%
-Multi-start methods define a loop around a given local search procedure so that it starts from many different starting points, perform local search, and record the best optimum found so far as they explore the search space randomly.
-For example, any of the methods described in Sect.~\ref{s:incomplete} can be embedded in a multi-start framework as follows:
-%
-\begin{enumerate}
- \item initialize a ``best solution so far'' $x^\ast$
- %
- \item sample a starting point $x'$ uniformly at random from a given compact set known to contain a global optimum; \label{step:ms2}
- %
- \item run a local search method from $x'$ to yield an improved (feasible) point $x$
- %
- \item if $x$ improves on $x^\ast$ with respect to the objective function value, replace $x^\ast$ with $x$
- %
- \item repeat from Step \ref{step:ms2} until a given termination condition is satisfied. \label{step:ms5}
-\end{enumerate}
-%
-The method is asymptotically complete if the termination condition in Step \ref{step:ms2} is a certificate of global optimality for $x^\ast$, which is usually hard to obtain.
-In practice, typically some bound on the permitted total CPU time or number of function evaluations, or any other criteria that makes sense for the application at hand, is needed, which renders the method incomplete.
-
-In general, the applicability of meta-heuristics to a given problem relies on availability of a local search that exploit the structural properties of the problem.
-Depending on the local search employed, multi-start methods can address MINLP of the most general class.
+The solvers in Table \ref{t:solvers} which implement asymptotically complete methods are \oqnlp and \knitro (which apply to \textit{QGQ}) as well as \msnlp and certain sub-solvers of \lgo (which apply to \textit{QCQ}).
 
@@
 
-%Complete methods are often referred to as \emph{exact} in a large part of the mathematical optimization community.
-This nomenclature has to be used with care, as it implicitly makes assumptions on the underlying computational model that may not be acceptable in all cases.
-To see that, consider that, as already mentioned, QPs (more precisely, \textit{LIQ}) are generally undecidable \cite{jeroslow}; and yet, there exists a general decision method for deciding feasibility of systems of polynomial equations and inequalities \cite{tarski-reals}, including the solution of \textit{LCQ} with zero objective function.
-This apparent contradiction is due to the fact that the two statements refer to different computational models: the former is based on the Turing Machine (TM), whereas the latter is based on the Real RAM (RRAM) machine \cite{blum}.
+Complete methods are often referred to as \emph{exact} in a large part of the mathematical optimization community.
+This term has to be used with care, as it implicitly makes assumptions on the underlying computational model that may not be acceptable in all cases.
+For example, te decision version of \textit{QCL} is known to be in the complexity class \textbf{NP} \cite{vavasis90a}, whereas the same is not known about \textit{LCQ}, even with zero objective.
+On the other hand, there exists a method for deciding feasibility of systems of polynomial equations and inequalities \cite{tarski-reals}, including the solution of \textit{LCQ} with zero objective function.
+
+To explain this apparent contradiction, we remark that the two statements refer to different computational models: the former is based on the Turing Machine (TM), whereas the latter is based on a computational model that allows operations or real numbers, e.g.~the Real RAM (RRAM) machine \cite{blum}.
 Due to the potentially infinite nature of exact real arithmetic computations, exact computations on the RRAM necessarily end up being approximate on the TM.
@@
 
-\paragraph{Branch-and-Bound.}
-%
-Nearly all of the complete solvers in Table \ref{t:solvers} that address $\mathcal{NP}$-hard problems (i.e.~those in \textit{QGQ}$\smallsetminus$\textit{CCC}) are based on Branch-and-Bound (BB).
-This is an implicit but exhaustive search process based on exploring a \emph{branching tree} of the problem, where each node in the tree represents a subset of the feasible region.
-Guaranteed lower and upper bounds to the objective function value relative to nodes are computed in various ways.
-Nodes are discarded when: (a) they can be shown to be empty; (b) their bound in the optimization direction is worse than an opposite global bound; (c) a global optimum limited to the node can be found (this happens when the two bounds are closer than a given $\varepsilon$ tolerance); (d) they are selected for branching, which means expanding the tree constructing at least two new nodes, children of the current one.
-Branching takes place by identifying one or more branching directions, which are usually a coordinate axes, and one or more branching point per direction, in various common sense fashions.
-The algorithm is driven by a queue of active nodes, usually endowed with a priority to select the most promising node from which to continue exploration of the tree (such as ``most promising bound''); the BB algorithm terminates when the queue is empty.
-
-Typically, bounds in the optimization direction are computed by means of convex relaxations \cite{mccormick,BeLeLiMaWa08}, which replace nonconvex terms $t(x)$ with linearization variables $\hat{t}$, and then replace the corresponding defining constraints $\hat{t} = t(x)$ by means of lower and upper (respectively, convex and concave) bounding functions $\hat{t} \ge \underline{t}(x)$ and $\hat{t} \le\bar{t}(x)$.
-This is actually where finite (and tight) bounds on the variables are crucial, which differentiate also in practice the bounded case from the unbounded one.
-Different strategies are used when the nonconvexities are only quadratic \cite{MiFl13,bliek}.
-
-When the BB algorithm is allowed to select coordinate directions corresponding to continuous variables, it is called \emph{spatial} BB (sBB).
-Branching on continuous (rather than integer or binary) variables becomes necessary in the presence of nonconvex nonlinearities, as it happens e.g.~in \textit{QCQ}, since the quality of the bounds improves as the feasible set in the current node gets smaller.
-
+Nearly all of the complete solvers in Table \ref{t:solvers} that address $\mathcal{NP}$-hard problems (i.e.~those in \textit{QGQ}$\smallsetminus$\textit{CCC}) are based on Branch-and-Bound (BB) \cite{land_doig}.
+When the BB algorithm is allowed to branch on coordinate directions corresponding to continuous variables, it is called \emph{spatial} BB (sBB) \cite{dakin,couenne}.
 BB algorithms are exponential time in the worst case, and their exponential behavior unfortunately often shows up in practice.
-They can also be used heuristically (forsaking their completeness guarantee) by either terminating them early, or by using non-guaranteed bounds.
-
+They can also be used heuristically (forsaking their completeness guarantee) in a number of ways, e.g.~by either terminating them early.
 The following solvers from Table \ref{t:solvers} implement complete BB algorithms for \textit{QGQ} or some subclasses:
@@
 We remark that the latter category can be used as incomplete solvers for \textit{QGQ}.
-%We also remark that CPLEX can currently only target problems with linear constraints when the objective function is nonconvex \cite{bliek}.
 We also note that \lgo implements an incomplete BB algorithm for \textit{QCQ} by using bounds obtained from sampling.
 
-\paragraph{Cutting plane approaches.}
-%
 Cutting plane approaches construct and iteratively improve a MILP (\textit{LIL}) relaxation of the problem \cite{DuGr86,WePoe02}.
@@
 Additionally, incomplete methods can be used to provide local solutions.
-Therefor, these methods are complete on \textit{CGC} if a complete method is used to solve the MILP.
+Therefore, these methods are complete on \textit{CGC} if a complete method is used to solve the MILP.
 The latter is typically based on BB, which is therefore a crucial technique also for this class of approaches.
-
 Solvers in Table \ref{t:solvers} that implement complete cutting plane methods for \textit{CGC} are \alphaecp, \bonmin (in the algorithmic mode B-OA), and \dicopt.
-
-%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-%  End QPLIB-2.3.tex
-%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

### 2018-01-24 AnswerToTheReferees.tex
Update AnswerToTheReferees.tex

--- 2017-08-16 AnswerToTheReferees.tex
@@
 \usepackage{amsmath}
+\usepackage{amsfonts}
+\usepackage{booktabs}
+\usepackage{multirow}
 
@@
 For that, it would be better to simply provide some references for the benefit of readers who are interested in understanding the algorithms in the solvers.
+}
+
+\rep{We agree with the AE that the description of the solution methods goes beyond the scope of the paper.
+For this reason we have streamlined Section 2.3 which now contains only a quick overview of the solvers and their solution methods.
+%\leo{Togliamo la spiegazione degli algoritmi e sostituiamola con una citazione di un book-tutorial-survey-introduction.}
+}
+
+{\it
 I also agree with the reviewer's concern that some of the instance classes have only a single instance in them.
@@
 
-\rep{[\dots] }
+\rep{
+We performed an internal call for instances among the members of the committee. 
+Thanks to this new research, we gathered $86$ new instances and we added them to the library. The updated division of the instances is as follows:\\
+\begin{tabular}{lllr}
+\toprule
+Obj. Fun. & Variables & Constraints & \#\\
+\cmidrule(lr){1-4}
+%
+\multirow{6}*{Linear}
+          & \multirow{2}*{Binary}
+%                    & None      &   \\
+%          &         & Linear    &  \\
+                    & Convex    &   0 +\textcolor{red}{xx}\\[1.2 ex]
+           &         & Quadratic &   9 \\[1.2 ex]
+\cmidrule(lr){2-4}
+          & \multirow{2}*{Mixed}
+                    & Convex    &   2 +\textcolor{red}{xx}\\[1.2 ex]
+          &         & Quadratic &    118\\[1.2 ex]
+\cmidrule(lr){2-4}
+          & \multirow{1}*{Integer}
+%                    & Linear    &    \\
+                   & Quadratic &    2\\[1.2 ex]
+\cmidrule(lr){2-4}
+          & \multirow{1}*{General}
+%                    & Linear    &    \\
+                   & Quadratic &    3\\[1.2 ex]
+\cmidrule(lr){1-4}
+%\multirow{5}*{Linear}
+%          & Binary  & Quadratic &   9\\
+%          & \multirow{2}*{Mixed}
+%                    & Convex    &   2\\
+%          &         & Quadratic &  118\\
+%%          & \multirow{2}*{Integer}
+%%                    & Convex    &  15 \\
+%%          &         & Quadratic &   5 \\
+%          & {General} & Quadratic &   3 \\
+%\hline
+\multirow{1}*{Convex (if min)}
+          & Binary  & Linear    &  2  +\textcolor{red}{xx}\\[1.2 ex]
+\cmidrule(lr){2-4}
+\multirow{1}*{or}
+          & \multirow{2}*{Mixed}
+                    & Linear    &   12\\[1.2 ex]
+\multirow{1}*{Concave (if max)}
+          &         & Quadratic &    1 +\textcolor{red}{xx}\\[1.2 ex]
+%          & General & Linear    &    \\
+%\cmidrule(lr){2-4}
+%          & \multirow{1}*{Integer}
+%                    & Linear    &   1\\[1.2 ex]
+%\hline
+\cmidrule(lr){1-4}
+\multirow{6}*{Quadratic}
+          & \multirow{3}*{Binary}
+                    & None      &   23\\[1.2 ex]
+          &         & Linear    &  59\\[1.2 ex]
+          &         & Quadratic &   5 \\[1.2 ex]
+\cmidrule(lr){2-4}
+          & \multirow{2}*{Mixed}
+                    & Linear    &   11\\[1.2 ex]
+          &         & Quadratic &    1\\[1.2 ex]
+\cmidrule(lr){2-4}
+          & Integer & Linear    &    2\\[1.2 ex]
+\cmidrule(lr){2-4}
+          & \multirow{1}*{General}
+                    & Quadratic    &    1\\[1.2 ex]
+%          &         & Quadratic &    \\
+\hline
+Total Discrete     &         &           & 251 +\textcolor{red}{xx}\\
+%
+\bottomrule
+\end{tabular}\\\\
+%
+%
+\begin{tabular}{llr}
+\toprule
+Obj. Fun. & Constraints & \#\\
+\cmidrule(lr){1-3}
+%
+\multirow{2}*{Linear}    & Convex    &   13\\[1.2 ex]
+                         & Quadratic &   48\\[1.2 ex]
+\cmidrule(lr){1-3}
+\multirow{1}*{Convex (if min)}
+                         & Box       &   3 \\[1.2 ex]
+\multirow{1}*{or}
+                         & Linear    &   11\\[1.2 ex]
+                        % & Convex    &    2\\[1.2 ex]
+\multirow{1}*{Concave (if max)}
+                         & Quadratic &    6\\[1.2 ex]
+\cmidrule(lr){1-3}
+\multirow{3}*{Quadratic}
+                         & Linear    &   6\\[1.2 ex]
+                         & Convex    &   1\\[1.2 ex]
+                         & Quadratic &   28\\[1.2 ex]
+\hline
+Total Continuous                   &           & 116 +\textcolor{red}{18} \\
+%
+\bottomrule
+\end{tabular}\\
+All the new instances 
+%%
+%\leo{Sono d'accordo, ma come/dove troviamo le istanze supplementari?
+Sono le categorie con  $<10$ istanze nelle tavole 4 e 5, giusto?
+Dire che ce ne sono parecchie.
+15 categorie su 23 hanno meno di 10 istanze (il numero 10 e` magico perche' le nostre mani hanno dieci dita, ovviamente).}
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{table}
+ \centering
+ %\scriptsize
+ \setlength{\tabcolsep}{18pt}
+ \renewcommand \arraystretch{1.1}
+\begin{tabular}{llr}
+\toprule
+Variables & Convexity & \#\\
+\cmidrule(lr){1-3}
+%
+ Continuous	& Convex		&  27+\textcolor{red}{yy} \\[1.2 ex]
+ Continuous	& Non Convex	&  89+\textcolor{red}{yy} \\[1.2 ex]
+ Discrete	& Convex		&  16+\textcolor{red}{63} \\[1.2 ex]
+ Discrete	& Non Convex	&  235+\textcolor{red}{5} \\[1.2 ex]
+\cmidrule(lr){1-3}
+Total		&				& 367+\textcolor{red}{86}\\
+%
+\bottomrule
+\end{tabular}
+\caption{Classification of the final set of discrete instances}
+\label{tab:DD}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\rep{[\dots] }
+\rep{TODO:Fabio and Emiliano}
 
@@
 
-\rep{[\dots] }
+\rep{ TODO: Leo 
+%\leo{secondo me il secondo commento si riferisce a pagina 19...
+a pagina 7 riga -4 non trovo niente su SOCP.
+Pero` sono d'accordo che a pagina 19 l'ultimo paragrafo della sezione 3.1 potrebbe essere migliorato.
+Cosi` e` un po' criptico.
+Perche' non inserire un bell'esempio di una forma bilineare nonconvessa (scrivendola esplicitamente come xQx e dando esplicitamente lo spettro di Q con la parte definita negativa) inserita in un vincolo che definisce una regione convessa?
+E poi io direi che facciamo la scelta di mettere queste istanze come "nonconvesse" perche' dimostrare convessita` in generale e` difficile (a questo proposito, io conosco i risultati del PhD di Amir Ali Ahmadi sulla difficolta` di riconoscere la convessita` in polinomi di grado $>= 4$, ma non ho mai visto una dimostrazione di difficolta` di riconoscere la convessita` della feasible region in vincoli quadratici; spero che qualcuno dei nostri co-autori ne sia al corrente)...
+Il referee ricorda che riconoscere i casi di forme quadratiche con una negative eigenvalue inserite in un vincolo convesso e` facile e  da` un metodo, di cui sarebbe bene ritrovare il primo caso in cui viene menzionato, oppure darne una breve dimostrazione...
+io non ne ero a conoscenza (e nemmeno mi viene in mente sui due piedi come dimostrarlo).}
+}
 
@@
 
-\rep{[\dots] }
+\rep{ TODO:Fabio and Emiliano
+%\leo{rispondiamo al referee che CPLEX e` classificato 'C' (complete) in tavola 2 rispetto a QGC e ricordiamogli che pero` CPLEX non risolve (a mia conoscenza e per il momento) QGQ.
+Tra l'altro, c'e` a p.
+17 una menzione di "Table 2 below", tranne che la table 2 appare above.
+Diciamo solo "Table 2" senza il below e l'above, dato che la posizione di tavole e figure puo` comunque essere modificata sia da LaTeX sia dal typesetting finale.}
+}
 
@@
 
-\rep{[\dots] }
+\rep{TODO: Leo
+% \leo{sono d'accordo col referee (ovvero: massi`, facciamolo contento)} \fabio{va qui? o e' un commento al punto di sotto?}
+ }
 
@@
 
-\rep{[\dots] }
+\rep{TODO:FABIO and Emiliano.}
 
@@
 
-\rep{[\dots] }
+\rep{TODO:FABIO and Emiliano.
+%\leo{ basta dargli una risposta secondo me} 
+}
 
@@
 
-\rep{[\dots] }
+\rep{
+TODO:Fabio and Emiliano. 
+%\leo{sono d'accordo, dovremmo aggiungere delle istanze da trust-region subproblem e delle istanze da problemi PDE-constrained.
+Chiediamo a Gould per le istanze trust-region e a Gould, Mittelmann e Sahinidis per PDE-constrained.
+Specifichiamo che vorremmo istanze NON gia` in CUTE.}
+}
 
@@
 
-\rep{[\dots] }
+\rep{TODO:Antonio. 
+%\leo{cosa vuol dire? Questa condizione si puo` testare in polytime? Sicuramente puo` entrare nel discorso a risposta dei punti 2-4 del referee \#1.} 
+}
 
@@
 
-\rep{[\dots] }
+\rep{TODO: Antonio.
+%\leo{cancelliamo tutto!} 
+%As requested, aliminated the discussion from the manuscript
+}
 
-\item What do you mean by ``exact solution of the instances'' (p.10, l.-18)? I don’t believe we can even
+\item What do you mean by ``exact solution of the instances'' (p.10, l.-18)? I don't believe we can even
 solve $Ax = b$ ``exactly'', except in some special cases. Please clarify this statement.
 
-\rep{[\dots] }
+\rep{TODO: Leo. 
+%\leo{possiamo inserire qualche nota sui modelli di computazione per i numeri reali, razionali e floating point: (a) real numbers con real RAM model, in pratica tutti i calcoli tra numeri reali sono O(1) (la gran parte del lavoro fatto da Blum, Shub e Smale e` sulla definizione di calcolabilita` e complessita` in questo 'ambiente');
+%(b) hic sunt leones, non si sa nemmeno se la existential theory of rationals (cioe` risolvere sistemi polinomiali su $\mathbb{Q}$) sia decidibile o no [Pheidas 2000]; 
+%(c) quello che la gente fa davvero: approssimare optimality e feasibility a $\epsilon>=0$ dato, nel migliore dei casi si dimostra anche che il tempo di calcolo o numero di iterazioni e` $O(1/poly(\epsilon)) o O(log(1/\epsilon))$.
+%} 
+}
 
@@
 
-\rep{[\dots] }
+\rep{TODO: Stefan and Ambros
+%\leo{correggiamo gli errori e ringraziamo il referee} 
+}
 
@@
 
-\rep{[\dots] }
+\rep{
+TODO: Stefan and Ambros.
+%\leo{stesso commento dell'AE -- dobbiamo mettere piu` istanze} 
+}
 
@@
 
-\rep{[\dots] }
+\rep{
+TODO: Stefan and Ambros
+%\leo{mettiamo la label al y axis  (duh)} 
+}
 
@@
 
-\rep{[\dots] }
+\rep{
+TODO: Stefan and Ambros.
+% \leo{non so nulla di "hard eigenvalues" nel senso del trust region; chiediamo agli esperti} 
+ }
 
@@
 
-\rep{[\dots] }
+\rep{TODO: Stefan and Ambros}
 
@@
 
-\rep{[\dots] }
+\rep{
+We thank the refereee for having raised this delicate point. We definetly agree that formats are an issue.
+For this reason we have translated all the instances also into the AMPL format and we have uploaded them on the website.
+%\leo{Possiamo spiegare che proporre NOI le istanze in formati multipli darebbe la responsabilita` del traduttore a noi, che invece non vogliamo accettarla.
+D'altro canto penso che, sebbene sia stata una scelta discussa e effettuata democraticamente, non sia necessariamente la scelta giusta per ogni criterio.
+Per esempio, se il nostro criterio fosse la visibilita` della QPLIB (il numero di gente che la cita), dovremmo ASSOLUTAMENTE tradurre le istanze nel maggior numero possibile di formati.
+E se qualcuno ci fara` una correzione in futuro chi se ne frega, ringrazieremo e correggeremo.
+Francamente, se io fossi il dittatore della QPLIB, sceglierei di tradurre le istanze in quanti piu` formati sia possibile (almeno AMPL e SIF), e poi distribuirle con un warning.} \fabio{Adesso abbiamo anche aggiunto il formato AMPL, secondo me con questo il referee si mette tranquillo!}
+}
 

### 2018-01-24 AnswerToTheReferees.tex
Update AnswerToTheReferees.tex

@@
 
-\rep{TODO:Fabio and Emiliano}
+\rep{We updated the manuscript by adding the suggest link to the Conic library.}
 
@@
 
-\rep{TODO:FABIO and Emiliano.}
+\rep{The beta version of the library, which containt some of the filtered instances is available online at  \url{http://www.lamsade.dauphine.fr/QPlib2014}. We added this link to the qplib webpage.}
 
@@
 
-\rep{TODO:FABIO and Emiliano.
-%\leo{ basta dargli una risposta secondo me} 
-}
+\rep{Yes. The numbering is due to the filtering. We added a comment on this point in the Appendix of the paper.}
 

### 2018-01-29 QPLIB-3.tex
Update QPLIB-3.tex
update tables 4 and 5 regarding new instances

@@
 specific QP problems as Max-Cut, Quadratic Assignment, Portfolio
-Optimization, and several others.
+Optimization, and several others. \textcolor{red}{In addition, we mention
+that other generic libraries exist, e.g., 
+Conic library CBLIB (\url{http://cblib.zib.de}) and 
+MIPLIB 2010 (http://miplib.zib.de/), to mention just a few.}
 
@@
 allowed us to identify the final set of $251$ discrete instances and
-$116$ continuous instances.
+$116$ continuous instances.\todo{needs update due to newly added instances ($\to$ 319 discrete 134 cont.)}
 
@@
           & \multirow{2}*{Mixed}
-                    & Convex    &   2\\[1.2 ex]
-          &         & Quadratic &    118\\[1.2 ex]
+                    & Convex    &   \textcolor{red}{15}\\[1.2 ex]
+          &         & Quadratic &  \textcolor{red}{151}\\[1.2 ex]
 \cmidrule(lr){2-4}
@@
 \multirow{1}*{Convex (if min)}
-          & Binary  & Linear    &  2 \\[1.2 ex]
+          & Binary  & Linear    &  \textcolor{red}{4} \\[1.2 ex]
 \cmidrule(lr){2-4}
@@
 \multirow{1}*{Concave (if max)}
-          &         & Quadratic &    1\\[1.2 ex]
+          &         & Quadratic &    \textcolor{red}{6}\\[1.2 ex]
 %          & General & Linear    &    \\
@@
                     & None      &   23\\[1.2 ex]
-          &         & Linear    &  59\\[1.2 ex]
+          &         & Linear    &  \textcolor{red}{74}\\[1.2 ex]
           &         & Quadratic &   5 \\[1.2 ex]
@@
 \hline
-Total     &         &           & 251\\
+Total     &         &           & \textcolor{red}{319}\\
 %
@@
 \multirow{2}*{Linear}    & Convex    &   13\\[1.2 ex]
-                         & Quadratic &   48\\[1.2 ex]
+                         & Quadratic &   \textcolor{red}{52}\\[1.2 ex]
 \cmidrule(lr){1-3}
@@
 \multirow{1}*{or}
-                         & Linear    &   11\\[1.2 ex]
+                         & Linear    &   \textcolor{red}{16}\\[1.2 ex]
                         % & Convex    &    2\\[1.2 ex]
 \multirow{1}*{Concave (if max)}
-                         & Quadratic &    6\\[1.2 ex]
+                         & Quadratic &    \textcolor{red}{11}\\[1.2 ex]
 \cmidrule(lr){1-3}
@@
                          & Linear    &   6\\[1.2 ex]
-                         & Convex    &   1\\[1.2 ex]
-                         & Quadratic &   28\\[1.2 ex]
+                         & Convex    &   \textcolor{red}{3}\\[1.2 ex]
+                         & Quadratic &   \textcolor{red}{30}\\[1.2 ex]
 \hline
-Total                    &           & 116 \\
+Total                    &           & \textcolor{red}{134} \\
 %

### 2018-01-30 QPLIB-3.tex
extend caption of Figure 5
- Hessian is only half
- explain what red means

@@
 \caption{Example for the sparsity pattern of the Jacobian of the constraint functions (left) and
-  of the Hessian of the Lagrangian function (right) for instance QPLIB\_2967.
+  of \textcolor{red}{the upper-right triangle of} the Hessian of the Lagrangian function (right) for instance QPLIB\_2967.
   The gradient of the objective function is displayed as the first row
-  of the Jacobian matrix.}
+  of the Jacobian matrix. \textcolor{red}{Non-constant entries are shown in red.}}
 \label{fig:sparsitypattern}

### 2018-01-30 methods.tex
Update methods.tex
fixup references and add a question as todo
fix typos, add todo where unsure

@@
   \end{itemize}
+
+\textcolor{red}{We refer the interested reader to 
+\cite{book:2077007} and 
+\cite{Michae-2007}\todo{was it really mean to cite the book review instead of the book here?}\ for further details on the solution methods.}
 
@@
 
-To explain this apparent contradiction, we remark that the two statements refer to different computational models: the former is based on the Turing Machine (TM), whereas the latter is based on a computational model that allows operations or real numbers, e.g.~the Real RAM (RRAM) machine \cite{blum}.
+To explain this apparent contradiction, we remark that the two statements refer to different computational models: the former is based on the Turing Machine (TM), whereas the latter is based on a computational model that allows operations on real numbers, e.g.~the Real RAM (RRAM) machine \cite{blum}.
 Due to the potentially infinite nature of exact real arithmetic computations, exact computations on the RRAM necessarily end up being approximate on the TM.
@@
 When the BB algorithm is allowed to branch on coordinate directions corresponding to continuous variables, it is called \emph{spatial} BB (sBB) \cite{dakin,couenne}.
-BB algorithms are exponential time in the worst case, and their exponential behavior unfortunately often shows up in practice.
-They can also be used heuristically (forsaking their completeness guarantee) in a number of ways, e.g.~by either terminating them early.
+BB algorithms are\todo{require?}\ exponential time in the worst case, and their exponential behavior unfortunately often shows up in practice.
+They can also be used heuristically (forsaking their completeness guarantee) in a number of ways, e.g.~by terminating them early.
 The following solvers from Table \ref{t:solvers} implement complete BB algorithms for \textit{QGQ} or some subclasses:

### 2018-02-01 QPLIB-2.1.tex
{\textstyle \half}
Update QPLIB-2.1.tex

@@
  \min \textrm{or} \max \;\;
- & \half x^\top Q^0 x + b^0 x + q^0 \\
+ & {\textstyle \half} x^\top Q^0 x + b^0 x + q^0 \\
  \textrm{such that}\;\;
- & c^i_l \leq \half x^\top Q^i x + b^i x \leq c^i_u & i \in \Mcm, \\
+ & c^i_l \leq {\textstyle \half} x^\top Q^i x + b^i x \leq c^i_u & i \in \Mcm, \\
  %
@@
 
+\textcolor{red}{In the rest of this paper, we shall sometimes refer to {\it exact solutions} of quadratic programs.
+In view of the fact that their solutions may be irrational, this notation deserves a comment.
+If the decision version of the problem being referred to is in NP (e.g.~LP, MILP, QP \cite{vavasis90a}), then the assumption is that all rational numbers can be represented exactly by a Turing Machine (TM).
+%This assumption has the same limitation as the assumption that TMs defined on integers are a faithful abstraction of a real computer, namely that the range of integers needing a representation is bounded.
+If there is no known proof that the problem being solved (or its decision version) is in NP, then there are four main approaches:
+\begin{enumerate}
+\item finding a representable solution $x'$ such that $\|x'-x^\ast\|_\infty\le\varepsilon$, where $x^\ast$ is the true solution, $\varepsilon>0$ is given, and representable means having a polynomially sized description length
+% ``
+(in function of the instance size) \cite{hochbaum2};
+\item using the {\it Thom encoding} of an algebraic number \cite[Prop.~2.28]{pollack} (limited to problems involving polynomials);
+\item using the optimality gap: finding a representable solution $x'$ such that $|f(x')-f(x^\ast)|\le\varepsilon$, where $f$ is the objective function, $x^\ast$ is the true solution, $\varepsilon>0$ is given
+%, and representable is as above 
+(limited to optimization problems);\label{optgap}
+\item using a computational model according to which every elementary computation on the reals takes $O(1)$ and returns a precise result \cite[p.~24]{blum}.
+\end{enumerate}
+Approach \ref{optgap} in the list above is the one most often used in computational papers, including the present one.}
+
 %- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

### 2018-02-02 QPLIB-2.2.tex
Update QPLIB-2.2.tex

@@
   Note that (\textit{D}) and (\textit{C}) are intended to mean that the feasible set of the constraint is convex, i.e., that either $Q^i$ is positive semidefinite and $c_l^i = -\infty$, or $Q^i$ is negative semidefinite and $c_u^i = \infty$.
+    \textcolor{red}{Note that (positive or negative) definiteness of $Q^i$ is a sufficient, but not in general necessary, condition for convexity.
+As detailed in \S\ref{subsec:final set}, in our taxonomy we mark the constraints ``$C$'' based on the sufficient condition alone, the rationale of this choice being discussed in \S\ref{ssec:reform}.}
   Quadratic constraints with both finite bounds cannot ever be convex (unless $Q^i = 0$, i.e., they are not ``truly'' quadratic constraints).

### 2018-02-02 QPLIB-2.2.tex
Update QPLIB-2.2.tex

@@
 
+\textcolor{red}{
+Furthermore, as previously remarked, a not (positive or negative) definite $Q^i$ does not necessarily correspond to a nonconvex feasible region.
+For instance, it is well-known that Second-Order Cone Programs have convex feasible regions; when represented in terms of quadratic constraints, however, they correspond to $Q^i$ in one negative eigenvalue.
+In our taxonomy we still consider the corresponding instances as $**Q$ ones, with no attempt to detect the different special structures that actually correspond to convex feasible regions.
+Although this may lead to classify as ``potentially nonconvex'' some instances that are in fact convex, our choice is justified by the fact that not all QP solvers are capable of detecting and exploiting these structures, which means that the instance can actually be treated as a nonconvex one even if it is not.
+}
+
 % -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -

### 2018-02-02 AnswerToTheReferees.tex
Update AnswerToTheReferees.tex

@@
 \rep{
-TODO:Fabio and Emiliano. 
+TODO:Nick G. 
 %\leo{sono d'accordo, dovremmo aggiungere delle istanze da trust-region subproblem e delle istanze da problemi PDE-constrained.
@@
 
-\rep{TODO:Antonio. 
+\rep{Thank you for having raised this point. We addressed in Sections 2.2.1 and 2.2.2.
+%TODO:Antonio. 
+%lo dicevamo già a p. 10, ma grazie per avercelo ricordato: adesso lo richiamiamo brevemente in §2.2.1 e lo discutiamo in §2.2.2
+%
 %\leo{cosa vuol dire? Questa condizione si puo` testare in polytime? Sicuramente puo` entrare nel discorso a risposta dei punti 2-4 del referee \#1.} 
@@
 
-\rep{TODO: Antonio.
+\rep{
+Although all the facts recalled in §2.2.2 are in fact well known, we still feel they need be discussed in this context because they illustrate how the taxonomy (as almost all of them) has some grey areas whereby the same underlying problem can be described by instances that would be classified differently.
+This is, as the header of the section recalls, strongly tied to the concept of reformulation.
+One relevant example is the relationship between definiteness of the Hessians and convexity, correctly pointed out by the referee, which is naturally discussed in §2.2.2.
+One of the nontrivial choices in our library is that we made no effort to reformulate the instances (although of course we could have), and inserted them in the library in the very same format they have been provided to us by the original contributors.
+Section 2.2.2 is crucial in justifying this choice, as it shows that there are several degrees of freedom to move the instances from one class to another.
+Doing so, however, is typically justified by the fact that this makes one specific solver more efficient in solving them, a bias that we don't want to add.
+Hence, we have chosen to keep the instances in their ``natural'' format, this being the one in which the original contributor initially wrote them.
+%TODO: Antonio.
 %\leo{cancelliamo tutto!} 
@@
 
-\rep{TODO: Leo. 
+\rep{We added a comment about this issue in Section 2.1 
 %\leo{possiamo inserire qualche nota sui modelli di computazione per i numeri reali, razionali e floating point: (a) real numbers con real RAM model, in pratica tutti i calcoli tra numeri reali sono O(1) (la gran parte del lavoro fatto da Blum, Shub e Smale e` sulla definizione di calcolabilita` e complessita` in questo 'ambiente');
@@
 
-\rep{
-TODO: Stefan and Ambros.
-%\leo{stesso commento dell'AE -- dobbiamo mettere piu` istanze} 
-}
+\rep{We followed your suggestion and we gathered several additional instances and we included to the final version of the library (see the reply to the AE for further details).}
 

### 2018-02-02 QPLIB-3.tex
Update QPLIB-3.tex

@@
 
+\textcolor{red}{
+One of the nontrivial choices in our library is that we made no effort to reformulate the instances, and inserted them in the library in the very same format they have been provided to us by the original contributors.
+Section \ref{ssec:reform} is crucial in justifying this choice, as it shows that there are several degrees of freedom to move the instances from one class to another.
+Tailoring the structure of a problem to the solver is, however, a bias we do not want to add.
+}
+
 We now report some graphs that help in illustrating the main features

### 2018-02-02 AnswerToTheReferees.tex
Update AnswerToTheReferees.tex

@@
 
-\rep{TODO: Leo
+\rep{We thank the referee for bringing up this point. We addressed it in two comments in Sections 2.2.1 and 2.2.2. 
+%TODO: Leo
 % \leo{sono d'accordo col referee (ovvero: massi`, facciamolo contento)} \fabio{va qui? o e' un commento al punto di sotto?}
@@
 \rep{
-Although all the facts recalled in §2.2.2 are in fact well known, we still feel they need be discussed in this context because they illustrate how the taxonomy (as almost all of them) has some grey areas whereby the same underlying problem can be described by instances that would be classified differently.
+Although all the facts recalled in \S 2.2.2 are in fact well known, we still feel they need to be discussed in this context because they illustrate how the taxonomy (as almost all of them) has some grey areas whereby the same underlying problem can be described by instances that would be classified differently.
 This is, as the header of the section recalls, strongly tied to the concept of reformulation.
-One relevant example is the relationship between definiteness of the Hessians and convexity, correctly pointed out by the referee, which is naturally discussed in §2.2.2.
+One relevant example is the relationship between definiteness of the Hessians and convexity, correctly pointed out by the referee, which is naturally discussed in \S 2.2.2.
 One of the nontrivial choices in our library is that we made no effort to reformulate the instances (although of course we could have), and inserted them in the library in the very same format they have been provided to us by the original contributors.

### 2018-02-02 QPLIB-2.2.tex
Update QPLIB-2.2.tex

@@
  \item constraints: (\textit{N})one, (\textit{B})ox, (\textit{L})inear, (\textit{D})iagonal convex quadratic, (\textit{C})on\-vex quadratic, nonconvex (\textit{Q})uadratic.
-  Note that (\textit{D}) and (\textit{C}) are intended to mean that the feasible set of the constraint is convex, i.e., that either $Q^i$ is positive semidefinite and $c_l^i = -\infty$, or $Q^i$ is negative semidefinite and $c_u^i = \infty$.
+  Note that (\textit{D}) and (\textit{C}) are intended to mean that either $Q^i$ is positive semidefinite and $c_l^i = -\infty$, or $Q^i$ is negative semidefinite and $c_u^i = \infty$. 
     \textcolor{red}{Note that (positive or negative) definiteness of $Q^i$ is a sufficient, but not in general necessary, condition for convexity.

### 2018-02-02 AnswerToTheReferees.tex
Update AnswerToTheReferees.tex

@@
 
-\rep{ TODO: Leo 
+\rep{We removed a sentence from this paragraph to address this concern
+%TODO: Leo 
 %\leo{secondo me il secondo commento si riferisce a pagina 19...
@@
 \rep{
-TODO: Stefan and Ambros.
+We renamed ``hard" to ``problematic".
+However, as explained in \S 2.2.1 and 2.2.2, we refrain from reformulating the instance, and we do not tailor problems to specific solvers (moreover, not all QP solvers would resort to reduced Hessians).
+In the same vein, our analysis can only be based on the given instance structure.
+%TODO: Stefan and Ambros.
 % \leo{non so nulla di "hard eigenvalues" nel senso del trust region; chiediamo agli esperti} 

### 2018-02-02 AnswerToTheReferees.tex
Update AnswerToTheReferees.tex

@@
 \rep{
-TODO:Nick G. 
+Thank you for pointing out the missing categories. As requested, we updated Table 1 with the two required Applications.
+We also included PDE-constrained instances in the updated version of the library.
+We have deliberately excluded the highly-specialized class of trust-region subproblems, since there are dedicated, very fast solvers for this class. See for example:
+\begin{itemize}
+\item
+Minimizing a quadratic over a sphere. \\
+W. W. Hager. 
+SIAM Journal on Optimization,12(1):188-208, 2001.
+\item
+Solving the trust-region subproblem using the Lanczos method. \\
+N. I. M. Gould, S. Lucidi, M. Roma, and Ph. L. Toint. 
+SIAM Journal on Optimization, 9(2):504-525, 1999.
+\item
+A semidefinite framework for trust region subproblems with applications to large scale minimization.\\
+F Rendl, H Wolkowicz.
+Mathematical Programming 77 (1), 273-299
+\item
+A Subspace Minimization Method for the Trust-Region Step.\\
+J. B. Erway and P. E. Gill.
+SIAM J. Optim. 20-3 (2010), pp. 1439-1461
+\item
+Solving the Trust-Region Subproblem By a Generalized Eigenvalue Problem.\\
+S. Adachi, S. Iwata, Y. Nakatsukasa and A Takeda.
+SIAM Journal on Optimization 27 (1), 269-291
+\item
+On solving trust-region and other regularised subproblems in optimization.\\
+N.I.M. Gould, D.P. Robinson and H.S. Thorne.
+Mathematical Programming Computation 2 (1), 21-57
+\end{itemize}
 %\leo{sono d'accordo, dovremmo aggiungere delle istanze da trust-region subproblem e delle istanze da problemi PDE-constrained.

### 2018-02-03 QPLIB-2.2.tex
Update QPLIB-2.2.tex

@@
 
-\Leo{\sout{Following the remark of one of our anonymous referees,} We note that, although the left-to-right progression in the above text marks ``harder problems'', this does \textit{not} mean that \textit{every} instance of a hard problem is itself ``difficult to solve''.
+\Leo{We note that, although the left-to-right progression in the above text marks ``harder problems'', this does \textit{not} mean that \textit{every} instance of a hard problem is itself ``difficult to solve''.
 As is well known, problems as infinite collections of instances; and it is always possible to find subclasses that can be solved efficiently.}

### 2018-02-03 methods.tex
Update methods.tex

@@
 \cite{book:2077007} and 
-\cite{Michae-2007}\todo{was it really mean to cite the book review instead of the book here?}\ for further details on the solution methods.}
+\cite{lee2006quadratic} for further details on the solution methods.}
 
@@
 When the BB algorithm is allowed to branch on coordinate directions corresponding to continuous variables, it is called \emph{spatial} BB (sBB) \cite{dakin,couenne}.
-BB algorithms are\todo{require?}\ exponential time in the worst case, and their exponential behavior unfortunately often shows up in practice.
+BB algorithms require exponential time in the worst case, and their exponential behavior unfortunately often shows up in practice.
 They can also be used heuristically (forsaking their completeness guarantee) in a number of ways, e.g.~by terminating them early.

### 2018-02-03 QPLIB-3.tex
Update QPLIB-3.tex

@@
 In Table~\ref{tab:filters} we summarize the two filter steps, which
-allowed us to identify the final set of $251$ discrete instances and
-$116$ continuous instances.\todo{needs update due to newly added instances ($\to$ 319 discrete 134 cont.)}
+allowed us to identify the final set of $319$ discrete instances and
+$134$ continuous instances.
 
@@
 % & 600 Discr. Inst.  & 250  Cont. inst. & \\
-  & 251 Discr. Inst.  & 116  Cont. inst. & \\
+  & 319 Discr. Inst.  & 134  Cont. inst. & \\
 \end{tabular}

### 2018-02-03 QPLIB-3.tex
Update QPLIB-3.tex

@@
 Conic library CBLIB (\url{http://cblib.zib.de}) and 
-MIPLIB 2010 (http://miplib.zib.de/), to mention just a few.}
+MIPLIB 2010 (\url{http://miplib.zib.de/}), to mention just a few.}
 

### 2018-02-03 AnswerToTheReferees.tex
Update AnswerToTheReferees.tex

@@
  Following the Area Editor suggestions, we 
- performed a deep revision of our manuscript. We carefully took into account all the comments made by the Referees and the Associate Editor.
- [\dots]
+ performed a revision of our manuscript. We carefully took into account all the comments made by the Referees and the Associate Editor.
  The new submitted document highlights in red the text that has been
@@
 We performed an internal call for instances among the members of the committee. 
-Thanks to this new research, we gathered $86$ new instances and we added them to the library. The updated division of the instances is as follows:\\
-\begin{tabular}{lllr}
-\toprule
-Obj. Fun. & Variables & Constraints & \#\\
-\cmidrule(lr){1-4}
-%
-\multirow{6}*{Linear}
-          & \multirow{2}*{Binary}
-%                    & None      &   \\
-%          &         & Linear    &  \\
-                    & Convex    &   0 +\textcolor{red}{xx}\\[1.2 ex]
-           &         & Quadratic &   9 \\[1.2 ex]
-\cmidrule(lr){2-4}
-          & \multirow{2}*{Mixed}
-                    & Convex    &   2 +\textcolor{red}{xx}\\[1.2 ex]
-          &         & Quadratic &    118\\[1.2 ex]
-\cmidrule(lr){2-4}
-          & \multirow{1}*{Integer}
-%                    & Linear    &    \\
-                   & Quadratic &    2\\[1.2 ex]
-\cmidrule(lr){2-4}
-          & \multirow{1}*{General}
-%                    & Linear    &    \\
-                   & Quadratic &    3\\[1.2 ex]
-\cmidrule(lr){1-4}
-%\multirow{5}*{Linear}
-%          & Binary  & Quadratic &   9\\
-%          & \multirow{2}*{Mixed}
-%                    & Convex    &   2\\
-%          &         & Quadratic &  118\\
-%%          & \multirow{2}*{Integer}
-%%                    & Convex    &  15 \\
-%%          &         & Quadratic &   5 \\
-%          & {General} & Quadratic &   3 \\
-%\hline
-\multirow{1}*{Convex (if min)}
-          & Binary  & Linear    &  2  +\textcolor{red}{xx}\\[1.2 ex]
-\cmidrule(lr){2-4}
-\multirow{1}*{or}
-          & \multirow{2}*{Mixed}
-                    & Linear    &   12\\[1.2 ex]
-\multirow{1}*{Concave (if max)}
-          &         & Quadratic &    1 +\textcolor{red}{xx}\\[1.2 ex]
-%          & General & Linear    &    \\
-%\cmidrule(lr){2-4}
-%          & \multirow{1}*{Integer}
-%                    & Linear    &   1\\[1.2 ex]
-%\hline
-\cmidrule(lr){1-4}
-\multirow{6}*{Quadratic}
-          & \multirow{3}*{Binary}
-                    & None      &   23\\[1.2 ex]
-          &         & Linear    &  59\\[1.2 ex]
-          &         & Quadratic &   5 \\[1.2 ex]
-\cmidrule(lr){2-4}
-          & \multirow{2}*{Mixed}
-                    & Linear    &   11\\[1.2 ex]
-          &         & Quadratic &    1\\[1.2 ex]
-\cmidrule(lr){2-4}
-          & Integer & Linear    &    2\\[1.2 ex]
-\cmidrule(lr){2-4}
-          & \multirow{1}*{General}
-                    & Quadratic    &    1\\[1.2 ex]
-%          &         & Quadratic &    \\
-\hline
-Total Discrete     &         &           & 251 +\textcolor{red}{xx}\\
-%
-\bottomrule
-\end{tabular}\\\\
-%
-%
-\begin{tabular}{llr}
-\toprule
-Obj. Fun. & Constraints & \#\\
-\cmidrule(lr){1-3}
-%
-\multirow{2}*{Linear}    & Convex    &   13\\[1.2 ex]
-                         & Quadratic &   48\\[1.2 ex]
-\cmidrule(lr){1-3}
-\multirow{1}*{Convex (if min)}
-                         & Box       &   3 \\[1.2 ex]
-\multirow{1}*{or}
-                         & Linear    &   11\\[1.2 ex]
-                        % & Convex    &    2\\[1.2 ex]
-\multirow{1}*{Concave (if max)}
-                         & Quadratic &    6\\[1.2 ex]
-\cmidrule(lr){1-3}
-\multirow{3}*{Quadratic}
-                         & Linear    &   6\\[1.2 ex]
-                         & Convex    &   1\\[1.2 ex]
-                         & Quadratic &   28\\[1.2 ex]
-\hline
-Total Continuous                   &           & 116 +\textcolor{red}{18} \\
-%
-\bottomrule
-\end{tabular}\\
-All the new instances 
+We gathered $86$ new instances (belonging to the categories less that were not well represented) and we added them to the library.
+We added a macro classification of the instances, based on convexity and nature of the variables.
+We believe that the majority of the algorithms used in the literature falls in one of these category.
+This macro classification shows that the library is now well-balanced.
+We kept also the detailed classification since it allows a fine grained analysis of the instances.
 %%
@@
 
-\begin{table}
- \centering
- %\scriptsize
- \setlength{\tabcolsep}{18pt}
- \renewcommand \arraystretch{1.1}
-\begin{tabular}{llr}
-\toprule
-Variables & Convexity & \#\\
-\cmidrule(lr){1-3}
-%
- Continuous	& Convex		&  27+\textcolor{red}{yy} \\[1.2 ex]
- Continuous	& Non Convex	&  89+\textcolor{red}{yy} \\[1.2 ex]
- Discrete	& Convex		&  16+\textcolor{red}{63} \\[1.2 ex]
- Discrete	& Non Convex	&  235+\textcolor{red}{5} \\[1.2 ex]
-\cmidrule(lr){1-3}
-Total		&				& 367+\textcolor{red}{86}\\
-%
-\bottomrule
-\end{tabular}
-\caption{Classification of the final set of discrete instances}
-\label{tab:DD}
-\end{table}
+%\begin{table}[h1]
+%{
+% \centering
+% %\scriptsize
+% \setlength{\tabcolsep}{18pt}
+% \renewcommand \arraystretch{1.1}
+%\begin{tabular}{llr}
+%\toprule
+%Variables & Convexity & \#\\
+%\cmidrule(lr){1-3}
+%%
+% \textit{continuous}	& \textit{convex}		&  32 \\[1.2 ex]
+% \textit{continuous}	& \textit{non-convex} 	&  102 \\[1.2 ex]
+% \textit{discrete}	& \textit{convex}		&  31 \\[1.2 ex]
+% \textit{discrete}	& \textit{non-convex}	&  288 \\[1.2 ex]
+%\cmidrule(lr){1-3}
+%Total		&				& 453\\
+%%
+%\bottomrule
+%\end{tabular}
+%\caption{Macro classification of the final set of instances}
+%\label{tab:DDDDD}
+%}
+%\end{table}
 
@@
 In the following we present the improvements we made in accordance with the comments of the referees.
-[\dots] }
+}
 
@@
 
-\rep{[\dots] }
+\rep{We thank the referee for his/her positive opinion on our manuscript. We hope that the new library will serve our scientific community.}
 
@@
 
-\rep{We updated the manuscript by adding the suggest link to the Conic library.}
+\rep{We updated the manuscript by adding the suggest link.}
 
@@
 
-\rep{We removed a sentence from this paragraph to address this concern
+\rep{We removed a sentence from this paragraph to address this concern and we added a paragraph on this issue.
 %TODO: Leo 
@@
 
-\rep{ TODO:Fabio and Emiliano
+\rep{We wrote an email to the development team of Cplex and they confirmed that their solver is able to tackle QGL and CGC 
+% TODO:Fabio and Emiliano
 %\leo{rispondiamo al referee che CPLEX e` classificato 'C' (complete) in tavola 2 rispetto a QGC e ricordiamogli che pero` CPLEX non risolve (a mia conoscenza e per il momento) QGQ.
@@
 
-\rep{[\dots] }
+\rep{We followed the suggestion of the referee ad the AI and we have shortened unnecessary sections of the paper.}
 
@@
 
-\item The description of NLP solvers in Section 2.3.2 is at somewhat misleading. For example, KNI-
-TRO is also a primal-dual interior-point method, not just a barrier code (in fact primal-dual
+\item The description of NLP solvers in Section 2.3.2 is at somewhat misleading. For example, KNITRO is also a primal-dual interior-point method, not just a barrier code (in fact primal-dual
 methods are preferred). The description of the “active-set” methods is wrong, unless you refer
@@
 
-\rep{TODO: Stefan and Ambros
+\rep{The complete description of the solver goes beyond the scope of the paper. For this reason we have streamlined the related Section.
 %\leo{correggiamo gli errori e ringraziamo il referee} 
@@
 \rep{
-TODO: Stefan and Ambros
+Added.
 %\leo{mettiamo la label al y axis  (duh)} 
@@
 
-\rep{TODO: Stefan and Ambros}
+\rep{Corrected, thanks.}
 
@@
 \rep{
-We thank the refereee for having raised this delicate point. We definetly agree that formats are an issue.
+We thank the referee for having raised this delicate point. We definitely agree that formats are an issue.
 For this reason we have translated all the instances also into the AMPL format and we have uploaded them on the website.

### 2018-02-03 QPLIB-3.tex
Update QPLIB-3.tex

@@
 
-We now analyse the features of the instances selected to be part of the
-instance library. The characteristics of the instances are presented in Table
+\textcolor{red}{We now analyse the features of the instances selected to be part of the
+ library.
+In Table \ref{tab:DDDDD}, we provide a global overview.
+The instances have been divided in \textit{continuous} vs \textit{discrete} and \textit{convex} vs \textit{non-convex}, forming in this way, a  classification of 4 macro categories.
+As previously mentioned, an instance is classified \textit{discrete} if it contains at least a binary or integer variable, and \textit{continuous} otherwise.
+On the other hand, an instance is classified as \textit{non-convex} if the objective function is non-convex and/or at least one of the constraints is  non-convex, and \textit{convex} otherwise.
+}
+
+\begin{table}
+\textcolor{red}{
+ \centering
+ %\scriptsize
+ \setlength{\tabcolsep}{18pt}
+ \renewcommand \arraystretch{1.1}
+\begin{tabular}{llr}
+\toprule
+Variables & Convexity & \#\\
+\cmidrule(lr){1-3}
+%
+ \textit{continuous}	& \textit{convex}		&  32 \\[1.2 ex]
+ \textit{continuous}	& \textit{non-convex} 	&  102 \\[1.2 ex]
+ \textit{discrete}	& \textit{convex}		&  31 \\[1.2 ex]
+ \textit{discrete}	& \textit{non-convex}	&  288 \\[1.2 ex]
+\cmidrule(lr){1-3}
+Total		&				& 453\\
+%
+\bottomrule
+\end{tabular}
+\caption{Macro classification of the final set of instances}
+\label{tab:DDDDD}
+}
+\end{table}
+
+ The detailed characteristics of the instances are presented in Table
 \ref{tab:DD} for \emph{discrete} instances
@@
 On the theme of nonconvexity, Figure~\ref{fig:pic_neg_eig} (left) focuses on
-the instances with a quadratic objective function, ordered by the
-the percentage ``hard'' eigenvalues in the Hessian $Q^0$ (vertical axis), by which
+the instances with a quadratic objective function, ordered by 
+ percentage of ``problematic'' \textcolor{red}{(defined using a tollerance of XXX)} \todo{please define the tolerance used to identify the ``Problematic'' eigenvalues and change the label of the plot}\ eigenvalues in the Hessian $Q^0$ (vertical axis), by which
 we mean negative eigenvalues in case of a minimization problem and positive eigenvalues in case of a maximization problem.
 The instances are mostly clustered around two values. About 25\% of the instances have a convex (if minimization) or concave (if maximization) objective function,
-i.e., they have 0\% of ``hard'' eigenvalues. Among the others, a vast
-majority has around 50\% of ``hard'' eigenvalues. However, instances
-with high or low percentages of ``hard'' eigenvalues are present, too.
+i.e., they have 0\% of ``problematic'' eigenvalues. Among the others, a vast
+majority has around 50\% of ``problematic'' eigenvalues. However, instances
+with high or low percentages of ``problematic'' eigenvalues are present, too.
 
@@
   \includegraphics[width=0.49\textwidth]{pic_density.pdf}
-  \caption{``Hard'' eigenvalues (left) and density (right) of the Hessian $Q^0$ for QPLIB instances with a quadratic objective function. \label{fig:pic_neg_eig}}
+  \caption{``Problematic'' eigenvalues (left) and density (right) of the Hessian $Q^0$ for QPLIB instances with a quadratic objective function. \label{fig:pic_neg_eig}}
 \end{figure}

### 2018-02-03 QPLIB-2.2.tex
Update QPLIB-2.2.tex

@@
 In the following discussion we will repeatedly exploit this by assuming that, unless otherwise mentioned, when a method can be applied to a given problem, it can be applied as well to all simpler problems where the value of each field is arbitrarily replaced with a value denoting a less-general class.
-
-\Leo{We note that, although the left-to-right progression in the above text marks ``harder problems'', this does \textit{not} mean that \textit{every} instance of a hard problem is itself ``difficult to solve''.
-As is well known, problems as infinite collections of instances; and it is always possible to find subclasses that can be solved efficiently.}
 

### 2018-02-04 QPLIB-2.2.tex
consistent title case in section headers

@@
 % -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -
-\subsubsection{Examples and reformulations}\label{ssec:reform}
+\subsubsection{Examples and Reformulations}\label{ssec:reform}
 
@@
 %
-i.e., a \textit{L*Q} (\textit{L*C}) pronlem.
+i.e., a \textit{L*Q} (\textit{L*C}) problem.
 Hence, it is clear that quadratic constraints are, in a well-defined sense, the most general situation (cf.~also the result above about hardness of \textit{LIQ}).
@@
 % -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -
-\subsubsection{QP classes}\label{ssec:classes}
+\subsubsection{QP Classes}\label{ssec:classes}
 
@@
 In this sense, like in the linear case, there is a significant distinction between solvers that need all the data of QP to work, and those that are ``matrix-free'', i.e., only require the application of simple operations (typically, matrix-vector products) with the problem data.
-While when building our instance library we never exploited such characteristics, since they are not amenable to standard modelling tools, but this may be relevant for the solution of very-large-scale \textit{CIC}.
+While when building our instance library we never exploited such characteristics, since they are not amenable to standard modeling tools, but this may be relevant for the solution of very-large-scale \textit{CIC}.
  %

### 2018-02-04 QPLIB-3.tex
consistent title case in section headers

@@
 optimization and numerical analysis communities, reaching in this way
-the largest possible set of interested researchers and practitioners.
-The call remained open for 10 months, during which we received a large
+the largest\todo{AG: a large?} possible set of interested researchers and practitioners.
+The call remained open for ten months, during which we received a large
 number of contributions of different nature. The instances we gathered
@@
 
-In addition to spontaneous contribution we analysed the other generic
-libraries of instances available  on internet and containing QP
+In addition to spontaneous contribution we analyzed the other generic
+libraries of instances available  on the internet and containing QP
 instances. Namely, the libraries from which we gathered instances are
@@
 Conic library CBLIB (\url{http://cblib.zib.de}) and 
-MIPLIB 2010 (\url{http://miplib.zib.de/}), to mention just a few.}
+MIPLIB~2010 (\url{http://miplib.zib.de/}), to mention just a few.}\todo{AG: maybe cite CBLIB paper \cite{Friberg2016}}
 
@@
 Note that this might occasionally lead us to classify some instances that have conic constraints as non-convex ones, although their feasible region is in fact convex---fortunately, only some solvers are capable of properly exploiting this property.
- All this information allowed us to analyse the gathered instances and to perform the
+ All this information allowed us to analyze the gathered instances and to perform the
 filters described in the the next paragraph.
@@
        The goal of the second filter was to eliminate ``similar'' instances.
-       We carefully analysed the instances one by one, and we clustered them
+       We carefully analyzed the instances one by one, and we clustered them
        according to their features; for each cluster we kept only a few
@@
 %- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-\subsection{Analysis of the final set of instances}\label{subsec:final set}
+\subsection{Analysis of the Final Set of Instances}\label{subsec:final set}
 
-\textcolor{red}{We now analyse the features of the instances selected to be part of the
+\textcolor{red}{We now analyze the features of the instances selected to be part of the
  library.
@@
 the instances with a quadratic objective function, ordered by 
- percentage of ``problematic'' \textcolor{red}{(defined using a tollerance of XXX)} \todo{please define the tolerance used to identify the ``Problematic'' eigenvalues and change the label of the plot}\ eigenvalues in the Hessian $Q^0$ (vertical axis), by which
+ percentage of ``problematic'' \textcolor{red}{(defined using a tolerance of XXX)} \todo{please define the tolerance used to identify the ``Problematic'' eigenvalues and change the label of the plot}\ eigenvalues in the Hessian $Q^0$ (vertical axis), by which
 we mean negative eigenvalues in case of a minimization problem and positive eigenvalues in case of a maximization problem.

### 2018-02-04 QPLIB-1.0.tex
Update QPLIB-1.0.tex
fix typos

@@
 It is, therefore, perhaps not surprising that, unlike for ``simpler'' problems classes \cite{Koch2011},  there has been, to date, no single library devoted to all different kinds of instances of QP.
-While several specialised libraries devoted to particular cases of QP are available, each of them is either focussed on a particular application (a specific problem that can be modelled as a QP), or on QPs with specific structural properties that make them suitable for solution by some given class of algorithmic approaches.
+While several specialized libraries devoted to particular cases of QP are available, each of them is either focused on a particular application (a specific problem that can be modeled as a QP), or on QPs with specific structural properties that make them suitable for solution by some given class of algorithmic approaches.
 To the best of our knowledge, collecting a set of QP instances that is at the same time not overwhelmingly onerous but sufficiently challenging for the many different interested communities has not been attempted.
@@
 %
-%\multicolumn{3}{l}{Process networks optimisation problems typically involve two} \\
+%\multicolumn{3}{l}{Process networks optimization problems typically involve two} \\
 %
@@
 %
-Controlled Tabular Adjustment & \checkmark & \cite{CaFG14}
+Controlled Tabular Adjustment & \checkmark & \cite{CaFG14} \\
+\midrule
+%
+\multicolumn{3}{l}{\textbf{Trust-region methods}} \\[1pt]
+%
+Trust-region subproblem & & \cite{hager2001,Gould1999,Rendl1997,Erway2010,Adachi2017,Gould2010}\\
+\midrule
+%
+\multicolumn{3}{l}{\textbf{PDE-constrained optimization}} \\[1pt]
+%
+Optimal control problem & & \cite{DIPILLO1983101,Stojanovic1991,Schittkowski1979}
 \end{longtable}

### 2018-02-04 conclusion.tex
fix typos

@@
 
- This manuscript describes the first comprehensive library of instances for Quadratic Programming (QP).
-Since QP comprises  different and ``varied'' categories of problems, we proposed a classification and we briefly discuss the main classes of solution methods for QP.
+This manuscript described the first comprehensive library of instances for Quadratic Programming (QP).
+Since QP comprises  different and ``varied'' categories of problems, we proposed a classification and we briefly discussed the main classes of solution methods for QP.
 
-We then describe the steps of the adopted process used to filter the gathered instances  in order to build the new library.
+We then described the steps of the adopted process used to filter the gathered instances  in order to build the new library.
 Our design goals were to build a library which is computationally challenging and as broad as possible, i.e., it represents the largest possible categories of QP problems, while remaining of manageable size.
@@
 
-We want to stress once again that we intentionally avoid to perform a computational comparison of the performances of the different solvers.
-Our goal is instead to provide a common test-bed of instances for practitioners and researchers in the field.
+We want to stress once again that we intentionally avoided to perform a computational comparison of the performances of the different solvers.
+Our goal was instead to provide a common test bed of instances for practitioners and researchers in the field.
 This new library will hopefully serve as a point of reference to test new ideas and algorithms for QP problems.

### 2018-02-04 methods.tex
fix typos

@@
 This term has to be used with care, as it implicitly makes assumptions on the underlying computational model that may not be acceptable in all cases.
-For example, te decision version of \textit{QCL} is known to be in the complexity class \textbf{NP} \cite{vavasis90a}, whereas the same is not known about \textit{LCQ}, even with zero objective.
+For example, the decision version of \textit{QCL} is known to be in the complexity class \textbf{NP} \cite{vavasis90a}, whereas the same is not known about \textit{LCQ}, even with zero objective.
 On the other hand, there exists a method for deciding feasibility of systems of polynomial equations and inequalities \cite{tarski-reals}, including the solution of \textit{LCQ} with zero objective function.

### 2018-02-04 conclusion.tex
remarks on conclusion

@@
 
-This manuscript described the first comprehensive library of instances for Quadratic Programming (QP).
+\todo [inline]{AG: We do not draw many conclusions here, maybe call it ``Final Remarks''?}
+
+This manuscript\todo{AG: paper/article?
+Manuscript usually refers to pre-final versions I think} described the first comprehensive library of instances for Quadratic Programming (QP).
 Since QP comprises  different and ``varied'' categories of problems, we proposed a classification and we briefly discussed the main classes of solution methods for QP.
-
 We then described the steps of the adopted process used to filter the gathered instances  in order to build the new library.
 Our design goals were to build a library which is computationally challenging and as broad as possible, i.e., it represents the largest possible categories of QP problems, while remaining of manageable size.
-We have also proposed a stand-alone QP format that is intended
+We also proposed a stand-alone QP format that is intended
 for the convenient exchange and use of our QP instances.
 
-We want to stress once again that we intentionally avoided to perform a computational comparison of the performances of the different solvers.
+We want to stress once again that we intentionally avoided to perform a computational comparison of the performances of different solvers\todo{AG: maybe ``solution methods'' instead?
+Since not all algorithms described are actually available as what I would call a solver.}.
 Our goal was instead to provide a common test bed of instances for practitioners and researchers in the field.
-This new library will hopefully serve as a point of reference to test new ideas and algorithms for QP problems.
+This new library will hopefully serve as a point of reference to inspire and test new ideas and algorithms for QP problems.
+
+\todo[inline]{AG: Maybe as last point add that it is supposed to be a
+  dynamically growing library that will be updated as new instance types become
+  available, especially from currently under- or not-represented classes.}

### 2018-02-04 QPLIB-1.0.tex
suggestions

@@
 This is illustrated by the fact that the class of problems that we simply refer to here as ``QP'' might more accurately be called Mixed-Integer Quadratically-Constrained Quadratic Programming (MIQCQP) in the most general case.
-It is, therefore, perhaps not surprising that, unlike for ``simpler'' problems classes \cite{Koch2011},  there has been, to date, no single library devoted to all different kinds of instances of QP.
-While several specialized libraries devoted to particular cases of QP are available, each of them is either focused on a particular application (a specific problem that can be modeled as a QP), or on QPs with specific structural properties that make them suitable for solution by some given class of algorithmic approaches.
+It is, therefore, perhaps not surprising that, unlike for ``simpler'' problems classes like Mixed-Integer Linear Programming \cite{Koch2011},  there has been, to date, no single library devoted to all different kinds of instances of QP.\todo{AG: many commas in this sentence} While several specialized libraries devoted to particular cases of QP are available, each of them is either focused on a particular application (a specific problem that can be modeled as a QP), or on QPs with specific structural properties that make them suitable for solution by some given class of algorithmic approaches.
 To the best of our knowledge, collecting a set of QP instances that is at the same time not overwhelmingly onerous but sufficiently challenging for the many different interested communities has not been attempted.

### 2018-02-04 QPLIB.tex
more accurate description of the advantages of .qplib provided
remove todonotes include
packages
fixup references and add a question as todo
fix typos
suggestions

@@
 \usepackage{microtype}
-
-% TODO remove these before submission
+\usepackage{graphicx}
+\usepackage[normalem]{ulem}
 \usepackage{todonotes}
-%\usepackage{showlabels}
 
@@
 \newcommand{\half}{\frac{1}{2}}
+\newcommand{\Leo}[1]{{\color{red}#1}}
 
@@
 
-%\listoftodos
 \newpage
@@
  This paper describes a new instance library for Quadratic Programming (QP), i.e., the family of continuous and (mixed)-integer optimization problems where the objective function, the constrains, or both are quadratic.
- QP is a very ``varied'' class of problems, comprising sub-classes of problems ranging from
- trivial to undecidable. Solution methods for QP are very diverse, ranging from
+ QP is a very diverse class of problems, comprising sub-classes of problems ranging from
+ trivial to undecidable. This diversity is reflected in the variety of solution methods for QP, ranging from
  entirely combinatorial ones to completely continuous ones, including many for
@@
 %- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-\section{Acknowledgements}
+\section{Acknowledgments}
 
@@
 The work of the fifth and twelfth author was supported by the Research Campus MODAL \emph{Mathematical Optimization and Data Analysis Laboratories} funded by the Federal Ministry of Education and Research (BMBF Grant~05M14ZAM).
-All responsibilty for the content of this publication is assumed by the authors.
+The work of the sixth author was supported by the EPSRC grant EP/M025179/1.
+All responsibility for the content of this publication is assumed by the authors.
 

### 2018-02-05 QPLIB-3.tex
fix typos
suggestions
update on qplib statistics after fixing issue in conversion of some new instances
- for some of the new instances, the original quadratic objective made it into a constraint
- 1 instance moved from LMC to CBL and 17 instances moved from LMQ to QBL

--- 2018-02-03 QPLIB-3.tex
@@
           & \multirow{2}*{Mixed}
-                    & Convex    &   \textcolor{red}{15}\\[1.2 ex]
-          &         & Quadratic &  \textcolor{red}{151}\\[1.2 ex]
+                    & Convex    &   \textcolor{red}{14}\\[1.2 ex]
+          &         & Quadratic &  \textcolor{red}{134}\\[1.2 ex]
 \cmidrule(lr){2-4}
@@
 \multirow{1}*{Convex (if min)}
-          & Binary  & Linear    &  \textcolor{red}{4} \\[1.2 ex]
+          & Binary  & Linear    &  \textcolor{red}{5} \\[1.2 ex]
 \cmidrule(lr){2-4}
@@
                     & None      &   23\\[1.2 ex]
-          &         & Linear    &  \textcolor{red}{74}\\[1.2 ex]
+          &         & Linear    &  \textcolor{red}{91}\\[1.2 ex]
           &         & Quadratic &   5 \\[1.2 ex]

### 2018-02-05 QPLIB-3.tex
add eigval tolerance

@@
 the instances with a quadratic objective function, ordered by 
- percentage of ``problematic'' \textcolor{red}{(defined using a tollerance of XXX)} \todo{please define the tolerance used to identify the ``Problematic'' eigenvalues and change the label of the plot}\ eigenvalues in the Hessian $Q^0$ (vertical axis), by which
-we mean negative eigenvalues in case of a minimization problem and positive eigenvalues in case of a maximization problem.
+ percentage of ``problematic'' eigenvalues in the Hessian $Q^0$ (vertical axis), by which
+we mean \textcolor{red}{eigenvalues below $-10^{-12}$} in case of a minimization problem and \textcolor{red}{eigenvalues above $10^{-12}$} in case of a maximization problem.
 The instances are mostly clustered around two values. About 25\% of the instances have a convex (if minimization) or concave (if maximization) objective function,

### 2018-02-05 AnswerToTheReferees.tex
some typoes corrected

@@
 We added a macro classification of the instances, based on convexity and nature of the variables.
-We believe that the majority of the algorithms used in the literature falls in one of these category.
+We believe that the majority of the algorithms used in the literature falls in one of these categories.
 This macro classification shows that the library is now well-balanced.
@@
 
-\rep{We followed the suggestion of the referee ad the AI and we have shortened unnecessary sections of the paper.}
+\rep{We followed the suggestion of the referee and the AE and we have shortened unnecessary sections of the paper.}
 
@@
 \rep{
-Thank you for pointing out the missing categories. As requested, we updated Table 1 with the two required Applications.
+Thank you for pointing out the missing categories. As requested, we updated Table 1 with the two required applications.
 We also included PDE-constrained instances in the updated version of the library.
@@
 
-\rep{Thank you for having raised this point. We addressed in Sections 2.2.1 and 2.2.2.
+\rep{Thank you for having raised this point. We addressed it in Sections 2.2.1 and 2.2.2.
 %TODO:Antonio. 
@@
 
-\rep{We followed your suggestion and we gathered several additional instances and we included to the final version of the library (see the reply to the AE for further details).}
+\rep{We followed your suggestion and we gathered several additional instances which we included in the final version of the library (see the reply to the AE for further details).}
 

### 2018-02-05 QPLIB-1.0.tex
Update QPLIB-1.0.tex

--- 2017-02-03 QPLIB-1.0.tex
@@
 A consequence of our focus is that this paper is \emph{not} concerned with the performance of the very diverse available set of QP solvers; we will \emph{not} report any data comparing them.
-The only reason that solvers are used (and, therefore, described) in this context is to ensure that the instances of the library are nontrivial---at least for a significant fraction of the current solution methods.
+The only reason that solvers are used (and, therefore, described) in this context is to ensure that the instances of the library are nontrivial, at least for a significant fraction of the current solution methods.
 Providing guidance about which solvers are most suited to some specific class of QPs is entirely outside the scope of our work.
@@
 This is, in principle, true for MILP as well, but at the cost of much larger and much more complicated formulations.
-Hence, for many applications QP may represent the ``sweet spot'' between the effectiveness, but lower expressive power, of MILP and the higher expressive power, but much higher computational cost of MINLP.
+Hence, for many applications QP may represent the ``sweet spot'' between the effectiveness, but lower expressive power, of MILP and the higher expressive power, but much higher computational cost, of MINLP.
 

### 2018-02-05 QPLIB-2.1.tex
consistent title case in section headers
Update QPLIB-2.1.tex

@@
 (limited to optimization problems);\label{optgap}
-\item using a computational model according to which every elementary computation on the reals takes $O(1)$ and returns a precise result \cite[p.~24]{blum}.
+\item using a computational model according to which every elementary computation on the reals takes $O(1)$ and returns an exact result \cite[p.~24]{blum}.
 \end{enumerate}

### 2018-02-05 QPLIB-2.2.tex
fix typos
Update QPLIB-2.2.tex

--- 2018-02-03 QPLIB-2.2.tex
@@
  \item constraints: (\textit{N})one, (\textit{B})ox, (\textit{L})inear, (\textit{D})iagonal convex quadratic, (\textit{C})on\-vex quadratic, nonconvex (\textit{Q})uadratic.
-  Note that (\textit{D}) and (\textit{C}) are intended to mean that either $Q^i$ is positive semidefinite and $c_l^i = -\infty$, or $Q^i$ is negative semidefinite and $c_u^i = \infty$. 
+  %Note that (\textit{D}) and (\textit{C}) are intended to mean that either $Q^i$ is positive semidefinite and $c_l^i = -\infty$, or $Q^i$ is negative semidefinite and $c_u^i = \infty$. 
     \textcolor{red}{Note that (positive or negative) definiteness of $Q^i$ is a sufficient, but not in general necessary, condition for convexity.
@@
 Furthermore, as previously remarked, a not (positive or negative) definite $Q^i$ does not necessarily correspond to a nonconvex feasible region.
-For instance, it is well-known that Second-Order Cone Programs have convex feasible regions; when represented in terms of quadratic constraints, however, they correspond to $Q^i$ in one negative eigenvalue.
+For instance, it is well-known that Second-Order Cone Programs have convex feasible regions; when represented in terms of quadratic constraints, however, they correspond to $Q^i$ with one negative eigenvalue.
 In our taxonomy we still consider the corresponding instances as $**Q$ ones, with no attempt to detect the different special structures that actually correspond to convex feasible regions.
-Although this may lead to classify as ``potentially nonconvex'' some instances that are in fact convex, our choice is justified by the fact that not all QP solvers are capable of detecting and exploiting these structures, which means that the instance can actually be treated as a nonconvex one even if it is not.
+Although this may lead to classify as ``potentially nonconvex'' some instances that are in fact convex, our choice is justified by the fact that not all QP solvers are capable of detecting and exploiting these structures, which means that the instance can actually be treated as a nonconvex one even if it is not.\\
+}
+\textcolor{red}{
+One of the nontrivial choices in our library is that we made no effort to reformulate the instances, and inserted them in the library in the very same format they have been provided to us by the original contributors.
+The rationale of this choice is that reformulation techniques, like the ones discussed here and others, are typically motivated by the fact that they make the instance easier to solve for one specific class of solvers.
+This being a bias that we don’t want to add we have chosen to keep the instances in their ``natural'' format, this being the one in which the original contributor initially wrote them.
 }
@@
 When written as QPs the corresponding $Q^i$ may not be positive semidefinite, but nonetheless such problems can be efficiently solved.
-Of course just as for \textit{LCL}, these problems may still require considerable computational effort when the size of the instance grows.
+Of course, just as for \textit{LCL}, these problems may still require considerable computational effort when the size of the instance grows.
 In this sense, like in the linear case, there is a significant distinction between solvers that need all the data of QP to work, and those that are ``matrix-free'', i.e., only require the application of simple operations (typically, matrix-vector products) with the problem data.
-While when building our instance library we never exploited such characteristics, since they are not amenable to standard modelling tools, but this may be relevant for the solution of very-large-scale \textit{CIC}.
+When building our instance library we never exploited such characteristics, since they are not amenable to standard modelling tools, but this may be relevant for the solution of very-large-scale \textit{CIC}.
  %

### 2018-02-05 QPLIB-3.tex
Update QPLIB-3.tex

@@
 two sided, finite bounds is non-convex.
-Note that this might occasionally lead us to classify some instances that have conic constraints as non-convex ones, although their feasible region is in fact convex---fortunately, only some solvers are capable of properly exploiting this property.
+Note that this might occasionally lead us to classify some instances that have conic constraints as non-convex ones, although their feasible region is in fact convex; however, only some solvers are capable of properly exploiting this property.
  All this information allowed us to analyse the gathered instances and to perform the
@@
        according to their features; for each cluster we kept only a few
-       representatives (e.g., very similar size, same donor,\ldots). Finally, in order to only
+       representatives, e.g. by eliminating all but a few of those with very similar size and coming from the same donor. Finally, in order to only
        keep computationally challenging instances we ran a a complete solver for QGQ with a time limit of 120 seconds; all the

### 2018-02-08 QPLIB-2.2.tex

@@
 In this sense, like in the linear case, there is a significant distinction between solvers that need all the data of QP to work, and those that are ``matrix-free'', i.e., only require the application of simple operations (typically, matrix-vector products) with the problem data.
-When building our instance library we never exploited such characteristics, since they are not amenable to standard modelling tools, but this may be relevant for the solution of very-large-scale \textit{CIC}.
+When building our instance library we never exploited such characteristics, since they are not amenable to standard modeling tools, but this may be relevant for the solution of very-large-scale \textit{CIC}.
  %

### 2018-02-08 QPLIB-3.tex

--- 2018-02-05 QPLIB-3.tex
@@
 
-In addition to spontaneous contribution we analysed the other generic
-libraries of instances available  on internet and containing QP
+In addition to these spontaneous contributions we analyzed existing generic
+libraries of instances available  on the internet that contain QP
 instances. Namely, the libraries from which we gathered instances are
@@
 %
-\item the \texttt{Meszaros} library
+\item the \texttt{Maros-M{\'e}sz{\'a}ros} library
  \url{http://www.doc.ic.ac.uk/~im/00README.QP};
@@
        The goal of the second filter was to eliminate ``similar'' instances.
-       We carefully analysed the instances one by one, and we clustered them
+       We carefully analyzed the instances one by one, and we clustered them
        according to their features; for each cluster we kept only a few
-       representatives (e.g., very similar size, same donor,\ldots). Finally, in order to only
+       representatives, e.g. by eliminating all but a few of those with very similar size and coming from the same donor. Finally, in order to only
        keep computationally challenging instances we ran a a complete solver for QGQ with a time limit of 120 seconds; all the

### 2018-02-08 QPLIB-1.0.tex
reword to reduce by one comma

@@
 This is illustrated by the fact that the class of problems that we simply refer to here as ``QP'' might more accurately be called Mixed-Integer Quadratically-Constrained Quadratic Programming (MIQCQP) in the most general case.
-It is, therefore, perhaps not surprising that, unlike for ``simpler'' problems classes like Mixed-Integer Linear Programming \cite{Koch2011},  there has been, to date, no single library devoted to all different kinds of instances of QP.\todo{AG: many commas in this sentence} While several specialized libraries devoted to particular cases of QP are available, each of them is either focused on a particular application (a specific problem that can be modeled as a QP), or on QPs with specific structural properties that make them suitable for solution by some given class of algorithmic approaches.
+Therefore, it is perhaps not surprising that, unlike for ``simpler'' problems classes like Mixed-Integer Linear Programming \cite{Koch2011},  there has been, to date, no single library devoted to all different kinds of instances of QP.\todo{AG: many commas in this sentence SV: I removed one.}\ While several specialized libraries devoted to particular cases of QP are available, each of them is either focused on a particular application (a specific problem that can be modeled as a QP), or on QPs with specific structural properties that make them suitable for solution by some given class of algorithmic approaches.
 To the best of our knowledge, collecting a set of QP instances that is at the same time not overwhelmingly onerous but sufficiently challenging for the many different interested communities has not been attempted.

### 2018-02-08 QPLIB-1.0.tex
update end of intro
- update number of minlplib2 instances
- add question on credibility of mittelmann page
- fix some typos

@@
 In particular, \S\ref{subsec:notation} sets out the notation, \S\ref{sec:classification} proposes a new taxonomy of QP that helps us in discussing the (very) different classes of QPs, and \S\ref{sec:algo} very briefly reviews the solution methods for QP and the solvers we have employed.
-Next \S\ref{sec:lib} describes the process used to obtain the library and its results.
+Next, \S\ref{sec:lib} describes the process used to obtain the library and its results.
 %; the software tools that we have used, and that are freely released together with the library, are discussed in \S\ref{subsec:tools}.
@@
 While no performance issues of solvers for QP problems are considered in
-this paper, we refer to the comprehensive benchmark site \url{http://plato.asu.edu/bench.html}. Of the result on this site, three deal exclusively with
-QP problems, namely the (1) large SOCP, (2) MISOCP, and the (3) MIQ(C)P benchmarks, while three others contain have partial results for such problems,
+this paper, we refer to the comprehensive benchmark site \url{http://plato.asu.edu/bench.html}.\todo[inline]{SV: Should one really point to this site?
+The benchmarks there are often faulty.
+Until recently, the SCIP 4 benchmark was using SCIP 3.1.
+Currently, the GAMS 25.0.2 benchmark partly uses 25.0.1.
+I wouldn't suggest anyone to take this site much serious.} Of the result on this site, three deal exclusively with
+QP problems, namely the (1) large SOCP, (2) MISOCP, and the (3) MIQ(C)P benchmarks, while three others contain partial results for such problems,
 namely those for (4) parallel barrier solvers on large LP/QP problems, (5) AMPL-NLP and (6) MINLP.
@@
 Benchmark (6) is based on MINLPLib~2 \cite{Vigerske2014}, a
-collection of currently 1388 instances. In order to give a first representative
+collection of currently 1527 instances. In order to give a first representative
 impression of solver performance, care was taken there to reduce the number of

### 2018-02-08 QPLIB-2.2.tex
move wildcard paragraph down
remove artificial space

@@
 \end{itemize}
-%
-The wildcard ``*'' will be used below to indicate any possible choice, and lists of the form ``\{\textit{X}, \textit{Y}, \textit{Z}\}'' will indicate that the value of the given field can freely attain any of the specified values.
 
@@
 In the following discussion we will repeatedly exploit this by assuming that, unless otherwise mentioned, when a method can be applied to a given problem, it can be applied as well to all simpler problems where the value of each field is arbitrarily replaced with a value denoting a less-general class.
+
+The wildcard ``*'' will be used below to indicate any possible choice, and lists of the form ``\{\textit{X}, \textit{Y}, \textit{Z}\}'' will indicate that the value of the given field can freely attain any of the specified values.
 
@@
 
-\smallskip
+%\smallskip
 It is important to note that the relationships between the different classes can be somehow blurred because reformulation techniques may allow one to move an instance from one class to another.

### 2018-02-08 QPLIB-2.2.tex
replace T by \top for transposition

@@
 When a $Q^i$ is positive semidefinite (PSD), i.e., the corresponding constraint/objective function is convex, general Hessians are in fact equivalent to diagonal ones.
-In particular, since every PSD matrix can be factorized as $Q^i = L^i (L^i)^T$, e.g.~by the (incomplete) Cholesky factorization, the term $\half x^T Q^i x \equiv \half \sum_{j \in \Mcn} z^{i\;2}_j$ where $z^{i\;T} = x^T L^i$.
+In particular, since every PSD matrix can be factorized as $Q^i = L^i (L^i)^\top$, e.g.~by the (incomplete) Cholesky factorization, the term $\half x^\top Q^i x \equiv \half \sum_{j \in \Mcn} z^{i\;2}_j$ where $z^{i\;\top} = x^\top L^i$.
 Hence, one might maintain that D** problems need not be distinguished from C** ones.
-However in reality, this is only true for ``complicated'' constraints but not for ``simple'' ones, because the above reformulation technique introduces additional linear constraints, $L^{i\;T} x - z^i = 0$.
+However in reality, this is only true for ``complicated'' constraints but not for ``simple'' ones, because the above reformulation technique introduces additional linear constraints, $L^{i\;\top} x - z^i = 0$.
 Indeed, while \textit{C*L} (and, a fortiori, \textit{C*}\{\textit{C},\textit{Q}\}) can always be brought to \textit{D*L} (\textit{D*}\{\textit{C},\textit{Q}\}), using the above technique \textit{C*B} becomes \textit{D*L}, which may be significantly different from \textit{D*B}.

### 2018-02-08 QPLIB-2.2.tex
fix formatting of **Q

@@
 For instance, it is well-known that Second-Order Cone Programs have convex feasible regions; when represented in terms of quadratic constraints, however, they correspond to $Q^i$ with one negative eigenvalue.
-In our taxonomy we still consider the corresponding instances as $**Q$ ones, with no attempt to detect the different special structures that actually correspond to convex feasible regions.
+In our taxonomy we still consider the corresponding instances as \textit{**Q} ones, with no attempt to detect the different special structures that actually correspond to convex feasible regions.
 Although this may lead to classify as ``potentially nonconvex'' some instances that are in fact convex, our choice is justified by the fact that not all QP solvers are capable of detecting and exploiting these structures, which means that the instance can actually be treated as a nonconvex one even if it is not.\\

### 2018-02-08 QPLIB-2.2.tex
replace format by form, fix some grammar
- format could mean that we include the original files, which we do not do

@@
 \textcolor{red}{
-One of the nontrivial choices in our library is that we made no effort to reformulate the instances, and inserted them in the library in the very same format they have been provided to us by the original contributors.
+One of the nontrivial choices in our library is that we made no effort to reformulate the instances, and inserted them in the library in the very same form as they have been provided to us by the original contributors.
 The rationale of this choice is that reformulation techniques, like the ones discussed here and others, are typically motivated by the fact that they make the instance easier to solve for one specific class of solvers.
-This being a bias that we don’t want to add we have chosen to keep the instances in their ``natural'' format, this being the one in which the original contributor initially wrote them.
+This being a bias that we do not want to add we have chosen to keep the instances in their ``natural'' form, this being the one in which the original contributor initially wrote them.
 }

### 2018-02-08 QPLIB-2.2.tex
bugfix and minor changes in QP classes

@@
 The general view is that \textit{CGC} are not, all other things being equal,  substantially more difficult than \textit{LGL} to solve, especially if the objective function and/or the constraints have specific properties (e.g., \textit{DGL}, \textit{CGL}).
-Often integer variables are in fact binary ones, so several \textit{CCC} models are \textit{C}\{\textit{B},\textit{M}\}\textit{C} ones.
-In practice binary variables are considered to lead to somewhat easier problems than general integer ones (cf.~the cited result about hardness of unbounded integer quadratic programs), and several algorithmic techniques have been specifically developed for this special case.
+Often, integer variables are in fact binary ones, so several \textit{CGC} models are \textit{C}\{\textit{B},\textit{M}\}\textit{C} ones.
+In practice, binary variables are considered to lead to somewhat easier problems than general integer ones (cf.~the cited result about hardness of unbounded integer quadratic programs) and several algorithmic techniques have been specifically developed for this special case.
 However, the general approaches for \textit{CBC} are basically the same as for \textit{CGC}, so there is seldom the need to distinguish between the two classes as far as solvability is concerned, although matters can be different regarding actual solution cost.
-Programs with only binary variables \textit{CBC} can be easier than mixed-binary or integer ones \textit{C}\{\textit{M},\textit{I}\}\textit{C} because some techniques are specifically known for the binary-only case, cf.~the next point \cite{BEP09}.
-Absence of continuous variables, even in the presence of integer ones \textit{CIC}, can also lead to specific techniques \cite{BEL13}.
+Programs with only binary variables (\textit{CBC}) can be easier than mixed-binary or integer ones (\textit{C}\{\textit{M},\textit{I}\}\textit{C}) because some techniques are specifically known for the binary-only case, cf.~the next point \cite{BEP09}.
+Absence of continuous variables, even in the presence of integer ones (\textit{CIC}), can also lead to specific techniques \cite{BEL13}.
   %
-  \item \emph{Nonconvex Binary Quadratic Programs} \textit{QB}\{\textit{B}, \textit{N}, \textit{L}\} obviously are $\mathcal{NP}$-hard.
+  \item \emph{Nonconvex Binary Quadratic Programs} \textit{QB}\{\textit{B},\textit{N},\textit{L}\} obviously are $\mathcal{NP}$-hard.
 However, the special nature of binary variables combined with quadratic forms allows for quite specific techniques to be developed, one of which is the reformulation of the problem as a \textit{LBL}.
@@
   \item \emph{Nonconvex Integer Quadratic Programs} \textit{QGQ} is the most general, and therefore is the most difficult, class.
-Due to the lack of convexity even when integrality requirements are removed, solution methods must typically combine several algorithmic ideas, such as enumeration (distinguishing the role of integral variables from that of continuous ones involved into nonconvex terms) and techniques (e.g., outer approximation, semidefinite programming relaxation, \ldots) that allow the efficient computation of bounds.
+Due to the lack of convexity even when integrality requirements are removed, solution methods must typically combine several algorithmic ideas, such as enumeration (distinguishing the role of integral variables from that of continuous ones involved in nonconvex terms) and techniques that allow the efficient computation of bounds (e.g., outer approximation, semidefinite programming relaxation, \ldots).
 As in the convex case, \textit{QBQ}, \textit{QMQ}, and \textit{QIQ} can benefit from more specific properties of the variables \cite{Buchheim2013,Dong2016}.

### 2018-02-08 QPLIB-2.2.tex
add doubt on choice of references

@@
 These can have a significant algorithmic impact, and therefore can be of interest to researchers.
-Common structures are, e.g., network flows \cite{FrFG16,FrGaSc14,FrGaSt16a,FGGP11,tadayon-smith:2013} or knapsack-type linear constraints \cite{FrFG16,FrGo13b,FGGP11}, and semi-continuous variables \cite{FrFG16,FrGaSc14,FrGaSt16a,FrGe06a,FrGe07a,FrGe09a,FGGP11}, or the fact that a nonconvex quadratic objective function/constraint can be reformulated as a second-order cone (hence, convex) one \cite{FrFG16,FrGaSc14,FrGaSt16a,FrGe09a,FGGP11}.
+Common structures are, e.g., network flows \cite{FrFG16,FrGaSc14,FrGaSt16a,FGGP11,tadayon-smith:2013} or knapsack-type linear constraints \cite{FrFG16,FrGo13b,FGGP11}, and semi-continuous variables \cite{FrFG16,FrGaSc14,FrGaSt16a,FrGe06a,FrGe07a,FrGe09a,FGGP11}\todo{SV: do we really need 7 Frangioni et.al citations for semi-continuous variables?}, or the fact that a nonconvex quadratic objective function/constraint can be reformulated as a second-order cone (hence, convex) one \cite{FrFG16,FrGaSc14,FrGaSt16a,FrGe09a,FGGP11}\todo{SV: 5 Frangioni et.al citations for SOC?}.
 It would be very hard to collect a comprehensive list of all types of structures that might be of interest to any individual researcher, since these are as varied as the different possible approaches for specialized sub-classes of QP.

### 2018-02-08 QPLIB-2.1.tex
consistent formatting of NP

@@
 %- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-\section{Quadratic Programming in a nutshell}\label{sec:QPbasic}
+\section{Quadratic Programming in a Nutshell}\label{sec:QPbasic}
 
@@
 In view of the fact that their solutions may be irrational, this notation deserves a comment.
-If the decision version of the problem being referred to is in NP (e.g.~LP, MILP, QP \cite{vavasis90a}), then the assumption is that all rational numbers can be represented exactly by a Turing Machine (TM).
+If the decision version of the problem being referred to is in $\mathcal{NP}$ (e.g.~LP, MILP, QP \cite{vavasis90a}), then the assumption is that all rational numbers can be represented exactly by a Turing Machine (TM).
 %This assumption has the same limitation as the assumption that TMs defined on integers are a faithful abstraction of a real computer, namely that the range of integers needing a representation is bounded.
-If there is no known proof that the problem being solved (or its decision version) is in NP, then there are four main approaches:
+If there is no known proof that the problem being solved (or its decision version) is in $\mathcal{NP}$, then there are four main approaches:
 \begin{enumerate}

### 2018-02-08 methods.tex
consistent formatting of NP

--- 2018-02-03 methods.tex
@@
 
-\subsubsection{Incomplete methods}
+\subsubsection{Incomplete Methods}
 
@@
 % -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -
-\subsubsection{Asymptotically complete methods}\label{sss:asymc}
+\subsubsection{Asymptotically Complete Methods}\label{sss:asymc}
 
@@
 % -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -
-\subsubsection{Complete methods}\label{s:complete}
+\subsubsection{Complete Methods}\label{s:complete}
 
@@
 This term has to be used with care, as it implicitly makes assumptions on the underlying computational model that may not be acceptable in all cases.
-For example, te decision version of \textit{QCL} is known to be in the complexity class \textbf{NP} \cite{vavasis90a}, whereas the same is not known about \textit{LCQ}, even with zero objective.
+For example, the decision version of \textit{QCL} is known to be in the complexity class $\mathcal{NP}$ \cite{vavasis90a}, whereas the same is not known about \textit{LCQ}, even with zero objective.
 On the other hand, there exists a method for deciding feasibility of systems of polynomial equations and inequalities \cite{tarski-reals}, including the solution of \textit{LCQ} with zero objective function.

### 2018-02-08 conclusion.tex
revised conclusion

--- 2018-02-04 conclusion.tex
@@
-\section{Conclusions}\label{sec:conclusions}
+\section{Final Remarks}\label{sec:conclusions}
 
-This manuscript described the first comprehensive library of instances for Quadratic Programming (QP).
+This paper described the first comprehensive library of instances for Quadratic Programming (QP).
 Since QP comprises different and ``varied'' categories of problems, we proposed a classification and we briefly discussed the main classes of solution methods for QP.
-
 We then described the steps of the adopted process used to filter the gathered instances  in order to build the new library.
 Our design goals were to build a library which is computationally challenging and as broad as possible, i.e., it represents the largest possible categories of QP problems, while remaining of manageable size.
-We have also proposed a stand-alone QP format that is intended
+We also proposed a stand-alone QP format that is intended
 for the convenient exchange and use of our QP instances.
 
-We want to stress once again that we intentionally avoided to perform a computational comparison of the performances of the different solvers.
-Our goal was instead to provide a common test bed of instances for practitioners and researchers in the field.
-This new library will hopefully serve as a point of reference to test new ideas and algorithms for QP problems.
+We want to stress once again that we intentionally avoided to perform a
+computational comparison of the performances of different solution methods or
+solver implementations. Our goal was instead to provide a broad test bed of
+instances for researchers and practitioners in the field. This new library will
+hopefully serve as a point of reference to inspire and test new ideas and
+algorithms for QP problems.
+
+Finally, we want to emphasize that this QP collection can only be a snapshot of
+the types of problems that researchers and practitioners have worked on in the
+past.  With the growing interest in this area, we hope that new applications and
+instances will become available and that the library can be extended dynamically
+in the future.

### 2018-02-08 QPLIB-3.tex
ambros is right: it's not sure that the largest possible set has been reached
- further, there is no clear definition what *the* main intl. mailing lists are

@@
 instances. In January $2014$, we issued an online call for instances
-using the main international mailing lists of the mathematical
+using main international mailing lists of the mathematical
 optimization and numerical analysis communities, reaching in this way
-the largest\todo{AG: a large?} possible set of interested researchers and practitioners.
+a \textcolor{red}{large set of possibly interested} researchers and practitioners.
 The call remained open for ten months, during which we received a large

### 2018-02-08 QPLIB-3.tex
fix double "and" and other minor fixes

@@
 we gathered $\approx 1800$ Quadratic Binary Linear (QBL) instances,
-$\approx 2000$ Quadratic Continuous Quadratic (QCQ) instances, and
+$\approx 2000$ Quadratic Continuous Quadratic (QCQ) instances,
 and $\approx 2500$ Quadratic General Quadratic (QGQ) instances. We
@@
 Continuous Convex (CCC) and Convex Mixed Convex (CMC) instances,
-($\approx 150$, $\approx 200$ and $\approx 200$ instances respectively).
+($\approx 150$, $\approx 200$, and $\approx 200$ instances, respectively).
 Finally, we found only $17$ Quadratic Mixed
-Linear (QML) instances. In the call for instances, no specific formats
+Linear (QML) instances. In the call for instances, no specific format
 requirements were imposed for the submissions.

### 2018-02-08 QPLIB-3.tex
fix another double-and and a double-the

@@
 Hessians have only non-negative eigenvalues (when $c_u^i < \infty$) and
-and non-positive eigenvalues (when $c_l^i > - \infty$)
+non-positive eigenvalues (when $c_l^i > - \infty$)
 are classified as convex constraints; thus, a quadratic constraint with
@@
  All this information allowed us to analyse the gathered instances and to perform the
-filters described in the the next paragraph.
+filters described in the next paragraph.
 

### 2018-02-08 QPLIB-3.tex
replace non-convex by nonconvex
- as nonconvex seems to be used much more often

@@
 are classified as convex constraints; thus, a quadratic constraint with
-two sided, finite bounds is non-convex.
-Note that this might occasionally lead us to classify some instances that have conic constraints as non-convex ones, although their feasible region is in fact convex---fortunately, only some solvers are capable of properly exploiting this property.
+two sided, finite bounds is nonconvex.
+Note that this might occasionally lead us to classify some instances that have conic constraints as nonconvex ones, although their feasible region is in fact convex---fortunately, only some solvers are capable of properly exploiting this property.
  All this information allowed us to analyse the gathered instances and to perform the
@@
 In Table \ref{tab:DDDDD}, we provide a global overview.
-The instances have been divided in \textit{continuous} vs \textit{discrete} and \textit{convex} vs \textit{non-convex}, forming in this way, a  classification of 4 macro categories.
+The instances have been divided in \textit{continuous} vs \textit{discrete} and \textit{convex} vs \textit{nonconvex}, forming in this way, a  classification of 4 macro categories.
 As previously mentioned, an instance is classified \textit{discrete} if it contains at least a binary or integer variable, and \textit{continuous} otherwise.
-On the other hand, an instance is classified as \textit{non-convex} if the objective function is non-convex and/or at least one of the constraints is  non-convex, and \textit{convex} otherwise.
+On the other hand, an instance is classified as \textit{nonconvex} if the objective function is nonconvex and/or at least one of the constraints is nonconvex, and \textit{convex} otherwise.
 }
@@
  \textit{continuous}	& \textit{convex}		&  32 \\[1.2 ex]
- \textit{continuous}	& \textit{non-convex} 	&  102 \\[1.2 ex]
+ \textit{continuous}	& \textit{nonconvex} 	&  102 \\[1.2 ex]
  \textit{discrete}	& \textit{convex}		&  31 \\[1.2 ex]
- \textit{discrete}	& \textit{non-convex}	&  288 \\[1.2 ex]
+ \textit{discrete}	& \textit{nonconvex}	&  288 \\[1.2 ex]
 \cmidrule(lr){1-3}

### 2018-02-08 QPLIB-3.tex
update numbers in text of 3.3
- remove wrong parenthesis
- fixed some typos and a double-a

@@
 the expected goals that we wished to achieve. The following four goals
-where finally identified:
+were finally identified:
 %
@@
        representatives, e.g. by eliminating all but a few of those with very similar size and coming from the same donor. Finally, in order to only
-       keep computationally challenging instances we ran a a complete solver for QGQ with a time limit of 120 seconds; all the
+       keep computationally challenging instances we ran a complete solver for QGQ with a time limit of 120 seconds; all the
        instances which have been solved to proven optimality within this time limit
@@
 As previously mentioned, an instance is classified \textit{discrete} if it contains at least a binary or integer variable, and \textit{continuous} otherwise.
-On the other hand, an instance is classified as \textit{nonconvex} if the objective function is nonconvex and/or at least one of the constraints is nonconvex, and \textit{convex} otherwise.
+On the other hand, an instance is classified as \textit{nonconvex} if the objective function is nonconvex (if minimization) or nonconcave (if maximization) and/or at least one of the constraints is nonconvex, and \textit{convex} otherwise.
 }
@@
 instances among the different categories. Indeed, the discrete categories
-(\textit{LMQ}) or (\textit{QBL}) are well represented by $118$ and $59$
+\textit{LMQ} and \textit{QBL} are well represented by \textcolor{red}{134} and \textcolor{red}{91}
 instances, respectively. Similarly, the continuous categories
-(\textit{LCQ}) and (\textit{QCQ}) are well
-represented by $50$ and $17$ instances, respectively. Moreover, the library
+\textit{LCQ} and \textit{QCQ} are well
+represented by \textcolor{red}{52} and \textcolor{red}{30} instances, respectively. Moreover, the library
 actually covers the large majority of all possible categories of instances.

### 2018-02-08 QPLIB-3.tex
revise repetitive paragraph on not-reformulating
- format to form
- other minor fixes

@@
 \textcolor{red}{
-One of the nontrivial choices in our library is that we made no effort to reformulate the instances, and inserted them in the library in the very same format they have been provided to us by the original contributors.
+One of the nontrivial choices in our library is that we made no effort to reformulate the instances, and inserted them in the library in the very same form as they have been provided to us by the original contributors.
 Section \ref{ssec:reform} is crucial in justifying this choice, as it shows that there are several degrees of freedom to move the instances from one class to another.
-Tailoring the structure of a problem to the solver is, however, a bias we do not want to add.
+Tailoring the structure of a problem to a solver is, however, a bias that we did not want to add.
 }

### 2018-02-08 QPLIB-3.tex
update maximum on vars and conss

@@
 that the library contains a quite diverse set of instances in
-terms of number of variables and constraints. The maximal number of constraints
-is 100000, while the maximal number of
-variables is almost 40000.
+terms of number of variables and constraints. \textcolor{red}{The record on the maximal number of variables and constraints (both $\approx 1,000,000$) is set by the instances 8547 and 9008.}
 Figure~\ref{fig:distribution}~(right) plots the number of nonzero elements in the gradient of the objective function and the Jacobian and the number of these nonzeros corresponding to nonlinear variables, that is, it counts the appearances of variables in objectives and constraints and how often such an appearance is in a quadratic term.

### 2018-02-08 QPLIB-3.tex
update number of total and quad conss in text

@@
 and the total number of quadratic constraints
-with a ``$\times$''. Also, in this case, different percentages of linear and
+with a ``$\times$''. Also in this case, different percentages of linear and
 quadratic constraints are present and well-distributed across the spectrum
@@
 from Figure~\ref{fig:pic_constr} (left) we learn that while the maximum number
-of linear constraints exceeds 100000, the maximum number of quadratic
-constraints tops up at around 10000. This is, however, reasonable, considering
+of linear constraints exceeds 1,000,000, the maximum number of quadratic
+constraints tops up at 140,000. This is, however, reasonable, considering
 how quadratic constraints can, in general, be expected to be much more

### 2018-02-08 QPLIB-3.tex
revise website section
- add question on "(provided)"
- sparsity pattern not on all instances
- mention solution points
- hard ev -> problematic ev
- contributor information has been included by now

@@
 format offers three main advantages: it is easier to read by a stand-alone
-parser (provided), it typically produces smaller files, and it permits the
+parser (provided)\todo{SV: what is meant by ``(provided)''?}, it typically produces smaller files, and it permits the
 inclusion of two-sided inequalities without needless repetition of data.
@@
 Beyond the instances, the website provides a rich set of metadata for
-each instance: the three letter problem classification (as described in \S\ref{subsec:final set}), basic properties such as the number of variables and constraints of
+each instance: the three letter problem classification (as described in \S\ref{subsec:final set}), \textcolor{red}{the contributor of the instance}, basic properties such as the number of variables and constraints of
 different types, the sense and convexity/concavity of the objective function, and information on the nonzero structure of the
@@
 In addition, we display a visualization of the sparsity patterns of the Jacobian and the Hessian matrix of the
-Lagrangian function.  In the plots of the Jacobian nonzero pattern, the linear and nonlinear entries are distinguished
+Lagrangian function, \textcolor{red}{if the instance size allows}.  In the plots of the Jacobian nonzero pattern, the linear and nonlinear entries are distinguished
 by color.  Figure \ref{fig:sparsitypattern} shows an example for instance QPLIB\_2967.
+\textcolor{red}{Finally, feasible solution points are provided for most instances.}
 
@@
 The entire set of instances can be explored in a searchable and sortable table of selected instance features: problem
-classification, convexity of the continuous relaxation, number of (all, binary, integer) variables, (all, quadratic) constraints, nonzeros, hard eigenvalues in~$Q^0$, and density of~$Q^0$.
+classification, convexity of the continuous relaxation, number of (all, binary, integer) variables, (all, quadratic) constraints, nonzeros, \textcolor{red}{problematic} eigenvalues in~$Q^0$, and density of~$Q^0$.
 %
@@
 %
-In the future, we plan to extend the website by the addition of contributor information and references to the literature.
+In the future, we plan to extend the website by references to the literature.
 

### 2018-02-08 QPLIB-3.tex
add QPLIB_ prefix on instance names

@@
 that the library contains a quite diverse set of instances in
-terms of number of variables and constraints. \textcolor{red}{The record on the maximal number of variables and constraints (both $\approx 1,000,000$) is set by the instances 8547 and 9008.}
+terms of number of variables and constraints.
+\textcolor{red}{The record on the maximal number of variables and constraints (both $\approx 1,000,000$) is set by the instances QPLIB\_8547 and QPLIB\_9008.}
 Figure~\ref{fig:distribution}~(right) plots the number of nonzero elements in the gradient of the objective function and the Jacobian and the number of these nonzeros corresponding to nonlinear variables, that is, it counts the appearances of variables in objectives and constraints and how often such an appearance is in a quadratic term.

### 2018-02-08 QPLIB.tex
there are many gams licenses

@@
 We are grateful to all the donors who provided instances for the library.
-We gratefully acknowledge the financial support of the Gaspard Monge Program for Optimization and operations research (PGMO)  and the logistic support of GAMS for having provided us with the license of their software.
+We gratefully acknowledge the financial support of the Gaspard Monge Program for Optimization and operations research (PGMO)  and the logistic support of GAMS for having provided us with a license for their software.
 Finally, we would like to acknowledge the financial and networking support by the COST Action TD1207.

### 2018-02-08 QPLIB-2.2.tex
uniform wording for b&b
revise bibliography
- remove two duplicates
- fix some capitalizations
- no space in pages
- full journal names
- add missing issue/number
- fix some typos

@@
  \item \emph{Nonconvex Continuous Quadratic Programs} \textit{QCQ} are generally $\mathcal{NP}$-hard, even if the constraints are very specific (\textit{QCB}) and only a single eigenvalue of $Q^0$ is negative \cite{Hemmecke2010}.
-They therefore require enumerative techniques, such as spatial Branch\&Bound \cite{FV90,BeLeLiMaWa08}, to be solved to optimality.
+They therefore require enumerative techniques, such as spatial Branch-and-Bound \cite{FV90,BeLeLiMaWa08}, to be solved to optimality.
 Of course, local approaches are available that are able to efficiently provide saddle points (hopefully, local optima) of the \textit{CQC}, but providing global guarantees about the quality of the obtained solutions is challenging.
@@
 However, the special nature of binary variables combined with quadratic forms allows for quite specific techniques to be developed, one of which is the reformulation of the problem as a \textit{LBL}.
-Also, many well-known combinatorial problems can be naturally reformulated as problems of this class, and therefore a considerable number of results have been obtained by exploiting specific properties of the set of constraints \cite{Rendl2008,Loiola2007657}.
+Also, many well-known combinatorial problems can be naturally reformulated as problems of this class, and therefore a considerable number of results have been obtained by exploiting specific properties of the set of constraints \cite{Rendl2008,loiola-etal:2007}.
   %

### 2018-02-08 methods.tex
mosek, gurobi, xpress don't run on QGQ
xpress and gurobi can also to nonconvex objective if binary variables only
alphabetic order of solvers
revise bibliography
- remove two duplicates
- fix some capitalizations
- no space in pages
- full journal names
- add missing issue/number
- fix some typos

@@
 Nearly all of the complete solvers in Table \ref{t:solvers} that address $\mathcal{NP}$-hard problems (i.e.~those in \textit{QGQ}$\smallsetminus$\textit{CCC}) are based on Branch-and-Bound (BB) \cite{land_doig}.
-When the BB algorithm is allowed to branch on coordinate directions corresponding to continuous variables, it is called \emph{spatial} BB (sBB) \cite{dakin,couenne}.
+When the BB algorithm is allowed to branch on coordinate directions corresponding to continuous variables, it is called \emph{spatial} BB (sBB) \cite{dakin,BeLeLiMaWa08}.
 BB algorithms require exponential time in the worst case, and their exponential behavior unfortunately often shows up in practice.
@@
 \begin{itemize}
- \item \antigone, \baron, \couenne, \lindo, \scip for \textit{QGQ};
+ \item \antigone, \baron, \couenne, \lindo, and \scip for \textit{QGQ};
  %
@@
 %
- \item \knitro, \bonmin, \sbb, \xpress, \gurobi, and \mosek for \textit{CGC}.
+ \item \textcolor{red}{\gurobi and \xpress for \textit{QBC};}
+%
+ \item \bonmin, \gurobi, \knitro, \mosek, \sbb, and \xpress for \textit{CGC}.
 \end{itemize}
 %
-We remark that the latter category can be used as incomplete solvers for \textit{QGQ}.
+We remark that \textcolor{red}{the solvers \bonmin, \knitro, and \sbb from} the latter category can be used as incomplete solvers for \textit{QGQ}.
 We also note that \lgo implements an incomplete BB algorithm for \textit{QCQ} by using bounds obtained from sampling.

### 2018-02-08 QPLIB-3.tex
problem types should be formatted in italics

@@
 the remainder contained only continuous variables. In more detail,
-we gathered $\approx 1800$ Quadratic Binary Linear (QBL) instances,
-$\approx 2000$ Quadratic Continuous Quadratic (QCQ) instances,
-and $\approx 2500$ Quadratic General Quadratic (QGQ) instances. We
-also received $\approx 1000$ Convex General Convex (CGC) instances. We
-obtained relatively fewer Quadratic Binary Quadratic (QBQ), Convex
-Continuous Convex (CCC) and Convex Mixed Convex (CMC) instances,
+we gathered $\approx 1800$ Quadratic Binary Linear (\textit{QBL}) instances,
+$\approx 2000$ Quadratic Continuous Quadratic (\textit{QCQ}) instances,
+and $\approx 2500$ Quadratic General Quadratic (\textit{QGQ}) instances. We
+also received $\approx 1000$ Convex General Convex (\textit{CGC}) instances. We
+obtained relatively fewer Quadratic Binary Quadratic (\textit{QBQ}), Convex
+Continuous Convex (\textit{CCC}) and Convex Mixed Convex (\textit{CMC}) instances,
 ($\approx 150$, $\approx 200$, and $\approx 200$ instances, respectively).
 Finally, we found only $17$ Quadratic Mixed
-Linear (QML) instances. In the call for instances, no specific format
+Linear (\textit{QML}) instances. In the call for instances, no specific format
 requirements were imposed for the submissions.
@@
        representatives, e.g. by eliminating all but a few of those with very similar size and coming from the same donor. Finally, in order to only
-       keep computationally challenging instances we ran a complete solver for QGQ with a time limit of 120 seconds; all the
+       keep computationally challenging instances we ran a complete solver for \textit{QGQ} with a time limit of 120 seconds; all the
        instances which have been solved to proven optimality within this time limit

### 2018-02-08 format.tex
consistent title case in section headers
problem types should be formatted in italics

@@
-\section{The file format}\label{sec:format}
+\section{The File Format}\label{sec:format}
 
@@
 \item [{[2]}]
-For problems of type \texttt{**N} or \texttt{**B}, these lines/sections
+For problems of type \textit{**N} or \textit{**B}, these lines/sections
 are omitted.
 \item [{[3]}]
-For problems of type \texttt{L**}, this section is omitted.
+For problems of type \textit{L**}, this section is omitted.
 \item [{[4]}]
-For problems of type \texttt{**N}, \texttt{**B} or \texttt{**L},
+For problems of type \textit{**N}, \textit{**B} or \textit{**L},
 this section is omitted.
 \item [{[5]}]
-For problems of type \texttt{*C*}, \texttt{*B*} or \texttt{*I*},
-this section is omitted. For problems of type \texttt{*I*}, binary variables
+For problems of type \textit{*C*}, \textit{*B*} or \textit{*I*},
+this section is omitted. For problems of type \textit{*I*}, binary variables
 should be specified as integer variables with lower and upper bounds 0 and 1.
 \item [{[6]}]
-For problems of type \texttt{*B*}, this section is omitted.
+For problems of type \textit{*B*}, this section is omitted.
 %Bounds for binary variables must not be specified in this section.
 \end{description}
-Binary variables defined either implicitly via the type \texttt{*B*}
+Binary variables defined either implicitly via the type \textit{*B*}
 or explicitly in the variable type section will be assumed to

### 2018-02-08 QPLIB-1.0.tex
remove one todo again

@@
 While no performance issues of solvers for QP problems are considered in
-this paper, we refer to the comprehensive benchmark site \url{http://plato.asu.edu/bench.html}.\todo[inline]{SV: Should one really point to this site?
+this paper, we refer to the comprehensive benchmark site \url{http://plato.asu.edu/bench.html}.%\todo[inline]{SV: Should one really point to this site?
 The benchmarks there are often faulty.
@@
 Currently, the GAMS 25.0.2 benchmark partly uses 25.0.1.
-I wouldn't suggest anyone to take this site much serious.} Of the result on this site, three deal exclusively with
+I wouldn't suggest anyone to take this site much serious.
+SV: With Hans as co-author, I guess we have to leave this in.}
+Of the result on this site, three deal exclusively with
 QP problems, namely the (1) large SOCP, (2) MISOCP, and the (3) MIQ(C)P benchmarks, while three others contain partial results for such problems,

### 2018-02-08 AnswerToTheReferees.tex
update answers to referees
- gurobi, xpress also can do QBL, I think
- format to form
- rewrite why we don't do reduced hessian: say it's the same reason as for soc
- mention availability of lp files, too

@@
 
-\rep{We wrote an email to the development team of Cplex and they confirmed that their solver is able to tackle QGL and CGC 
+\rep{We wrote an email to the development team of Cplex and they confirmed that their solver is able to tackle QGL (which contains QBL) and CGC.
+This is now mentioned in Section 2.3.4, additionally to QBC support for Gurobi and Xpress.
 % TODO:Fabio and Emiliano
@@
 One relevant example is the relationship between definiteness of the Hessians and convexity, correctly pointed out by the referee, which is naturally discussed in \S 2.2.2.
-One of the nontrivial choices in our library is that we made no effort to reformulate the instances (although of course we could have), and inserted them in the library in the very same format they have been provided to us by the original contributors.
+One of the nontrivial choices in our library is that we made no effort to reformulate the instances (although of course we could have), and inserted them in the library in the very same form as they have been provided to us by the original contributors.
 Section 2.2.2 is crucial in justifying this choice, as it shows that there are several degrees of freedom to move the instances from one class to another.
 Doing so, however, is typically justified by the fact that this makes one specific solver more efficient in solving them, a bias that we don't want to add.
-Hence, we have chosen to keep the instances in their ``natural'' format, this being the one in which the original contributor initially wrote them.
+Hence, we have chosen to keep the instances in their ``natural'' form, this being the one in which the original contributor initially wrote them.
 %TODO: Antonio.
@@
 We renamed ``hard" to ``problematic".
-However, as explained in \S 2.2.1 and 2.2.2, we refrain from reformulating the instance, and we do not tailor problems to specific solvers (moreover, not all QP solvers would resort to reduced Hessians).
+However, as explained in \S 2.2.1 and 2.2.2, the analysis of our instances has some limitations.
+Similar as with nonconvex quadratic constraints that could be reformulated as SOC, also resorting to a reduced Hessian for the objective function is a technique that is applied only by a subset of current QP solvers.
+It might be interesting, though, to extend our analysis in these directions in the future.
+%However, as explained in \S 2.2.1 and 2.2.2, we refrain from reformulating the instance, and we do not tailor problems to specific solvers (moreover, not all QP solvers would resort to reduced Hessians).
 In the same vein, our analysis can only be based on the given instance structure.
@@
 For this reason we have translated all the instances also into the AMPL format and we have uploaded them on the website.
+Further, also the additionally available LP format is understood by a number of solvers and is relatively easy to parse.
 %\leo{Possiamo spiegare che proporre NOI le istanze in formati multipli darebbe la responsabilita` del traduttore a noi, che invece non vogliamo accettarla.

### 2018-02-12 QPLIB-3.tex
comment on todo; correct MINLPLib name

@@
 %
-\item the \texttt{MINLP} library
+\item the \texttt{MINLPLib} library
  \url{http://www.gamsworld.org/minlp/minlplib.htm};
@@
 Conic library CBLIB (\url{http://cblib.zib.de}) and 
-MIPLIB~2010 (\url{http://miplib.zib.de/}), to mention just a few.}\todo{AG: maybe cite CBLIB paper \cite{Friberg2016}}
+MIPLIB~2010 (\url{http://miplib.zib.de/}), to mention just a few.}\todo{AG: maybe cite CBLIB paper \cite{Friberg2016} SV: then we would have to add citations for all mentioned libraries (where available)}
 

### 2018-02-12 QPLIB-1.0.tex
Update QPLIB-1.0.tex

@@
 This is illustrated by the fact that the class of problems that we simply refer to here as ``QP'' might more accurately be called Mixed-Integer Quadratically-Constrained Quadratic Programming (MIQCQP) in the most general case.
-Therefore, it is perhaps not surprising that, unlike for ``simpler'' problems classes like Mixed-Integer Linear Programming \cite{Koch2011},  there has been, to date, no single library devoted to all different kinds of instances of QP.\todo{AG: many commas in this sentence SV: I removed one.}\ While several specialized libraries devoted to particular cases of QP are available, each of them is either focused on a particular application (a specific problem that can be modeled as a QP), or on QPs with specific structural properties that make them suitable for solution by some given class of algorithmic approaches.
+Therefore, it is perhaps not surprising that, unlike for ``simpler'' problems classes like Mixed-Integer Linear Programming \cite{Koch2011},  there has been no single library devoted to all different kinds of instances of QP.
+%\todo{AG: many commas in this sentence SV: I removed one.}\ 
+While several specialized libraries devoted to particular cases of QP are available, each of them is either focused on a particular application (a specific problem that can be modeled as a QP), or on QPs with specific structural properties that make them suitable for solution by some given class of algorithmic approaches.
 To the best of our knowledge, collecting a set of QP instances that is at the same time not overwhelmingly onerous but sufficiently challenging for the many different interested communities has not been attempted.

### 2018-02-12 QPLIB-2.1.tex
Update QPLIB-2.1.tex

@@
 
-\textcolor{red}{In the rest of this paper, we shall sometimes refer to {\it exact solutions} of quadratic programs.
-In view of the fact that their solutions may be irrational, this notation deserves a comment.
-If the decision version of the problem being referred to is in $\mathcal{NP}$ (e.g.~LP, MILP, QP \cite{vavasis90a}), then the assumption is that all rational numbers can be represented exactly by a Turing Machine (TM).
-%This assumption has the same limitation as the assumption that TMs defined on integers are a faithful abstraction of a real computer, namely that the range of integers needing a representation is bounded.
-If there is no known proof that the problem being solved (or its decision version) is in $\mathcal{NP}$, then there are four main approaches:
-\begin{enumerate}
-\item finding a representable solution $x'$ such that $\|x'-x^\ast\|_\infty\le\varepsilon$, where $x^\ast$ is the true solution, $\varepsilon>0$ is given, and representable means having a polynomially sized description length
-% ``
-(in function of the instance size) \cite{hochbaum2};
-\item using the {\it Thom encoding} of an algebraic number \cite[Prop.~2.28]{pollack} (limited to problems involving polynomials);
-\item using the optimality gap: finding a representable solution $x'$ such that $|f(x')-f(x^\ast)|\le\varepsilon$, where $f$ is the objective function, $x^\ast$ is the true solution, $\varepsilon>0$ is given
-%, and representable is as above 
-(limited to optimization problems);\label{optgap}
-\item using a computational model according to which every elementary computation on the reals takes $O(1)$ and returns an exact result \cite[p.~24]{blum}.
-\end{enumerate}
-Approach \ref{optgap} in the list above is the one most often used in computational papers, including the present one.}
-
 %- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

### 2018-02-12 QPLIB-2.2.tex
problem types should be formatted in italics
Update QPLIB-2.2.tex

@@
 Some problem classes are actually ``too simple'' to make sense in our context.
-For instance, D*B problems have only diagonal quadratic (hence separable) objective function and bound constraints; as such, they read
+For instance, \textit{D*B} problems have only diagonal quadratic (hence separable) objective function and bound constraints; as such, they read
 \[
@@
 These can have a significant algorithmic impact, and therefore can be of interest to researchers.
-Common structures are, e.g., network flows \cite{FrFG16,FrGaSc14,FrGaSt16a,FGGP11,tadayon-smith:2013} or knapsack-type linear constraints \cite{FrFG16,FrGo13b,FGGP11}, and semi-continuous variables \cite{FrFG16,FrGaSc14,FrGaSt16a,FrGe06a,FrGe07a,FrGe09a,FGGP11}\todo{SV: do we really need 7 Frangioni et.al citations for semi-continuous variables?}, or the fact that a nonconvex quadratic objective function/constraint can be reformulated as a second-order cone (hence, convex) one \cite{FrFG16,FrGaSc14,FrGaSt16a,FrGe09a,FGGP11}\todo{SV: 5 Frangioni et.al citations for SOC?}.
+Common structures are, e.g., network flows \cite{FrFG16,FrGaSc14,FrGaSt16a,FGGP11,tadayon-smith:2013} or knapsack-type linear constraints \cite{FrFG16,FrGo13b,FGGP11}, and semi-continuous variables \cite{FrFG16,FrGaSc14,FrGaSt16a,FrGe06a,FrGe07a,FrGe09a,FGGP11},
+%\todo{SV: do we really need 7 Frangioni et.al citations for semi-continuous variables?}
+or the fact that a nonconvex quadratic objective function/constraint can be reformulated as a second-order cone (hence, convex) one \cite{FrFG16,FrGaSc14,FrGaSt16a,FrGe09a,FGGP11}.
+%\todo{SV: 5 Frangioni et.al citations for SOC?}. 
 It would be very hard to collect a comprehensive list of all types of structures that might be of interest to any individual researcher, since these are as varied as the different possible approaches for specialized sub-classes of QP.

### 2018-02-12 QPLIB-3.tex
Update QPLIB-3.tex

--- 2018-02-08 QPLIB-3.tex
@@
 Conic library CBLIB (\url{http://cblib.zib.de}) and 
-MIPLIB~2010 (\url{http://miplib.zib.de/}), to mention just a few.}\todo{AG: maybe cite CBLIB paper \cite{Friberg2016}}
+MIPLIB~2010 (\url{http://miplib.zib.de/}), to mention just a few.}
+%\todo{AG: maybe cite CBLIB paper \cite{Friberg2016} SV: then we would have to add citations for all mentioned libraries (where available)}
 
@@
 format offers three main advantages: it is easier to read by a stand-alone
-parser (provided)\todo{SV: what is meant by ``(provided)''?}, it typically produces smaller files, and it permits the
+parser, 
+%(provided)\todo{SV: what is meant by ``(provided)''?}, 
+it typically produces smaller files, and it permits the
 inclusion of two-sided inequalities without needless repetition of data.

### 2018-02-12 QPLIB-2.2.tex
update todos

@@
 These can have a significant algorithmic impact, and therefore can be of interest to researchers.
-Common structures are, e.g., network flows \cite{FrFG16,FrGaSc14,FrGaSt16a,FGGP11,tadayon-smith:2013} or knapsack-type linear constraints \cite{FrFG16,FrGo13b,FGGP11}, and semi-continuous variables \cite{FrFG16,FrGaSc14,FrGaSt16a,FrGe06a,FrGe07a,FrGe09a,FGGP11},
-%\todo{SV: do we really need 7 Frangioni et.al citations for semi-continuous variables?}
-or the fact that a nonconvex quadratic objective function/constraint can be reformulated as a second-order cone (hence, convex) one \cite{FrFG16,FrGaSc14,FrGaSt16a,FrGe09a,FGGP11}.
-%\todo{SV: 5 Frangioni et.al citations for SOC?}. 
+Common structures are, e.g., network flows \cite{FrFG16,FrGaSc14,FrGaSt16a,FGGP11,tadayon-smith:2013} or knapsack-type linear constraints \cite{FrFG16,FrGo13b,FGGP11}, and semi-continuous variables \cite{FrFG16,FrGaSc14,FrGaSt16a,FrGe06a,FrGe07a,FrGe09a,FGGP11}
+\todo{SV: reduce to 1-3 citations here},
+or the fact that a nonconvex quadratic objective function/constraint can be reformulated as a second-order cone (hence, convex) one \cite{FrFG16,FrGaSc14,FrGaSt16a,FrGe09a,FGGP11}
+\todo{SV: reduce to 1-3 citations here}.
 It would be very hard to collect a comprehensive list of all types of structures that might be of interest to any individual researcher, since these are as varied as the different possible approaches for specialized sub-classes of QP.

### 2018-02-12 QPLIB-3.tex
update todos

@@
 MIPLIB~2010 (\url{http://miplib.zib.de/}), to mention just a few.}
-%\todo{AG: maybe cite CBLIB paper \cite{Friberg2016} SV: then we would have to add citations for all mentioned libraries (where available)}
 
@@
 parser,
-%(provided)\todo{SV: what is meant by ``(provided)''?}, 
 it typically produces smaller files, and it permits the

### 2018-07-26 AnswerToTheReferees2_second.tex
Add files via upload

\documentclass[11pt]{article}
\usepackage[american]{babel}
%\usepackage{a4wide,paralist,url,xcolor,xspace}
\usepackage{a4wide,url,xcolor,xspace}
\definecolor{acblue}{rgb}{0,0,1}
\definecolor{acgreen}{rgb}{0,1,0}
\definecolor{acred}{rgb}{1,0,0}
\newcommand{\rep}[1]{{\textcolor{acblue}{#1}}}
%\newcommand{\comm}[1]{{\textcolor{acgreen}{#1}}}
\newcommand{\comm}[1]{}
\newcommand{\todo}[1]{{\textcolor{acred}{#1}}}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{booktabs}
\usepackage{multirow}

\newcommand{\eg}{e.g.,\xspace}
\newcommand{\ie}{i.e.,\xspace}

\parindent0pt
\parskip.8ex

\newcommand{\leo}[1]{{\color{red}#1}}
\newcommand{\fabio}[1]{{\color{cyan}#1}}
\newcommand{\emiliano}[1]{{\color{blue}#1}}   
\newcommand{\antonio}[1]{{\color{ForestGreen}#1}}     
\newcommand{\andrea}[1]{{\color{blue}{\bf #1}}}

\begin{document}

\section*{QPLIB: a Library of a Quadratic Programming Instances (MPC-022017-00003)-- Comments of the minor revision}

\rep{
 Following the Associate and Area Editor suggestions, we 
 performed a minor revision of our manuscript. We carefully took into account all the comments made by the AE.
 The new submitted document highlights in red the text that has been
added or changed.
In the following, the comments of the AE, together with our point-by-point replies.
}

\subsection*{Reply to the Associate Editor}

{\it
I thank the authors for their thorough revision. One of the two reviewers is now satisfied with the revision and is recommending acceptance. The other reviewer has some remaining concerns. I disagree with most of those concerns, however, and am nearly ready to recommend this paper for acceptance in MPC. I would like to ask the authors to make one additional minor revision, addressing the following minor comments (some are from the referee, others are my own):  
}

\rep{We would like to thank the AE for his/her positive opinion on the revision we performed of our manuscript. 
%\leo{Togliamo la spiegazione degli algoritmi e sostituiamola con una citazione di un book-tutorial-survey-introduction.}
}

\bigskip
\textbf{Main points:}

\begin{itemize}
\item 
{\it
[from referee, paraphrased] 8. Please try to find a more precise definition of asypmtotically complete algorithm to 
cite.
}

\rep{\dots}

\item 
{\it
[from referee, paraphrased] 9. Please add more explanation of how instances were determined to be similar. 
}

\rep{\dots}

\item 
{\it
The paragraph toward the bottom of page 19, beginning "One of the nontrivial choices..." is very nearly identical to a paragraph on page 10. Please remove this repitition. 
}

\rep{\dots}

\item Page 18, first instances filter: The way this is written it is not clear if the stated rule (30\% of complete solvers within a time limit of 30 seconds) was what was actually used, or if it an example of what might have been used. Please clarify.

\rep{\dots}

\end{itemize}

\bigskip
\textbf{Errata:}

 \begin{itemize}

\item 
{\it
- p9, l4. Please remove the word "downright" here. While I appreciate the colloquial style, I'm afraid this might actually cause some confusion. 
}

\rep{\dots}

\item 
{\it
- Please consider searching for the word "obvious" in the paper, and removing it. 
}

\rep{\dots}

\item 
{\it

- p17, l12. Something wrong with this sentence: perhaps ``to use GAMS'' should be ``we use GAMS''? 
}

\rep{\dots}

\item 
{\it

- p18, l5-6. You could perhaps leave out the part about a "discussion ensued" and simply start this section with something like: "We chose instances based on the following four goals:" 
}

\rep{\dots}

\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%55
\subsection*{Point-by-point reply to the Referee 2}

\bigskip
\textbf{
Detailed comments on this first revision: 
}

 \begin{enumerate}

\item 
{\it
 How do you propose to update the library? The criteria for problems include 
the solution time required for these problem. Presumably, as better algorithms 
are developed, these criteria change. Please comment on how the library will 
be updated. 
}

\rep{\dots}

\item 
{\it
 P.6. Please change the first sentence in Section 2.1. to be something like 
"We refer to quadratic programs (QPs) as ..." This is YOUR definition of QPs, 
which does not agree with any current textbook definitions, so you need to 
make it clear that you change QPs to means MIQCQPs, or drop QPs (better). 
}

\rep{\dots}

\item 
{\it
 The statement on p.8, that "QPs with both finite bounds cannot ever be convex" 
is wrong. Consider the set $$\{ x | 1 <= x^2 <= 2, x=1.5\}$$ which is trivially 
convex and has $Q<>0$. I would say that quadratic constraints with finite lower 
and upper bounds are typically nonconvex. 
}

\rep{\dots}

\item 
{\it
 P.9. The statement that "there is no loss of generality in assuming a linear 
objective" is somewhat inaccurate, because algorithms behave differently if 
the objective is linearized in this way. 
}

\rep{\dots}

\item 
{\it
 The fact that the library does not support matrix-free solvers is somewhat 
disappointing. Ideally, a library should be agnostic to these solver requirements, 
as AMPL is, for example. 
}

\rep{\dots}

\item 
{\it
 You must justify the statement (p.11) that QGQ "is the most general and therefore 
the most difficult class". Generality alone does not increase difficulty. Instead, 
I would add a reference to the classical result by Jeroslov that these problems 
are undecidable, which makes them more difficult than NP-hard problems. 
}

\rep{\dots}

\item 
{\it
 Add references to surveys on QP methods to your review of methods in Section 2.3. 
}

\rep{\dots}

\item 
{\it
 The definition of asymptotically complete methods in Section 2.3.3 is very sloppy, 
and must be made more precise (mathematical). Something like: for every $\epsilon>0$ 
there exist $T>0$ such that solution is $\epsilon$ optimal after time $T$ and 
$\lim_{T\to\infty}$ ... Either provide a reference to a rigorous definition, or create 
your own. 
}

\rep{\dots}

\item 
{\it
 Section 3.2 on Instance Selection is not reproducible, and somewhat sloppy. The authors 
must describe properly how the instances were selected. The current lack of transparency 
is not acceptable (it might leave contributors with the feeling that their problem was 
not selected for obscure reasons). 
}

\rep{\dots}

\item 
{\it
Please address how you decided that problems were "similar" in the second phase? 
In particular, what were the "features" in the clustering approach. If these features 
are the characteristics described earlier, then this selection is fundamentally flawed, 
because problems with the same number of variables and constraints are not even remotely 
similarly difficult to solve ... just look at the HS selection in CUTEst. 
}

\rep{\dots}

\item 
{\it
 Following on from the previous point is my most serious concern. There are seven classes 
of problems that have three or fewer instances. Some classes have only a single instance. 
Testing any algorithm on a single instance is totally inadequate. This imbalance of 
test problems must be addressed. You can either go back to your filtering and find more 
instances, or contact the people who submitted these problems, or create new ones. 
Tables 5 and 6 on p.20 should show a healthy number of test problems in each class (I 
would like to see a dozen, but I am OK with fewer, 1-3 is not OK).
}

\rep{\dots}

 \end{enumerate}

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2018-08-08 QPLIB-3.tex
Update QPLIB-3.tex

@@
 
-\textcolor{red}{
-One of the nontrivial choices in our library is that we made no effort to reformulate the instances, and inserted them in the library in the very same form as they have been provided to us by the original contributors.
-Section \ref{ssec:reform} is crucial in justifying this choice, as it shows that there are several degrees of freedom to move the instances from one class to another.
+%\textcolor{red}{
+%One of the nontrivial choices in our library is that we made no effort to reformulate the instances, and inserted them in the %library in the very same form as they have been provided to us by the original contributors.
+Section \ref{ssec:reform} is %crucial in justifying this choice, as it shows that there are several degrees of freedom to move the instances from one class %to another.
 Tailoring the structure of a problem to a solver is, however, a bias that we did not want to add.
-}
+%}
 

### 2018-08-08 QPLIB-3.tex
Update QPLIB-3.tex

@@
        of the instance under consideration, cf.~Table \ref{t:solvers}.
-       We then filtered according to a first
-       measure of computational difficulty, i.e., we discarded all instances
+       \textcolor{green}{Thanks to these extensive preliminary tests, we discarded all instances
        that are solved by at least 30\% of the complete solvers within a time
-       limit of 30 seconds.
+       limit of 30 seconds.}
+       %We then filtered according to a first
+       %measure of computational difficulty, i.e., we discarded all instances
+       %that are solved by at least 30\% of the complete solvers within a time
+       %limit of 30 seconds.
  %
@@
        according to their features; for each cluster we kept only a few
-       representatives, e.g. by eliminating all but a few of those with very similar size and coming from the same donor. Finally, in order to only
+       representatives, e.g. by eliminating all but a few of those with very similar size and coming from the same donor.
+       \textcolor{green}{The instances discarded by this second filter are instances of the same specific problem, e.g., we gathered many Max Cut Problem instances and we kept in the library only a representative small set of them.
+The selected representative instances are the larger and computationally harder ones.}
+       Finally, in order to only
        keep computationally challenging instances we ran a complete solver for \textit{QGQ} with a time limit of 120 seconds; all the

### 2018-08-08 methods.tex
Update methods.tex

@@
   \item \textit{Incomplete} methods are only able to identify solutions, often locally optimal according to a suitable notion, and may even fail to find one even when one exists; in particular, they are typically unable to determine that an instance has no solution.
-  \item \textit{Asymptotically complete} methods can find a globally optimal solution with probability one in infinite time, but again they cannot prove that a given instance is infeasible.
+  \item \textit{Asymptotically complete} methods can find a globally optimal solution with probability one in infinite time, but  they cannot prove that a given instance is infeasible \textcolor{green}{(see \S\ref{sss:asymc} below))}.
   \item \textit{Complete} methods find an approximate globally optimal solution within a prescribed optimality tolerance within finite time, or prove that none such exists (but see \S\ref{s:complete} below); they are often referred to as \emph{exact} methods in the computational optimization community.

### 2018-08-08 AnswerToTheReferees2_second.tex
Update AnswerToTheReferees2_second.tex

@@
 
-\rep{\dots}
+\rep{We think that a formal definition of these class of algorithms goes beyond the scope of the paper. We instead added a pointer to the subsection in which this class of algorithms are presented.}
 
@@
 
-\rep{\dots}
+\rep{We added a sentence to better clarify this point.}
 
@@
 
-\rep{\dots}
+\rep{The paragraph at page 19 has been removed.}
 
@@
 
-\rep{\dots}
+\rep{We have rephrased the sentence.}
 

### 2018-08-08 QPLIB-3.tex
Update QPLIB-3.tex

@@
 
-To evaluate the instances we decided, for practical reasons, to use
+To evaluate the instances we decided, for practical reasons, \textcolor{green}{we use}
 GAMS as common platform for all our final selection computations.
@@
 
-During the development of the library, a discussion ensued about
-the expected goals that we wished to achieve. The following four goals
-were finally identified:
+\textcolor{green}{We chose instances based on the following four goals:}
+%During the development of the library, a discussion ensued about
+%the expected goals that we wished to achieve. The following four goals
+%were finally identified:
 %

### 2018-08-08 QPLIB-2.2.tex
Update QPLIB-2.2.tex

@@
 \]
-is not only $\mathcal{NP}$-hard, but downright undecidable \cite{jeroslow}. Hence so are the ``harder'' \{\textit{C},\textit{Q}\}\textit{IQ}.
+is not only $\mathcal{NP}$-hard, but undecidable \cite{jeroslow}. Hence so are the ``harder'' \{\textit{C},\textit{Q}\}\textit{IQ}.
 
@@
   %
-  \item \emph{Nonconvex Binary Quadratic Programs} \textit{QB}\{\textit{B},\textit{N},\textit{L}\} obviously are $\mathcal{NP}$-hard.
+  \item \emph{Nonconvex Binary Quadratic Programs} \textit{QB}\{\textit{B},\textit{N},\textit{L}\} are $\mathcal{NP}$-hard.
 However, the special nature of binary variables combined with quadratic forms allows for quite specific techniques to be developed, one of which is the reformulation of the problem as a \textit{LBL}.

### 2018-08-08 QPLIB-2.2.tex
Update QPLIB-2.2.tex

@@
 Hence, their solution only requires the independent minimization of a convex quadratic univariate function in each single variable $x_j$ over a box constraint and possibly integrality requirements, which can be attained trivially in $O(1)$ operations (per variable) by closed-form formul{\ae}, projection and rounding arguments.
-\emph{A fortiori}, the even simpler cases \textit{L*B}, \textit{D*N} and \textit{L*N} (the latter obviously unbounded unless $b^0 = 0$) will not be discussed here.
+\emph{A fortiori}, the even simpler cases \textit{L*B}, \textit{D*N} and \textit{L*N} (the latter unbounded unless $b^0 = 0$) will not be discussed here.
 Similarly, \textit{CCN} are immediately solved by linear algebra techniques, and therefore are of no interest in this context.

### 2018-08-08 AnswerToTheReferees2_second.tex
Update AnswerToTheReferees2_second.tex

--- 2018-07-26 AnswerToTheReferees2_second.tex
@@
 
-\rep{\dots}
+\rep{We think that a formal definition of this class of algorithms goes beyond the scope of the paper. We instead added a pointer to the subsection in which the class is presented.}
 
@@
 
-\rep{\dots}
+\rep{Done.}
 
@@
 
-\rep{\dots}
+\rep{We followed the suggestion and we have revomed it.}
 
@@
 
-\rep{\dots}
+\rep{Corrected, thanks.}
 
@@
 
-\rep{\dots}
+\rep{Updated according to your suggestion.}
 

### 2018-08-08 methods.tex
Update methods.tex

--- 2018-02-08 methods.tex
@@
   \item \textit{Incomplete} methods are only able to identify solutions, often locally optimal according to a suitable notion, and may even fail to find one even when one exists; in particular, they are typically unable to determine that an instance has no solution.
-  \item \textit{Asymptotically complete} methods can find a globally optimal solution with probability one in infinite time, but again they cannot prove that a given instance is infeasible.
+  \item \textit{Asymptotically complete} methods can find a globally optimal solution with probability one in infinite time, but  they cannot prove that a given instance is infeasible \textcolor{green}{(see \S\ref{sss:asymc} below)}.
   \item \textit{Complete} methods find an approximate globally optimal solution within a prescribed optimality tolerance within finite time, or prove that none such exists (but see \S\ref{s:complete} below); they are often referred to as \emph{exact} methods in the computational optimization community.

### 2018-08-08 QPLIB-3.tex
Update QPLIB-3.tex

@@
        The goal of the second filter was to eliminate ``similar'' instances.
-       We carefully analyzed the instances one by one, and we clustered them
-       according to their features; for each cluster we kept only a few
-       representatives, e.g. by eliminating all but a few of those with very similar size and coming from the same donor.
-       \textcolor{green}{The instances discarded by this second filter are instances of the same specific problem, e.g., we gathered many Max Cut Problem instances and we kept in the library only a representative small set of them.
+       \textcolor{green}{We carefully analyzed the instances one by one,
+       %, and we clustered them
+       %according to their features; for each cluster we kept only a few
+       %representatives, e.g. 
+        eliminating all but a few of those with very similar size and coming from the same donor.
+       The instances discarded by this second filter are instances of the same specific problem, e.g., we gathered many Max Cut Problem instances and we kept in the library only a representative small set of them.
 The selected representative instances are the larger and computationally harder ones.}

### 2018-08-08 QPLIB-3.tex
Update QPLIB-3.tex

@@
        complete solver (cf.~\S\ref{sec:algo}) to solve it to
-       global optimality. Accordingly, for each of the gathered instance we
-       ran the complete solvers in GAMS, which number depends on the category
+       global optimality. Accordingly, for each of the gathered instances we
+       ran the complete solvers in GAMS, whose number depends on the category
        of the instance under consideration, cf.~Table \ref{t:solvers}.

### 2018-08-08 QPLIB-2.2.tex
Update QPLIB-2.2.tex

@@
   %Note that (\textit{D}) and (\textit{C}) are intended to mean that either $Q^i$ is positive semidefinite and $c_l^i = -\infty$, or $Q^i$ is negative semidefinite and $c_u^i = \infty$. 
-    \textcolor{red}{Note that (positive or negative) definiteness of $Q^i$ is a sufficient, but not in general necessary, condition for convexity.
+    {Note that (positive or negative) definiteness of $Q^i$ is a sufficient, but not in general necessary, condition for convexity.
 As detailed in \S\ref{subsec:final set}, in our taxonomy we mark the constraints ``$C$'' based on the sufficient condition alone, the rationale of this choice being discussed in \S\ref{ssec:reform}.}
@@
 
-\textcolor{red}{
+{
 Furthermore, as previously remarked, a not (positive or negative) definite $Q^i$ does not necessarily correspond to a nonconvex feasible region.
@@
 }
-\textcolor{red}{
+{
 One of the nontrivial choices in our library is that we made no effort to reformulate the instances, and inserted them in the library in the very same form as they have been provided to us by the original contributors.

### 2018-08-08 methods.tex
Update methods.tex

@@
 
-\textcolor{red}{We refer the interested reader to 
+{We refer the interested reader to 
 \cite{book:2077007} and 
@@
 %
-We remark that \textcolor{red}{the solvers \bonmin, \knitro, and \sbb from} the latter category can be used as incomplete solvers for \textit{QGQ}.
+We remark that {the solvers \bonmin, \knitro, and \sbb from} the latter category can be used as incomplete solvers for \textit{QGQ}.
 We also note that \lgo implements an incomplete BB algorithm for \textit{QCQ} by using bounds obtained from sampling.

### 2018-08-08 QPLIB-3.tex
Update QPLIB-3.tex

@@
 optimization and numerical analysis communities, reaching in this way
-a \textcolor{red}{large set of possibly interested} researchers and practitioners.
+a {large set of possibly interested} researchers and practitioners.
 The call remained open for ten months, during which we received a large
@@
 specific QP problems as Max-Cut, Quadratic Assignment, Portfolio
-Optimization, and several others. \textcolor{red}{In addition, we mention
+Optimization, and several others. {In addition, we mention
 that other generic libraries exist, e.g., 
@@
 
-\textcolor{red}{We now analyze the features of the instances selected to be part of the
+{We now analyze the features of the instances selected to be part of the
  library.
@@
           & \multirow{2}*{Mixed}
-                    & Convex    &   \textcolor{red}{14}\\[1.2 ex]
-          &         & Quadratic &  \textcolor{red}{134}\\[1.2 ex]
+                    & Convex    &   {14}\\[1.2 ex]
+          &         & Quadratic &  {134}\\[1.2 ex]
 \cmidrule(lr){2-4}
@@
 \multirow{1}*{Convex (if min)}
-          & Binary  & Linear    &  \textcolor{red}{5} \\[1.2 ex]
+          & Binary  & Linear    &  {5} \\[1.2 ex]
 \cmidrule(lr){2-4}
@@
 \multirow{1}*{Concave (if max)}
-          &         & Quadratic &    \textcolor{red}{6}\\[1.2 ex]
+          &         & Quadratic &    {6}\\[1.2 ex]
 %          & General & Linear    &    \\
@@
                     & None      &   23\\[1.2 ex]
-          &         & Linear    &  \textcolor{red}{91}\\[1.2 ex]
+          &         & Linear    &  {91}\\[1.2 ex]
           &         & Quadratic &   5 \\[1.2 ex]
@@
 \hline
-Total     &         &           & \textcolor{red}{319}\\
+Total     &         &           & {319}\\
 %
@@
 \multirow{2}*{Linear}    & Convex    &   13\\[1.2 ex]
-                         & Quadratic &   \textcolor{red}{52}\\[1.2 ex]
+                         & Quadratic &   {52}\\[1.2 ex]
 \cmidrule(lr){1-3}
@@
 \multirow{1}*{or}
-                         & Linear    &   \textcolor{red}{16}\\[1.2 ex]
+                         & Linear    &   {16}\\[1.2 ex]
                         % & Convex    &    2\\[1.2 ex]
 \multirow{1}*{Concave (if max)}
-                         & Quadratic &    \textcolor{red}{11}\\[1.2 ex]
+                         & Quadratic &    {11}\\[1.2 ex]
 \cmidrule(lr){1-3}
@@
                          & Linear    &   6\\[1.2 ex]
-                         & Convex    &   \textcolor{red}{3}\\[1.2 ex]
-                         & Quadratic &   \textcolor{red}{30}\\[1.2 ex]
+                         & Convex    &   {3}\\[1.2 ex]
+                         & Quadratic &   {30}\\[1.2 ex]
 \hline
-Total                    &           & \textcolor{red}{134} \\
+Total                    &           & {134} \\
 %
@@
 that the library contains a quite diverse set of instances in
-terms of number of variables and constraints.
-\textcolor{red}{The record on the maximal number of variables and constraints (both $\approx 1,000,000$) is set by the instances QPLIB\_8547 and QPLIB\_9008.}
+terms of number of variables and constraints. {The record on the maximal number of variables and constraints (both $\approx 1,000,000$) is set by the instances QPLIB\_8547 and QPLIB\_9008.}
 Figure~\ref{fig:distribution}~(right) plots the number of nonzero elements in the gradient of the objective function and the Jacobian and the number of these nonzeros corresponding to nonlinear variables, that is, it counts the appearances of variables in objectives and constraints and how often such an appearance is in a quadratic term.
@@
  percentage of ``problematic'' eigenvalues in the Hessian $Q^0$ (vertical axis), by which
-we mean \textcolor{red}{eigenvalues below $-10^{-12}$} in case of a minimization problem and \textcolor{red}{eigenvalues above $10^{-12}$} in case of a maximization problem.
+we mean {eigenvalues below $-10^{-12}$} in case of a minimization problem and {eigenvalues above $10^{-12}$} in case of a maximization problem.
 The instances are mostly clustered around two values. About 25\% of the instances have a convex (if minimization) or concave (if maximization) objective function,
@@
 Beyond the instances, the website provides a rich set of metadata for
-each instance: the three letter problem classification (as described in \S\ref{subsec:final set}), \textcolor{red}{the contributor of the instance}, basic properties such as the number of variables and constraints of
+each instance: the three letter problem classification (as described in \S\ref{subsec:final set}), {the contributor of the instance}, basic properties such as the number of variables and constraints of
 different types, the sense and convexity/concavity of the objective function, and information on the nonzero structure of the
@@
 In addition, we display a visualization of the sparsity patterns of the Jacobian and the Hessian matrix of the
-Lagrangian function, \textcolor{red}{if the instance size allows}.  In the plots of the Jacobian nonzero pattern, the linear and nonlinear entries are distinguished
+Lagrangian function, {if the instance size allows}.  In the plots of the Jacobian nonzero pattern, the linear and nonlinear entries are distinguished
 by color.  Figure \ref{fig:sparsitypattern} shows an example for instance QPLIB\_2967.
-\textcolor{red}{Finally, feasible solution points are provided for most instances.}
+{Finally, feasible solution points are provided for most instances.}
 
@@
 \caption{Example for the sparsity pattern of the Jacobian of the constraint functions (left) and
-  of \textcolor{red}{the upper-right triangle of} the Hessian of the Lagrangian function (right) for instance QPLIB\_2967.
+  of {the upper-right triangle of} the Hessian of the Lagrangian function (right) for instance QPLIB\_2967.
   The gradient of the objective function is displayed as the first row
-  of the Jacobian matrix. \textcolor{red}{Non-constant entries are shown in red.}}
+  of the Jacobian matrix. {Non-constant entries are shown in red.}}
 \label{fig:sparsitypattern}
@@
 The entire set of instances can be explored in a searchable and sortable table of selected instance features: problem
-classification, convexity of the continuous relaxation, number of (all, binary, integer) variables, (all, quadratic) constraints, nonzeros, \textcolor{red}{problematic} eigenvalues in~$Q^0$, and density of~$Q^0$.
+classification, convexity of the continuous relaxation, number of (all, binary, integer) variables, (all, quadratic) constraints, nonzeros, {problematic} eigenvalues in~$Q^0$, and density of~$Q^0$.
 %

### 2018-08-08 QPLIB-3.tex
Update QPLIB-3.tex

@@
 \begin{table}
-\textcolor{red}{
+{
  \centering
@@
 instances among the different categories. Indeed, the discrete categories
-\textit{LMQ} and \textit{QBL} are well represented by \textcolor{red}{134} and \textcolor{red}{91}
+\textit{LMQ} and \textit{QBL} are well represented by {134} and {91}
 instances, respectively. Similarly, the continuous categories
 \textit{LCQ} and \textit{QCQ} are well
-represented by \textcolor{red}{52} and \textcolor{red}{30} instances, respectively. Moreover, the library
+represented by {52} and {30} instances, respectively. Moreover, the library
 actually covers the large majority of all possible categories of instances.

### 2018-08-10 QPLIB-2.2.tex
reduce references to common structures

@@
 These can have a significant algorithmic impact, and therefore can be of interest to researchers.
-Common structures are, e.g., network flows \cite{FrFG16,FrGaSc14,FrGaSt16a,FGGP11,tadayon-smith:2013} or knapsack-type linear constraints \cite{FrFG16,FrGo13b,FGGP11}, and semi-continuous variables \cite{FrFG16,FrGaSc14,FrGaSt16a,FrGe06a,FrGe07a,FrGe09a,FGGP11}
-\todo{SV: reduce to 1-3 citations here},
-or the fact that a nonconvex quadratic objective function/constraint can be reformulated as a second-order cone (hence, convex) one \cite{FrFG16,FrGaSc14,FrGaSt16a,FrGe09a,FGGP11}
-\todo{SV: reduce to 1-3 citations here}.
+Common structures are, e.g., network flows \cite{FrFG16,FrGaSc14,tadayon-smith:2013} or knapsack-type linear constraints \cite{FrFG16,FGGP11}, and semi-continuous variables \cite{FrGaSc14,FrGaSt16a,FGGP11},
+or the fact that a nonconvex quadratic objective function/constraint can be reformulated as a second-order cone (hence, convex) one \cite{FrGaSt16a,FrGe09a,FGGP11}.
 It would be very hard to collect a comprehensive list of all types of structures that might be of interest to any individual researcher, since these are as varied as the different possible approaches for specialized sub-classes of QP.

### 2018-08-18 QPLIB-1.0.tex
update todos
Section 1 typographical suggestions

@@
 This is not surprising, given the vast scope of practical applications of such problems, and of solution methods designed to solve them \cite{GoulToin00a}.
-Depending on the specifics of the formulation, solving a QP may require primarily combinatorial techniques, ideas rooted in nonlinear optimization principles, or a mix of the two.
-In this sense, QP is arguably one of the classes of problems where collaboration between the communities interested in combinatorial and in nonlinear optimization is most needed, and potentially fruitful.
+Depending on the formulation specifics, solving a QP may require primarily combinatorial techniques, ideas rooted in nonlinear optimization principles, or a mix of the two.
+In this sense, QP is a class of problems where collaboration between the communities interested in combinatorial and in nonlinear optimization is potentially fruitful.
 
@@
 Therefore, it is perhaps not surprising that, unlike for ``simpler'' problems classes like Mixed-Integer Linear Programming \cite{Koch2011},  there has been no single library devoted to all different kinds of instances of QP.
-%\todo{AG: many commas in this sentence SV: I removed one.}\ 
-While several specialized libraries devoted to particular cases of QP are available, each of them is either focused on a particular application (a specific problem that can be modeled as a QP), or on QPs with specific structural properties that make them suitable for solution by some given class of algorithmic approaches.
+While several specialized libraries devoted to particular classes of QP are available, each of them is either focused on a particular application (a specific problem that can be modeled as a QP), or on QPs with specific structural properties that make them suitable for solution by some given class of algorithmic approaches.
 To the best of our knowledge, collecting a set of QP instances that is at the same time not overwhelmingly onerous but sufficiently challenging for the many different interested communities has not been attempted.
@@
 
-In this paper, we report the steps that have been done to collect what we consider to be a quality library of QP instances, filtering a much larger set of currently available (or specifically provided) instances in order to end up with a manageable set that still contains a meaningful sample of possible QP types.
+This paper reports our steps towards collecting what we consider to be a quality QP instance library, filtering a much larger set of currently available (or specifically provided) instances and proposing a manageable set that still contains a meaningful sample of possible QP types.
 A particularly thorny issue in this process was how to select instances that are ``interesting''.
@@
 Putting this approach into practice requires a nontrivial number of technical steps and decisions that are detailed in the paper.
-We hope that our work can provide useful guidelines for other researchers interested in the constructions of benchmarks for mathematical optimization problems.
+We hope that our work can provide useful guidelines for other researchers interested in constructing benchmarks for mathematical optimization problems.
 
 A consequence of our focus is that this paper is \emph{not} concerned with the performance of the very diverse available set of QP solvers; we will \emph{not} report any data comparing them.
-The only reason that solvers are used (and, therefore, described) in this context is to ensure that the instances of the library are nontrivial, at least for a significant fraction of the current solution methods.
+The only reason that solvers are used (and, therefore, described) in this context is to ensure that the library instances are nontrivial, at least for a significant fraction of the current solution methods.
 Providing guidance about which solvers are most suited to some specific class of QPs is entirely outside the scope of our work.

### 2018-08-18 QPLIB-3.tex
fix commented text
Section 1 typographical suggestions

@@
 
-%\textcolor{red}{
-%One of the nontrivial choices in our library is that we made no effort to reformulate the instances, and inserted them in the %library in the very same form as they have been provided to us by the original contributors.
-Section \ref{ssec:reform} is %crucial in justifying this choice, as it shows that there are several degrees of freedom to move the instances from one class %to another.
+%One of the nontrivial choices in our library is that we made no effort to reformulate the instances, and inserted them in the library in the very same form as they have been provided to us by the original contributors.
+Section \ref{ssec:reform} is crucial in justifying this choice, as it shows that there are several degrees of freedom to move the instances from one class to another.
 Tailoring the structure of a problem to a solver is, however, a bias that we did not want to add.
-%}
 
@@
 of the instances. In Figure~\ref{fig:distribution}~(left) we plot the number
-of variables (horizontal axis) versus the number of constrains
+of variables (horizontal axis) versus the number of constraints
 (vertical axis), both in logarithmic scale. Continuous instances are

### 2018-08-18 QPLIB.tex
Section 1 typographical suggestions

@@
 \begin{abstract}
- This paper describes a new instance library for Quadratic Programming (QP), i.e., the family of continuous and (mixed)-integer optimization problems where the objective function, the constrains, or both are quadratic.
- QP is a very diverse class of problems, comprising sub-classes of problems ranging from
- trivial to undecidable. This diversity is reflected in the variety of solution methods for QP, ranging from
- entirely combinatorial ones to completely continuous ones, including many for
+ This paper describes a new instance library for Quadratic Programming (QP), i.e., the family of continuous and (mixed)-integer optimization problems where the objective function and/or the constraints are quadratic.
+ QP is a very diverse class of problems, comprising sub-classes ranging from
+ trivial to undecidable. This diversity is reflected in the variety of QP solution methods, ranging from
+ entirely combinatorial approaches to completely continuous algorithms, including many methods for
  which both aspects are fundamental. Selecting a set of instances of QP that is
  at the same time not overwhelmingly onerous but sufficiently challenging for the
- many different interested communities is therefore important. We propose a simple
- taxonomy for QP instances that leads to a systematic problem selection mechanism.
+ different, interested communities is therefore important. We propose a simple
+ taxonomy for QP instances leading to a systematic problem selection mechanism.
  We then briefly survey the field of QP, giving an overview of theory, methods and

### 2018-08-18 QPLIB-1.0.tex
More typographical items

@@
 
-Optimization problems with quadratic constraints and/or objective function (QP) have been the subject of a considerable amount of research over the last almost seventy years.
-At least some of the rationale for this interest is likely due to the fact that QPs are the ``least-nonlinear nonlinear problems''.
+Optimization problems with quadratic constraints and/or objective function (QP) have been the subject of a considerable amount of research for almost seventy years.
+Part of the rationale for this interest is that QPs are the ``least-nonlinear nonlinear problems''.
 Hence, in particular for the convex case, tools and techniques that have been honed during decades of research for Linear Programming (LP), typically with integrality constraints (MILP), can often be extended to the quadratic case more easily than would be required to tackle general (Mixed-Integer) Nonlinear Programming ((MI)NLP) problems.
-This has indeed happened over-and-over again with different algorithmic techniques, such as interior-point methods, active-set methods (of which the simplex method is a prototypical example), enumeration methods, cut-generation techniques, reformulation techniques, and many others \cite{BS09}.
+This has indeed happened over-and-over again with different algorithmic techniques, such as interior-point methods, active-set methods, e.g., the simplex method, enumeration methods, cut-generation techniques, reformulation techniques, and many others \cite{BS09}.
 Similarly, nonconvex continuous QP is perhaps the ``simplest'' class of problems that require features such as spatial enumeration techniques for their solution.
-Hence, QPs are both a natural basis for the development of general techniques for nonconvex NLP, and a very specific class so that specialized approaches can be developed \cite{Dur2010,Burer2012}.
+Hence, QPs are both a natural basis for developing general techniques for nonconvex NLP, and a very specific class enabling the development of specialized approaches \cite{Dur2010,Burer2012}.
 
@@
 Quadratic expressions are found, either naturally or after appropriate reformulations, in very many optimization problems \cite{Kochenberger2014}.
-Table \ref{tbl:application} provides a certainly non-exhaustive collection of applications that lead to formulations with quadratic constraints, quadratic objective function, or both.
+Table \ref{tbl:application} provides a non-exhaustive collection of applications that lead to formulations with quadratic constraints, quadratic objective function, or both.
 In general, any continuous function can be approximated with arbitrary accuracy (over a compact set) by a polynomial of arbitrary degree.
@@
 In \S\ref{sec:QPbasic} we review the basic notion of QP.
-In particular, \S\ref{subsec:notation} sets out the notation, \S\ref{sec:classification} proposes a new taxonomy of QP that helps us in discussing the (very) different classes of QPs, and \S\ref{sec:algo} very briefly reviews the solution methods for QP and the solvers we have employed.
+In particular, \S\ref{subsec:notation} sets out the notation, \S\ref{sec:classification} proposes a new QP taxonomy that helps discuss the (very) different QP classes, and \S\ref{sec:algo} very briefly reviews the QP solution methods and the solvers we have employed.
 Next, \S\ref{sec:lib} describes the process used to obtain the library and its results.

### 2018-08-18 QPLIB-2.1.tex
More typographical items

@@
 %
-Due to the presence of integrality requirements on the variables and of quadratic constraints, this class of problems is often referred to as Mixed-Integer Quadratically Constraint Quadratic Program (MIQCQP).
-It will be sometimes useful to refer to the (sub)set $\Mcb =  \{ \, j \in \Mcz : l_j = 0 , u_j = 1 \, \} \subseteq \Mcz$ of the binary variables, and to $\Mcr = \Mcn \setminus \Mcz$ as the set of continuous ones.
-Similarly, it will be sometimes useful to distinguish the (sub)set $\Mcx = \{ \, j : l_j > -\infty \lor u_j < \infty \, \}$ of the box-constrained variables from $\Mcu = \Mcn \setminus \Mcx$ of the unconstrained ones (in the sense that finite bounds are not explicitly provided in the data of the problem, although they may be implied by the other constraints).
+Due to the quadratic constraints and the integrality requirements on the variables, this class is often referred to as Mixed-Integer Quadratically Constraint Quadratic Program (MIQCQP).
+It will be sometimes useful to refer to the (sub)set $\Mcb =  \{ \, j \in \Mcz : l_j = 0 , u_j = 1 \, \} \subseteq \Mcz$ of the binary variables, and to $\Mcr = \Mcn \setminus \Mcz$ as the set of continuous variables.
+Similarly, it will be sometimes useful to distinguish the (sub)set $\Mcx = \{ \, j : l_j > -\infty \lor u_j < \infty \, \}$ of the box-constrained variables from $\Mcu = \Mcn \setminus \Mcx$ of the unconstrained ones (in the sense that finite bounds are not explicitly provided in the problem data, although bounds may be implied by the other constraints).
 
-The relative flexibility offered by quadratic functions, as opposed e.g.~to linear ones, allows several reformulation techniques to be applicable to this family of problems in order to emphasize different properties of the various components.
+The relative flexibility offered by quadratic functions, as opposed, e.g., to linear ones, allows several reformulation techniques to be applicable to this family of problems in order to emphasize different properties of the various components.
 Some of these reformulation techniques will be commented later on; here we remark that, for instance, integrality requirements, in particular in the form of binary variables could always be ``hidden'' by introducing (nonconvex) quadratic constraints utilizing the celebrated relationship $x_j \in \{0, 1\} \iff x_j^2 = x_j$.
-Therefore, when discussing these problems some effort has to be made to distinguish between features that come from the original model, and those that can be introduced by reformulation techniques in order to extract (and algorithmically exploit) specific properties.
+Therefore, when discussing these problems, some effort has to be made to distinguish between features that come from the original model, and those that can be introduced by reformulation techniques in order to extract (and algorithmically exploit) specific properties.
 

### 2018-08-18 QPLIB-2.2.tex
More typographical items

@@
 
-Despite the apparent simplicity of the definition given in \S\ref{subsec:notation}, Quadratic Programming instances can be of several rather different ``types'' in practice, depending on fine details of the data.
-In particular, many algorithmic approaches can only be applied to QP when the data of the problem has specific properties.
-A taxonomy of QP instances should thus strive to identify the set of properties that an instance should have in order to apply the most relevant computational methods.
+Despite the apparent simplicity of the \S\ref{subsec:notation} definition, Quadratic Programming instances can be of several rather different ``types'' in practice, depending on fine details of the data.
+In particular, many algorithmic approaches can only be applied to QP when the problem data has specific properties.
+A taxonomy of QP instances should thus strive to identify a set of properties that an instance should have in order to apply the most relevant computational methods.
 However, the sheer number of different existing approaches, and the fact that new ones are frequently proposed, makes it hard to provide a taxonomy that is both simple and covers all possible special cases.
@@
 
-The ordering of the values in the previous lists is not irrelevant; in general, problems become ``harder'' when going from left to right.
+The ordering in the preceding lists is relevant; in general, problems become ``harder'' when going from left to right.
 More specifically, for the \textit{O} and \textit{C} fields the order is that of \emph{strict} containment between problem classes: for instance, linear objective functions are strictly a special case of diagonal convex quadratic ones (by allowing the diagonal elements all to be zero), the latter are a strict subset of general convex quadratic objectives (by allowing the off-diagonal elements all to be zero), and these are strictly subsets of general nonconvex quadratic ones (since these permit any symmetric Hessian including positive semidefinite ones).
@@
 hold.
-In the following discussion we will repeatedly exploit this by assuming that, unless otherwise mentioned, when a method can be applied to a given problem, it can be applied as well to all simpler problems where the value of each field is arbitrarily replaced with a value denoting a less-general class.
+The following discussion repeatedly exploits this ordering by assuming that, unless otherwise mentioned, when a method can be applied to a given problem, it can also be applied to all simpler problems where the value of each field is arbitrarily replaced with a value denoting a less-general class.
 
@@
 We now give a general discussion about the different problem classes that our proposed taxonomy defines.
-For simplicity, we will assume minimization problems for the remaining of this section.
+For simplicity, this section assumes minimization problems.
 Some problem classes are actually ``too simple'' to make sense in our context.

### 2018-08-18 QPLIB-3.tex
More typographical items

@@
 
-In this section we present all the steps we performed in order to build
+This section presents all the steps we performed to build
 the new instance library. In \S\ref{subsec:instColl}, we describe the set
@@
 
-In this section we describe the procedure we adopted to gather the
+This section describes the procedure we adopted to gather the
 instances. In January $2014$, we issued an online call for instances
@@
 
-In addition to these spontaneous contributions we analyzed existing generic
-libraries of instances available  on the internet that contain QP
-instances. Namely, the libraries from which we gathered instances are
+In addition to these spontaneous contributions, we analyzed existing generic
+instance libraries available  containing QP
+instances. The libraries from which we gathered instances are
 %
@@
 
-At the end of this process we had gathered more than eight thousand
+At the end of this process, we had gathered more than eight thousand
 instances. Three quarters of them contained discrete variables, while
@@
  library.
-In Table \ref{tab:DDDDD}, we provide a global overview.
+Table \ref{tab:DDDDD} provides a global overview.
 The instances have been divided in \textit{continuous} vs \textit{discrete} and \textit{convex} vs \textit{nonconvex}, forming in this way, a  classification of 4 macro categories.

### 2018-08-18 methods.tex
Update methods.tex
More typographical items

@@
 
-In this section we provide a quick overview of existing solution methods for QP, restricting ourselves to these implemented by the set of solvers considered in this paper (see \S\ref{subsec:solver}).
-For each approach we briefly describe  the formulation they address according to the classification set out in \S\ref{sec:classification}.
-We remark that many solvers implement more than one algorithm, which the user can choose at runtime.
+This section provides a quick overview of existing solution methods for QP, restricting ourselves to these implemented by the specific solvers considered in this paper (see \S\ref{subsec:solver}).
+For each approach, we briefly describe the formulation they address according to the \S\ref{sec:classification} classification.
+Many solvers implement more than one algorithm, which the user can choose at runtime.
 Moreover, algorithms are typically implemented in different ways within different solvers, so that the same conceptual algorithm can sometimes yield different results or performance measures on the same instances.
@@
 
-The solvers in Table \ref{t:solvers} which implement incomplete methods for NLPs (a problem class containing \textit{QCQ}) are \conopt, \ipopt, \minos, \snopt, and \knitro.
-Note that all these solvers tackle the more general class of NLP, while we use them only for the considerably more restricted class of QP.
+The Table \ref{t:solvers} solvers which implement incomplete methods for NLPs (a problem class containing \textit{QCQ}) are \conopt, \ipopt, \minos, \snopt, and \knitro.
+Note that all these solvers tackle the more general class of NLP, while we use them only for the considerably more restricted QP class.
 Aside from solvers provided by GAMS, there are a number of other, specialized, incomplete QP solvers, such as
@@
 %
- \item \textcolor{red}{\gurobi and \xpress for \textit{QBC};}
+ \item {\gurobi and \xpress for \textit{QBC};}
 %

### 2018-08-23 methods.tex
Update methods.tex

@@
 
-Asymptotically complete methods do not usually require a starting point, and, if given sufficient time (infinite in the worst case) will identify a globally optimal solution with probability one.
+\textcolor{green}{An asymptotically complete method reaches a global minimum with certainty or
+at least with probability one if allowed to run indefinitely long, but has no means to know when a global minimizer has been found (see \cite{neumaier}).}
+%
+%Asymptotically complete methods do not usually require a starting point, and, if given sufficient time (infinite in the worst case) will identify a globally optimal solution with probability one. 
+%
 Most often, these methods are meta-heuristics, involving an element of random choice, which exploit a given (heuristic) local search procedure.

### 2018-08-23 AnswerToTheReferees2_second.tex
Update AnswerToTheReferees2_second.tex

--- 2018-07-26 AnswerToTheReferees2_second.tex
@@
 
-\rep{\dots}
+\rep{We have added the definition in paragraph 2.3.3 and it is taken from: 
+\begin{itemize}
+\item Neumaier, A.: Complete search in continuous global optimization and
+constraint satisfaction. Acta Numerica 13, 271–369 (2004)
+\end{itemize}
+}
 
@@
 
-\rep{\dots}
+\rep{We have better clarified this point.}
 
@@
 
-\rep{\dots}
+\rep{Updated according to your suggestion.}
 
 \end{itemize}
+
+\end{document}
 

### 2018-08-24 QPLIB-3.tex
proofreading, mainly Section 3

@@
 To achieve such goals, we performed the following two filters, applied in
-a cascade.
+a cascade:
 %
 \begin{itemize}
- \item \emph{First Instances Filter.}\\
+ \item \emph{First Instance Filter.}\\
        The first filter was designed to drastically reduce the number of
@@
  %
- \item \emph{Second Instances Filter.}\\
+ \item \emph{Second Instance Filter.}\\
        The goal of the second filter was to eliminate ``similar'' instances.
@@
 \begin{tabular}{cccc}
-Starting set& \multicolumn{ 2}{c}{ $\approx$ 8500 Instances }& \\
-& \multicolumn{ 2}{c}{$\Downarrow$}& \\
-& $\approx$ 6000 Discr. Inst.  & $\approx$ 2500 Cont. Inst. & \\
-First Filter  & $\Downarrow$  & $\Downarrow$ & \\
- & $\approx$ 3000 Discr. Inst.  & $\approx$ 1000 Cont. Inst. & \\
-Second Filter & $\Downarrow$  & $\Downarrow$  & \\
+Starting set& \multicolumn{ 2}{c}{ $\approx$ 8500 instances }& \\
+& \multicolumn{ 2}{c}{\rotatebox{-45}{$\Downarrow$} \quad \raisebox{-1.2ex}{\rotatebox{45}{$\Downarrow$}}}& \\
+& $\approx$ 6000 discr. inst.  & $\approx$ 2500 cont. inst. & \\
+First filter  & $\Downarrow$  & $\Downarrow$ & \\
+ & $\approx$ 3000 discr. inst.  & $\approx$ 1000 cont. inst. & \\
+Second filter & $\Downarrow$  & $\Downarrow$  & \\
 % & 600 Discr. Inst.  & 250  Cont. inst. & \\
-  & 319 Discr. Inst.  & 134  Cont. inst. & \\
+  & 319 discr. inst.  & 134  cont. inst. & \\
 \end{tabular}
@@
 The instances have been divided in \textit{continuous} vs \textit{discrete} and \textit{convex} vs \textit{nonconvex}, forming in this way, a  classification of 4 macro categories.
-As previously mentioned, an instance is classified \textit{discrete} if it contains at least a binary or integer variable, and \textit{continuous} otherwise.
+As previously mentioned, an instance is classified \textit{discrete} if it contains at least one binary or integer variable, and \textit{continuous} otherwise.
 On the other hand, an instance is classified as \textit{nonconvex} if the objective function is nonconvex (if minimization) or nonconcave (if maximization) and/or at least one of the constraints is nonconvex, and \textit{convex} otherwise.
@@
 For each category, the tables
-report in column ``$\#$'' the corresponding number of instances. It can be seen
+report the corresponding number of instances in column ``$\#$''. It can be seen
 that the final set well respects the original distribution of the gathered
@@
 (vertical axis), both in logarithmic scale. Continuous instances are
-marked with ``$+$'', and discrete ones with ``$\times$''. The figure shows
+marked with ``$+$'' and discrete ones with ``$\times$''.
+\textcolor{green}{Box constraints are not counted as constraints.} \todo{Ambros: From text, tables, and Figure~1 did I correctly infer that variable bounds are not counted as constraints?
+The 23 ``none'' in the discrete table seem to be the blue crosses on the x-axis, but then where do the 3 red ``+'' on the x-axis come from?
+Are these continuous instances with box constraints and some boxes are infinite?
+Should there be a ``none'' in the continuous table?}
+The figure shows
 that the library contains a quite diverse set of instances in

### 2018-08-24 QPLIB-3.tex
remove todo

@@
 marked with ``$+$'' and discrete ones with ``$\times$''.
-\textcolor{green}{Box constraints are not counted as constraints.} \todo{Ambros: From text, tables, and Figure~1 did I correctly infer that variable bounds are not counted as constraints?
-The 23 ``none'' in the discrete table seem to be the blue crosses on the x-axis, but then where do the 3 red ``+'' on the x-axis come from?
-Are these continuous instances with box constraints and some boxes are infinite?
-Should there be a ``none'' in the continuous table?}
+\textcolor{green}{Box constraints are not counted as constraints.}
 The figure shows

### 2018-08-28 AnswerToTheReferees2_second.tex
Update AnswerToTheReferees2_second.tex

@@
  performed a minor revision of our manuscript. We carefully took into account all the comments made by the AE.
- The new submitted document highlights in red the text that has been
+ The new submitted document highlights in green the text that has been
 added or changed.
@@
 
-\rep{We have added the definition in paragraph 2.3.3 and it is taken from: 
+\rep{We have added the definition of asypmtotically complete algorithms to Section 2.3.3. The definition has been taken from: 
 \begin{itemize}
